### 2016-12-24 abstractShort.tex
maintexfiles added

Optimization problems involving uncertainties are common in a variety
of engineering disciplines such as transportation systems, manufacturing,
communication networks, healthcare and finance. The large number of input
variables and the lack of a system model prohibit a precise analytical
solution and a viable alternative is to employ simulation-based
optimization. The idea here is to simulate a few times the stochastic
system under consideration while updating the system parameters until a
good enough solution is obtained.

 Formally, given only noise-corrupted measurements of an
objective function, we wish to find a parameter which minimises the
objective function. Iterative algorithms using statistical methods search
the feasible region to improve upon the candidate parameter. Stochastic
approximation algorithms are best suited, most studied and applied
algorithms for finding solutions when the feasible region is a
continuously valued set. One can use information on the gradient/Hessian
of the objective to aid the search process. However, due to the lack of
knowledge of the noise distribution, one needs to estimate the
gradient/Hessian from noisy samples of the cost function obtained from
simulation. Simple gradient search schemes take many iterations to
converge to a local minimum and are heavily dependent on the choice of
step-sizes. Stochastic Newton methods, on the other hand, can counter the
ill-conditioning of the objective function as they incorporate
second-order information into the stochastic updates. Stochastic Newton
methods are often more accurate than simple gradient search schemes.

We propose enhancements to the Hessian estimation scheme used in two recently proposed stochastic Newton methods, based on the ideas of random directions stochastic approximation (2RDSA) \cite{prashanth2015rdsa} and simultaneous perturbation stochastic approximation (2SPSA-3\footnote{The 3 in the abberviation of the algorithm is used to indicate that the algorithms in \cite{bhatnagar2015simultaneous} require 3 function evaluations per iteration.}) \cite{bhatnagar2015simultaneous}, respectively. The proposed scheme, inspired by \cite{spall-jacobian}, reduces the error in the Hessian estimate by 
\begin{inparaenum}[\bfseries (i)]
	\item incorporating a zero-mean feedback term; and
	\item optimizing the step-sizes used in the Hessian recursion.
\end{inparaenum}
We prove that 2RDSA and 2SPSA-3 with our Hessian improvement scheme converges asymptotically to the true Hessian.
 %Moreover, for the special of a quadratic objective, we provide a convergence rate result for 2RDSA-IH that is of the same order - in the number of iterations of Hessian recursion - as that of 2SPSA with improved Hessian estimation (2SPSA-IH) \cite{spall-jacobian}. 
The advantage with 2RDSA and 2SPSA-3  is that it requires only 75\% of the simulation cost per-iteration for 2SPSA with improved Hessian estimation (2SPSA-IH) \cite{spall-jacobian}.
Numerical experiments show that 2RDSA-IH outperforms both 2SPSA-IH and 2RDSA without the improved Hessian estimation scheme.

### 2016-12-24 ack.tex
maintexfiles added

I take this opportunity to express my profound gratitude to the people who have helped and supported me throughout my PhD. I thank my guide, Prof. Shalabh Bhatnagar, for his constant and valuable suggestions during my research. Without his constant guidance and support, this thesis would not have been possible. Next I thank Prof. Borkar for several interesting and useful discussions on our collaborative work. I thank Sunil, Hemanth and Sindhu for enthusiastically collaborating on the research problems. I thank Prof. Narahari, Electrical Sciences Divisional Chairman, who has always been keen to know how my thesis work is progressing. I am grateful to all the professors of CSA Dept for sharing their knowledge and for their encouragement.  I, also, thank the CSA Dept office staff for their invaluable support.  I would like to particularly acknowledge Mrs. Suguna and Mrs. Lalitha who promptly helped me on many occasions. Last but not the least, I express my heartfelt gratitude to the Almighty, my parents, younger sister, relatives and friends for their love and blessings.

### 2016-12-24 conc.tex
maintexfiles added

\chapter{Conclusions and Future work}\label{conc}
\label{sec:conclusions}
We presented an improved Hessian estimation scheme for the 2RDSA algorithm \cite{prashanth2015rdsa}. The proposed scheme was shown to be provably convergent to the true Hessian. We proposed generalisation of RDSA algorithm proposed in \cite{prashanth2015rdsa}. As a result of generalisation we are able to use much larger class of distributions  for perturbation random variables. The proposed generalisation inherits all the improved Hessian estimation  schemes proposed earlier. We also presented an improved Hessian estimation scheme for 2SPSA-3 algorithm \cite{bhatnagar2015simultaneous} and as a special case, for a quadratic objective resulted in a  convergence rate that is on par with the corresponding rate for 2SPSA with Hessian estimation improvements (2SPSA-IH) \cite{spall-jacobian}. The advantage with 2RDSA-IH, 2SPSA-3-IH is that it requires only 75\% of the simulation cost per-iteration for 2SPSA with Hessian estimation improvements (2SPSA-IH) \cite{spall-jacobian}. Numerical experiments demonstrated that 2RDSA-IH, 2SPSA-3-IH outperforms both 2SPSA-IH and 2SPSA without the improved Hessian estimation scheme. They also indicate that schemes with improved Hessian estimation outperforms their respective regular schemes without the improved Hessian estimation scheme. 2RDSA-IH with asymmetric Bernoulli distribution performing the best overall.

As future work, it would be interesting to derive finite time bounds that show a lower Hessian estimation error for 2RDSA-IH when compared to 2RDSA and 2SPSA. Stochastic Newton methods  methods converge faster and  are often more accurate than simple gradient search schemes. However they are computationally much costlier than first-order methods. Since first-order methods exhibits slower convergence rate and heavily depended on the step-size selection one has to resort to second-order methods. As a future work, it would be interesting to look at the methods like momentum descent, conjugate gradient methods, Hessian free optimization schemes which result in faster convergence for the first-order methods in deterministic optimization case. It would be interesting to extend those methods to stochastic optimization case and prove that they exhibits  faster convergence than first-oder methods. From application point of view it would be interesting to apply these methods to some interesting real world applications.

### 2016-12-24 experiments.tex
maintexfiles added

\chapter{Simulation experiments}
\label{sec:expts}
\section{Implementation}
We test the performance of 2RDSA-Unif, 2RDSA-AsymBer and 2SPSA, with/without improved Hessian estimation. 
2SPSA algorithm uses Bernoulli $\pm 1$-valued perturbations, while 2RDSA/2RDSA-IH come in two variants - one that uses $U[-1,1]$ distributed perturbations (referred to as 2RDSA-Unif/2RDSA-IH-Unif) and the other that uses asymmetric Bernoulli perturbations (referred to as 2RDSA-AsymBer/2RDSA-IH-AsymBer)\footnote{The implementation is available at \url{https://github.com/prashla/RDSA/archive/master.zip}.}.

For the empirical evaluations, we use the following two loss functions in $N=10$ dimensions:
\paragraph{Quadratic loss}
\begin{align}
f(x) = x\tr A x + b\tr x.\label{eq:quadratic}
\end{align} 
The optimum $x^*$ for the above $f$ is such that each coordinate of $x^*$ is $-0.9091$, with $f(x^*) = -4.55$.

\paragraph{Fourth-order loss}
\begin{align} 
f(x) = x\tr A\tr A x + 0.1 \sum_{j=1}^N (Ax)^3_j + 0.01 \sum_{j=1}^N (Ax)^4_j.\label{eq:4thorder}
 \end{align} 
The optimum $x^*$ for above $f$ is $x^*=0$, with $f(x^*) = 0$.

In both functions, $A$ is such that $NA$ is an upper triangular matrix with each entry one, $b$ is the $N$-dimensional vector of ones and the noise structure is similar to that used in \cite{spall_adaptive}. For any $x$, the noise is $[x\tr, 1]z$, where $z \approx \N(0,\sigma^2 I_{11\times11})$. We perform experiments for noisy as well as noise-less settings, with $\sigma=0.1$ for the noisy case.

For all algorithms, we set $\delta_n = 3.8/n^{0.101}$ and $a_n = 1/n^{0.6}$, while $b_n$ are set according to \eqref{eq:wieghts}. These choices have been used  for 2SPSA implementations before (see \cite{spall_adaptive}) and have demonstrated good finite-sample performance empirically, while satisfying the theoretical requirements needed for asymptotic convergence.  For all the algorithms, the initial point $x_0$ is the $N$-dimensional vector of ones.  For both 2SPSA and 2RDSA/2RDSA-IH, an initial $20\%$ of the  simulation budget was used up by 1SPSA/1RDSA and the resulting iterate was used to initialize 2SPSA/2RDSA. The distribution parameter $\epsilon$ is set to $0.0001$ for 2RDSA and to $0.01$ for 1RDSA.

\section{Results}
We use normalized loss and normalized MSE (NMSE) as performance metrics for evaluating the algorithms. 
NMSE is the ratio $\l x_{n_\text{end}} - x^* \r^2 / \l x_0 - x^*\r^2$, while normalized loss is the ratio $f(x_{n_\text{end}})/f(x_0)$.  Here $n_\text{end}$ denotes the iteration number when the algorithm stopped updating its parameter. Note that $n_\text{end}$ is a function of the simulation budget. 2RDSA/2RDSA-IH use only three simulations per-iteration and hence, $n_\text{end}$ is $1/3$rd of the simulation budget, while it is $1/4$th of the simulation budget for 2SPSA, since the latter algorithm uses four simulations per-iteration.

\begin{table}
\centering
 \caption{Normalized loss values for fourth-order  objective \eqref{eq:4thorder} with and without noise: standard error from $500$ replications shown after $\pm$}
\label{tab:norloss-4thf}
\begin{tabular}{|c|c|c|}
\toprule
\rowcolor{gray!20}
\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0.1$}}}\\[1em]
%&&&\\
\midrule
  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
  & & \textbf{ estimation} \\
 \midrule
\textbf{2SPSA} & $0.132 \pm 0.0267$ & $0.104 \pm 0.0355$\\
&&\\
\textbf{2RDSA-Unif} &$0.115 \pm 0.0214$ & $0.0271 \pm 0.0538$\\ 
&&\\
\textbf{2RDSA-AsymBer}& $0.0471 \pm 0.021$& $\bm{0.0099 \pm 0.0014}$\\
 \bottomrule
 %%%%%%%%%%%%2nd ordr
\rowcolor{gray!20}
\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0$}}}\\[1em]
%&&&\\
\midrule
  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
  & & \textbf{ estimation} \\
 \midrule
\textbf{2SPSA} & $0.0795 \pm 0.0234$ & $0.0628 \pm 0.0234$\\
&&\\
\textbf{2RDSA-Unif} &$0.0813 \pm 0.0275$ & $0.0214 \pm 0.00376$\\ 
&&\\
\textbf{2RDSA-AsymBer}& $0.0199 \pm 0.0114$& $\bm{0.0098 \pm 0.00147}$\\
 \bottomrule
\end{tabular}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%% Quadratic
\begin{table}
\centering
 \caption{Normalized loss values for quadratic objective \eqref{eq:quadratic} with and without noise: standard error from $500$ replications shown after $\pm$}
\label{tab:norloss-quadratic}
\begin{tabular}{|c|c|c|}
\toprule
\rowcolor{gray!20}
\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0.1$}}}\\[1em]
%&&&\\
\midrule
  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
  & & \textbf{ estimation} \\
 \midrule
\textbf{2SPSA} & $-0.0062 \pm 0.1164$ & $-0.1229 \pm 0.1374$\\
&&\\
\textbf{2RDSA-Unif} &$0.0485 \pm 0.1465$ & $-0.259 \pm 0.0398$\\ 
&&\\
\textbf{2RDSA-AsymBer}& $-0.2564 \pm 0.068$& $\bm{-0.2877 \pm 0.0051}$\\
 \bottomrule
 %%%%%%%%%%%%2nd ordr
\rowcolor{gray!20}
\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0$}}}\\[1em]
%&&&\\
\midrule
  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
  & & \textbf{ estimation} \\
 \midrule
\textbf{2SPSA} & $-0.0785 \pm 0.1178$ & $-0.1716 \pm 0.1339$\\
&&\\
\textbf{2RDSA-Unif} &$0.0326 \pm 0.1599$ & $-0.2672 \pm 0.0299$\\ 
&&\\
\textbf{2RDSA-AsymBer}& $-0.2777 \pm 0.0488$& $\bm{-0.2881 \pm 0.0012}$\\
 \bottomrule
\end{tabular}
\end{table}

%%%%%%%%%%%%%% NMSE for quadratic
\begin{table}
\centering
 \caption{NMSE values for quadratic objective \eqref{eq:quadratic} with and without noise: standard error from $500$ replications shown after $\pm$}
\label{tab:nmse-quadratic}
\begin{tabular}{|c|c|c|}
\toprule
\rowcolor{gray!20}
\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0.1$}}}\\[1em]
%&&&\\
\midrule
  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
  & & \textbf{ estimation} \\
 \midrule
\textbf{2SPSA} & $0.9491 \pm 0.0131$ & $0.5495 \pm 0.0217$\\
&&\\
\textbf{2RDSA-Unif} &$1.0073 \pm 0.0140$ & $0.1953 \pm 0.0095$\\ 
&&\\
\textbf{2RDSA-AsymBer}& $0.1667 \pm 0.0095$& $\bm{0.0324 \pm 0.0007}$\\
 \bottomrule
 %%%%%%%%%%%%2nd ordr
\rowcolor{gray!20}
\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0$}}}\\[1em]
%&&&\\
\midrule
  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
  & & \textbf{ estimation} \\
 \midrule
\textbf{2SPSA} & $0.7325 \pm 0.0180$ & $0.3939 \pm 0.0230$\\
&&\\
\textbf{2RDSA-Unif} &$0.9834 \pm 0.0170$ & $0.1623 \pm 0.0086$\\ 
&&\\
\textbf{2RDSA-AsymBer}& $0.0686 \pm 0.0078$& $\bm{0.0316 \pm 0.0006}$\\
 \bottomrule
\end{tabular}
\end{table}

\newcommand{\errorband}[5][]{ % x column, y column, error column, optional argument for setting style of the area plot
\pgfplotstableread[col sep=comma, skip first n=2]{#2}\datatable
    % Lower bound (invisible plot)
    \addplot [draw=none, stack plots=y, forget plot] table [
        x={#3},
        y expr=\thisrow{#4}-2*\thisrow{#5}
    ] {\datatable};

    % Stack twice the error, draw as area plot
    \addplot [draw=none, fill=gray!40, stack plots=y, area legend, #1] table [
        x={#3},
        y expr=4*\thisrow{#5}
    ] {\datatable} \closedcycle;

    % Reset stack using invisible plot
    \addplot [forget plot, stack plots=y,draw=none] table [x={#3}, y expr=-(\thisrow{#4}+2*\thisrow{#5})] {\datatable};
}

 \begin{figure}
    \centering
\tabl{c}{\scalebox{0.9}{\begin{tikzpicture}
      \begin{axis}[
	xlabel={number of simulations},
	ylabel={Normalized loss},
       legend entries={
	 ,2SPSA,,2SPSA-IH,,2RDSA-Unif,,2RDSA-Unif-IH,,2RDSA-AsymBer,,2RDSA-AsymBer-IH },
        %legend pos=outer north east,
			legend style={
at={(0,0)},
anchor=north,at={(axis description cs:0.5,-0.2)},legend columns=2}	]
      % 2SPSA
      \errorband[red!50!white, opacity=0.3]{norloss_vs_simulations_noisycase.csv}{0}{1}{2}
      \addplot [thick, red,mark=square*,mark options={scale=0.7}] table [x index=0, y index=1,col sep=comma] {norloss_vs_simulations_noisycase.csv};
			%%% With IH
      % 2SPSA
      \errorband[darkgray!50!white, opacity=0.3]{norloss_vs_simulations_noisycase.csv}{0}{7}{8}
      \addplot [thick, darkgray,mark=*,mark options={scale=0.75}] table [x index=0, y index=7,col sep=comma] {norloss_vs_simulations_noisycase.csv};

      % 2RDSA-Unif
      \errorband[blue!50!white, opacity=0.3]{norloss_vs_simulations_noisycase.csv}{0}{3}{4}
      \addplot [thick, blue,mark=square*,mark options={scale=0.7}] table [x index=0, y index=3,col sep=comma] {norloss_vs_simulations_noisycase.csv};
			%%% With IH
      % 2RDSA-Unif
      \errorband[magenta!50!white, opacity=0.3]{norloss_vs_simulations_noisycase.csv}{0}{9}{10}
      \addplot [thick, magenta,mark=*,mark options={scale=0.75}] table [x index=0, y index=9,col sep=comma] {norloss_vs_simulations_noisycase.csv};

      % 2RDSA-AsymBer
      \errorband[green!50!white, opacity=0.3]{norloss_vs_simulations_noisycase.csv}{0}{5}{6}
      \addplot [thick, darkgreen,mark=square*,mark options={scale=0.7}] table [x index=0, y index=5,col sep=comma] {norloss_vs_simulations_noisycase.csv};
			%%% With IH	
      % 2RDSA-AsymBer
      \errorband[violet!50!white, opacity=0.3]{norloss_vs_simulations_noisycase.csv}{0}{11}{12}
      \addplot [thick, violet, mark=*,mark options={scale=0.75}] table [x index=0, y index=11,col sep=comma] {norloss_vs_simulations_noisycase.csv};
      \end{axis}
      \end{tikzpicture}}\\}
			\caption{Normalized loss vs. number of simulations for fourth-order loss \eqref{eq:4thorder} with $\sigma=0.1$ for 2SPSA, 2RDSA-Unif and 2RDSA-AsymBer algorithms with/without improved Hessian estimation: bands around the curves represent standard error from $500$ replications. }
      \label{fig:norlossvssims} 
\end{figure}

Tables \ref{tab:norloss-4thf}--\ref{tab:norloss-quadratic} present the normalized loss values observed for the three algorithms - 2SPSA, 2RDSA-Unif and 2RDSA-AsymBer - with/without improved Hessian estimation scheme and for the fourth-order and quadratic loss functions, respectively. Table \ref{tab:nmse-quadratic} presents the NMSE values obtained for the aforementioned algorithms with the quadratic loss. The results in Tables \ref{tab:norloss-4thf}--\ref{tab:nmse-quadratic} are obtained after running all the algorithms with a budget of $10000$ function evaluations.
 Figure \ref{fig:norlossvssims} plots the normalized loss as a function of the simulation budget with the fourth-order loss objective with $\sigma=0.1$ 
%  (see Figures \ref{fig:norlossvssimsNoiseless}--\ref{fig:norlossvssimsNoiseless} in the appendix for similar results with $\sigma = 0$ for fourth-order loss and $\sigma=0.1$ for quadratic loss). 
 From the results in Tables \ref{tab:norloss-4thf}--\ref{tab:nmse-quadratic} and Fig \ref{fig:norlossvssims}, we make the following observations:

\textit{\textbf{Observation 1:} Among 2RDSA schemes, 2RDSA-IH performs better than regular 2RDSA, for both perturbation choices.}

\textit{\textbf{Observation 2:} 2RDSA-IH variants outperform both 2SPSA and 2SPSA-IH, with 2RDSA-IH-AsymBer performing the best overall.}

\begin{figure}[h]
    \centering
		\begin{tabular}{c}
		\begin{subfigure}{0.5\textwidth}
\tabl{c}{\scalebox{0.8}{\begin{tikzpicture}
      \begin{axis}[
	xlabel={number of simulations},
	ylabel={Normalized loss},
       legend entries={
	 ,2SPSA,,2SPSA-IH,,2RDSA-Unif,,2RDSA-Unif-IH,,2RDSA-AsymBer,,2RDSA-AsymBer-IH },
        %legend pos=outer north east,
			legend style={
at={(0,0)},
anchor=north,at={(axis description cs:0.5,-0.2)},legend columns=2}	]
      % 2SPSA
      \errorband[red!50!white, opacity=0.3]{norloss_vs_simulations.csv}{0}{1}{2}
      \addplot [thick, red,mark=square*,mark options={scale=0.7}] table [x index=0, y index=1,col sep=comma] {norloss_vs_simulations.csv};
			%%% With IH
      % 2SPSA
      \errorband[darkgray!50!white, opacity=0.3]{norloss_vs_simulations.csv}{0}{7}{8}
      \addplot [thick, darkgray,mark=*,mark options={scale=0.75}] table [x index=0, y index=7,col sep=comma] {norloss_vs_simulations.csv};

      % 2RDSA-Unif
      \errorband[blue!50!white, opacity=0.3]{norloss_vs_simulations.csv}{0}{3}{4}
      \addplot [thick, blue,mark=square*,mark options={scale=0.7}] table [x index=0, y index=3,col sep=comma] {norloss_vs_simulations.csv};
			%%% With IH
      % 2RDSA-Unif
      \errorband[magenta!50!white, opacity=0.3]{norloss_vs_simulations.csv}{0}{9}{10}
      \addplot [thick, magenta,mark=*,mark options={scale=0.75}] table [x index=0, y index=9,col sep=comma] {norloss_vs_simulations.csv};

      % 2RDSA-AsymBer
      \errorband[green!50!white, opacity=0.3]{norloss_vs_simulations.csv}{0}{5}{6}
      \addplot [thick, darkgreen,mark=square*,mark options={scale=0.7}] table [x index=0, y index=5,col sep=comma] {norloss_vs_simulations.csv};
			%%% With IH	
      % 2RDSA-AsymBer
      \errorband[violet!50!white, opacity=0.3]{norloss_vs_simulations.csv}{0}{11}{12}
      \addplot [thick, violet, mark=*,mark options={scale=0.75}] table [x index=0, y index=11,col sep=comma] {norloss_vs_simulations.csv};
      \end{axis}
      \end{tikzpicture}}\\}
			\caption{Fourth-order loss \eqref{eq:4thorder} with $\sigma=0$. }
      \label{fig:norlossvssimsNoiseless} 
\end{subfigure} 
\\
\\
%%%%%%% Norloss-Quadratic-noisy %%%%%%%%%%%%%%
\begin{subfigure}{0.5\textwidth}
\tabl{c}{\scalebox{0.8}{\begin{tikzpicture}
      \begin{axis}[
	xlabel={number of simulations},
	ylabel={Normalized loss},
       legend entries={
	 ,2SPSA,,2SPSA-IH,,2RDSA-Unif,,2RDSA-Unif-IH,,2RDSA-AsymBer,,2RDSA-AsymBer-IH },
        %legend pos=outer north east,
			legend style={
at={(0,0)},
anchor=north,at={(axis description cs:0.5,-0.2)},legend columns=2}	]
      % 2SPSA
      \errorband[red!50!white, opacity=0.3]{quad_norloss_vs_simulations_noisycase.csv}{0}{1}{2}
      \addplot [thick, red,mark=square*,mark options={scale=0.7}] table [x index=0, y index=1,col sep=comma] {quad_norloss_vs_simulations_noisycase.csv};
			%%% With IH
      % 2SPSA
      \errorband[darkgray!50!white, opacity=0.3]{quad_norloss_vs_simulations_noisycase.csv}{0}{7}{8}
      \addplot [thick, darkgray,mark=*,mark options={scale=0.75}] table [x index=0, y index=7,col sep=comma] {quad_norloss_vs_simulations_noisycase.csv};

      % 2RDSA-Unif
      \errorband[blue!50!white, opacity=0.3]{quad_norloss_vs_simulations_noisycase.csv}{0}{3}{4}
      \addplot [thick, blue,mark=square*,mark options={scale=0.7}] table [x index=0, y index=3,col sep=comma] {quad_norloss_vs_simulations_noisycase.csv};
			%%% With IH
      % 2RDSA-Unif
      \errorband[magenta!50!white, opacity=0.3]{quad_norloss_vs_simulations_noisycase.csv}{0}{9}{10}
      \addplot [thick, magenta,mark=*,mark options={scale=0.75}] table [x index=0, y index=9,col sep=comma] {quad_norloss_vs_simulations_noisycase.csv};

      % 2RDSA-AsymBer
      \errorband[green!50!white, opacity=0.3]{quad_norloss_vs_simulations_noisycase.csv}{0}{5}{6}
      \addplot [thick, darkgreen,mark=square*,mark options={scale=0.7}] table [x index=0, y index=5,col sep=comma] {quad_norloss_vs_simulations_noisycase.csv};
			%%% With IH	
      % 2RDSA-AsymBer
      \errorband[violet!50!white, opacity=0.3]{quad_norloss_vs_simulations_noisycase.csv}{0}{11}{12}
      \addplot [thick, violet, mark=*,mark options={scale=0.75}] table [x index=0, y index=11,col sep=comma] {quad_norloss_vs_simulations_noisycase.csv};
      \end{axis}
      \end{tikzpicture}}\\}
			\caption{Quadratic loss \eqref{eq:quadratic} with $\sigma=0.1$. }
      \label{fig:QuadnorlossvssimsNoisy} 
\end{subfigure} 
		\end{tabular}
			\caption{Normalized loss vs. number of simulations in two different loss settings for all the algorithms. }
      \label{fig:extras} 
		\end{figure}

%%%%%% Norloss-Quadratic-noisefree %%%%%%%%%%%%%%
\begin{figure}
    \centering
\tabl{c}{\scalebox{0.9}{\begin{tikzpicture}
      \begin{axis}[
	xlabel={number of simulations},
	ylabel={Normalized loss},
       legend entries={
	 ,2SPSA,,2SPSA-IH,,2RDSA-Unif,,2RDSA-Unif-IH,,2RDSA-AsymBer,,2RDSA-AsymBer-IH },
        %%legend pos=outer north east,
			legend style={
at={(0,0)},
anchor=north,at={(axis description cs:0.5,-0.2)},legend columns=2}	]
      % 2SPSA
      \errorband[red!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{1}{2}
      \addplot [thick, red,mark=square*,mark options={scale=0.7}] table [x index=0, y index=1,col sep=comma] {quad_norloss_vs_simulations.csv};
			%%%% With IH
      %2SPSA
      \errorband[darkgray!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{7}{8}
      \addplot [thick, darkgray,mark=*,mark options={scale=0.75}] table [x index=0, y index=7,col sep=comma] {quad_norloss_vs_simulations.csv};
			%
      % 2RDSA-Unif
      \errorband[blue!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{3}{4}
      \addplot [thick, blue,mark=square*,mark options={scale=0.7}] table [x index=0, y index=3,col sep=comma] {quad_norloss_vs_simulations.csv};
			%%%% With IH
      % 2RDSA-Unif
      \errorband[magenta!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{9}{10}
      \addplot [thick, magenta,mark=*,mark options={scale=0.75}] table [x index=0, y index=9,col sep=comma] {quad_norloss_vs_simulations.csv};

      % 2RDSA-AsymBer
      \errorband[green!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{5}{6}
      \addplot [thick, darkgreen,mark=square*,mark options={scale=0.7}] table [x index=0, y index=5,col sep=comma] {quad_norloss_vs_simulations.csv};
			%%%% With IH	
      % 2RDSA-AsymBer
      \errorband[violet!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{11}{12}
      \addplot [thick, violet, mark=*,mark options={scale=0.75}] table [x index=0, y index=11,col sep=comma] {quad_norloss_vs_simulations.csv};
      \end{axis}
      \end{tikzpicture}}\\}
			\caption{Normalized loss vs. number of simulations for quadratic loss \eqref{eq:quadratic} with $\sigma=0$ for 2SPSA, 2RDSA-Unif and 2RDSA-AsymBer algorithms with/without improved Hessian estimation. }
      \label{fig:QuadnorlossvssimsNoiseless} 
\end{figure}

### 2016-12-24 intro.tex
maintexfiles added

\chapter{Introduction}
Optimization problems can be classified into several branches depending upon the nature of the problem. Consider the following optimization problem which is to find a $\theta^{*}$ that minimizes the objective function $f(\theta)$:
%\begin{equation} 
\begin{align}\label{eq:opt}
 \underset{\theta \in \Theta}{\min} ~f(\theta),
\end{align}
%\end{equation}
where $f \colon \mathbb{R}^N \to \mathbb{R}$ is called the objective function, $\theta$ is tunable N-dimensional parameter and $\Theta \subseteq \mathbb{R}^N$ is the constraint set in which $\theta$ takes values. 
\section{Classification of optimization problems based on objective function}\label{sc:objclass}
If we have complete information about $f$ and its derivatives etc., and about the set $\Theta$ then \eqref{eq:opt} would be a \emph{deterministic optimization} problem. In the real world unfortunately, many problems do not fall in this class. However, Optimization problems involving uncertainties are very common in a variety of engineering disciplines such as transportation systems, manufacturing, networks, healthcare and finance.

The setting in \emph{stochastic optimization}, however presumes that we have little knowledge on the structure of $f$ and moreover $f$ cannot be obtained directly, but rather is an expectation of another quantity $h(\theta,\xi)$, to which we have access, i.e.,
\begin{align}\label{eq:stopt}
 f(\theta) \equiv E_{\xi}[h(\theta,\xi)], 
\end{align}
 where $\xi$ comprises the randomness in the system and one is allowed to observe only these $h(\theta,\xi)$ samples, largely  because one dose not have access to distribution of noise $\xi$. This kind of optimization problems are more challenging because of the added complexity of not knowing $f$ and to find $\theta^{*}$ only on the basis of aforementioned noisy samples. The large number of input variables and the lack of precise system model may prohibit analytical solution approaches and a viable alternative is to employ a simulation-based optimization approach. As illustrated in Figure \ref{fig:so}, the idea here is to simulate a few times the stochastic system under consideration until a good enough solution is obtained. A natural solution approach is to devise an algorithm that incrementally updates the parameter, say $\theta_n$, in the descent direction using the gradient and/or Hessian of the objective $f$. However, in practice, one can only obtain estimates of the function $f$ through black-box simulation and the challenge is to estimate the gradient and/or Hessian of $f$ from function samples. \par

\tikzstyle{block} = [draw, fill=white, rectangle,
   minimum height=3em, minimum width=6em]
\tikzstyle{sum} = [draw, fill=white, circle, node distance=1cm]
\tikzstyle{input} = [coordinate]
\tikzstyle{output} = [coordinate]
\tikzstyle{pinstyle} = [pin edge={to-,thin,black}]

 \begin{figure}[t]
    \centering
\scalebox{0.8}{\begin{tikzpicture}[auto, node distance=2cm,>=latex']
% We start by placing the blocks
\node (theta) {$\boldsymbol{\theta_n}$};
\node [block, fill=blue!20,right=0.6cm of theta,align=center] (sample) {\makecell{\textbf{Simulator}}}; 
\node [right=0.6cm of sample] (end) {$\boldsymbol{\mathbf{f(\theta_n) + \xi_n}}$};
\node [ above right= 0.6cm of end] (bias) {\textbf{Zero mean}};
\draw [->] (theta) --  (sample);
\draw [->] (sample) -- (end);
\path [darkgreen,->] (bias) edge [bend left] (end);
\end{tikzpicture}}
\caption{Simulation optimization}
\label{fig:so}
\end{figure}

Suppose the objective function $f$ has the form $f(\theta) = \sum_{i=1}^{n} E[h_{i}(X_{i})]$ , where $n$ denotes the number of stages and $X_{i}$ is the state of the underlying process in stage $i$ and $h_{i}$ denotes a stage and state dependent cost function. Then this type of optimization problems are called \emph{multi-stage problems}. Let $\theta = (\theta_{1},\ldots,\theta_{n})^{T}$,  with each $\theta_{j}$ being scalar and Let $X_{i}$ depend on $\theta_{1},\ldots,\theta_{i}$. The idea is that optimization can be done one stage at a time over $n$ stages after observing the state $X_{i}$ in each stage $i$. The value $\theta_{i}$ in stage $i$ has bearing on the cost of all subsequent stages $i+1,...,n$. This is  a problem of dynamic optimization. Approaches such as Dynamic programming can be used to solve this optimization problem.\par
Now we consider sub-class of multi-stage problems having infinite number of stages and having \emph{long-run average cost function} as objective. Objective function here is of the form:
\begin{align}\label{eq:long}
 f(\theta) = \lim_{n \to \infty} \frac{1}{n} ~ E \left[ \sum_{i=1}^{n} h_{i}(X_{i})\right ] ,
\end{align}
where $X_{i}$ is defined same as before and we assume it depends on parameter $\theta$. Objective function \eqref{eq:long} in most cases is not known analytically. In such cases it makes difficult for the search procedure to update the search parameter with out estimating the cost over infinite long trajectory. Another subclassifications of optimization problems is based on feasible region $\Theta$, see \eqref{eq:opt}.

\section{Classification of optimization problems based on feasible region}
Consider the basic optimization problem in \eqref{eq:opt}. Optimization problems can be classified based on the structure of feasible region $\Theta$. Where set $\Theta$ can be of following types:

\begin{inparaenum}[\bfseries (i)]
 \item \emph{Discrete}
\item \emph{Continuous}
\item \emph{Hybrid} (continuous in some dimensions and discrete in others).
\end{inparaenum}
Hence corresponding optimization problems are called discrete, continuous, hybrid optimization problems. These subclassifications are common to both deterministic and stochastic optimization problems presented in section \ref{sc:objclass}.

\section{Optimization methods via simulation }\label{sc:dis}
Optimization via simulation is fast developing area for both research and practice. As mentioned before, for stochastic optimization problems simulation is a viable approach compared to solving them analytically. Multiple simulations replications must be performed in order to get good estimate of $E_{\xi}[h(\theta_i,\xi)]$, which is function value at $\theta_i$. The standard approach to estimate is by the sample mean
\begin{align}\label{eq:sm}
\overline f(\theta_i) = \frac{1}{n} ~  \sum_{j=1}^{n} h(\theta,\xi_j),
\end{align}
where $n$ is the number of simulation replications and $\xi_j$ is the j-th sample of randomness. As $n \to \infty$, $\overline f(\theta_i) \to E_{\xi}[h(\theta_i,\xi)]$. Primary concerns in stochastic optimization via simulation are 
\begin{inparaenum}[\bfseries (i)]
	\item $n$ must be large enough to get good estimate of  $E_{\xi}[h(\theta_i,\xi)]$, which has ultimate impact on the final solution that the optimization procedure need to find; and
	\item The above procedure must be applied for many different parameters ,i.e., $\theta$'s in order to find best $\theta$.
\end{inparaenum}
Hence one of the key criteria for comparing different simulation-based optimization approaches is by total number of simulations performed by them to attain specific function value. Note that different approaches may take different simulations per single iteration. In this section we present methods only for Continuous and Discrete optimization via simulation problems which often occur in practise. However, hybrid optimization problems are not frequently encountered. From  now on, we consider the setting of optimization problems mentioned in this section as the stochastic optimization setting presented in section \ref{sc:objclass}, i.e., when objective function is an expected value over certain noisy cost measurements. 
\subsection{Discrete optimization via simulation}
Optimization problems for which $\theta \in \Theta$ is discrete and integer ordered and $\Theta$ is a subset of $N$-dimensional integers are known as Discrete optimization via simulation problems. In the discrete case $\theta$ can take only countable (finite or infinite) number of values. Discrete optimization problems are some times called as \emph{combinatorial optimization} and have applications in resource allocation, network routing, policy planning etc.
In this section we consider the case of $\Theta$ being finite i.e., $\Theta = \{\theta_{1},\theta_{2},\ldots,\theta_{k}\}$ where $k \in \mathbb{N}$. Discrete optimization problems are divided into two sub-classes when feasible region is  either small (often less than 100 ) or large. we present optimization methods in literature separately for each class.\\
%\begin{enumerate}
\subsubsection{$\Theta$ is small}
In this case we can simulate all possible solutions and select the best among them. However, unlike in deterministic optimization, simulating only once for each parameter is not enough since the objective function is noisy in stochastic optimization setting. Hence the main question is how to conduct multiple simulation effectively for each parameter in order to determine best $\theta$.  Two popular methods in literature for this case are Ranking and Selection (R\&S)\cite{bech} and Multiple comparison procedures (MCP) \cite{mcp}. These methods are well suited for simulation since under lying assumptions such as normality and independence of observations can be met by careful selection procedures.
Traditional methods are applicable for maximum size of 20 while some recent techniques allow more. 
\begin{enumerate}

\item \textbf {Ranking and selection }
R\&S procedures are statistical procedures specially developed to select the best parameter or subset that contains the best parameter. Majority of work in R\&S is classified into  two categories. The first one is called \emph{Indifferent-zone ranking}\cite{bech,izrs} which is aimed  to select the best and another one is \emph{Subset selection}\cite{ssrs} where the procedures are designed to find subset consisting of best. Again there are two approaches for selecting the best problem : \emph{the frequentist approach} and \emph{ Bayesian approach}. One of the popular streams of Bayesian approach procedures is  optimal computing budget allocation (OCBA) procedures \cite{ocbabook,chen2000} in which simulation budget is allocated in a manner that maximises the posterior probability of correct selection. This method is widely considered the best procedure for small scale discrete optimization.

\item \textbf {MCPs}
Like R\&S, MCPs attempt to identify the best parameter. But MCPs approach optimization problem as a statistical inference problem and do not guarantee a decision. Three main classes of MCPs that are used in practise are \emph{Multiple Comparison approach} (MCA) , \emph{Multiple Comparison with the Best} (MCB) and \emph{Multiple Comparisons with Control} (MCC). The most popular among these is MCB. In particular, MCB looks at $f(\theta_{j}) - opt_{i \neq j} f(\theta_{i})$ for $j = 1,2,\ldots,k$ to determine that $j_{*}$ with $f(\theta_{j_{*}}) - opt_{ i \neq j_{*}} f(\theta_{i}) > 0$. Simultaneous confidence intervals $f(\theta_{j}) - opt_{i \neq j} f(\theta_{i})$ for $j = 1,2,\ldots,k$ can be used to determine $j_{*}$ by looking  for the confidence interval with lower confidence limit that is zero.
\end{enumerate}
For a comprehensive treatment of these methods one can refer \cite{swish,swish2}.
\subsubsection{$\Theta$ is large}
When $\Theta$ is large enumeration becomes too expensive to conduct. Some sort of search techniques like the ones in deterministic optimization must be applied to avoid simulating for all the parameters while ensuring a high chance of finding the best or good parameters. Some approaches in this category include the following : 
\begin{enumerate}
\item \emph{Model-based approaches:} Iterative algorithms using statistical methods search the feasible region to improve upon the candidate parameter. One can  use \emph{gradient/Hessian information} with respect to parameter to help the search similar to deterministic case. But due to lack of availability of function analytical form and distribution of noise one needs to estimate the gradient/Hessian from samples of function $f$. However gradient/Hessian  estimation could be  quite noisy due to stochastic nature involved  in these problems. There are several approaches  which are used in continuous case can also be used in this setting to estimate the gradient/Hessian when simulation model is treated as black box. We will present detailed literature survey of these approaches in sections \ref{sec:gradest} and \ref{sec:hesest}. 
\item \emph{Metaheuristic:} These are gradient free approaches. They include approaches such as genetic algorithms, evolutionary algorithms, simulated annealing \cite{simann}, tabu search \cite{tabu} and cross entropy \cite{crossent}. Most of these iterative algorithms start with an initial population of parameters and in each iteration elite parameters are selected from previous population and generates  better parameter population as search progresses. 
\end{enumerate}

\subsection{Continuous optimization via simulation}
Optimization problems for which $\theta \in \Theta$ is a vector of continuous decision variables and $\Theta$ is a convex subset of $\mathcal{R}^N$ are called continuous optimization via simulation  problems. In these problems $\theta$ can take uncountable number of values. These kind of problems frequently occur in applications such as model fitting, adaptive control, neural network training, signal processing etc. A natural solution approach is to devise an algorithm that incrementally updates the parameter, say $\theta_n$, in the descent direction using the gradient and/or Hessian of the objective $f$. Stochastic approximation algorithms are best suited and most studied and used algorithms for solving continuous optimization problems via simulation.\par
The \textbf{stochastic approximation algorithm} (SA) takes the following iterative form:
\begin{align}\label{saalg}
\theta_{n+1} = \Gamma_{\Theta} \left [\theta_{n} - a_{n} \widehat{\nabla} f(\theta_{n}) \right],
\end{align}
where $\theta_{n}$ is the solution found at iteration $n$. $ \widehat{\nabla} f(\theta_{n})$ is an estimate of the gradient $ {\nabla} f(\theta_{n})$ and $\{a_{n}\}$ is a sequence of positive reals satisfy following properties: $\sum_{n=1}^{\infty} a_{n}  = \infty$,  $\sum_{n = 1}^{\infty} a_{n}^{2} < \infty$, and $\Gamma_{\Theta}$ denotes the projection of a point outside $\Theta$ onto $\Theta$. Under appropriate conditions, as number of iterations goes to infinity one can guarantee convergence to the local minimum with probability one. The SA algorithm can be considered as a stochastic analog of gradient decent method which seeks the next solution along negative gradient direction. The main difference between the SA algorithm and steepest descent algorithm is even though gradient estimate in former algorithm is a noisy estimate it requires only weak assumptions (like $E \left[\widehat{\nabla} f(\theta_{n})\right] - \nabla f(\theta_{n}) \to 0 $ at a certain rate) for converging to local minimum. In practice, the performance of the SA algorithm is quite sensitive to the sequence $\{a_{n}\}$ for example in the case $ a_n = a/n$, the convergence is highly dependent on the choice of $a$. However, in practice, one can only obtain estimates of the function $f$ through black-box simulation and the challenge is to estimate the gradient and/or Hessian of $f$ from function samples.  In the following sections we will present a brief survey of existing methods to estimate gradient and Hessian from function samples.
%There are several Gradient estimation methods present in the literature. Some of them are Finite difference approximation (FDA) and SPSA techniques, Perturbation analysis (PA) and Likelihood ratio / Score function (LF/SF) are popular among those.
%
%Formally, given only noise-corrupted measurements of an objective function $f$, we want to solve the following problem:
%\begin{align}
%\mbox{Find } x^* = \arg\min_{x \in \R^N} f(x). \label{eq:pb}
%\end{align}
\section{Gradient estimation}\label{sec:gradest}
\subsection{Finite-difference stochastic approximation (FDSA)}
When the simulation model is treated as a block box, the traditional means of forming the estimate of gradient is by using finite-difference stochastic approximation (FDSA) method. There are mainly two variations in the FDSA scheme. First one is  two-sided gradient approximation which involve function measurements $f(\theta_n + \delta_n e_i)$ and $f(\theta_n - \delta_n e_i)$, where $e_i$ denotes $i$th column of a identity matrix of size $N \times N$. Denote these respective values by $y_{ni}^+$ and $y_{ni}^-$, i.e., 
$$
y_{ni}^+ = f(\theta_n+\delta_n e_i) + \xi_{ni}^+,\quad y_{ni}^- = f(\theta_n-\delta_n e_i) + \xi_{ni}^-.
$$
In the above, we assume the noise vector $(\xi_{ni}^+ - \xi_{ni}^-, i=1,2,\ldots,N)\tr$ is a martingale difference sequence for every $n \ge 0$, the sequence of the perturbation constants $\{\delta_n, n\ge 0\}$ is a positive and asymptotically vanishing sequence. The gradient estimate of objective function using this scheme is given by
\begin{align}\label{eq:twofdsa}
 \widehat{\nabla} f(\theta_{n}) = \left(
\begin{array}{c}
\frac{y_{n1}^+ - y_{n1}^-}{2\delta_n}\\
\vdots\\
\frac{y_{nN}^+ - y_{nN}^-}{2\delta_n}\\
\end{array}
\right)
\end{align}
Second one is  one-sided gradient approximations which involve function measurements $f(\theta_n+\delta_n e_i)$ and $f(\theta_n)$. Let us denote these values by  $y_{ni}^+$, $y_{n}$  respectively, i.e., 
$y_{ni}^+ = f(\theta_n+\delta_n e_i) + \xi_{ni}^+$, $y_{n} = f(\theta_n) + \xi_n$,  where we assume the noise terms $(\xi_{ni}^+ - \xi_{n}, i=1,2,\ldots,N)\tr$ satisfy the martingale difference sequence property for $n \ge 0$. The gradient estimate of objective function using this scheme is given by
\begin{align}\label{eq:onefdsa}
 \widehat{\nabla} f(\theta_{n}) = \left(
\begin{array}{c}
\frac{y_{n1}^+ - y_{n}}{\delta_n}\\
\vdots\\
\frac{y_{nN}^+ - y_{n}}{\delta_n}\\
\end{array}
\right)
\end{align}
It is important to determine conditions under which $\theta_n$ converges to $\theta^*$, when one uses estimates shown in \eqref{eq:twofdsa} or \eqref{eq:onefdsa} used in \eqref{saalg}. The convergence theory for FDSA algorithm is similar to convergence theory of Robbins and Monro root finding SA algorithm \cite{rm}. However difficulties arise due to a bias in gradient approximation, i.e. $\E \left[\widehat{\nabla} f(\theta_{n})|\F_n \right] - \nabla f(\theta_{n}) $, where $\F_n = \sigma(x_m, m\le n)$ denotes the underlying sigma-field. The conditions required for showing convergence of this algorithm is as follows
\begin{enumerate}[label=(\textbf{A\arabic*})]
\item $f:\R^N\rightarrow \R$ is three-times continuously differentiable\footnote{Here $\nabla^3 f(\theta) = \dfrac{\partial^3 f (\theta)}{\partial \theta\tr \partial \theta\tr \partial \theta\tr}$ denotes the third derivate of $f$ at $\theta$ and $\nabla^3_{i_1 i_2 i_3} f(\theta)$ denotes the $(i_1 i_2 i_3)$th entry of $\nabla^3 f(\theta)$, for $i_1, i_2, i_3=1,\ldots, N$.}  with $\left|\nabla^3_{i_1 i_2 i_3} f(\theta) \right| < \alpha_0 < \infty$, for $i_1, i_2, i_3=1,\ldots, N$ and for all $\theta\in \R^N$. 
\item $\{\xi_n^+,\xi_n^-, n=1,2,\ldots\}$ satisfy $\E\left[\left.\xi_n^+ - \xi_n^- \right| \F_n\right] = 0$.
\item For some $\alpha_1, \alpha_2, \zeta >0$ and for all $n,i$, 
$\E \left|\xi_n^{\pm}\right|^{2+\zeta} \le \alpha_1$, $\E \left|f(\theta_n\pm \delta_n e_i)\right|^{2+\zeta} \le \alpha_2$. 
%\item $\{d_n^i, i=1,\ldots,N, n=1,2,\ldots\}$ are i.i.d. and independent of $\F_n$.
\item The step-sizes $a_n$ and perturbation constants $\delta_n$ are positive, for all $n$ and satisfy
$$a_n, \delta_n \rightarrow 0\text{ as } n \rightarrow \infty, 
\sum_n a_n=\infty \text{ and } \sum_n \left(\frac{a_n}{\delta_n}\right)^2 <\infty.$$
\item $\sup_n \left\| \theta_n \right\| < \infty$ w.p. $1$.
\end{enumerate}
Note that two-sided gradient estimation scheme requires $2N$ simulations of objective function in order to obtain single gradient estimate, where as one-sided gradient estimation scheme requires only $N+1$ simulations, where $N$ is the dimension of the vector $\theta$. Dependence on $N$ for number of simulations to get one gradient estimate makes this procedure highly computationally expensive when $N$ is large. Hence one requires gradient estimate procedure that is independent of dimension of $\theta$.
\subsection{Simultaneous perturbation stochastic approximation (SPSA)}
Simultaneous perturbation (SP) methods are a popular and efficient approach for estimating gradient/Hessian from function samples, especially in high dimensional problems - see \cite{bhatnagar-book} for a comprehensive treatment of this subject matter. Simultaneous perturbation stochastic approximation (SPSA) is a popular SP method. The first-order SPSA algorithm, henceforth referred to as 1SPSA, was proposed in \cite{spall2005introduction}.  1SPSA scheme for approximating gradient requires only two simulations of function measurements irrespective of parameter dimension. The idea in this scheme is to perturb all components of parameter randomly for getting two function measurements. The function measurements required for this scheme corresponds to $\theta_n + \delta_n \Delta_n$ and $\theta_n - \delta_n \Delta_n$, respectively, where $\Delta_n = (\Delta_{n1},\ldots,\Delta_{nN})\tr$ is any vector consists of i.i.d, mean-zero, symmetric random variables whose inverse moments are bounded. The simplest and most commonly used perturbation distribution of random variables being used are symmetric Bernoulli distribution with $\Delta_{ni} = \pm 1 \,\,w.p.\,\,1/2$, $i=1,\ldots,N,\,\, n\ge 0$. The gradient estimate of objective function using this scheme is as follows
\begin{align}\label{eq:spsa}
 \widehat{\nabla} f(\theta_{n}) = \left(
\begin{array}{c}
\frac{f(\theta_n+\delta_n \Delta_n)-f(\theta_n-\delta_n \Delta_n)}{2\delta_n \Delta_{n1}} + \frac{\xi_{n}^+ - \xi_{n}^-}{2\delta_n \Delta_{n1}}\\
\vdots\\
\frac{f(\theta_n+\delta_n \Delta_n)-f(\theta_n-\delta_n \Delta_n)}{2\delta_n \Delta_{nN}} +  \frac{\xi_{n}^+ - \xi_{n}^-}{2\delta_n \Delta_{nN}}\\
\end{array}
\right)
\end{align}
The reason why 1SPSA algorithm is valid gradient estimation scheme can be seen easily from Taylor expansions of $f(\theta_n+\delta_n \Delta_n$ and $f(\theta_n-\delta_n \Delta_n)$.
\begin{align*}
&f(\theta_n + \delta_n \Delta_n) = f(\theta_n) + \delta_n \Delta_n\tr \nabla f(\theta_n) + \frac{\delta_n^2}{2} \Delta_n\tr \nabla^2 f(\theta_n) \Delta_n + o(\delta_n^2)\\
&f(\theta_n - \delta_n \Delta_n) = f(\theta_n) - \delta_n \Delta_n\tr \nabla f(\theta_n) + \frac{\delta_n^2}{2} \Delta_n\tr \nabla^2 f(\theta_n) \Delta_n + o(\delta_n^2)
\end{align*}
From above two equations, the $i$th component of gradient approximation \eqref{eq:spsa} is given by
\begin{align}\label{eq:spsaexp}
\frac{f(\theta_n + \delta_n \Delta_n) - \theta_n - \delta_n \Delta_n)}{2\delta_n \Delta_{ni}} = \nabla_i f(\theta_n) + \sum\limits_{j=1, j \neq i}^N \frac{\Delta_{nj} }{\Delta_{ni}}  \nabla^2_{j} f(\theta_n)  + o(\delta_n)
\end{align}
Now the conditional expectation of the 1SPSA gradient estimate is given by 
\begin{align}\label{eq:spsacondmean}
\E\left[\frac{f(\theta_n + \delta_n \Delta_n) - \theta_n - \delta_n \Delta_n)}{2\delta_n \Delta_{ni}} | \F_n \right] = \nabla_i f(\theta_n) + o(\delta_n)
\end{align}

The asymptotic convergence of 1SPSA algorithm requires one extra assumption other than the conditions specified in FDSA scheme above. The extra condition for convergence of $\theta_n$ to $\theta^*$ as $n \to \infty$ is as follows
\begin{enumerate}[label=(\textbf{A\arabic*}),resume]
\item $\{\Delta_{ni}, i=1,\ldots,N, n=1,2,\ldots\}$ are i.i.d., independent of $\F_n$ and for some $\alpha_0 > 0$ and $\forall n$, $\E (\Delta_{ni})^{-2} \le \alpha_0$, for $i = 1,2,\ldots,N$.
\end{enumerate}
Since 1SPSA randomly perturbing the parameter vector $\theta$, number of function measurements required is two, irrespective of the dimension of $\theta$, see that the numerator of \eqref{eq:spsa} is same in all $N$ components of $\widehat{\nabla} f(\theta_{n})$. However, Inspite of the fact that the 1SPSA requires only two function measurements compared to  $2N$  measurements required by FDSA, the asymptotic convergence rate is the same as the FDSA scheme.
\subsection{Random directions stochastic approximation (RDSA)}
A closely related algorithm to SPSA is random directions stochastic approximation (RDSA) \cite[pp.~58-60]{kushcla}. The gradient estimate in RDSA differs from that in SPSA, both in the construction as well as in the choice of random perturbations.  In \cite{kushcla}, the random perturbations for 1RDSA were generated by picking samples uniformly on the surface of a sphere and the resulting 1RDSA scheme was found to be  inferior to 1SPSA from an asymptotic convergence rate viewpoint - see \cite{chin1997comparative}.  
The RDSA estimate of the gradient is given by
\begin{align}
\label{eq:grad-unif}
\widehat\nabla f(\theta_n) = d_n \left[ \dfrac{y_n^+ - y_n^-}{2\delta_n}\right].
\end{align}
where $y_{n}^+ = f(\theta_n+\delta_n d_n) + \xi_{n}^+,\quad y_{n}^- = f(\theta_n-\delta_n d_n) + \xi_{n}^-$ ,$d_n = (d_n^i,\ldots,d_n^N)\tr$, and $d_n^i$, $ i=1,\ldots,N$ are i.i.d random perturbations distributed uniformly on $N$-dimensional sphere with radius 1. Like 1SPSA, 1RDSA also requires only two function measurements irrespective of dimension of parameter $\theta$. 
The asymptotic convergence of 1RDSA algorithm also requires one extra assumption other than the conditions specified in FDSA scheme. 
\begin{enumerate}[label=(\textbf{A\arabic*}),resume]
\item $\{ d_{n}^i, i=1,\ldots,N, n=1,2,\ldots\}$ are i.i.d., independent of $\F_n$, and $\forall n$, $\E (d_{n} d_{n}\tr) = I_{N \times N}$, and for some $\alpha_0 > 0$, $\E {d_{n}^i}^2 f(\theta_n \pm \delta_n d_n) \le \alpha_0$ for $i=1,2,\ldots,N$.
\end{enumerate}
Recent work in \cite{prashanth2015rdsa} attempts to bridge the gap between 1RDSA and 1SPSA in terms of improving the performance by incorporating random perturbations based on an asymmetric Bernoulli
distribution as well as those with the  i.i.d uniform distribution. However,  1SPSA was found to be still marginally better than 1RDSA.  
% On the other hand, results in \cite{prashanth2015rdsa} show that a second order RDSA approach (referred to as 2RDSA hereafter) can considerably outperform the corresponding higher order SPSA algorithm \cite{spall_adaptive} (referred to as 2SPSA hereafter). 
\section{Hessian estimation}\label{sec:hesest}
Stochastic Newton methods can counter the ill-conditioning of the objective $f$ as they incorporate second-order information into the update iteration given by   
\begin{align}
\label{eq:newton}
\theta_{n+1} = \theta_n - a_n (\overline H_n)^{-1}\widehat\nabla f(\theta_n), 
\end{align}
where $a_n$ is the step-size that satisfies standard stochastic approximation conditions (see (C5) in Section \ref{sec:2rdsa-results}), $\widehat\nabla f(\theta_n)$ and $\overline H_n$ are estimates of the gradient and Hessian, respectively. Thus, \eqref{eq:newton} can be considered as  the stochastic version of the well-known Newton method for optimization. Stochastic Newton methods are often more accurate than simple gradient search schemes.
%which are sensitive to the choice of the constant $a_0$ in the canonical step-size, $a_n= a_0/n$. The optimal (asymptotic) convergence rate is obtained only if $a_0 > 1/3 \lambda_0$, where $\lambda_0$ is the minimum eigenvalue of the Hessian of the objective function (see \cite{fabian1967stochastic}). However, this dependency is problematic, as $\lambda_0$ is unknown in a \textit{simulation optimization} setting. Hessian-based methods get rid of this dependency, while attaining the optimal rate (one can set $a_0=1$). An alternative approach to achieve the same effect is to employ Polyak-Ruppert averaging, which uses larger step-sizes and averages the iterates. However, iterate averaging is optimal only in an asymptotic sense. Finite-sample analysis (see Theorem 2.4 in \cite{fathi2013transport}) shows that the initial error (that depends on the starting point $x_0$ of the algorithm) is not forgotten sub-exponentially fast, but at the rate $1/n$ where $n$ is the number of iterations. 
%Thus, 
%the effect of averaging kicks in only after enough iterations have passed and the bulk of the iterates are centered around the optimum.

In \cite{fabian}, an estimation scheme for $\overline H_n$ that uses $O(N^2)$ function samples per-iteration of \eqref{eq:newton} was proposed.
%while in \cite{ruppert} the Hessian is estimated assuming knowledge of objective function gradients.

% In \cite{bhat1}, three different  Hessian estimation schemes 
% that require three, two, and one simulation(s)  have been proposed
% in the context of long-run average cost objectives. 
%The resulting algorithms incorporate two-timescale stochastic approximation, see Chapter 6 of \cite{borkar}.
% Certain three-simulation balanced simultaneous perturbation Hessian estimates have been
% proposed in \cite{sbpla}. In addition, certain procedures for Hessian inversion  that require
% lower computational effort have also been proposed, see also \cite{bhatnagar-book}.
%In \cite{spall-jacobian}, certain enhancements to the four-simulation Hessian estimates of
%\cite{spall_adaptive} using some feedback and weighting mechanisms
%have been proposed. 
% In \cite{bhat2}, Newton-based smoothed functional algorithms based on Gaussian perturbations
% have been proposed.  An overview of random search approaches (both gradient and Newton-based) involving both theory and application of these techniques is available in \cite{bhatnagar-book}.

\subsection{Second-order SPSA (2SPSA)}
The number of samples per-iteration for estimating Hessian was brought down to four, irrespective of dimension $N$, by the second-order SPSA algorithm (henceforth referred to as 2SPSA). The Hessian estimator is projected to the space of positive definite and symmetric matrices at each iterate for the algorithm to progress along a descent direction. In this scheme two independent perturbation sequences $\Delta_n = (\Delta_{n1},\ldots,\Delta_{nN})\tr$, $\widehat \Delta_n = (\widehat \Delta_{n1},\ldots,\widehat \Delta_{nN})\tr$, each consists of random variables satisfying conditions specified in 1SPSA scheme are used.
The 2SPSA algorithm obtains four function samples $y_n^{++}$, $y_n^+$,  $y_n^{-+}$ and $y_n^{-}$ at $\theta_n + \delta_n \Delta_n + \widehat \delta_n \widehat \Delta_n $, $\theta_n +\delta_n \Delta_n$, $\theta_n - \delta_n \Delta_n + \widehat \delta_n \widehat \Delta_n$ and $\theta_n - \delta_n \Delta_n$, where  the sequences of the perturbation constants $\{\delta_n, n\ge 0\}, \{\widehat \delta_n, n\ge 0\}$ is a positive and asymptotically vanishing sequence and
the random perturbations $\Delta_n, \widehat \Delta_n$ are such that $\{\Delta_{ni},\widehat\Delta_{ni}, i=1,\ldots,N, n=1,2,\ldots\}$ are i.i.d. and independent of the noise sequence. And $y_n^{++} = f(\theta_n + \delta_n \Delta_n + \widehat \delta_n \widehat \Delta_n ) + \xi_n^{++}$, $y_n^+ = f(\theta_n+\delta_n d_n) + \xi_n^+$, $y_n^{-+} = f(\theta_n - \delta_n \Delta_n + \widehat \delta_n \widehat \Delta_n) + \xi_n^{+-}$ and $y_n^- = f(\theta_n-\delta_n d_n) + \xi_n^-$, where the noise terms $\xi_n^{++}, \xi_n^+, \xi_n^{-+}, \xi_n^-$ satisfy $\E\left[\left.\xi_n^{++} - \xi_n^+ - \xi_n^{-+} +  \xi_n^-\right| \F_n\right] = 0$ with $\F_n = \sigma(\theta_m,m\le n)$ denoting the underlying sigma-field. The $i,j$th entry of the Hessian estimate $\widehat H_n$ in this case is given by
\begin{align}
\label{eq:2spsahhat}
&\left(\widehat H_n\right)_{ij} = \left[\frac{1}{\Delta_{ni}\widehat \Delta_{nj}} + \frac{1}{\Delta_{nj}\widehat \Delta_{ni}} \right] \left[\dfrac{y_n^{++} - y_n^+ - y_n^{-+} + y_n^-}{4 \delta_n \widehat \delta_n}\right], 
\end{align}
The second-order SPSA algorithm performs an update iteration as follows:
\begin{align}
\label{eq:2spsa}
\theta_{n+1} = \theta_n - a_n \Upsilon(\overline H_n)^{-1}\widehat\nabla f(\theta_n), \\
\overline H_n = \frac{n}{n+1} \overline H_{n-1} + \frac{1}{n+1} \widehat H_n.\label{eq:2spsa-H}
\end{align}
In the above, 
\begin{itemize}
 \item $\widehat\nabla f(\theta_n)$ is the estimate of $\nabla f(\theta_n)$ and this corresponds to \eqref{eq:spsa} 
  \item $\widehat H_n$ is an estimate of the true Hessian ${\nabla}^2 f(\cdot)$ at $\theta_n$. 
 \item $\overline H_n$ is a smoothed version of $\widehat H_n$, which is crucial to ensure convergence. 
 \item $\Upsilon$ is an operator that projects a matrix onto the set of positive definite matrices. Update \eqref{eq:2spsa-H} does not necessarily ensure that $\overline H_n$ is invertible and without $\Upsilon$, the parameter update \eqref{eq:2spsa} may not move along a descent direction - see conditions (C7)  in Section \ref{sec:2rdsa-results}  for the precise requirements on the matrix projection operator.
\end{itemize}
It is important to determine conditions under which $\theta_n$ converges to $\theta^*$ and $\overline H_n \to H(\theta^*)$. Convergence theory for second-order methods varies significantly to that of first-order methods. One can see \cite{spall_adaptive} for detailed convergence results.
\subsection{Second-order RDSA (2RDSA)}
The basic algorithm in \eqref{eq:2spsa}--\eqref{eq:2spsa-H} is similar to the adaptive scheme analyzed by \cite{prashanth2015rdsa}. However, they have used RDSA by incorporating random perturbations based on an asymmetric Bernoulli
distribution as well as those with the  i.i.d uniform distribution for the gradient and Hessian estimates (see \eqref{eq:grad-ber} and \eqref{eq:2rdsa-estimate-ber}, \eqref{eq:grad-unif} and \eqref{eq:2rdsa-estimate-unif} for their respective estimates), while \cite{spall_adaptive} employs SPSA. Though 1SPSA was found to be still marginally better than 1RDSA, results in \cite{prashanth2015rdsa} show that a second order RDSA approach (referred to as 2RDSA hereafter) can considerably outperform the corresponding second order SPSA algorithm \cite{spall_adaptive}, while requiring only three simulations per iteration of \eqref{eq:newton}.
\section{Contributions of this thesis}
The following are the brief description of the contributions of this thesis:
\begin{enumerate}
\item Our work in chapter \ref{sec:2rdsa-ih} is centred on improving the 2RDSA scheme of \cite{prashanth2015rdsa} by 
\begin{enumerate}[label={\bf\Roman*}]
\item reducing the error in the Hessian estimate through a feedback term; and
\item optimizing the step-sizes used in the Hessian estimation recursion, again with the objective of improving the quality of the Hessian estimate.
\end{enumerate}

Items (I) and (II) are inspired by the corresponding improvements to the Hessian estimation recursion in the enhanced 2SPSA from \cite{spall-jacobian}. We shall refer to the latter algorithm as 2SPSA-IH. While item (II) above is a relatively straightforward migration to the 2RDSA setting, item (I) is a non-trivial contribution, primarily because the Hessian estimate in 2RDSA is entirely different from that in 2SPSA and the feedback term that we incorporate in 2RDSA to improve the Hessian estimate neither correlates with that in 2SPSA nor follows from the analysis in \cite{spall-jacobian}. 
The advantage with 2RDSA scheme along with proposed improvement to Hessian estimation (henceforth referred to as 2RDSA-IH) is that it requires only 75\% of the simulation cost per-iteration for 2SPSA-IH.

We establish that the proposed improvements to Hessian estimation in 2RDSA are such  that the resulting 2RDSA-IH algorithm is provably convergent, in particular, the Hessian estimate $\overline H_n$ of 2RDSA-IH converges almost surely to the true Hessian. 
%Moreover, we show, for the special case of a quadratic objective, that	2RDSA-IH results in a convergence rate that is on par with the corresponding rate for 2SPSA with Hessian estimation improvements (2SPSA-IH) \cite{spall-jacobian}. 
Further, we show empirically that 2RDSA-IH outperforms both 2SPSA-IH of \cite{spall-jacobian} and regular 2RDSA of \cite{prashanth2015rdsa}. Our contribution is important because 2RDSA-IH, like 2RDSA, has lower simulation cost per iteration than 2SPSA and unlike 2RDSA, has an improved Hessian estimation scheme.
\item Our work in chapter  \ref{sec:2rdsa-gen} is centred on generalising the 2RDSA scheme of \cite{prashanth2015rdsa}. In \cite{prashanth2015rdsa}, 2RDSA scheme involving only two perturbation distributions was proposed, those are uniform and asymmetric Bernoulli distributions. In chapter \ref{sec:2rdsa-gen} , we proposed gradient and Hessian estimation schemes which are independent of the perturbation distributions. However, perturbations have to satisfy the i.i.d, mean-zero assumptions. Apart from these common assumptions  one additional assumption like difference between the second moment square  and fourth moment should be non zero for Hessian estimation is also required. This generalisation of distributions of random variables for perturbations is important because it enhances the scope of improving the performance of the 1RDSA and 2RDSA schemes by incorporating several other distributions. Some well known distribution that can used for these schemes as a result of generalisation are Normal (0,1) and Cauchy distributions.
\item Our work in \ref{sec:2spsa-3-ih} is centred on improving the second order SPSA scheme with three simulations ( referred to as 2SPSA-3 hereafter), which was proposed in  \cite{bhatnagar2015simultaneous}. This improvements are as a result of incorporating the zero mean feedback term in Hessian estimate and optimizing the step-sizes used in the Hessian estimate recursion. Though these ideas are same as the ideas presented in this section for 2RDSA scheme, the contribution is non-trivial because the derivation of feedback term and Hessian estimate is entirely different from 2SPSA, 2RDSA improvements. Moreover, we show, for the special case of a quadratic objective and when there is no noise, that 2SPSA-3 scheme along with proposed improvement to Hessian estimation (henceforth referred to as 2SPSA-3-IH) results in a convergence rate that is on par with the corresponding rate for 2SPSA with Hessian estimation improvements (2SPSA-IH) \cite{spall-jacobian}. The advantage with 2SPSA-3-IH is that it requires only 75\% of the simulation cost per-iteration for 2SPSA-IH.
\end{enumerate}

The rest of the thesis is organised as follows: In Chapter~\ref{sec:2rdsa-ih}, we describe the improved Hessian estimation scheme, which is incorporated into the  2RDSA algorithm from \cite{prashanth2015rdsa}. We also presented the theoretical results for the 2RDSA algorithm with  improved Hessian estimation and the results from numerical experiments. In Chapter~\ref{sec:2rdsa-gen}, we describe the generalised RDSA scheme for both the gradient and Hessian estimates and also present the theoretical convergence results as well. The work related to improving the Hessian estimation scheme 2SPSA-3 was presented in Chapter~\ref{sec:2spsa-3-ih} and finally, in  Chapter~\ref{sec:conclusions}, the concluding remarks were provided.

### 2016-12-24 second-order-rdsa-gen.tex
maintexfiles added

\chapter{Generalised RDSA }
\label{sec:2rdsa-gen}
Our work in chapter  is centred on generalising the 2RDSA scheme of \cite{prashanth2015rdsa}. In \cite{prashanth2015rdsa}, 2RDSA scheme involving only two perturbation distributions was proposed, those are uniform and asymmetric Bernoulli distributions. We propose gradient and Hessian estimation schemes which are independent of the perturbation distributions. However, perturbations have to satisfy the i.i.d, mean-zero assumptions. Apart from these common assumptions  one additional assumption like difference between the second moment square  and fourth moment should be non zero for Hessian estimation is also required. This generalisation of distributions of random variables for perturbations is important because it enhances the scope of improving the performance of the 1RDSA and 2RDSA schemes by incorporating several other distributions. Some well known distribution that can used for these schemes as a result of generalisation are Normal (0,1) and Cauchy distributions. We will prove the  gradient and Hessian estimates as a result of generalisation are indeed asymptotically unbiased estimates.
 The improved Hessian estimation scheme proposed in chapter \ref{sec:2rdsa-ih} for 2RDSA algorithm with uniform and asymmetric Bernoulli distributions can be easily extended to the generalised RDSA scheme.

\section{Function evaluations}
Let $\delta_n, n\geq 0$ denote a sequence of diminishing positive real numbers and $d_n = (d_n^1,\ldots,d_n^N)\tr$ denote a random perturbation vector at instant $n$,
where the perturbations $\{d_n^i, i=1,\ldots,N, n=1,2,\ldots\}$ are any i.i.d., mean zero random varibles. 
%\begin{equation}
%\label{eq:det-proj}
% d_n^i =
%  \begin{cases}
%   -1 &  \text{ w.p. } \dfrac{(1+\epsilon)}{(2+\epsilon)}, \\
%   1+\epsilon &  \text{ w.p. } \dfrac{1}{(2+\epsilon)},
%  \end{cases}
%\end{equation}
%with $\epsilon>0$ being a constant that can be chosen to be arbitrarily small.

The 2RDSA-IH algorithm obtains three function samples $y_n$, $y_n^+$ and $y_n^-$ at $\theta_n$, $\theta_n+\delta_n d_n$ and $\theta_n - \delta_n d_n$, respectively, i.e.,
$y_n = f(\theta_n) + \xi_n$, $y_n^+ = f(\theta_n+\delta_n d_n) + \xi_n^+$ and 
$y_n^- = f(\theta_n-\delta_n d_n) + \xi_n^-$,
where the noise terms $\xi_n, \xi_n^+, \xi_n^-$ satisfy $\E\left[\left.\xi_n^+ + \xi_n^- - 2\xi_n \right| \F_n\right] = 0$ with $\F_n = \sigma(\theta_m,m\le n)$ denoting the underlying sigma-field.

\section{Generalised RDSA gradient estimate}
The RDSA estimate of the gradient $\nabla f(\theta_n)$ is given by
\begin{align}
\label{eq:grad-gen}
\widehat\nabla f(\theta_n) = \frac1{\lambda} d_n \left[ \dfrac{y_n^+ - y_n^-}{2\delta_n}\right],
\end{align}
where $\lambda = \E(d_n^i)^2 $ and the perturbations $d_n^i$, $i=1,\ldots,N$ are i.i.d and zero-mean random variables. 
%Note that $E d_n^i = 0$, $E (d_n^i)^2 = 1+\epsilon$ and $E (d_n^i)^4 = \dfrac{(1+\epsilon)(1+(1+\epsilon)^3)}{(2+\epsilon)}$.

\section{Generalised RDSA Hessian estimate}
\begin{align}
\label{eq:2rdsa-estimate-gen}
&\widehat H_n = M_n \left(\dfrac{y_n^+ + y_n^- - 2 y_n}{\delta_n^2}\right), \text{ where }\\
& M_n =
\left[
\begin{array}{ccc}
\frac{1}{\kappa}\left((d_n^1)^2\!-\lambda\right) & \cdots & \frac{1}{2 \lambda^2}d_n^1 d_n^N\\
\frac{1}{2 \lambda^2}d_n^2 d_n^1  &  \cdots & \frac{1}{2 \lambda^2}d_n^2 d_n^N\\
\cdots&\cdots&\cdots\\
\frac{1}{2 \lambda^2}d_n^N d_n^1 & \cdots &  \frac{1}{\kappa}\left((d_n^N)^2-\lambda\right) \\
\end{array}
\right],\nonumber
\end{align}
where $\lambda = \E(d_n^i)^2 $ , $\tau = E (d_n^i)^4$, and $\kappa = \left(\tau - \lambda^2\right)$ for any $i=1,\ldots,N$.
From the above Hessian estimation scheme it is easy to see that perturbations $d_n^i$, $i=1,\ldots,N$ should satisfy  $\kappa \neq 0$. However, many distributions of the random variables satisfy this requirement.

%%%%%%%%%%%%%%%%%%%% LEMMA 
\begin{enumerate}[label=(\textbf{A\arabic*}),resume]
\item For some $\alpha_1, \alpha_2,\zeta >0$ and for all $n$, 
$\E \left|\xi_n^{\pm}\right|^{2+\zeta} \le \alpha_1$, $\E \left|f(x_n\pm \delta_n d_n)\right|^{2+\zeta} \le \alpha_2$. 
\item $\{d_n^i, i=1,\ldots,N, n=1,2,\ldots\}$ are i.i.d., zero-mean and independent of $\F_n$. For some $\eta_1,\eta_2,\zeta >0$ and for all $n$, $\E(d_n^i)^{2+\zeta} \le \eta_1,\E(d_n^i)^{4+\zeta} \le \eta_2$.
\end{enumerate}

\begin{lemma}(\textbf{Bias in the gradient estimate})
\label{lemma:1rdsa-bias}
%Almost surely, as $\delta_n \rightarrow 0$, we have that
Under (A1),(A2),(A4),(A5),(A13),(A14) for $\widehat\nabla f(x_n)$ defined according to  \eqref{eq:grad-gen} , we have a.s. that\footnote{Here $\widehat\nabla_i f(x_n)$ and $\nabla_i f(x_n)$ denote the $i$th coordinates in the gradient estimate $\widehat\nabla f(x_n)$ and true gradient $\nabla f(x_n)$, respectively.}
\begin{align}
 \left| \E\left[\left.\widehat\nabla_i f(x_n)\right| \F_n \right] - \nabla_i f(x_n)\right| = O(\delta_n^2),  \quad \text{ for } i=1,\ldots,N.
\end{align} 
\end{lemma}

\begin{proof}
We use the proof techniques of \cite{spall},\cite{prashanth2015rdsa} (see Lemma 1) in order to prove the main claim here.

Notice that
\begin{align*}
&\E\left[\left.\dfrac{y_n^+ - y_n^-}{2\delta_n} \right| \F_n\right] \\
%= &\E\left[d_n\left.\left(\dfrac{f(x_n+\delta_n d_n) - f(x_n-\delta_n d_n)}{2\delta_n}\right) \right| \F_n\right] + \E\left[  d_n\left.\left(\dfrac{\xi_n^+ - \xi_n^-}{2\delta_n}\right) \right| \F_n\right]\\
= & \E\left[d_n\left.\left(\dfrac{f(x_n+\delta_n d_n) - f(x_n-\delta_n d_n)}{2\delta_n}\right) \right| \F_n\right].
\end{align*}
The last equality above follows from the fact that $\E\left[  d_n\left.\left(\dfrac{\xi_n^+ - \xi_n^-}{2\delta_n}\right) \right| \F_n\right]=0$ from (A2) and (A13). We now analyse the term on the RHS above.
Let $\nabla^2 f(\cdot)$ denote the Hessian of $f$.
By Taylor's series expansions, we obtain, a.s.,
\begin{align*}
f(x_n \pm \delta_n d_n) =& f(x_n) \pm \delta_n d_n\tr \nabla f(x_n) + \frac{\delta_n^2}{2} d_n\tr \nabla^2 f(x_n) d_n \\
& \pm  \frac{\delta_n^3}{6} \nabla^3 f(\tilde  x_n^+)(d_n \otimes d_n \otimes d_n),
\end{align*}
where $\otimes$ denotes the Kronecker product and $\tilde x_n^+$ (resp. $\tilde x_n^-$) are on the line segment between $x_n$ and $(x_n + \delta_n d_n)$ (resp. $(x_n - \delta_n d_n)$).
Hence,
\begin{align}
&\E\left[d_n\left.\left(\dfrac{f(x_n+\delta_n d_n) - f(x_n-\delta_n d_n)}{2\delta_n}\right)\right| \F_n \right] \nonumber\\
&= \E\left[d_n d_n\tr \left.\nabla f(x_n)\right| \F_n\right]\nonumber \\
& +   \E\left[\left.\frac{\delta_n^2}{12} d_n (\nabla^3 f(\tilde  x_n^+)+\nabla^3 f(\tilde  x_n^-))(d_n \otimes d_n \otimes d_n)\right| \F_n\right]. \label{eq:l1}
\end{align}
The first term on the RHS above can be simplified as follows:
\begin{align}
\E\left[d_n d_n\tr \left.\nabla f(x_n)\right| \F_n\right] = &\E\left[d_n d_n\tr \right] \nabla f(x_n) \nonumber\\
% = & \E\left[
%\begin{array}{cccc}
%(d_n^1)^2 & d_n^1 d_n^2 & \cdots & d_n^1 d_n^N\\
%d_n^2 d_n^1 &(d_n^2)^2 &  \cdots & d_n^2 d_n^N\\
%d_n^N d_n^1 & d_n^N d_n^2 & \cdots &  (d_n^N)^2 \\
%\end{array}
%\right]\nabla f(x_n) 
= &  \lambda \nabla f(x_n). \label{eq:l1-a}
\end{align}
In the above, the first equality follows from (A14) and the last equality in \eqref{eq:l1-a} follows from $\E[(d_n^i)^2] = \lambda$ and $\E[d_n^i d_n^j] = \E[d_n^i] \E[d_n^j] = 0$ for $i\ne j$.

Now, the $l$th coordinate of the second term in the RHS of \eqref{eq:l1} can be upper-bounded as follows:
\begin{align}
&\E\left[\left.\frac{\delta_n^2}{12} d_n^l (\nabla^3 f(\tilde  x_n^+)+\nabla^3 f(\tilde  x_n^-))(d_n \otimes d_n \otimes d_n)\right| \F_n\right]\nonumber\\
\le & \dfrac{\alpha_0 \delta_n^2}{6} \sum_{i_1=1}^N\sum_{i_2=1}^N\sum_{i_3=1}^N \E\left( d_n^l d_n^{i_1} d_n^{i_2} d_n^{i_3}\right)\nonumber\\
\le & \dfrac{\alpha_0 \delta_n^2\eta N^3}{6}.\label{eq:l2}
\end{align}
The first inequality above follows from (A1), while the second inequality follows from (A14) and considering $\eta = \max \{\eta_1^2, \eta_2\}$. The claim follows by plugging \eqref{eq:l1-a} and \eqref{eq:l2} into \eqref{eq:l1} .\\
\end{proof}

\begin{enumerate}[label=(\textbf{C\arabic*}),resume]
\item $\{d_n^i, i=1,\ldots,N, n=1,2,\ldots\}$ are i.i.d., zero-mean and independent of $\F_n$. For some $\eta_1,\eta_2,\zeta >0$ and for all $n$, $\E(d_n^i)^{2+\zeta} \le \eta_1,\E(d_n^i)^{4+\zeta} \le \eta_2$.
\end{enumerate}

\begin{lemma}(\textbf{Bias in Hessian estimate})
\label{lemma:2rdsa-bias}
Under (C1)-(C3),(C5)-(C10),(C15) with $\widehat H_n$ defined according to \eqref{eq:2rdsa-estimate-gen} , we have a.s. that\footnote{Here $\widehat H_n(i,j)$ and $\nabla^2_{ij}f(\cdot)$ denote the $(i,j)$th entry in the Hessian estimate $\widehat H_n$ and the true Hessian $\nabla^2 f(\cdot)$, respectively.}, for $i,j = 1,\ldots,N$,
\begin{align}
\left|\E\left[
\left. \widehat H_n(i,j) \right| \F_n \right] - \nabla^2_{ij} f(\theta_n)\right| = O(\delta_n^2).
\end{align} 
\end{lemma}
\begin{proof}
We use the proof techniques of \cite{prashanth2015rdsa} (see Lemma 4) in order to prove the main claim here. However, we provide proof for general RDSA scheme in which perturbations need not be asymmetric Bernoulli or uniform as in the case of \cite{prashanth2015rdsa}.

By a Taylor's series expansion, we obtain
\begin{align*}
f(\theta_n \pm \delta_n d_n) = f(\theta_n) \pm \delta_n d_n\tr \nabla f(\theta_n) + \frac{\delta_n^2}{2} d_n\tr \nabla^2 f(\theta_n) d_n &\pm \frac{\delta_n^3}{6} \nabla^3 f(\theta_n) (d_n \otimes d_n \otimes d_n) \\
 & +  \frac{\delta_n^4}{24} \nabla^4 f(\tilde  \theta_n^+)(d_n \otimes d_n \otimes d_n \otimes d_n).
\end{align*}
The fourth-order term in each of the expansions above can be shown to be of order $O(\delta_n^4)$ using (C1) and arguments similar to that in Lemma 3.1. Hence,
\begin{align*}
\dfrac{f(\theta_n+\delta_n d_n) + f(\theta_n-\delta_n d_n) - 2 f(\theta_n)}{\delta_n^2}
 = &\,\, d_n\tr \nabla^2 f(\theta_n) d_n +  O(\delta_n^2)\\
= & \sum\limits_{i=1}^N\sum\limits_{j=1}^N d_n^i d_n^j \nabla^2_{ij} f(\theta_n) + O(\delta_n^2)\\
= & \sum\limits_{i=1}^N (d_n^i)^2 \nabla^2_{ii} f(\theta_n) + 2\sum\limits_{i=1}^{N-1}\sum\limits_{j=i+1}^N d_n^i d_n^j \nabla^2_{ij} f(\theta_n) + O(\delta_n^2).
\end{align*}
Now, taking the conditional expectation of the Hessian estimate $\widehat{H_n}$ and observing that $\E[\xi_n^+ + \xi_n^- - 2\xi_n \mid \F_n, d_n] = 0$ by (C3), we obtain the following:
\begin{align}
\E[\widehat H_n \mid \F_n] =  \E\left[\left. M_n \left(\sum\limits_{i=1}^{N-1} (d_n^i)^2 \nabla^2_{ii} f(\theta_n) \right.\right.\right.\left.\left.\left.+ 2\sum\limits_{i=1}^N\sum\limits_{j=i+1}^N d_n^i d_n^j \nabla^2_{ij} f(\theta_n) + O(\delta_n^2)\right)\right| \F_n\right]. \label{eq:h1}
\end{align}

Note that the $O(\delta_n^2)$ term inside the conditional expectation above remains $O(\delta_n^2)$ even after the multiplication with $M_n$.
We analyse the diagonal and off-diagonal terms in the multiplication of the matrix $M_n$ with the scalar above, ignoring the $O(\delta_n^2)$ term.

\subsection*{Diagonal terms in \eqref{eq:h1}:}

Consider the $lth$ diagonal term inside the conditional expectation in \eqref{eq:h1}:
\begin{align}
\dfrac{1}{\kappa} & \left((d_n^l)^2-\lambda\right) \left(\sum\limits_{i=1}^N (d_n^i)^2 \nabla^2_{ii} f(\theta_n)\right. \left. + 2\sum\limits_{i=1}^{N-1}\sum\limits_{j=i+1}^N d_n^i d_n^j \nabla^2_{ij} f(\theta_n)\right)
=\dfrac{1}{\kappa} (d_n^l)^2 \sum\limits_{i=1}^N (d_n^i)^2 \nabla^2_{ii} f(\theta_n) \nonumber\\
& + \dfrac{2}{\kappa} (d_n^l)^2\sum\limits_{i=1}^{N-1}\sum\limits_{j=i+1}^N d_n^i d_n^j \nabla^2_{ij} f(\theta_n) - \dfrac{ \lambda}{\kappa} \sum\limits_{i=1}^N (d_n^i)^2 \nabla^2_{ii} f(\theta_n) - \dfrac{2 \lambda}{ \kappa} \sum\limits_{i=1}^{N-1}\sum\limits_{j=i+1}^N d_n^i d_n^j \nabla^2_{ij} f(\theta_n).\label{eq:h2}
\end{align}
From the distributions of $d_n^i,d_n^j$ and the fact that $d_n^i$ is independent of $d_n^j$ for $i<j$, it is easy to see that $\E\left(\left. (d_n^l)^2\sum\limits_{i=1}^{N-1}\sum\limits_{j=i+1}^N d_n^i d_n^j \nabla^2_{ij} f(\theta_n) \right| \F_n\right) = 0$ and $\E\left(\left.\sum\limits_{i=1}^{N-1}\sum\limits_{j=i+1}^N d_n^i d_n^j \nabla^2_{ij} f(\theta_n) \right| \F_n\right) = 0$. Thus, the conditional expectations of the second and fourth terms on the RHS of \eqref{eq:h2} are both zero.

The first term on the RHS of \eqref{eq:h2} with the conditional expectation can be simplified as follows:
\begin{align}\label{eq:term1lem1}
\dfrac{1}{\kappa} \E\left(\left. (d_n^l)^2 \sum\limits_{i=1}^N (d_n^i)^2 \nabla^2_{ii} f(\theta_n) \right| \F_n\right)
= & \dfrac{1}{\kappa} \E\left((d_n^l)^4 \nabla^2_{ll} f(\theta_n) + \sum\limits_{i=1,i\ne l}^N (d_n^l)^2(d_n^i)^2 \nabla^2_{ii} f(\theta_n) \right)\nonumber\\
= & \dfrac{1}{\kappa} \left( \tau \nabla^2_{ll} f(\theta_n) + \lambda^2 \sum\limits_{i=1,i\ne l}^N  \nabla^2_{ii} f(\theta_n) \right), \text{ a.s.}
\end{align} 
For the second equality above, we have used the fact that $\E[(d_n^l)^4] =  \tau$ and $\E[(d_n^l)^2 (d_n^i)^2] = \E[(d_n^l)^2] \E[(d_n^i)^2] = \lambda^2$, $\forall l \ne i$.

The third term in \eqref{eq:h2} with the conditional expectation and without the negative sign can be simplified as follows: 
\begin{align}\label{eq:term3lem1}
\dfrac{ \lambda}{\kappa} \E\left(\left. \sum\limits_{i=1}^N (d_n^i)^2 \nabla^2_{ii} f(\theta_n) \right| \F_n\right)
= &\dfrac{ \lambda}{\kappa} \sum\limits_{i=1}^N \E \left[(d_n^i)^2\right] \nabla^2_{ii} f(\theta_n) \nonumber \\
=  &\dfrac{\lambda^2}{\kappa} \sum\limits_{i=1}^N \nabla^2_{ii} f(\theta_n), \text{ a.s.}
\end{align} 
Combining the above followed by some algebra, i.e., $\eqref{eq:term1lem1}-\eqref{eq:term3lem1}$, we obtain
\begin{align*}
\frac{1}{\kappa} (\tau - \lambda^2) \nabla^2_{ll} f(\theta_n)
\end{align*} 
By using the fact that $\kappa = \tau - \lambda^2$, we obtain
\begin{align*}
\dfrac{1}{\kappa}  \E\left[\left. \left((d_n^l)^2-\lambda\right) \left(\sum\limits_{i=1}^N (d_n^i)^2 \nabla^2_{ii} f(\theta_n) 
+ 2\sum\limits_{i=1}^{N-1}\sum\limits_{j=i+1}^N d_n^i d_n^j \nabla^2_{ij} f(\theta_n)\right)\right| \F_n\right] = \nabla^2_{ll} f(\theta_n), \text{ a.s.}
\end{align*}

\subsection*{Off-diagonal terms in \eqref{eq:h1}:}

We now consider the $(k,l)$th term in \eqref{eq:h1}: Assume w.l.o.g that $k<l$. Then,
\begin{align}
 &\dfrac{1}{2 \lambda^2} \E\left[\left.d_n^k d_n^l \left(\sum\limits_{i=1}^N (d_n^i)^2 \nabla^2_{ii} f(\theta_n)  
+ 2\sum\limits_{i=1}^{N-1}\sum\limits_{j=i+1}^N d_n^i d_n^j \nabla^2_{ij} f(\theta_n)\right)\right| \F_n \right] \nonumber\\
&=\dfrac{1}{2 \lambda^2} \sum\limits_{i=1}^N \E \left(d_n^k d_n^l (d_n^i)^2 \right)\nabla^2_{ii} f(\theta_n) + \dfrac{1}{\lambda^2}\sum\limits_{i=1}^{N-1}\sum\limits_{j=i+1}^N \E\left(d_n^k d_n^l d_n^i d_n^j\right) \nabla^2_{ij} f(\theta_n) \label{eq:crossh}\\
&=  \nabla^2_{kl} f(\theta_n).\nonumber
\end{align}
The last equality follows from the fact that the first term in \eqref{eq:crossh} is $0$ since $k\ne l$, while the second term in \eqref{eq:crossh} can be seen to be equal to $\dfrac{1}{\lambda^2} \E\left((d_n^k)^2 (d_n^l)^2\right) \nabla^2_{kl} f(\theta_n) = \nabla^2_{kl} f(\theta_n)$. \\
\end{proof}

### 2016-12-24 second-order-spsa-3.tex
maintexfiles added

 \chapter{Second-order spsa with improved hessian estimation (2SPSA-3-IH)}
\label{sec:2spsa-3-ih}
The second-order SPSA-3 with improved Hessian estimate performs an update iteration similar to that of 2RDSA-IH proposed in section \ref{sec:2rdsa-ih} (See \eqref{eq:e2rdsa} and \eqref{eq:2rdsa-H}). The main difference in the iterates is that 2SPSA-3-IH uses gradient and Hessian estimates according to \eqref{eq:grad-spsa} and \eqref{eq:2spsa-ber}. In this section we present the improvements to 2SPSA-3 Hessian recursion \eqref{eq:2spsa-3-ih} by incorporating the zero-mean feedback term and general step-size. Note that though these ideas of improving the Hessian recursion is similar to the ideas presented in section \ref{sec:2rdsa-ih} for 2RDSA-IH, the derivation of feedback term does not follow from it. We show that the proposed improvements to Hessian estimation in 2SPSA-3 are such  that the resulting 2SPSA-3-IH algorithm is provably convergent, in particular, the Hessian estimate $\overline H_n$ of 2SPSA-3-IH converges almost surely to the true Hessian. Moreover, we show, for the special case of a quadratic objective in noise-free setting, that 2SPSA-3 scheme along with proposed improvement to Hessian estimation (henceforth referred to as 2SPSA-3-IH) results in a convergence rate that is on par with the corresponding rate for 2SPSA with Hessian estimation improvements (2SPSA-IH) \cite{spall-jacobian}. The advantage with 2SPSA-3-IH is that it requires only 75\% of the simulation cost per-iteration for 2SPSA-IH.

%The recursion \eqref{eq:e2rdsa} is identical to that in 2RDSA, while the Hessian estimation recursion \eqref{eq:2rdsa-H} differs as follows:\\
%\begin{inparaenum}[\bfseries (i)]
%\item  $\widehat \Psi_n$ is a zero-mean feedback term that reduces the error in Hessian estimate; and\\
%\item $b_n$ is a general step-size that we optimize to improve the Hessian estimate.
%\end{inparaenum}
%
%On the other hand, $\widehat H_n$ is identical to that in 2RDSA, i.e., it estimates the true Hessian in each iteration using $3$ function evaluations. For the sake of completeness, we provide below the construction for $\widehat\nabla f(\theta_n)$ and $\widehat H_n$ using asymmetric Bernoulli perturbations, before we present the feedback term that reduces the error in $\widehat H_n$.
\begin{algorithm}[!htp]
\KwIn{initial parameter $\theta_0 \in \R^N$, perturbation constants $\delta_n>0$, step-sizes $\{a_n, b_n\}$, operator $\Upsilon$.}

\begin{enumerate}

 \item \textbf{Execution}: \\
 \For{$n \gets 0,1,2, \ldots, $} {
 \begin{itemize}
 \item Generate $\{\Delta_n^{i}, i=1,\ldots,N\}$,$\{\widehat\Delta_n^{i}, i=1,\ldots,N\}$ independent of $\{\Delta_m,\widehat\Delta_m, m=0,1,\ldots,n-1\}$.  
 \item For any $i=1,\ldots,N$, $\Delta_n^{i}$,$\widehat\Delta_n^{i}$  satisfy condition (C16) (most popular distribution being used is symmetric Bernoulli distribution, $\pm 1$ w.p. $1/2$).
\begin{itemize}
\item \textbf{\emph {Function evaluation 1} }\\ \hspace{4em} Obtain $y_n^+ = f(\theta_n+\delta_n \Delta_n +\delta_n \widehat\Delta_n) + \xi_n^+$.
\item \textbf{\emph {Function evaluation 2} }\\ \hspace{4em} Obtain $y_n^- = f(\theta_n -\delta_n\Delta_n-\delta_n\widehat\Delta_n) + \xi_n^-$.
\item \textbf{\emph {Function evaluation 3} }\\ \hspace{4em} Obtain $y_n = f(\theta_n) + \xi_n$.
\item \textbf{\emph {Newton step}}\\ \hspace{4.2em} Update the parameter and Hessian as follows:
%\begin{align*}
%		\theta_{n+1} = & \; \theta_n - a_n \Upsilon(\overline H_n)^{-1}\widehat\nabla f(\theta_n), \\
%\overline H_n = &\; (1-b_{n})  \overline H_{n-1} + b_{n} ( \widehat H_n - \widehat \Psi_n),
%\end{align*}
\begin{align} \label{eq:2spsa-3-theta}
\theta_{n+1} = \,& \theta_n - a_n \Upsilon(\overline H_n)^{-1}\widehat\nabla f(\theta_n),\\
\overline H_n =\,& (1-b_{n})  \overline H_{n-1} + b_{n} ( \widehat H_n - \widehat \Psi_n).\label{eq:2spsa-3-ih}
\end{align}
where $\widehat H_n$ and $\widehat \Psi_n$ are chosen according to \eqref{eq:2spsa-ber} and \eqref{eq:psinhatspsa3}, respectively. 
\end{itemize}

\end{itemize}
}
\Return{$\theta_n.$}
\end{enumerate}

\caption{Structure of 2SPSA-3-IH algorithm.}
\label{alg:spsa-3-ih-structure}
\end{algorithm}

Algorithm \ref{alg:spsa-3-ih-structure} presents the pseudocode and we describe the individual component of 2RDSA-IH below.
\section{Function evaluations}
Let $y_n$, $y_n^+$ and $y_n^-$ denote the function evaluations at $\theta_n$, $\theta_n+\delta_n \Delta_n+\delta_n \widehat \Delta_n$ and $\theta_n -\delta_n\Delta_n-\delta_n\widehat\Delta_n$ respectively, i.e., 
$y_n = f(\theta_n) + \xi_n$, $y_n^+ = f(\theta_n+\delta_n \Delta_n +\delta_n \widehat\Delta_n) + \xi_n^+$ and 
$y_n^- = f(\theta_n -\delta_n\Delta_n-\delta_n\widehat\Delta_n) + \xi_n^-$,
where the noise terms $\xi_n, \xi_n^+, \xi_n^-$ satisfy $\E\left[\left.\xi_n^+ + \xi_n^- - 2\xi_n \right| \F_n, \Delta_n, \widehat \Delta_n\right] = 0$ with $\F_n = \sigma(\theta_m,m <  n)$ denoting the underlying sigma-field.

Further, $\delta_n, n\geq 0$ is a sequence of diminishing positive real numbers and $\Delta_n = (\Delta_n^1,\ldots,\Delta_n^N)\tr$ , $\widehat\Delta_n = (\widehat\Delta_n^1,\ldots,\widehat\Delta_n^N)\tr$ are the perturbation random vector at instant $n$ with $\Delta_n^i,i=1,\ldots,N$, and $\widehat\Delta_n^i,i=1,\ldots,N$ are independent identically distributed (i.i.d), mean-zero random varibles having finite inverse movements of order greater than 2 and satisfy the conditions (C16), C(21).

\section{Gradient estimate}
The SPSA estimate of the gradient $\nabla f(\theta_n)$ is given by
\begin{align}
\label{eq:grad-spsa}
\widehat\nabla_{(i)} f(\theta_n) =  \left[ \dfrac{y_n^+ - y_n^-}{2\delta_n \Delta_n^{(i)}}\right],
\end{align}
where the perturbations $\Delta_n^i$, $i=1,\ldots,N$ are i.i.d. and as mentioned above.

%Note that $E d_n^i = 0$, $E (d_n^i)^2 = 1+\epsilon$ and $E (d_n^i)^4 = \dfrac{(1+\epsilon)(1+(1+\epsilon)^3)}{(2+\epsilon)}$.

\section{Hessian estimate}
The $i,j$th entry of the Hessian estimate in this case is given by
\begin{align}
\label{eq:2spsa-ber}
&\left(\widehat H_n\right)_{ij} =  \left(\dfrac{y_n^+ + y_n^- - 2 y_n}{2\delta_n^2 \Delta_n^{(i)} \widehat \Delta_n^{(j)}}\right), 
\end{align}
%%%%%% Feedback
\section{Feedback term $\widehat \Psi_n$}
The $i,j$th term of the Hessian estimate $\widehat H_n$ can be simplified as follows: 
\begin{align}
\left(\widehat H_n \right)_{ij} & =  \left(\dfrac{y_n^+ + y_n^- - 2 y_n}{2\delta_n^2 \Delta_n^{(i)} \widehat \Delta_n^{(j)}}\right) \nonumber \\
%& =  \left[\left(\dfrac{f(\theta_n+\delta_n \Delta_n+\delta_n \widehat \Delta_n) + f(\theta_n -\delta_n\Delta_n-\delta_n\widehat\Delta_n) - 2 f(\theta_n)}{\delta_n^2 \Delta_n^{(i)} \widehat \Delta_n^{(j)}}\right) \right.\nonumber \\&\hspace{10em}+\left. \left(\dfrac{\xi_n^+ + \xi_n^- - 2 \xi_n}{\delta_n^2}\right)\right] \nonumber \\
& = \left( \frac{(\Delta_n+ \widehat \Delta_n)\tr \nabla^2 f(\theta_n) (\Delta_n+ \widehat \Delta_n)}{2\Delta_n^{(i)} \widehat \Delta_n^{(j)}} +  O(\delta_n^2) + \right.\left.\left(\dfrac{\xi_n^+ + \xi_n^- - 2 \xi_n}{2\delta_n^2 \Delta_n^{(i)} \widehat \Delta_n^{(j)}} \right)\right).\label{eq:h3}
%& = M_n \left(d_n\tr \nabla^2 f(\theta_n) d_n +  O(\delta_n^2) + O(\delta_n^{-2})\right). \label{eq:h0}
\end{align}
For the first term on the RHS above, note that
\begin{align}
\E\left[\left.  \frac{(\Delta_n+ \widehat \Delta_n)\tr \nabla^2 f(\theta_n) (\Delta_n+ \widehat \Delta_n)}{2\Delta_n^{(i)} \widehat \Delta_n^{(j)}}  \right | \F_n\right] &=\E\left[\sum\limits_{l=1}^N\sum\limits_{m=1}^N \frac{\Delta_n^l \nabla^2_{lm} f(\theta_n) \Delta_n^m}{2\Delta_n^i \widehat\Delta_n^j} + \sum\limits_{l=1}^N\sum\limits_{m=1}^N \frac{\Delta_n^l \nabla^2_{lm} f(\theta_n) \widehat\Delta_n^m}{\Delta_n^i \widehat\Delta_n^j}  \right.\nonumber \\&\hspace{2.5cm}+\left.\sum\limits_{l=1}^N\sum\limits_{m=1}^N \left. \frac{\widehat\Delta_n^l \nabla^2_{lm} f(\theta_n)\widehat \Delta_n^m}{2\Delta_n^i \widehat\Delta_n^j}\right| \F_n\right].\label{eq:2spsaexp}
\end{align}
In analysing the RHS of the above expression, the first an the third terms are mean-zero (See Proposition 4.2 in \cite{bhatnagar2015simultaneous}):
\begin{align}
\E\left[\left.\sum\limits_{l=1}^N\sum\limits_{m=1}^N \frac{\Delta_n^l \nabla^2_{lm} f(\theta_n) \Delta_n^m}{\Delta_n^i \widehat\Delta_n^j} \right | \F_n\right] = 0 \hspace{0.2cm} and \hspace{0.2cm}\E\left[\left.\sum\limits_{l=1}^N\sum\limits_{m=1}^N \frac{\widehat\Delta_n^l \nabla^2_{lm} f(\theta_n) \widehat\Delta_n^m}{\Delta_n^i \widehat\Delta_n^j }\right | \F_n\right] = 0.
\end{align}
%\begin{align}
%\E\left[\left. \sum\limits_{l=1}^N\sum\limits_{m=1}^N \frac{\Delta_n^l \nabla^2_{lm} f(\theta_n) \widehat\Delta_n^m}{\Delta_n^i \widehat\Delta_n^j} \right | \F_n\right] = \nabla^2_{ij} f(\theta_n) \hspace{0.3cm}a.s..
%\end{align}
From the above equations the  terms that are in expectation, denoted by  $\Psi_n^1(\nabla^2 f(\theta_n))$, can be written in matrix form as follows:
\begin{align}
\Psi_n^1(\nabla^2 f(\theta_n)) =\frac{1}{2} M_n \left[\Delta_n\tr \nabla^2 f(\theta_n) \Delta_n + \widehat\Delta_n\tr \nabla^2 f(\theta_n) \widehat\Delta_n\right].
\end{align}
where $M_n = [1/\Delta_n^1, \ldots ,1/\Delta_n^N]\tr [1/\Delta_n^1, \ldots ,1/\Delta_n^N]$.
Now consider the second term in RHS of the \eqref{eq:2spsaexp}. It can be re-written as follows:
\begin{align}
\E\left[\left. \sum\limits_{l=1}^N\sum\limits_{m=1}^N \frac{\Delta_n^l \nabla^2_{lm} f(\theta_n) \widehat\Delta_n^m}{\Delta_n^i \widehat\Delta_n^j} \right | \F_n\right] = \E\left[\left.\nabla^2_{ij} f(\theta_n) + \frac{1}{\Delta_n^i \widehat\Delta_n^j} \mathop{\sum_{l=1}^N \sum_{m=1}^{N}}_{lm \neq ij} \Delta_n^l \nabla^2_{lm} f(\theta_n) \widehat\Delta_n^m \right | \F_n\right]
\end{align}
In analysing the RHS of the above expression, the second term in the expectation  is mean-zero term:  
\begin{align}
\E\left[\left. \frac{1}{\Delta_n^i \widehat\Delta_n^j} \mathop{\sum_{l=1}^N \sum_{m=1}^{N}}_{lm \neq ij} \Delta_n^l \nabla^2_{lm} f(\theta_n) \widehat\Delta_n^m \right | \F_n\right] = 0.
\end{align}
The term on the LHS above, can be denoted by $\Psi_n^2(\nabla^2 f(\theta_n))$, can be written in the matrix form as follows:
\begin{align}
\Psi_n^2(\nabla^2 f(\theta_n)) = \widehat N_n\tr \nabla^2 f(\theta_n) N_n  + \widehat N_n\tr \nabla^2 f(\theta_n) + \nabla^2 f(\theta_n) N_n.
\end{align}
where $N_n,\widehat N_n$ are defined as follows $N_n = \Delta_n \left[\frac{1}{\Delta_n^1},\ldots,\frac{1}{\Delta_n^N}\right] - I_N$ and $\widehat N_n = \widehat \Delta_n \left [\frac{1}{\widehat\Delta_n^1},\ldots,\frac{1}{\widehat \Delta_n^N}\right] - I_N$ and $I_N$ is  identity matrix of size $N \times N$.
From the foregoing, the per-iteration Hessian estimate $\widehat H_n$ can be re-written as follows:
\begin{align}
\widehat H_n = & \nabla^2 f(\theta_n) + \Psi_{n}(\nabla^2 f(\theta_n)) +  O(\delta_n^2) + O(\delta_n^{-2})  \label{eq:spsahnhat}
\end{align}
where, for any matrix $H$, 
\begin{align}
\Psi_{n}(H) &=  \Psi_{n}^{1}(H) + \Psi_{n}^{2}(H)\nonumber\\
&=  \frac{1}{2}M_n \left[\Delta_n\tr H \Delta_n + \widehat\Delta_n\tr H \widehat\Delta_n\right] + \widehat N_n\tr H N_n + \widehat N_n\tr H + H N_n.\label{eq:spsapsi}
\end{align}
Given that we operate in a simulation optimization setting, which implies $\nabla^2 f$ is not known, we construct the feedback term $\widehat \Psi_n$ in \eqref{eq:2spsa-3-ih} by using $\overline H_{n-1}$ as a proxy for $\nabla^2 f$, i.e.,
\begin{align}
\widehat \Psi_n = \Psi_{n} (\overline H_{n-1}).
\label{eq:psinhatspsa3}
\end{align}

\paragraph{\textbf{Optimizing the step-sizes }$b_n$}

The optimal choice for $b_n$ in \eqref{eq:2rdsa-H}  is the following:
\begin{align}
\label{eq:spsawieghts}
b_i  = \delta_i^{4}/\sum\limits_{j=0}^{i} \delta_j^{4}.
\end{align}
The main idea behind the above choice is provided below.
From \eqref{eq:spsahnhat}, we can infer that
\begin{align*}
\E \parallel \widehat H_n \parallel^2 \le \dfrac{C}{\delta_n^4} \text{ for some } C<\infty. 
\end{align*} 
This is because the third term in \eqref{eq:spsahnhat} vanishes asymptotically, while the fourth term there dominates asymptotically. Moreover, the noise factors in the fourth term in \eqref{eq:spsahnhat} are bounded above due to (C21)  and independent of $n$, leaving the $\delta_n^2$ term in the denominator there.

So, the optimization problem to be solved at instant $n$ is as follows:
\begin{align}
\sum \limits_{i=0}^{n} (\tilde b_k)^2 \delta_i^{-4}, \text{ subject to} \label{eq:spsawn-opt}\\
\tilde b_i \geq 0 ~\forall i \text{ and }\sum \limits_{i=0}^{n} \tilde b_i = 1.
\end{align}
The optimization variable $\tilde b_i$ from the above is related to the Hessian recursion \eqref{eq:2spsa-3-ih} as follows:
\begin{align}
\label{eq:spsahess}
\overline H_n = \sum\limits_{i=0}^{n} \tilde b_k(\widehat H_i -\widehat \Psi_i).
\end{align}
The solution to \eqref{eq:spsawn-opt} is achieved for $\tilde b_i^* = \delta_i^{4}/\sum \limits_{j=0}^{n} \delta_j^{4}, i=1,\ldots,n$. The optimal choice $\tilde b_i^*$ can be translated to the step-sizes $b_i$, leading to \eqref{eq:spsawieghts}.

\section{Convergence analysis for 2SPSA-3-IH}
\label{sec:2spsa-results}
We make the same assumptions as those used in the analysis of \cite{prashanth2015rdsa}, with a few minor alterations. The assumptions are listed below:
\begin{enumerate}[label=(\textbf{C\arabic*}),resume]
\item  The function
$f$ is four-times differentiable\footnote{Here $\nabla^4 f(\theta) = \dfrac{\partial^4 f (\theta)}{\partial \theta\tr \partial \theta\tr \partial \theta\tr \partial \theta\tr}$ denotes the fourth derivate of $f$ at $\theta$ and $\nabla^4_{i_1 i_2 i_3 i_4} f(\theta)$ denotes the $(i_1 i_2 i_3 i_4)$th entry of $\nabla^4 f(\theta)$, for $i_1, i_2, i_3,i_4=1,\ldots, N$.} with $\left|\nabla^4_{i_1 i_2 i_3 i_4} f(\theta) \right| < \infty$, for $i_1, i_2, i_3,i_4=1,\ldots, N$ and for all $\theta\in \R^N$.

%\item  For some $\rho>0$  and almost all $\theta_n$, the function $f$ is four-times differentiable with a uniformly (in $n$) bounded fourth derivative for all $x$ such that $\left\| \theta_n - x\right\| \le \rho$.

\item For each $n$ and all $x$, there exists a $\rho>0$ not dependent on $n$ and $\theta$, such that $(\theta-\theta^*)\tr \bar f_n(\theta) \ge \rho \left\| \theta_n - \theta\right\|$, where $\bar f_n(\theta) = \Upsilon(\overline H_n)^{-1} \nabla f(\theta)$.

\item $\{\xi_n, \xi_n^+,\xi_n^-, n=1,2,\ldots\}$ are such that, for all $n$, $\E\left[\left. \xi_n^+ + \xi_n^- - 2 \xi_n \right| \F_n, \Delta_n, \widehat \Delta_n \right] = 0$, where $\mathcal{F}_n = \sigma(\theta_m,m < n)$ denotes the underlying sigma-field..

\item $\{\Delta_n^i,\widehat\Delta_n^i, i=1,\ldots,N, n=1,2,\ldots\}$ are i.i.d independent of $\F_n$ and for some $\alpha_0 > 0$ and for all $n,l$ $|\Delta_{nl}|\le \alpha_0$, $\Delta_{nl}$ is symmetrically distributed about 0, $\Delta_{nl}$ are mutually independent across $n$ and $l$ and they satisfy $\E (\Delta_{nl}^{-2})$,$\E (\widehat\Delta_{nl}^{-2}) \le \alpha_0$.

\item  The step-sizes $a_n$ and perturbation constants $\delta_n$ are positive, for all $n$ and satisfy
$$\hspace{-1.7em} a_n, \delta_n \rightarrow 0\text{ as } n \rightarrow \infty, 
\sum_n a_n=\infty \text{ and } \sum_n \left(\frac{a_n}{\delta_n}\right)^2 <\infty.$$

\item For each $i=1,\ldots,N$ and any $\rho>0$, 
$P(\{ \bar f_{ni} (\theta_n) \ge 0 \text{ i.o}\} \cap \{ \bar f_{ni} (\theta_n) < 0 \text{ i.o}\} \mid \{ |\theta_{ni} - \theta^*_i| \ge \rho\quad \forall n\}) =0.$

\item The operator $\Upsilon$ satisfies $\delta_n^2 \Upsilon(H_n)^{-1} \rightarrow 0$ a.s. and  $E(\left\| \Upsilon(H_n)^{-1}\right\|^{2+\zeta}) \le \rho$ for some $\zeta, \rho>0$.

\item For any $\tau >0$ and nonempty $S \subseteq \{1,\ldots,N\}$, there exists a $\rho'(\tau,S)>\tau$ such that 
$$ \limsup_{n\rightarrow \infty} \left| \dfrac{\sum_{i \notin S} (\theta-\theta^*)_i \bar f_{ni}(\theta)}{\sum_{i \in S} (\theta-\theta^*)_i \bar f_{ni}(\theta)}               \right| < 1 \text{ a.s.}$$
for all $|(\theta-\theta^*)_i| < \tau$ when $i \notin S$ and $|(\theta-\theta^*)_i| \ge \rho'(\tau,S)$ when $i\in S$.
\item For some $\alpha_0, \alpha_1, \alpha_2 >0$ and for all $n,l,m$, $\E {\xi_n}^{2} \le \alpha_0$, $\E {\xi_n^{\pm}}^{2} \le \alpha_0$, $\E f(\theta_n)^{2} \le \alpha_1$,  $\E f(\theta_n + \delta_n \Delta_n + \delta_n \widehat \Delta_n)^{2},\E f(\theta_n - \delta_n \Delta_n - \delta_n \widehat \Delta_n)^{2} \le \alpha_1$, 
%$\E \left[ f(\theta_n + \delta_n \Delta_n + \delta_n \widehat \Delta_n)^{2}/ (\Delta_{nl}\widehat\Delta_{nm})^2 \mid \F_n\right]$,
$\E \left[| f(\theta_n + \delta_n \Delta_n + \delta_n \widehat \Delta_n)/ (\Delta_{nl}\widehat\Delta_{nm})|^{2+\alpha_2} \mid \F_n\right]$,
%$\E \left[ f(\theta_n - \delta_n \Delta_n - \delta_n \widehat \Delta_n)^{2}/ (\Delta_{nl}\widehat\Delta_{nm})^2 \mid \F_n\right]$,
$\E \left[ |f(\theta_n - \delta_n \Delta_n - \delta_n \widehat \Delta_n)/ (\Delta_{nl}\widehat\Delta_{nm})|^{2+\alpha_2} \mid \F_n\right]$,
 $\E\left[\left. (\xi_n^+ + \xi_n^- - 2 \xi_n)^2/(\Delta_{nl}\widehat\Delta_{nm})^2 \right| \F_n \right] \le \alpha_1$ 
and $\E \left(\left\| \Upsilon(\overline H_n) \right\|^2 \mid \F_n\right) \le \alpha_1$. 
\item  $\delta_n = \frac{\delta_0}{(n+1)^{\varsigma}}$, where $\delta_0 > 0$ and $0 < \varsigma \le 1/8$.
\end{enumerate}
The reader is referred to Section II-B of \cite{prashanth2015rdsa} for a detailed discussion of the above assumptions. We remark here that (C13)-(C20) are identical to that in \cite{prashanth2015rdsa}, while (C21) and (C22) introduce minor additional requirements on $\left\| \Upsilon(\overline H_n \right\|^2$ and $\delta_n$, respectively and these are inspired by \cite{spall-jacobian}.

\begin{lemma}(\textbf{2SPSA-3-IH Bias in Hessian estimate})
\label{lemma:2spsa-bias}
Under (C13)-(C22), with $\widehat H_n$ defined according to  \eqref{eq:2spsa-ber} , we have a.s. that\footnote{Here $\widehat H_n(i,j)$ and $\nabla^2_{ij}f(\cdot)$ denote the $(i,j)$th entry in the Hessian estimate $\widehat H_n$ and the true Hessian $\nabla^2 f(\cdot)$, respectively.}, for $i,j = 1,\ldots,N$,
\begin{align}
\left|\E\left[
\left. \widehat H_n(i,j) \right| \F_n \right] - \nabla^2_{ij} f(\theta_n)\right| = O(\delta_n^2).
\end{align} 
\end{lemma}
\begin{proof}
See Proposition 4.2 in \cite{bhatnagar2015simultaneous}.
\end{proof}

\begin{theorem}(\textbf{2SPSA-3-IH Strong Convergence of Hessian})
\label{thm:2spsa-H}
Under (C13)-(C22), we have that 
$$\theta_n \rightarrow \theta^*, \overline H_n \rightarrow \nabla^2 f(\theta^*) \text{ a.s. as } n\rightarrow \infty.$$ 
In the above, $\theta_n$ and $\overline H_n$ are updated according to \eqref{eq:2spsa-3-theta} and \eqref{eq:2spsa-3-ih}, respectively, $\widehat H_n$ defined according to \eqref{eq:2spsa-ber} and the step-sizes $b_n$ are chosen as suggested in \eqref{eq:spsawieghts}. 
\end{theorem}
\begin{proof}
This proof is similar to the proof of the theorem \ref{thm:2rdsa-H}
\end{proof}

We next present a convergence rate result for the special case of a quadratic objective function under the following additional assumptions:
\begin{enumerate}[label=(\textbf{C\arabic*}),resume]
\item  $f$ is quadratic and $\nabla^2 f(x^*) > 0$. 
\item The operator $\Upsilon$ is chosen such that $\E(\parallel \Upsilon(\overline H_n) - \overline H_n\parallel^2) = o(e^{-2wn^{1-r}/(1-r)})$ and $\parallel \Upsilon(H) - H \parallel^2 / (1+\parallel H \parallel^2)$ is uniformly bounded.
\end{enumerate}
\begin{theorem}(\textbf{2SPSA-3-IH Quadratic case - Convergence rate})
\label{thm:spsaquad-bound}
Assume (C16), (C22), (C23) and (C24) and also that the setting is noise-free. 
Let $b_n = b_0/n^r$, $n=1,2,\ldots,k$, where $1/2 < r < 1$ and $0 < b_0 \leq 1$. For notational simplicity, let $H^*=\nabla^2 f(x^*)$. Letting $\Lambda_k = \overline H_k - H^*$, we have 
\begin{align}
\text{trace}[\E (\Lambda_n \tr \Lambda_n)] = O(e^{-2b_0n^{1-r} / {1-r}}).
\label{eq:spsaquad-bigo}
\end{align}
\end{theorem}
\begin{proof}
Since the setting is noise-free with a quadratic objective, we can rewrite \eqref{eq:spsahnhat} as follows:
\begin{align}
 \widehat H_n =    \nabla^2 f(\theta_n) +\Psi_{n}(\nabla^2 f(\theta_n)),\label{eq:spsahnhat-ext}
\end{align}
where $\Psi_n(H)$ is defined in \eqref{eq:psinhatspsa3}.
The proof involves the following steps:
\begin{description}
  \item[Step 1:] Here we prove the MSE convergence of $\overline H_k$, i.e., $\E [\Lambda_n\tr \Lambda_n] \rightarrow 0$ a.s. as $n\rightarrow \infty$.
  \item[Step 2:] We unroll the recursion \eqref{eq:2rdsa-H} and then derive convenient representation for $\text{trace}[\E (\Lambda_n \tr \Lambda_n)]$.
  \item[Step 3:] We derive the main result in \eqref{eq:quad-bigo} using a proof by contradiction. 
\end{description}
\noindent\textbf{Step 1: MSE convergence of $\overline H_n$} \\
This part in exactly the same manner as part (i) in Theorem 3 of \cite{spall-jacobian}.\\
\noindent\textbf{Step 2: Representation of $\text{trace}[\E (\Lambda_n \tr \Lambda_n)]$} \\
This part in exactly the same manner as part (ii) in Theorem 3 of \cite{spall-jacobian}.\\
\noindent\textbf{Step 3: The big-O result on $\text{trace}[\E (\Lambda_n \tr \Lambda_n)]$ convergence} \\
This part in exactly the same manner as part (iii) in Theorem 3 of \cite{spall-jacobian}.\\
\end{proof}

### 2016-12-24 secorder-rdsa-ih.tex
maintexfiles added

\chapter{Second-order RDSA with improved hessian estimation (2RDSA-IH)}
\label{sec:2rdsa-ih}
The second-order RDSA with improved Hessian estimate performs an update iteration as follows:
\begin{align}
\label{eq:e2rdsa}
\theta_{n+1} = & \; \theta_n - a_n \Upsilon(\overline H_n)^{-1}\widehat\nabla f(\theta_n), \\
\overline H_n = & \; (1-b_{n})  \overline H_{n-1} + b_{n} ( \widehat H_n - \widehat \Psi_n),\label{eq:2rdsa-H}
\end{align}
where $\widehat\nabla f(\theta_n)$ is the estimate of $\nabla f(\theta_n)$, 
 %and this corresponds to  the uniform  and   asymmetric Bernoulli variant gradient estimates proposed in \cite{prashanth2015rdsa}.
$\overline H_n$ is an estimate of the true Hessian ${\nabla}^2 f(\cdot)$, $\Upsilon(\cdot)$ projects any matrix onto the set of positive definite matrices and $\{a_n, n\ge 0\}$ is a step-size sequence that satisfies standard stochastic approximation conditions. There are standard procedures such as Cholesky factorization, see \cite{bert22}, for projecting a given square matrix to set of positive definite matrices. Moreover, in the vicinity of a local minimum, one expects the Hessian to be positive definite. In such a case, $\Upsilon$ will represent the identity operator.

The recursion \eqref{eq:e2rdsa} is identical to that in 2RDSA, while the Hessian estimation recursion \eqref{eq:2rdsa-H} differs as follows:\\
\begin{inparaenum}[\bfseries (i)]
\item  $\widehat \Psi_n$ is a zero-mean feedback term that reduces the error in Hessian estimate; and\\
\item $b_n$ is a general step-size that we optimize to improve the Hessian estimate.
\end{inparaenum}

On the other hand, $\widehat H_n$ is identical to that in 2RDSA, i.e., it estimates the true Hessian in each iteration using $3$ function evaluations. %Further, $\widehat H_0 = I$. 
For the sake of completeness, we provide below the construction for $\widehat\nabla f(\theta_n)$ and $\widehat H_n$ using asymmetric Bernoulli perturbations, before we present the feedback term that reduces the error in $\widehat H_n$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\algblock{PEval}{EndPEval}
%\algnewcommand\algorithmicPEval{\textbf{\em Function evaluation 1}}
% \algnewcommand\algorithmicendPEval{}
%\algrenewtext{PEval}[1]{\algorithmicPEval\ #1}
%\algrenewtext{EndPEval}{\algorithmicendPEval}
%
%\algblock{PEvalPrime}{EndPEvalPrime}
%\algnewcommand\algorithmicPEvalPrime{\textbf{\em Function evaluation 2}}
% \algnewcommand\algorithmicendPEvalPrime{}
%\algrenewtext{PEvalPrime}[1]{\algorithmicPEvalPrime\ #1}
%\algrenewtext{EndPEvalPrime}{\algorithmicendPEvalPrime}
%
%\algblock{PImp}{EndPImp}
%\algnewcommand\algorithmicPImp{\textbf{\em Gradient descent}}
% \algnewcommand\algorithmicendPImp{}
%\algrenewtext{PImp}[1]{\algorithmicPImp\ #1}
%\algrenewtext{EndPImp}{\algorithmicendPImp}
%
%\algtext*{EndPEval}
%\algtext*{EndPEvalPrime}
%\algtext*{EndPImp}
%%%%%%%%%%%%%%%%% alg-custom-block %%%%%%%%%%%%
%\algblock{PEvalPrimeDouble}{EndPEvalPrimeDouble}
%\algnewcommand\algorithmicPEvalPrimeDouble{\textbf{\em Function evaluation 3}}
% \algnewcommand\algorithmicendPEvalPrimeDouble{}
%\algrenewtext{PEvalPrimeDouble}[1]{\algorithmicPEvalPrimeDouble\ #1}
%\algrenewtext{EndPEvalPrimeDouble}{\algorithmicendPEvalPrimeDouble}
%\algtext*{EndPEvalPrimeDouble}
%
%\algblock{PImpNewton}{EndPImpNewton}
%\algnewcommand\algorithmicPImpNewton{\textbf{\em Newton step}}
% \algnewcommand\algorithmicendPImpNewton{}
%\algrenewtext{PImpNewton}[1]{\algorithmicPImpNewton\ #1}
%\algrenewtext{EndPImpNewton}{\algorithmicendPImpNewton}
%
%\algtext*{EndPImpNewton}
%
%%%%%%%%%%%%%%%%%%%%
%
%\begin{algorithm}[t]
%\begin{algorithmic}
%\State {\bf Input:} 
%initial parameter $x_0 \in \R^N$, perturbation constants $\delta_n>0$, step-sizes $\{a_n, b_n\}$, operator $\Upsilon$.
%\For{$n = 0,1,2,\ldots$}	
%	\State Generate $\{d_n^{i}, i=1,\ldots,N\}$, independent of $\{d_m, m=0,1,\ldots,n-1\}$. 
%	\State For any $i=1,\ldots,N$, $d_n^{i}$ is distributed either as an asymmetric Bernoulli (see \eqref{eq:det-proj}) or Uniform $U[-\eta,\eta]$ for some $\eta >0$ (see Remark \ref{remark:unif}). 
%	\PEval
%	    \State Obtain $y_n^+ = f(x_n+\delta_n d_n) + \xi_n^+$.
%  \EndPEval
%	    \PEvalPrime
%	    \State Obtain $y_n^- = f(x_n-\delta_n d_n) + \xi_n^-$.
%	    \EndPEvalPrime
%	    	    \PEvalPrimeDouble
%	    \State Obtain $y_n = f(x_n) + \xi_n$.
%	    \EndPEvalPrimeDouble
%	    \PImpNewton
%		\State Update the parameter and Hessian as follows:
%		\begin{align*}
%		x_{n+1} = & \; x_n - a_n \Upsilon(\overline H_n)^{-1}\widehat\nabla f(x_n), \\
%\overline H_n = &\; (1-b_{n})  \overline H_{n-1} + b_{n} ( \widehat H_n - \widehat \Psi_n),
%\end{align*}
%where $\widehat H_n$ and $\widehat \Psi_n$ are chosen according to \eqref{eq:2rdsa-estimate-ber} and \eqref{eq:psinhat}, respectively. 
%		\EndPImpNewton
%\EndFor
%\State {\bf Return} $x_n.$
%\end{algorithmic}
%\caption{Structure of 2RDSA-IH algorithm.}
%\label{alg:structure}
%\end{algorithm}
%
\begin{algorithm}[!htp]
\KwIn{initial parameter $\theta_0 \in \R^N$, perturbation constants $\delta_n>0$, step-sizes $\{a_n, b_n\}$, operator $\Upsilon$.}
%\KwOut{Policy parameter $\theta$, value function parameter $r$ and feature $\Phi$ }
\begin{enumerate}

 \item \textbf{Execution}: \\
 \For{$n \gets 0,1,2, \ldots, $} {
 \begin{itemize}
 \item Generate $\{d_n^{i}, i=1,\ldots,N\}$, independent of $\{d_m, m=0,1,\ldots,n-1\}$.  
 \item For any $i=1,\ldots,N$, $d_n^{i}$ is distributed either as an asymmetric Bernoulli (see \eqref{eq:det-proj}) or Uniform $U[-\eta,\eta]$ for some $\eta >0$ (see Remark \ref{remark:unif}).
\begin{itemize}
\item \textbf{\emph {Function evaluation 1} }\\ \hspace{4em} Obtain $y_n^+ = f(\theta_n+\delta_n d_n) + \xi_n^+$.
\item \textbf{\emph {Function evaluation 2} }\\ \hspace{4em} Obtain $y_n^- = f(\theta_n-\delta_n d_n) + \xi_n^-$.
\item \textbf{\emph {Function evaluation 3} }\\ \hspace{4em} Obtain $y_n = f(\theta_n) + \xi_n$.
\item \textbf{\emph {Newton step}}\\ \hspace{4.2em} Update the parameter and Hessian as follows:
%\begin{align*}
%		x_{n+1} = & \; x_n - a_n \Upsilon(\overline H_n)^{-1}\widehat\nabla f(x_n), \\
%\overline H_n = &\; (1-b_{n})  \overline H_{n-1} + b_{n} ( \widehat H_n - \widehat \Psi_n),
%\end{align*}
\begin{align*} 
\theta_{n+1} = \,& \theta_n - a_n \Upsilon(\overline H_n)^{-1}\widehat\nabla f(\theta_n),\\
\overline H_n =\,& (1-b_{n})  \overline H_{n-1} + b_{n} ( \widehat H_n - \widehat \Psi_n).
\end{align*}
where $\widehat H_n$ and $\widehat \Psi_n$ are chosen according to \eqref{eq:2rdsa-estimate-ber} and \eqref{eq:psinhat}, respectively. 
\end{itemize}

\end{itemize}
}
\Return{$x_n.$}
\end{enumerate}

\caption{Structure of 2RDSA-IH algorithm.}
\label{alg:structure}
\end{algorithm}

Algorithm \ref{alg:structure} presents the pseudocode and we describe the individual component of 2RDSA-IH below.
\section{Function evaluations}
Let $\delta_n, n\geq 0$ denote a sequence of diminishing positive real numbers and $d_n = (d_n^1,\ldots,d_n^N)\tr$ denote a random perturbation vector at instant $n$,
where the perturbations $\{d_n^i, i=1,\ldots,N, n=1,2,\ldots\}$ are i.i.d. and distributed  as follows: 
\begin{equation}
\label{eq:det-proj}
 d_n^i =
  \begin{cases}
   -1 &  \text{ w.p. } \dfrac{(1+\epsilon)}{(2+\epsilon)}, \\
   1+\epsilon &  \text{ w.p. } \dfrac{1}{(2+\epsilon)},
  \end{cases}
\end{equation}
with $\epsilon>0$ being a constant that can be chosen to be arbitrarily small.

The 2RDSA-IH algorithm obtains three function samples $y_n$, $y_n^+$ and $y_n^-$ at $\theta_n$, $\theta_n+\delta_n d_n$ and $\theta_n - \delta_n d_n$, respectively, i.e.,
$y_n = f(\theta_n) + \xi_n$, $y_n^+ = f(\theta_n+\delta_n d_n) + \xi_n^+$ and 
$y_n^- = f(\theta_n-\delta_n d_n) + \xi_n^-$,
where the noise terms $\xi_n, \xi_n^+, \xi_n^-$ satisfy $\E\left[\left.\xi_n^+ + \xi_n^- - 2\xi_n \right| \F_n\right] = 0$ with $\F_n = \sigma(\theta_m,m\le n)$ denoting the underlying sigma-field.

\section{Gradient estimation}
The RDSA estimate of the gradient $\nabla f(\theta_n)$ is given by
\begin{align}
\label{eq:grad-ber}
\widehat\nabla f(\theta_n) = \frac1{1+\epsilon} d_n \left[ \dfrac{y_n^+ - y_n^-}{2\delta_n}\right],
\end{align}
%Note that $E d_n^i = 0$, $E (d_n^i)^2 = 1+\epsilon$ and $E (d_n^i)^4 = \dfrac{(1+\epsilon)(1+(1+\epsilon)^3)}{(2+\epsilon)}$.

\section{Hessian estimation}
\begin{align}
\label{eq:2rdsa-estimate-ber}
&\widehat H_n = M_n \left(\dfrac{y_n^+ + y_n^- - 2 y_n}{\delta_n^2}\right), \text{ where }\\
& M_n =
\left[
\begin{array}{ccc}
\frac{1}{\kappa}\left((d_n^1)^2\!-(1+\epsilon)\right) & \cdots & \frac{1}{2(1+\epsilon)^2}d_n^1 d_n^N\\
\frac{1}{2(1+\epsilon)^2}d_n^2 d_n^1  &  \cdots & \frac{1}{2(1+\epsilon)^2}d_n^2 d_n^N\\
\cdots&\cdots&\cdots\\
\frac{1}{2(1+\epsilon)^2}d_n^N d_n^1 & \cdots &  \frac{1}{\kappa}\left((d_n^N)^2-(1+\epsilon)\right) \\
\end{array}
\right],\nonumber
\end{align}
where $\kappa = \tau \left(1- \dfrac{(1+\epsilon)^2}{\tau}\right)$ and $\tau = E (d_n^i)^4= \dfrac{(1+\epsilon)(1+(1+\epsilon)^3)}{(2+\epsilon)}$, for any $i=1,\ldots,N$. 
%%%%%% Feedback

\section{Improved Hessian estimation}
The Hessian estimate $\widehat H_n$ can be simplified as follows: 
\begin{align}
\widehat H_n & = M_n \left(\dfrac{y_n^+ + y_n^- - 2 y_n}{\delta_n^2}\right) \nonumber \\
& =  M_n \left[\left(\dfrac{f(\theta_n+\delta_n d_n) + f(\theta_n-\delta_n d_n) - 2 f(\theta_n)}{\delta_n^2}\right) \right. +\left. \left(\dfrac{\xi_n^+ + \xi_n^- - 2 \xi_n}{\delta_n^2}\right)\right] \nonumber \\
& = M_n \left(d_n\tr \nabla^2 f(\theta_n) d_n +  O(\delta_n^2) + \left(\dfrac{\xi_n^+ + \xi_n^- - 2 \xi_n}{\delta_n^2} \right)\right).\label{eq:h0}
%& = M_n \left(d_n\tr \nabla^2 f(x_n) d_n +  O(\delta_n^2) + O(\delta_n^{-2})\right). \label{eq:h0}
\end{align}
For the first term on the RHS above, note that 
\begin{align}
&\E[M_n \left(d_n\tr \nabla^2 f(\theta_n) d_n\right) \mid \F_n] =   \E\left[\left. M_n \times\right.\right.\left.\left.\left(\sum\limits_{i=1}^{N-1} (d_n^i)^2 \nabla^2_{ii} f(\theta_n) + 2\sum\limits_{i=1}^N\sum\limits_{j=i+1}^N d_n^i d_n^j \nabla^2_{ij} f(\theta_n) \right)\right| \F_n\right]. \label{eq:h1}
\end{align}
In analyzing the $l$th diagonal term of the above expression, the following zero-mean term appears (see the proof of Lemma 4 in \cite{prashanth2015rdsa}):
\begin{align}
& \E\left[\left. \left([M_n]_{D}\right)_{l,l} \left(2\sum\limits_{i=1}^{N-1}\sum\limits_{j=i+1}^N d_n^i d_n^j \nabla^2_{ij} f(\theta_n)\right)\right| \F_{n}\right] = 0,\label{eq:diagzero}
\end{align}
where for any matrix $M$, $[M]_{D}$ refers to a matrix that retains only the diagonal entries of $M$ and replaces all the remaining  entries with zero, and $\left([M]_{D}\right)_{i,j}$ refers to the $(i,j)$th entry in $[M]_D$. We shall also use $[M]_{N}$ to refer to a matrix that retains only the off-diagonal entries of $M$, while replaces all the diagonal entries with zero.

The term on the LHS in \eqref{eq:diagzero}, denoted by $\Psi_{n}^{1}(\nabla^2 f(\theta_n))$, can be written in matrix form as follows: 
\begin{align}\label{eq:digmat}
\Psi_{n}^{1}(\nabla^2 f(\theta_n)) = [M_n]_{D}\left(d_{n}\tr \, [\nabla^2 f(\theta_n)]_{N} \, d_{n}\right).
\end{align}

In analyzing the off-diagonal term ($(k,l)$ where $k < l$) of \eqref{eq:h1}, the following zero-mean term appears:
\begin{align}
& \E\left[\left.\left([M_n]_{N}\right)_{k,l}   \left(\sum\limits_{i=1}^N (d_n^i)^2 \nabla^2_{ii} f(\theta_n)\right)\right| \F_n \right] = 0. 
\end{align}
The term on the LHS above, denoted by $\Psi_{n}^{2}(\nabla^2 f(\theta_n))$, can be written in matrix form as follows: 
\begin{align}\label{eq:ndmat}
\Psi_{n}^{2}(\nabla^2 f(\theta_n)) = [M_n]_{N}\left(d_{n}\tr \, [\nabla^2 f(\theta_n)]_{D} \, d_{n}\right).
\end{align}
From the foregoing, the per-iteration Hessian estimate $\widehat H_n$ can be re-written as follows:
\begin{align}
 \E\left[\left.\widehat H_n\right| \F_n\right] = & \nabla^2 f(\theta_n) + \E\left[\left.\Psi_{n}(\nabla^2 f(\theta_n))\right| \F_n\right]  +  O(\delta_n^2)+\E\left[\left.\left(\dfrac{\xi_n^+ + \xi_n^- - 2 \xi_n}{\delta_n^2} \right)\right| \F_n\right], \label{eq:hnhat}
\end{align}
where, for any matrix $H$, 
\begin{align}
\Psi_{n}(H) &= \Psi_{n}^{1}(H) + \Psi_{n}^{2}(H)\nonumber\\
&= [M_n]_{D}\left(d_{n}\tr \, [H]_{N} \, d_{n}\right) +  [M_n]_{N}\left(d_{n}\tr \, [H]_{D} \, d_{n}\right).\label{eq:psi}
\end{align}
In the RHS of \eqref{eq:hnhat}, it is easy to see that  the second term involving $\Psi_{n}$ and the last term involving the noise are zero-mean. Moreover, since the noise is bounded by assumption, the last term in \eqref{eq:hnhat} vanishes asymptotically at the rate $O(\delta_n^{-2})$. So, the error in estimating the Hessian is due to the second term, which involves the perturbations $d_n$. This motivates the term $\widehat \Psi_n$ in the update rule \eqref{eq:e2rdsa}. 
 %and using the fact that $\E\left[\left.\xi_n^+ + \xi_n^- - 2\xi_n \right| \F_n\right] = 0$

Given that we operate in a simulation optimization setting, which implies $\nabla^2 f$ is not known, we construct the feedback term $\widehat \Psi_n$ in \eqref{eq:e2rdsa} by using $\overline H_{n-1}$ as a proxy for $\nabla^2 f$, i.e.,
\begin{align}
\widehat \Psi_n = \Psi_{n} (\overline H_{n-1}).
\label{eq:psinhat}
\end{align}

\section{Step-size optimization}
Unlike the feedback term, adapting the idea of optimizing the step-sizes for 2RDSA is relatively straightforward from the corresponding approach for 2SPSA in \cite{spall-jacobian}. The difference here is that there exists only one $N$-dimensional perturbation vector $d_n$ in our setting, while 2SPSA required two such vectors. This in turn implies that only the perturbation constant $\delta_n$ is needed in optimizing $b_n$.

The optimal choice for $b_n$ in \eqref{eq:2rdsa-H} is the following:
\begin{align}
\label{eq:wieghts}
b_i  = \delta_i^{4}/\sum\limits_{j=0}^{i} \delta_j^{4}.
\end{align}
The main idea behind the above choice is provided below.
From \eqref{eq:hnhat}, we can infer that
\begin{align*}
\E \parallel \widehat H_n \parallel^2 \le \dfrac{C}{\delta_n^4} \text{ for some } C<\infty. 
\end{align*} 
This is because the third term in \eqref{eq:hnhat} vanishes asymptotically, while the fourth term there dominates asymptotically. Moreover, the noise factors in the fourth term in \eqref{eq:hnhat} are bounded above due to (C9) and independent of $n$, leaving the $\delta_n^2$ term in the denominator there.

So, the optimization problem to be solved at instant $n$ is as follows:
\begin{align}
\min_{ \{\tilde b_k\} } \sum \limits_{i=0}^{n} (\tilde b_k)^2 \delta_i^{-4}, \text{ subject to} \label{eq:wn-opt}\\
\tilde b_i \geq 0 ~\forall i \text{ and }\sum \limits_{i=0}^{n} \tilde b_i = 1.
\end{align}
The optimization variable $\tilde b_i$ from the above is related to the Hessian recursion \eqref{eq:2rdsa-H} as follows:
\begin{align}
\label{eq:hess}
\overline H_n = \sum\limits_{i=0}^{n} \tilde b_k(\widehat H_i -\widehat \Psi_i).
\end{align}
The solution to \eqref{eq:wn-opt} is achieved for $\tilde b_i^* = \delta_i^{4}/\sum \limits_{j=0}^{n} \delta_j^{4}, i=1,\ldots,n$. The optimal choice $\tilde b_i^*$ can be translated to the step-sizes $b_i$, leading to \eqref{eq:wieghts}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{remark}\textbf{\textit{(Uniform perturbations)}}
\label{remark:unif}
In \cite{prashanth2015rdsa}, the authors suggest two alternatives for the distribution of random perturbations $d_n$: the asymmetric Bernoulli, which we described earlier and uniform that we outline next.

Choose $d_n^i$, $ i=1,\ldots,N$ to be i.i.d. $U[-\eta,\eta]$ for some $\eta>0$, where $U[-\eta,\eta]$ denotes the uniform distribution on the interval $[-\eta,\eta]$.
Then, the RDSA estimate of the gradient is given by
\begin{align}
\label{eq:grad-unif}
\widehat\nabla f(\theta_n) = \frac3{\eta^2} d_n \left[ \dfrac{y_n^+ - y_n^-}{2\delta_n}\right].
\end{align}

The Hessian estimate in this case is given by
\begin{align}
\label{eq:2rdsa-estimate-unif}
&\widehat H_n =  M_n \left(\dfrac{y_n^+ + y_n^- - 2 y_n}{\delta_n^2}\right), \text{ where } \\
& M_n =
\dfrac{9}{2\eta^4}\left[
\begin{array}{cccc}
\frac{5}{2}\left((d_n^1)^2-\frac{\eta^2}{3}\right) & \cdots & d_n^1 d_n^N\\
d_n^2 d_n^1  &  \cdots & d_n^2 d_n^N\\
\cdots&\cdots&\cdots\\
d_n^N d_n^1 & \cdots &  \frac{5}{2}\left((d_n^N)^2-\frac{\eta^2}{3}\right) \\
\end{array}
\right].\nonumber
\end{align}

The feedback term in \eqref{eq:psinhat} can be easily extended to the case of uniform perturbations by using the $M_n$ as defined above instead of that for the asymmetric Bernoulli case.
\end{remark}
%Henceforth, we shall refer to algorithm \eqref{eq:2rdsa}--\eqref{eq:2rdsa-H} with Hessian estimate \eqref{eq:2rdsa-estimate} as 2RDSA.

\section{Convergence analysis}
\label{sec:2rdsa-results}
We make the same assumptions as those used in the analysis of \cite{prashanth2015rdsa}, with a few minor alterations. The assumptions are listed below:
\begin{enumerate}[label=(\textbf{C\arabic*})]
\item  The function
$f$ is four-times differentiable\footnote{Here $\nabla^4 f(\theta) = \dfrac{\partial^4 f (\theta)}{\partial \theta\tr \partial \theta\tr \partial \theta\tr \partial \theta\tr}$ denotes the fourth derivate of $f$ at $\theta$ and $\nabla^4_{i_1 i_2 i_3 i_4} f(\theta)$ denotes the $(i_1 i_2 i_3 i_4)$th entry of $\nabla^4 f(\theta)$, for $i_1, i_2, i_3,i_4=1,\ldots, N$.} with $\left|\nabla^4_{i_1 i_2 i_3 i_4} f(\theta) \right| < \infty$, for $i_1, i_2, i_3,i_4=1,\ldots, N$ and for all $\theta\in \R^N$.

%\item  For some $\rho>0$  and almost all $x_n$, the function $f$ is four-times differentiable with a uniformly (in $n$) bounded fourth derivative for all $x$ such that $\left\| x_n - x\right\| \le \rho$.

\item For each $n$ and all $\theta$, there exists a $\rho>0$ not dependent on $n$ and $\theta$, such that $(\theta-\theta^*)\tr \bar f_n(\theta) \ge \rho \left\| \theta_n - \theta\right\|$, where $\bar f_n(\theta) = \Upsilon(\overline H_n)^{-1} \nabla f(\theta)$.

\item $\{\xi_n, \xi_n^+,\xi_n^-, n=1,2,\ldots\}$ are such that, for all $n$, $\E\left[\left. \xi_n^+ + \xi_n^- - 2 \xi_n \right| \F_n\right] = 0$, where $\mathcal{F}_n = \sigma(\theta_m,m\le n)$ denotes the underlying sigma-field..

\item $\{d_n^i, i=1,\ldots,N, n=1,2,\ldots\}$ are i.i.d. and independent of $\F_n$.

\item  The step-sizes $a_n$ and perturbation constants $\delta_n$ are positive, for all $n$ and satisfy
$$\hspace{-1.7em} a_n, \delta_n \rightarrow 0\text{ as } n \rightarrow \infty, 
\sum_n a_n=\infty \text{ and } \sum_n \left(\frac{a_n}{\delta_n}\right)^2 <\infty.$$

\item For each $i=1,\ldots,N$ and any $\rho>0$, 
$P(\{ \bar f_{ni} (\theta_n) \ge 0 \text{ i.o}\} \cap \{ \bar f_{ni} (\theta_n) < 0 \text{ i.o}\} \mid \{ |\theta_{ni} - x^*_i| \ge \rho\quad \forall n\}) =0.$

\item The operator $\Upsilon$ satisfies $\delta_n^2 \Upsilon(H_n)^{-1} \rightarrow 0$ a.s. and  $E(\left\| \Upsilon(H_n)^{-1}\right\|^{2+\zeta}) \le \rho$ for some $\zeta, \rho>0$.

\item For any $\tau >0$ and nonempty $S \subseteq \{1,\ldots,N\}$, there exists a $\rho'(\tau,S)>\tau$ such that 
$$ \limsup_{n\rightarrow \infty} \left| \dfrac{\sum_{i \notin S} (x-x^*)_i \bar f_{ni}(x)}{\sum_{i \in S} (x-x^*)_i \bar f_{ni}(x)}               \right| < 1 \text{ a.s.}$$
for all $|(x-x^*)_i| < \tau$ when $i \notin S$ and $|(x-x^*)_i| \ge \rho'(\tau,S)$ when $i\in S$.
\item For some $\alpha_0, \alpha_1>0$ and for all $n$, $\E {\xi_n}^{2} \le \alpha_0$, $\E {\xi_n^{\pm}}^{2} \le \alpha_0$, $\E f(\theta_n)^{2} \le \alpha_1$,  $\E f(\theta_n\pm \delta_n d_n)^{2} \le \alpha_1$ and $\E \left(\left\| \Upsilon(\overline H_n) \right\|^2 \mid \F_n\right) \le \alpha_1$. 
\item  $\delta_n = \frac{\delta_0}{(n+1)^{\varsigma}}$, where $\delta_0 > 0$ and $0 < \varsigma \le 1/8$.
\end{enumerate}
The reader is referred to Section II-B of \cite{prashanth2015rdsa} for a detailed discussion of the above assumptions. We remark here that (C1)-(C8) are identical to that in \cite{prashanth2015rdsa}, while (C9) and (C10) introduce minor additional requirements on $\left\| \Upsilon(\overline H_n \right\|^2$ and $\delta_n$, respectively and these are inspired by \cite{spall-jacobian}.

\begin{lemma}(\textbf{Bias in Hessian estimate})
\label{lemma:2rdsa-bias}
Under (C1)-(C10), with $\widehat H_n$ defined according to \eqref{eq:2rdsa-estimate-ber}, we have a.s. that\footnote{Here $\widehat H_n(i,j)$ and $\nabla^2_{ij}f(\cdot)$ denote the $(i,j)$th entry in the Hessian estimate $\widehat H_n$ and the true Hessian $\nabla^2 f(\cdot)$, respectively.}, for $i,j = 1,\ldots,N$,
\begin{align}
\left|\E\left[
\left. \widehat H_n(i,j) \right| \F_n \right] - \nabla^2_{ij} f(\theta_n)\right| = O(\delta_n^2).
\end{align} 
\end{lemma}
\begin{proof}
See Lemma 4 in \cite{prashanth2015rdsa}.
\end{proof}

\begin{theorem}(\textbf{Strong Convergence of Hessian})
\label{thm:2rdsa-H}
Under (C1)-(C10), we have that 
$$\theta_n \rightarrow \theta^*, \overline H_n \rightarrow \nabla^2 f(\theta^*) \text{ a.s. as } n\rightarrow \infty.$$ 
In the above, $\theta_n$ and $\overline H_n$ are updated according to \eqref{eq:e2rdsa} and \eqref{eq:2rdsa-H}, respectively, $\widehat H_n$ defined according to \eqref{eq:2rdsa-estimate-ber} and the step-sizes $b_n$ are chosen as suggested in \eqref{eq:wieghts}. 
\end{theorem}
\begin{proof}
The first part of the claim regarding $\theta_n$ follows in exactly the same fashion as the proof of Theorem 5 in \cite{prashanth2015rdsa}.  
For proving the claim regarding $\overline H_n$, we closely follow the approach used to prove a corresponding result for 2SPSA (see Theorem 1 in \cite{spall-jacobian}). 
The first step is to prove the following:
\begin{align}
\sum_{k=0}^n \dfrac{\delta_k^4 \left(\widehat H_k - \widehat \Psi_k - \E(\widehat H_k \mid \F_k)\right)}{\sum_{i=0}^n \delta_i^4} \rightarrow 0.
\label{eq:step1}
\end{align}

By a completely parallel argument to that used in the proof of Theorem 1 in \cite{spall-jacobian}, we obtain: For any $i,j = 1,\ldots,N$,
\begin{align*}
\E \left[\left((\widehat {H}_k)_{i,j} - (\widehat {\Psi}_k)_{i,j} - \E((\widehat {H}_k)_{i,j} \mid \F_k)\right)^2\right] = O(\delta_k^{-4}).
\end{align*}
Now \eqref{eq:step1} follows by an application of Kronecker's Lemma along with the martingale convergence theorem (see Theorem 6.2.1 of \cite{lahaprobability}).

From Lemma \ref{lemma:2rdsa-bias}, we have 
$$ \E[ \widehat H_k \mid \F_k] = \nabla^2 f(\theta_n) + O(\delta_n^2) \text{ a.s.}$$
Since the Hessian is continuous near $\theta_n$ and $\theta_n$ converges almost surely to $\theta^*$, we have
% \footnote{Since assumptions (C1)-(C10) here are similar to that in \cite{prashanth2015rdsa}, Theorem 5 of \cite{prashanth2015rdsa} holds here as well. In particular, this implies almost sure convergence of $x_n$ to $x^*$.}
\begin{align*}
\sum_{k=0}^n \dfrac{\delta_k^4 \left(\E(\widehat H_k \mid \F_k)\right)}{\sum_{i=0}^n \delta_i^4} 
=&\sum_{k=0}^n \dfrac{\delta_k^4 \left(\nabla^2 f(\theta_n) + O(\delta_n^2)\right)}{\sum_{i=0}^n \delta_i^4}\\
=&\sum_{k=0}^n \dfrac{\delta_k^4 \left(\nabla^2 f(x^*) + o(1)\right)}{\sum_{i=0}^n \delta_i^4}\\
&\rightarrow \nabla^2 f(x^*) \text{ a.s. as } n \rightarrow \infty.
\end{align*}
The last step above follows from Toeplitz Lemma (see p. 89 of \cite{lahaprobability}) after observing that $\sum_{i=0}^n \delta_i^4 \rightarrow \infty$ due to (C10). 
The main claim now follows since 
$$ \overline H_n = \sum_{k=0}^n \dfrac{\delta_k^4 \left(\widehat H_k - \Psi_k \right)}{\sum_{i=0}^n \delta_i^4}.$$
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
%%%%%%%%%%%%%%%%%%%%%%% Quadratic case
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We next present a convergence rate result for the special case of a quadratic objective function under the following additional assumptions:
\begin{enumerate}[label=(\textbf{C\arabic*}),resume]
\item  $f$ is quadratic and $\nabla^2 f(x^*) > 0$. 
\item The operator $\Upsilon$ is chosen such that $\E(\parallel \Upsilon(\overline H_n) - \overline H_n\parallel^2) = o(e^{-2wn^{1-r}/(1-r)})$ and $\parallel \Upsilon(H) - H \parallel^2 / (1+\parallel H \parallel^2)$ is uniformly bounded.
\end{enumerate}
\begin{theorem}(\textbf{Quadratic case - Convergence rate})
\label{thm:quad-bound}
Assume (C4), (C10), (C11) and (C12) and also that the setting is noise-free. 
Let $b_n = b_0/n^r$, $n=1,2,\ldots,k$, where $1/2 < r < 1$ and $0 < b_0 \leq 1$. For notational simplicity, let $H^*=\nabla^2 f(x^*)$. Letting $\Lambda_k = \overline H_k - H^*$, we have 
\begin{align}
\text{trace}[\E (\Lambda_n \tr \Lambda_n)] = O(e^{-2b_0n^{1-r} / {1-r}}).
\label{eq:quad-bigo}
\end{align}
\end{theorem}
\begin{proof}
Since the setting is noise-free with a quadratic objective, we can rewrite \eqref{eq:hnhat} as follows:
\begin{align}
 \widehat H_n =    \Phi_n(\nabla^2 f(\theta_n)) &+\Psi_{n}(\nabla^2 f(\theta_n)),\label{eq:hnhat-ext}
\end{align}
where $\Psi_n(H)$ is defined in \eqref{eq:psi} and for any matrix $H$, 
$$\Phi_{n}(H) = [M_n]_{D}\left(d_{n}\tr \, [H]_{D} \, d_{n}\right) +  [M_n]_{N}\left(d_{n}\tr \, [H]_{N} \, d_{n}\right).$$
The proof involves the following steps:
\begin{description}
  \item[Step 1:] Here we prove the MSE convergence of $\overline H_k$, i.e., $\E [\Lambda_n\tr \Lambda_n] \rightarrow 0$ a.s. as $n\rightarrow \infty$.
  \item[Step 2:] We unroll the recursion \eqref{eq:2rdsa-H} and then derive convenient representation for $\text{trace}[\E (\Lambda_n \tr \Lambda_n)]$.
  \item[Step 3:] We derive the main result in \eqref{eq:quad-bigo} using a proof by contradiction. 
\end{description}
\noindent\textbf{Step 1: MSE convergence of $\overline H_n$} \\
This part in exactly the same manner as part (i) in Theorem 3 of \cite{spall-jacobian}.\\
\noindent\textbf{Step 2: Representation of $\text{trace}[\E (\Lambda_n \tr \Lambda_n)]$} \\
 From \eqref{eq:2rdsa-H} and \eqref{eq:hnhat-ext},  we have
 \begin{align}
 \Lambda_n &= \Lambda_{n-1} - b_n ( \overline H_{n-1} - \hat H_n + \hat \Psi_n) \nonumber\\
&= (1-b_n) \Lambda_{n-1} - b_n (H^* + \hat \Psi_n - \hat H_n ) \nonumber\\&= (1-b_n) \Lambda_{n-1} - b_n (H^*+\hat \Psi_n -\Phi_n(H^*) - \Psi_n(H^*)\nonumber\\&= (1-b_n) \Lambda_{n-1} - b_n \Psi_n(\Upsilon(\overline H_{n-1})) + b_n (\Phi_n(H^*) - H^*)\nonumber\\
&= (1-b_n) \Lambda_{n-1} - b_n\Psi_n(\Lambda_{n-1}')+ b_n (\Phi_n(H^*) - H^*),\label{eq:step1last}
 \end{align}
where $\Lambda_n' = \Upsilon(\overline H_n) - H^*$.
The equality in \eqref{eq:step1last} follows from \eqref{eq:psi}. 
Unrolling the recursion \eqref{eq:step1last}, we obtain 
 \begin{align}\label{lambda-exp}
  \Lambda_n  = \left[ \prod_{k=1}^n (1-b_k) \right]\Lambda_0 
&- \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right] b_k \Psi_k(\Lambda_{k-1}') \nonumber\\
&+ \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right] b_k (\Phi_k(H^*) - H^*) \hspace{0.2cm} a.s.
 \end{align}
 %%(note : $\prod_{j=n+1}^n (1-b_j) = 1$ for all n).
 %%Let us  characterise $trace[\E(\Lambda_n\tr \Lambda_n)]$ using \eqref{lambda-exp}. From independence of $d_k$ along k, \eqref{lambda-exp} represent martingale difference sequence, leading to 
Squaring on both sides and taking expectations, we obtain
 \begin{align}
  \E (\Lambda_n \tr \Lambda_n) =  &\left[ \prod_{k=1}^n (1-b_k) \right]^2 \E (\Lambda_0 \tr \Lambda_0) + \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2   \E (\Psi_k(\Lambda_{k-1}') \tr \Psi_k(\Lambda_{k-1}'))  \nonumber\\ &+ \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2  \times \E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*)) \nonumber\\
&-2 \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2 \E\left(\Psi_k(\Lambda_{k-1}) \tr (\Phi_k(H^*) - H^*)\right).\label{eq:exp-lambda2}
 \end{align}
The equality above uses the fact that $\E(\Psi_k(\Lambda_{k-1}'))=0$, $\E(\Phi_k(H^*) - H^*)=0$, which get rid of the corresponding cross terms with the first term on RHS of \eqref{lambda-exp}. Notice that there are only $n$ terms in the last term in the RHS of \eqref{eq:exp-lambda2} because $d_k$ are identically distributed for each $k$ and across $k$.  
% %
We now characterize $\text{trace}[\E(\Lambda_n\tr \Lambda_n)]$ using \eqref{eq:exp-lambda2} as follows:
%From the independence of $d_k$'s and $\Lambda_{k-1}'$ and using Cauchy-Schwartz inequality we get 
 \begin{align}\label{eq:trace}
 \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] = &\left[ \prod_{k=1}^n (1-b_k) \right]^2 \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  + \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2  \,\,\tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ &+ \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2   \times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right]  \nonumber\\ &- 2 \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2  \times \text{trace} \left[\E\left(\Psi_k(\Lambda_{k-1}) \tr (\Phi_k(H^*) - H^*)\right)\right].
 \end{align}
 % \begin{align}\label{eq:trace}
% & \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] \leq \left[ \prod_{k=1}^n (1-b_k) \right]^2 \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  \nonumber\\ &+ \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2  \,\,\tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ &+ \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2  \nonumber\\&  \hspace{0.5cm}\times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right] + \nonumber\\ &+ \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2   \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)^{1/2} \nonumber\\&  \times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right]^{1/2}.
% \end{align}

As in the proof of Theorem 3 of \cite{spall-jacobian}, observe that $1-b_k = e^{-b_k}(1-O(b_k^2))$ and since $0 < b_k <1$, we have that the $O(b_k^2)$ term is strictly positive. 
Letting $\Gamma_{ij} = \sum_{k=i}^j b_k$ with $\Gamma_{nn} = 1$ and $\beta_{kn} = \left[\prod_{i=k+1}^n (1- O(b_i^2))\right]^2$, we can simplify \eqref{eq:trace} as follows:
\begin{align}\label{eq:wij}
 \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] = &e^{-2 \Gamma_{1n}} \beta_{0n} \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  + e^{-2 \Gamma_{1n}} \sum_{k=1}^n e^{2 \Gamma_{1k}} \beta_{kn} b_k^2  \,\,\tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ &+ e^{-2 \Gamma_{1n}} \sum_{k=1}^n e^{2 \Gamma_{1k}} \beta_{kn} b_k^2  \times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right] \nonumber\\ &- 2 e^{-2 \Gamma_{1n}} \sum_{k=1}^n e^{2 \Gamma_{1k}} \beta_{kn} b_k^2   \times \text{trace} \left[\E\left(\Psi_k(\Lambda_{k-1}) \tr (\Phi_k(H^*) - H^*)\right)\right] .
 \end{align} 
% \begin{align}\label{eq:wij}
% & \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] \leq e^{-2 \Gamma_{1n}} \beta_{0n} \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  \nonumber\\ &+ e^{-2 \Gamma_{1n}} \sum_{k=1}^n e^{2 \Gamma_{1k}} \beta_{kn} b_k^2  \,\,\tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ &+ e^{-2 \Gamma_{1n}} \sum_{k=1}^n e^{2 \Gamma_{1k}} \beta_{kn} b_k^2  \nonumber\\&  \hspace{0.5cm}\times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right] \nonumber\\ &+ e^{-2 \Gamma_{1n}} \sum_{k=1}^n e^{2 \Gamma_{1k}} \beta_{kn} b_k^2  \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)^{1/2} \nonumber\\&  \times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right]^{1/2}.
% \end{align} 
Comparing the sum with integrals, we obtain
\begin{align*}
\Gamma_{ij} = \int_i^j \frac{b_0}{x^r} dx + O(1) &= \left(\frac{b_0}{1-r}\right)(j^{1-r}-i^{1-r}) +O(1),
\end{align*}
where we have used the facts that $0 < b_k < 1, \forall k \ge 2$ and 
$\sum_{k=i}^j b_k \to \infty$ as $j-i \to \infty$ since 
$b_k = b_0/k^r$ with $r > 0.5$.
Observing that $\beta_{kn}$ are uniformly upper-bounded, say by $\bar \beta_n$, we have
\begin{align}\label{eq:thmwts}
 & \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] =  e^{-2 \Gamma_{1n}} \beta_{0n} \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  \nonumber\\ 
&+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}  \times \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ 
&+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}} \times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right] \nonumber\\ &- 2 \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}   \times \text{trace} \left[\E\left(\Psi_k(\Lambda_{k-1}) \tr (\Phi_k(H^*) - H^*)\right)\right].
\end{align} 
\begin{align}\label{eq:thmcomb}
& \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] =  e^{-2 \Gamma_{1n}} \beta_{0n} \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  \nonumber\\ 
&+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}  \times \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ 
&+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}  \text{trace}\left[\E [((\Phi_k(H^*) - H^*) - 2 \Psi_k(\Lambda_{k-1}))\tr  (\Phi_k(H^*) - H^*)]\right] .
\end{align} 
%
%\begin{align}\label{eq:thmwts}
% & \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] =  e^{-2 \Gamma_{1n}} \beta_{0n} \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  \nonumber\\ 
%&+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}} \nonumber\\
%& \hspace{3.5cm} \times \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ 
%&+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}  \nonumber\\
%&  \hspace{0.5cm}\times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right] \nonumber\\ &+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}  \nonumber\\&\hspace{3.5cm}\times \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)^{1/2}\nonumber\\&\times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right]^{1/2}.
%\end{align}
\noindent\textbf{Step 3: The big-O result on $\text{trace}[\E (\Lambda_n \tr \Lambda_n)]$ convergence} \\
In comparison to Eq. (7.6) in \cite{spall-jacobian} that corresponds to \eqref{eq:thmwts} for the N-SPSA-4 setting, there are two extra terms - the third and the fourth - in the RHS of \eqref{eq:thmwts}. 
Consider the third term on the RHS of \eqref{eq:thmwts}. As a consequence of (C9) and the construction of random perturbations $d_k$, it can be seen that $\text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right]$ is uniformly bounded above, say by $\top$, independent of $k$. Thus, the third term is equivalent to the following:
\begin{align}
 \bar c_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \frac{b_0^2}{k^{2 r}} \top.
\label{eq:term3}
\end{align}
Now by moving out side exponential term to inside and changing the summation to integration, we can rewrite the above equation as follows
\begin{align}
  \overline{\overline{c_n}}  \int_1^n  \frac{1}{k^{2 r}} .
\label{eq:term3}
\end{align}
Where $\overline{\overline{c_n}}$ is a constant. By above equation we can conclude that the third term is of $O(\frac{1}{k^{2 r -1}})$
%Now, observe that $\lim_{n\rightarrow \infty} \sum_{k=1}^n \frac{1}{k^{2 r}} < \infty$ as $1/2 < r < 1$ and $e^{2 b_0 k^{1-r}/(1-r)}$ diverges as $n\rightarrow \infty$ and hence, by Kronecker's Lemma, the term in \eqref{eq:term3} vanishes asymptotically. 
%
%Now consider the fourth term. By using the uniform upper-bound $\top$ on $\text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right]$, we can rewrite  the fourth term in \eqref{eq:thmwts} as follows: 
% \begin{align}\label{eq:fourthterm}
% \bar c_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \frac{b_0^2}{k^{2 r}} \sqrt{\top} \nonumber \\ \hspace{3cm}\times \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)^{1/2},
% \end{align}
%where, as in the proof of Theorem 3 of \cite{spall-jacobian}, $\tau(\cdot)$ transforms the $\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')$ in a linear fashion and then returns the trace of the resulting $N\times N$ matrix.

\textbf{Part (iii) needed fix} 
%of Theorem 3 in \cite{spall-jacobian} uses a proof by contradiction to show that the first and second terms in \eqref{eq:thmwts} are both of the order $O(e^{-2b_0n^{1-r} / {1-r}})$. By a completely parallel argument to that used for the second term in \cite{spall-jacobian}, it can be seen that the fourth term in our setting vanishes at a faster rate than the second term.  The claim follows. 
\end{proof}

\section{Simulation experiments}
\label{sec:expts}
\subsection{Implementation}
We test the performance of 2RDSA-Unif, 2RDSA-AsymBer and 2SPSA, with/without improved Hessian estimation. 
2SPSA algorithm uses Bernoulli $\pm 1$-valued perturbations, while 2RDSA/2RDSA-IH come in two variants - one that uses $U[-1,1]$ distributed perturbations (referred to as 2RDSA-Unif/2RDSA-IH-Unif) and the other that uses asymmetric Bernoulli perturbations (referred to as 2RDSA-AsymBer/2RDSA-IH-AsymBer)\footnote{The implementation is available at \url{https://github.com/prashla/RDSA/archive/master.zip}.}.

For the empirical evaluations, we use the following two loss functions in $N=10$ dimensions:
\paragraph{Quadratic loss}
\begin{align}
f(x) = \theta\tr A \theta+ b\tr \theta.\label{eq:quadratic}
\end{align} 
The optimum $\theta^*$ for the above $f$ is such that each coordinate of $\theta^*$ is $-0.9091$, with $f(\theta^*) = -4.55$.

\paragraph{Fourth-order loss}
\begin{align} 
f(x) = \theta\tr A\tr A \theta + 0.1 \sum_{j=1}^N (A\theta)^3_j + 0.01 \sum_{j=1}^N (A\theta)^4_j.\label{eq:4thorder}
 \end{align} 
The optimum $\theta^*$ for above $f$ is $\theta^*=0$, with $f(\theta^*) = 0$.

In both functions, $A$ is such that $NA$ is an upper triangular matrix with each entry one, $b$ is the $N$-dimensional vector of ones and the noise structure is similar to that used in \cite{spall_adaptive}. For any $\theta$, the noise is $[\theta\tr, 1]z$, where $z \approx \N(0,\sigma^2 I_{11\times11})$. We perform experiments for noisy as well as noise-less settings, with $\sigma=0.1$ for the noisy case.

For all algorithms, we set $\delta_n = 3.8/n^{0.101}$ and $a_n = 1/n^{0.6}$, while $b_n$ are set according to \eqref{eq:wieghts}. These choices have been used  for 2SPSA implementations before (see \cite{spall_adaptive}) and have demonstrated good finite-sample performance empirically, while satisfying the theoretical requirements needed for asymptotic convergence.  For all the algorithms, the initial point $\theta_0$ is the $N$-dimensional vector of ones.  For both 2SPSA and 2RDSA/2RDSA-IH, an initial $20\%$ of the  simulation budget was used up by 1SPSA/1RDSA and the resulting iterate was used to initialize 2SPSA/2RDSA. The distribution parameter $\epsilon$ is set to $0.0001$ for 2RDSA and to $0.01$ for 1RDSA.

\subsection{Results}
We use normalized loss and normalized MSE (NMSE) as performance metrics for evaluating the algorithms. 
NMSE is the ratio $\l \theta_{n_\text{end}} - \theta^* \r^2 / \l \theta_0 - \theta^*\r^2$, while normalized loss is the ratio $f(\theta_{n_\text{end}})/f(\theta_0)$.  Here $n_\text{end}$ denotes the iteration number when the algorithm stopped updating its parameter. Note that $n_\text{end}$ is a function of the simulation budget. 2RDSA/2RDSA-IH use only three simulations per-iteration and hence, $n_\text{end}$ is $1/3$rd of the simulation budget, while it is $1/4$th of the simulation budget for 2SPSA, since the latter algorithm uses four simulations per-iteration.

\begin{table}
\centering
 \caption{Normalized loss values for fourth-order  objective \eqref{eq:4thorder} with and without noise: standard error from $500$ replications shown after $\pm$}
\label{tab:norloss-4thf}
\begin{tabular}{|c|c|c|}
\toprule
\rowcolor{gray!20}
\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0.1$}}}\\[1em]
%&&&\\
\midrule
  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
  & & \textbf{ estimation} \\
 \midrule
\textbf{2SPSA} & $0.132 \pm 0.0267$ & $0.104 \pm 0.0355$\\
&&\\
\textbf{2RDSA-Unif} &$0.115 \pm 0.0214$ & $0.0271 \pm 0.0538$\\ 
&&\\
\textbf{2RDSA-AsymBer}& $0.0471 \pm 0.021$& $\bm{0.0099 \pm 0.0014}$\\
 \bottomrule
 %%%%%%%%%%%%2nd ordr
\rowcolor{gray!20}
\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0$}}}\\[1em]
%&&&\\
\midrule
  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
  & & \textbf{ estimation} \\
 \midrule
\textbf{2SPSA} & $0.0795 \pm 0.0234$ & $0.0628 \pm 0.0234$\\
&&\\
\textbf{2RDSA-Unif} &$0.0813 \pm 0.0275$ & $0.0214 \pm 0.00376$\\ 
&&\\
\textbf{2RDSA-AsymBer}& $0.0199 \pm 0.0114$& $\bm{0.0098 \pm 0.00147}$\\
 \bottomrule
\end{tabular}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%% Quadratic
\begin{table}
\centering
 \caption{Normalized loss values for quadratic objective \eqref{eq:quadratic} with and without noise: standard error from $500$ replications shown after $\pm$}
\label{tab:norloss-quadratic}
\adjustbox{max height=\dimexpr\textheight-10cm\relax,
           max width=\textwidth}{
\begin{tabular}{|c|c|c|}
\toprule
\rowcolor{gray!20}
\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0.1$}}}\\[1em]
%&&&\\
\midrule
  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
  & & \textbf{ estimation} \\
 \midrule
\textbf{2SPSA} & $-0.0062 \pm 0.1164$ & $-0.1229 \pm 0.1374$\\
&&\\
\textbf{2RDSA-Unif} &$0.0485 \pm 0.1465$ & $-0.259 \pm 0.0398$\\ 
&&\\
\textbf{2RDSA-AsymBer}& $-0.2564 \pm 0.068$& $\bm{-0.2877 \pm 0.0051}$\\
 \bottomrule
 %%%%%%%%%%%%2nd ordr
\rowcolor{gray!20}
\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0$}}}\\[1em]
%&&&\\
\midrule
  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
  & & \textbf{ estimation} \\
 \midrule
\textbf{2SPSA} & $-0.0785 \pm 0.1178$ & $-0.1716 \pm 0.1339$\\
&&\\
\textbf{2RDSA-Unif} &$0.0326 \pm 0.1599$ & $-0.2672 \pm 0.0299$\\ 
&&\\
\textbf{2RDSA-AsymBer}& $-0.2777 \pm 0.0488$& $\bm{-0.2881 \pm 0.0012}$\\
 \bottomrule
\end{tabular}}

\end{table}

%%%%%%%%%%%%%% NMSE for quadratic
\begin{table}
\centering
 \caption{NMSE values for quadratic objective \eqref{eq:quadratic} with and without noise: standard error from $500$ replications shown after $\pm$}
\label{tab:nmse-quadratic}
\adjustbox{max height=\dimexpr\textheight-10cm\relax,
           max width=\textwidth}{
\begin{tabular}{|c|c|c|}
\toprule
\rowcolor{gray!20}
\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0.1$}}}\\[1em]
%&&&\\
\midrule
  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
  & & \textbf{ estimation} \\
 \midrule
\textbf{2SPSA} & $0.9491 \pm 0.0131$ & $0.5495 \pm 0.0217$\\
&&\\
\textbf{2RDSA-Unif} &$1.0073 \pm 0.0140$ & $0.1953 \pm 0.0095$\\ 
&&\\
\textbf{2RDSA-AsymBer}& $0.1667 \pm 0.0095$& $\bm{0.0324 \pm 0.0007}$\\
 \bottomrule
 %%%%%%%%%%%%2nd ordr
\rowcolor{gray!20}
\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0$}}}\\[1em]
%&&&\\
\midrule
  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
  & & \textbf{ estimation} \\
 \midrule
\textbf{2SPSA} & $0.7325 \pm 0.0180$ & $0.3939 \pm 0.0230$\\
&&\\
\textbf{2RDSA-Unif} &$0.9834 \pm 0.0170$ & $0.1623 \pm 0.0086$\\ 
&&\\
\textbf{2RDSA-AsymBer}& $0.0686 \pm 0.0078$& $\bm{0.0316 \pm 0.0006}$\\
 \bottomrule
\end{tabular}}
\end{table}

\newcommand{\errorband}[5][]{ % x column, y column, error column, optional argument for setting style of the area plot
\pgfplotstableread[col sep=comma, skip first n=2]{#2}\datatable
    % Lower bound (invisible plot)
    \addplot [draw=none, stack plots=y, forget plot] table [
        x={#3},
        y expr=\thisrow{#4}-2*\thisrow{#5}
    ] {\datatable};

    % Stack twice the error, draw as area plot
    \addplot [draw=none, fill=gray!40, stack plots=y, area legend, #1] table [
        x={#3},
        y expr=4*\thisrow{#5}
    ] {\datatable} \closedcycle;

    % Reset stack using invisible plot
    \addplot [forget plot, stack plots=y,draw=none] table [x={#3}, y expr=-(\thisrow{#4}+2*\thisrow{#5})] {\datatable};
}

 \begin{figure}
    \centering
\tabl{c}{\scalebox{0.9}{\begin{tikzpicture}
      \begin{axis}[
	xlabel={number of simulations},
	ylabel={Normalized loss},
       legend entries={
	 ,2SPSA,,2SPSA-IH,,2RDSA-Unif,,2RDSA-Unif-IH,,2RDSA-AsymBer,,2RDSA-AsymBer-IH },
        %legend pos=outer north east,
			legend style={
at={(0,0)},
anchor=north,at={(axis description cs:0.5,-0.2)},legend columns=2}	]
      % 2SPSA
      \errorband[red!50!white, opacity=0.3]{norloss_vs_simulations_noisycase.csv}{0}{1}{2}
      \addplot [thick, red,mark=square*,mark options={scale=0.7}] table [x index=0, y index=1,col sep=comma] {norloss_vs_simulations_noisycase.csv};
			%%% With IH
      % 2SPSA
      \errorband[darkgray!50!white, opacity=0.3]{norloss_vs_simulations_noisycase.csv}{0}{7}{8}
      \addplot [thick, darkgray,mark=*,mark options={scale=0.75}] table [x index=0, y index=7,col sep=comma] {norloss_vs_simulations_noisycase.csv};

      % 2RDSA-Unif
      \errorband[blue!50!white, opacity=0.3]{norloss_vs_simulations_noisycase.csv}{0}{3}{4}
      \addplot [thick, blue,mark=square*,mark options={scale=0.7}] table [x index=0, y index=3,col sep=comma] {norloss_vs_simulations_noisycase.csv};
			%%% With IH
      % 2RDSA-Unif
      \errorband[magenta!50!white, opacity=0.3]{norloss_vs_simulations_noisycase.csv}{0}{9}{10}
      \addplot [thick, magenta,mark=*,mark options={scale=0.75}] table [x index=0, y index=9,col sep=comma] {norloss_vs_simulations_noisycase.csv};

      % 2RDSA-AsymBer
      \errorband[green!50!white, opacity=0.3]{norloss_vs_simulations_noisycase.csv}{0}{5}{6}
      \addplot [thick, darkgreen,mark=square*,mark options={scale=0.7}] table [x index=0, y index=5,col sep=comma] {norloss_vs_simulations_noisycase.csv};
			%%% With IH	
      % 2RDSA-AsymBer
      \errorband[violet!50!white, opacity=0.3]{norloss_vs_simulations_noisycase.csv}{0}{11}{12}
      \addplot [thick, violet, mark=*,mark options={scale=0.75}] table [x index=0, y index=11,col sep=comma] {norloss_vs_simulations_noisycase.csv};
      \end{axis}
      \end{tikzpicture}}\\}
			\caption{Normalized loss vs. number of simulations for fourth-order loss \eqref{eq:4thorder} with $\sigma=0.1$ for 2SPSA, 2RDSA-Unif and 2RDSA-AsymBer algorithms with/without improved Hessian estimation: bands around the curves represent standard error from $500$ replications. }
      \label{fig:norlossvssims} 
\end{figure}

Tables \ref{tab:norloss-4thf}--\ref{tab:norloss-quadratic} present the normalized loss values observed for the three algorithms - 2SPSA, 2RDSA-Unif and 2RDSA-AsymBer - with/without improved Hessian estimation scheme and for the fourth-order and quadratic loss functions, respectively. Table \ref{tab:nmse-quadratic} presents the NMSE values obtained for the aforementioned algorithms with the quadratic loss. The results in Tables \ref{tab:norloss-4thf}--\ref{tab:nmse-quadratic} are obtained after running all the algorithms with a budget of $10000$ function evaluations.
 Figure \ref{fig:norlossvssims} plots the normalized loss as a function of the simulation budget with the fourth-order loss objective with $\sigma=0.1$ 
%  (see Figures \ref{fig:norlossvssimsNoiseless}--\ref{fig:norlossvssimsNoiseless} in the appendix for similar results with $\sigma = 0$ for fourth-order loss and $\sigma=0.1$ for quadratic loss). 
 From the results in Tables \ref{tab:norloss-4thf}--\ref{tab:nmse-quadratic} and Fig \ref{fig:norlossvssims}, we make the following observations:

\textit{\textbf{Observation 1:} Among 2RDSA schemes, 2RDSA-IH performs better than regular 2RDSA, for both perturbation choices.}

\textit{\textbf{Observation 2:} 2RDSA-IH variants outperform both 2SPSA and 2SPSA-IH, with 2RDSA-IH-AsymBer performing the best overall.}

\begin{figure}[h]
    \centering
		\begin{tabular}{c}
		\begin{subfigure}{0.5\textwidth}
\tabl{c}{\scalebox{0.8}{\begin{tikzpicture}
      \begin{axis}[
	xlabel={number of simulations},
	ylabel={Normalized loss},
       legend entries={
	 ,2SPSA,,2SPSA-IH,,2RDSA-Unif,,2RDSA-Unif-IH,,2RDSA-AsymBer,,2RDSA-AsymBer-IH },
        %legend pos=outer north east,
			legend style={
at={(0,0)},
anchor=north,at={(axis description cs:0.5,-0.2)},legend columns=2}	]
      % 2SPSA
      \errorband[red!50!white, opacity=0.3]{norloss_vs_simulations.csv}{0}{1}{2}
      \addplot [thick, red,mark=square*,mark options={scale=0.7}] table [x index=0, y index=1,col sep=comma] {norloss_vs_simulations.csv};
			%%% With IH
      % 2SPSA
      \errorband[darkgray!50!white, opacity=0.3]{norloss_vs_simulations.csv}{0}{7}{8}
      \addplot [thick, darkgray,mark=*,mark options={scale=0.75}] table [x index=0, y index=7,col sep=comma] {norloss_vs_simulations.csv};

      % 2RDSA-Unif
      \errorband[blue!50!white, opacity=0.3]{norloss_vs_simulations.csv}{0}{3}{4}
      \addplot [thick, blue,mark=square*,mark options={scale=0.7}] table [x index=0, y index=3,col sep=comma] {norloss_vs_simulations.csv};
			%%% With IH
      % 2RDSA-Unif
      \errorband[magenta!50!white, opacity=0.3]{norloss_vs_simulations.csv}{0}{9}{10}
      \addplot [thick, magenta,mark=*,mark options={scale=0.75}] table [x index=0, y index=9,col sep=comma] {norloss_vs_simulations.csv};

      % 2RDSA-AsymBer
      \errorband[green!50!white, opacity=0.3]{norloss_vs_simulations.csv}{0}{5}{6}
      \addplot [thick, darkgreen,mark=square*,mark options={scale=0.7}] table [x index=0, y index=5,col sep=comma] {norloss_vs_simulations.csv};
			%%% With IH	
      % 2RDSA-AsymBer
      \errorband[violet!50!white, opacity=0.3]{norloss_vs_simulations.csv}{0}{11}{12}
      \addplot [thick, violet, mark=*,mark options={scale=0.75}] table [x index=0, y index=11,col sep=comma] {norloss_vs_simulations.csv};
      \end{axis}
      \end{tikzpicture}}\\}
			\caption{Fourth-order loss \eqref{eq:4thorder} with $\sigma=0$. }
      \label{fig:norlossvssimsNoiseless} 
\end{subfigure} 
\\
\\
%%%%%%% Norloss-Quadratic-noisy %%%%%%%%%%%%%%
\begin{subfigure}{0.5\textwidth}
\tabl{c}{\scalebox{0.8}{\begin{tikzpicture}
      \begin{axis}[
	xlabel={number of simulations},
	ylabel={Normalized loss},
       legend entries={
	 ,2SPSA,,2SPSA-IH,,2RDSA-Unif,,2RDSA-Unif-IH,,2RDSA-AsymBer,,2RDSA-AsymBer-IH },
        %legend pos=outer north east,
			legend style={
at={(0,0)},
anchor=north,at={(axis description cs:0.5,-0.2)},legend columns=2}	]
      % 2SPSA
      \errorband[red!50!white, opacity=0.3]{quad_norloss_vs_simulations_noisycase.csv}{0}{1}{2}
      \addplot [thick, red,mark=square*,mark options={scale=0.7}] table [x index=0, y index=1,col sep=comma] {quad_norloss_vs_simulations_noisycase.csv};
			%%% With IH
      % 2SPSA
      \errorband[darkgray!50!white, opacity=0.3]{quad_norloss_vs_simulations_noisycase.csv}{0}{7}{8}
      \addplot [thick, darkgray,mark=*,mark options={scale=0.75}] table [x index=0, y index=7,col sep=comma] {quad_norloss_vs_simulations_noisycase.csv};

      % 2RDSA-Unif
      \errorband[blue!50!white, opacity=0.3]{quad_norloss_vs_simulations_noisycase.csv}{0}{3}{4}
      \addplot [thick, blue,mark=square*,mark options={scale=0.7}] table [x index=0, y index=3,col sep=comma] {quad_norloss_vs_simulations_noisycase.csv};
			%%% With IH
      % 2RDSA-Unif
      \errorband[magenta!50!white, opacity=0.3]{quad_norloss_vs_simulations_noisycase.csv}{0}{9}{10}
      \addplot [thick, magenta,mark=*,mark options={scale=0.75}] table [x index=0, y index=9,col sep=comma] {quad_norloss_vs_simulations_noisycase.csv};

      % 2RDSA-AsymBer
      \errorband[green!50!white, opacity=0.3]{quad_norloss_vs_simulations_noisycase.csv}{0}{5}{6}
      \addplot [thick, darkgreen,mark=square*,mark options={scale=0.7}] table [x index=0, y index=5,col sep=comma] {quad_norloss_vs_simulations_noisycase.csv};
			%%% With IH	
      % 2RDSA-AsymBer
      \errorband[violet!50!white, opacity=0.3]{quad_norloss_vs_simulations_noisycase.csv}{0}{11}{12}
      \addplot [thick, violet, mark=*,mark options={scale=0.75}] table [x index=0, y index=11,col sep=comma] {quad_norloss_vs_simulations_noisycase.csv};
      \end{axis}
      \end{tikzpicture}}\\}
			\caption{Quadratic loss \eqref{eq:quadratic} with $\sigma=0.1$. }
      \label{fig:QuadnorlossvssimsNoisy} 
\end{subfigure} 
		\end{tabular}
			\caption{Normalized loss vs. number of simulations in two different loss settings for all the algorithms. }
      \label{fig:extras} 
		\end{figure}

%%%%%% Norloss-Quadratic-noisefree %%%%%%%%%%%%%%
\begin{figure}
    \centering
\tabl{c}{\scalebox{0.9}{\begin{tikzpicture}
      \begin{axis}[
	xlabel={number of simulations},
	ylabel={Normalized loss},
       legend entries={
	 ,2SPSA,,2SPSA-IH,,2RDSA-Unif,,2RDSA-Unif-IH,,2RDSA-AsymBer,,2RDSA-AsymBer-IH },
        %%legend pos=outer north east,
			legend style={
at={(0,0)},
anchor=north,at={(axis description cs:0.5,-0.2)},legend columns=2}	]
      % 2SPSA
      \errorband[red!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{1}{2}
      \addplot [thick, red,mark=square*,mark options={scale=0.7}] table [x index=0, y index=1,col sep=comma] {quad_norloss_vs_simulations.csv};
			%%%% With IH
      %2SPSA
      \errorband[darkgray!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{7}{8}
      \addplot [thick, darkgray,mark=*,mark options={scale=0.75}] table [x index=0, y index=7,col sep=comma] {quad_norloss_vs_simulations.csv};
			%
      % 2RDSA-Unif
      \errorband[blue!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{3}{4}
      \addplot [thick, blue,mark=square*,mark options={scale=0.7}] table [x index=0, y index=3,col sep=comma] {quad_norloss_vs_simulations.csv};
			%%%% With IH
      % 2RDSA-Unif
      \errorband[magenta!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{9}{10}
      \addplot [thick, magenta,mark=*,mark options={scale=0.75}] table [x index=0, y index=9,col sep=comma] {quad_norloss_vs_simulations.csv};

      % 2RDSA-AsymBer
      \errorband[green!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{5}{6}
      \addplot [thick, darkgreen,mark=square*,mark options={scale=0.7}] table [x index=0, y index=5,col sep=comma] {quad_norloss_vs_simulations.csv};
			%%%% With IH	
      % 2RDSA-AsymBer
      \errorband[violet!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{11}{12}
      \addplot [thick, violet, mark=*,mark options={scale=0.75}] table [x index=0, y index=11,col sep=comma] {quad_norloss_vs_simulations.csv};
      \end{axis}
      \end{tikzpicture}}\\}
			\caption{Normalized loss vs. number of simulations for quadratic loss \eqref{eq:quadratic} with $\sigma=0$ for 2SPSA, 2RDSA-Unif and 2RDSA-AsymBer algorithms with/without improved Hessian estimation. }
      \label{fig:QuadnorlossvssimsNoiseless} 
\end{figure}

### 2017-01-05 ack.tex
experiments added

@@
-I take this opportunity to express my profound gratitude to the people who have helped and supported me throughout my PhD.
+I take this opportunity to express my profound gratitude to the people who have helped and supported me throughout my M.Sc(engg).
 I thank my guide, Prof.
@@
 Without his constant guidance and support, this thesis would not have been possible.
-Next I thank Prof.
-Borkar for several interesting and useful discussions on our collaborative work.
-I thank Sunil, Hemanth and Sindhu for enthusiastically collaborating on the research problems.
-I thank Prof.
-Narahari, Electrical Sciences Divisional Chairman, who has always been keen to know how my thesis work is progressing.
+Next I thank Prashanth L.A.
+for several interesting and useful discussions on our collaborative work.
+I thank Chandramouli K.
+for his help during the course work.
 I am grateful to all the professors of CSA Dept for sharing their knowledge and for their encouragement.
 I, also, thank the CSA Dept office staff for their invaluable support.
-I would like to particularly acknowledge Mrs.
-Suguna and Mrs.
-Lalitha who promptly helped me on many occasions.
-Last but not the least, I express my heartfelt gratitude to the Almighty, my parents, younger sister, relatives and friends for their love and blessings.
+Last but not the least, I express my heartfelt gratitude to  my parents, sister, brother-in-law, relatives and friends for their love and blessings.

### 2017-01-05 conc.tex
experiments added

@@
 
-As future work, it would be interesting to derive finite time bounds that show a lower Hessian estimation error for 2RDSA-IH when compared to 2RDSA and 2SPSA.
-Stochastic Newton methods  methods converge faster and  are often more accurate than simple gradient search schemes.
+As future work, it would be interesting to look into the following directions
+\begin{enumerate}
+\item To derive finite time bounds that show a lower Hessian estimation error for 2RDSA-IH when compared to 2RDSA and 2SPSA. 
+\item Stochastic Newton methods  methods converge faster and  are often more accurate than simple gradient search schemes.
 However they are computationally much costlier than first-order methods.
 Since first-order methods exhibits slower convergence rate and heavily depended on the step-size selection one has to resort to second-order methods.
-As a future work, it would be interesting to look at the methods like momentum descent, conjugate gradient methods, Hessian free optimization schemes which result in faster convergence for the first-order methods in deterministic optimization case.
-It would be interesting to extend those methods to stochastic optimization case and prove that they exhibits  faster convergence than first-oder methods.
-From application point of view it would be interesting to apply these methods to some interesting real world applications.
+As a future work, it would be interesting to look at the methods like momentum descent, conjugate gradient methods, Hessian free optimization schemes which result in faster convergence for the first-order methods and computationally cheaper than second-order methods in deterministic optimization case.
+It would be interesting to extend those methods to stochastic optimization case and prove that they exhibits  faster convergence than first-oder methods and results in lower computational cost.
+\item In the methods described in this thesis the perturbations used are random variables.
+In \cite{bhatnagar2003two}, for first-order SPSA method,  construction of two different deterministic  perturbation schemes were proposed and  under certain conditions on the perturbation vectors and noise sequences algorithm is shown to converge.
+The idea behind using deterministic perturbations is to reduce the bias in the gradient estimates.
+It would be interesting to extend this construction to RDSA scheme as well.
+And moreover this kind of construction for second-order methods is not available for both SPSA and RDSA.
+One can use the above ideas and show significant improvements to the existing methods.
+\item From application point of view it would be interesting to apply these methods to some interesting real world applications.
+Note that these methods can be easily extended to many of reinforcement learning algorithms.
+\end{enumerate}

### 2017-01-05 experiments.tex
experiments added

@@
 \section{Implementation}
-We test the performance of 2RDSA-Unif, 2RDSA-AsymBer and 2SPSA, with/without improved Hessian estimation. 
-2SPSA algorithm uses Bernoulli $\pm 1$-valued perturbations, while 2RDSA/2RDSA-IH come in two variants - one that uses $U[-1,1]$ distributed perturbations (referred to as 2RDSA-Unif/2RDSA-IH-Unif) and the other that uses asymmetric Bernoulli perturbations (referred to as 2RDSA-AsymBer/2RDSA-IH-AsymBer)\footnote{The implementation is available at \url{https://github.com/prashla/RDSA/archive/master.zip}.}.
+We test the performance of 2RDSA-Unif, 2RDSA-AsymBer, 2SPSA-3 and 2SPSA, with/without improved Hessian estimation. 
+2SPSA and 2SPSA-3 algorithm uses Bernoulli $\pm 1$-valued perturbations, while 2RDSA/2RDSA-IH come in two variants - one that uses $U[-1,1]$ distributed perturbations (referred to as 2RDSA-Unif/2RDSA-IH-Unif) and the other that uses asymmetric Bernoulli perturbations (referred to as 2RDSA-AsymBer/2RDSA-IH-AsymBer)\footnote{The implementation is available at \url{https://github.com/prashla/RDSA/archive/master.zip}.}.
 
@@
 \begin{align}
-f(x) = x\tr A x + b\tr x.\label{eq:quadratic}
+f(x) = \theta\tr A \theta+ b\tr \theta.\label{eq:quadratic}
 \end{align} 
-The optimum $x^*$ for the above $f$ is such that each coordinate of $x^*$ is $-0.9091$, with $f(x^*) = -4.55$.
+The optimum $\theta^*$ for the above $f$ is such that each coordinate of $\theta^*$ is $-0.9091$, with $f(\theta^*) = -4.55$.
 
@@
 \begin{align} 
-f(x) = x\tr A\tr A x + 0.1 \sum_{j=1}^N (Ax)^3_j + 0.01 \sum_{j=1}^N (Ax)^4_j.\label{eq:4thorder}
+f(x) = \theta\tr A\tr A \theta + 0.1 \sum_{j=1}^N (A\theta)^3_j + 0.01 \sum_{j=1}^N (A\theta)^4_j.\label{eq:4thorder}
  \end{align} 
-The optimum $x^*$ for above $f$ is $x^*=0$, with $f(x^*) = 0$.
+The optimum $\theta^*$ for above $f$ is $\theta^*=0$, with $f(\theta^*) = 0$.
 
 In both functions, $A$ is such that $NA$ is an upper triangular matrix with each entry one, $b$ is the $N$-dimensional vector of ones and the noise structure is similar to that used in \cite{spall_adaptive}.
-For any $x$, the noise is $[x\tr, 1]z$, where $z \approx \N(0,\sigma^2 I_{11\times11})$.
+For any $\theta$, the noise is $[\theta\tr, 1]z$, where $z \approx \N(0,\sigma^2 I_{11\times11})$.
 We perform experiments for noisy as well as noise-less settings, with $\sigma=0.1$ for the noisy case.
@@
 These choices have been used  for 2SPSA implementations before (see \cite{spall_adaptive}) and have demonstrated good finite-sample performance empirically, while satisfying the theoretical requirements needed for asymptotic convergence.
-For all the algorithms, the initial point $x_0$ is the $N$-dimensional vector of ones.
-For both 2SPSA and 2RDSA/2RDSA-IH, an initial $20\%$ of the  simulation budget was used up by 1SPSA/1RDSA and the resulting iterate was used to initialize 2SPSA/2RDSA.
+For all the algorithms, the initial point $\theta_0$ is the $N$-dimensional vector of ones.
+For  2SPSA, 2SPSA-3 and 2RDSA/2RDSA-IH, an initial $20\%$ of the  simulation budget was used up by 1SPSA/1RDSA and the resulting iterate was used to initialize 2SPSA, 2SPSA-3/2RDSA.
 The distribution parameter $\epsilon$ is set to $0.0001$ for 2RDSA and to $0.01$ for 1RDSA.
@@
 We use normalized loss and normalized MSE (NMSE) as performance metrics for evaluating the algorithms. 
-NMSE is the ratio $\l x_{n_\text{end}} - x^* \r^2 / \l x_0 - x^*\r^2$, while normalized loss is the ratio $f(x_{n_\text{end}})/f(x_0)$.
+NMSE is the ratio $\l \theta_{n_\text{end}} - \theta^* \r^2 / \l \theta_0 - \theta^*\r^2$, while normalized loss is the ratio $f(\theta_{n_\text{end}})/f(\theta_0)$.
 Here $n_\text{end}$ denotes the iteration number when the algorithm stopped updating its parameter.
 Note that $n_\text{end}$ is a function of the simulation budget.
-2RDSA/2RDSA-IH use only three simulations per-iteration and hence, $n_\text{end}$ is $1/3$rd of the simulation budget, while it is $1/4$th of the simulation budget for 2SPSA, since the latter algorithm uses four simulations per-iteration.
+2RDSA/2RDSA-IH, 2SPSA-3/2SPSA-3-IH use only three simulations per-iteration and hence, $n_\text{end}$ is $1/3$rd of the simulation budget, while it is $1/4$th of the simulation budget for 2SPSA, since the latter algorithm uses four simulations per-iteration.
 
@@
 \label{tab:norloss-4thf}
+\scalebox{0.9}{
 \begin{tabular}{|c|c|c|}
@@
 \end{tabular}
+}
 \end{table}
@@
 \label{tab:norloss-quadratic}
+\scalebox{0.9}{
 \begin{tabular}{|c|c|c|}
@@
  \bottomrule
-\end{tabular}
+\end{tabular}}
+
 \end{table}
@@
 \label{tab:nmse-quadratic}
+\scalebox{0.9}{
 \begin{tabular}{|c|c|c|}
@@
  \bottomrule
-\end{tabular}
+\end{tabular}}
 \end{table}
@@
 
-Tables \ref{tab:norloss-4thf}--\ref{tab:norloss-quadratic} present the normalized loss values observed for the three algorithms - 2SPSA, 2RDSA-Unif and 2RDSA-AsymBer - with/without improved Hessian estimation scheme and for the fourth-order and quadratic loss functions, respectively.
-Table \ref{tab:nmse-quadratic} presents the NMSE values obtained for the aforementioned algorithms with the quadratic loss.
-The results in Tables \ref{tab:norloss-4thf}--\ref{tab:nmse-quadratic} are obtained after running all the algorithms with a budget of $10000$ function evaluations.
- Figure \ref{fig:norlossvssims} plots the normalized loss as a function of the simulation budget with the fourth-order loss objective with $\sigma=0.1$ 
-%  (see Figures \ref{fig:norlossvssimsNoiseless}--\ref{fig:norlossvssimsNoiseless} in the appendix for similar results with $\sigma = 0$ for fourth-order loss and $\sigma=0.1$ for quadratic loss). 
- From the results in Tables \ref{tab:norloss-4thf}--\ref{tab:nmse-quadratic} and Fig \ref{fig:norlossvssims}, we make the following observations:
+%%%%%% Norloss-Quadratic-noisefree %%%%%%%%%%%%%%
+\begin{figure}
+    \centering
+\tabl{c}{\scalebox{0.9}{\begin{tikzpicture}
+      \begin{axis}[
+	xlabel={number of simulations},
+	ylabel={Normalized loss},
+       legend entries={
+	 ,2SPSA,,2SPSA-IH,,2RDSA-Unif,,2RDSA-Unif-IH,,2RDSA-AsymBer,,2RDSA-AsymBer-IH },
+        %%legend pos=outer north east,
+			legend style={
+at={(0,0)},
+anchor=north,at={(axis description cs:0.5,-0.2)},legend columns=2}	]
+      % 2SPSA
+      \errorband[red!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{1}{2}
+      \addplot [thick, red,mark=square*,mark options={scale=0.7}] table [x index=0, y index=1,col sep=comma] {quad_norloss_vs_simulations.csv};
+			%%%% With IH
+      %2SPSA
+      \errorband[darkgray!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{7}{8}
+      \addplot [thick, darkgray,mark=*,mark options={scale=0.75}] table [x index=0, y index=7,col sep=comma] {quad_norloss_vs_simulations.csv};
+			%
+      % 2RDSA-Unif
+      \errorband[blue!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{3}{4}
+      \addplot [thick, blue,mark=square*,mark options={scale=0.7}] table [x index=0, y index=3,col sep=comma] {quad_norloss_vs_simulations.csv};
+			%%%% With IH
+      % 2RDSA-Unif
+      \errorband[magenta!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{9}{10}
+      \addplot [thick, magenta,mark=*,mark options={scale=0.75}] table [x index=0, y index=9,col sep=comma] {quad_norloss_vs_simulations.csv};
 
-\textit{\textbf{Observation 1:} Among 2RDSA schemes, 2RDSA-IH performs better than regular 2RDSA, for both perturbation choices.}
-
-\textit{\textbf{Observation 2:} 2RDSA-IH variants outperform both 2SPSA and 2SPSA-IH, with 2RDSA-IH-AsymBer performing the best overall.}
+      % 2RDSA-AsymBer
+      \errorband[green!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{5}{6}
+      \addplot [thick, darkgreen,mark=square*,mark options={scale=0.7}] table [x index=0, y index=5,col sep=comma] {quad_norloss_vs_simulations.csv};
+			%%%% With IH	
+      % 2RDSA-AsymBer
+      \errorband[violet!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{11}{12}
+      \addplot [thick, violet, mark=*,mark options={scale=0.75}] table [x index=0, y index=11,col sep=comma] {quad_norloss_vs_simulations.csv};
+      \end{axis}
+      \end{tikzpicture}}\\}
+			\caption{Normalized loss vs.
+number of simulations for quadratic loss \eqref{eq:quadratic} with $\sigma=0$ for 2SPSA, 2RDSA-Unif and 2RDSA-AsymBer algorithms with/without improved Hessian estimation.
+}
+      \label{fig:QuadnorlossvssimsNoiseless} 
+\end{figure}
 
@@
 \begin{subfigure}{0.5\textwidth}
-\tabl{c}{\scalebox{0.8}{\begin{tikzpicture}
+\tabl{c}{\scalebox{0.9}{\begin{tikzpicture}
       \begin{axis}[
@@
 
-%%%%%% Norloss-Quadratic-noisefree %%%%%%%%%%%%%%
-\begin{figure}
-    \centering
-\tabl{c}{\scalebox{0.9}{\begin{tikzpicture}
-      \begin{axis}[
-	xlabel={number of simulations},
-	ylabel={Normalized loss},
-       legend entries={
-	 ,2SPSA,,2SPSA-IH,,2RDSA-Unif,,2RDSA-Unif-IH,,2RDSA-AsymBer,,2RDSA-AsymBer-IH },
-        %%legend pos=outer north east,
-			legend style={
-at={(0,0)},
-anchor=north,at={(axis description cs:0.5,-0.2)},legend columns=2}	]
-      % 2SPSA
-      \errorband[red!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{1}{2}
-      \addplot [thick, red,mark=square*,mark options={scale=0.7}] table [x index=0, y index=1,col sep=comma] {quad_norloss_vs_simulations.csv};
-			%%%% With IH
-      %2SPSA
-      \errorband[darkgray!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{7}{8}
-      \addplot [thick, darkgray,mark=*,mark options={scale=0.75}] table [x index=0, y index=7,col sep=comma] {quad_norloss_vs_simulations.csv};
-			%
-      % 2RDSA-Unif
-      \errorband[blue!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{3}{4}
-      \addplot [thick, blue,mark=square*,mark options={scale=0.7}] table [x index=0, y index=3,col sep=comma] {quad_norloss_vs_simulations.csv};
-			%%%% With IH
-      % 2RDSA-Unif
-      \errorband[magenta!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{9}{10}
-      \addplot [thick, magenta,mark=*,mark options={scale=0.75}] table [x index=0, y index=9,col sep=comma] {quad_norloss_vs_simulations.csv};
+Tables \ref{tab:norloss-4thf}--\ref{tab:norloss-quadratic} present the normalized loss values observed for the three algorithms - 2SPSA, 2RDSA-Unif and 2RDSA-AsymBer - with/without improved Hessian estimation scheme and for the fourth-order and quadratic loss functions, respectively.
+Table \ref{tab:nmse-quadratic} presents the NMSE values obtained for the aforementioned algorithms with the quadratic loss.
+The results in Tables \ref{tab:norloss-4thf}--\ref{tab:nmse-quadratic} are obtained after running all the algorithms with a budget of $10000$ function evaluations.
+ Figure \ref{fig:norlossvssims} plots the normalized loss as a function of the simulation budget with the fourth-order loss objective with $\sigma=0.1$ 
+%  (see Figures \ref{fig:norlossvssimsNoiseless}--\ref{fig:norlossvssimsNoiseless} in the appendix for similar results with $\sigma = 0$ for fourth-order loss and $\sigma=0.1$ for quadratic loss). 
+ From the results in Tables \ref{tab:norloss-4thf}--\ref{tab:nmse-quadratic} and Fig \ref{fig:norlossvssims}, we make the following observations:
 
-      % 2RDSA-AsymBer
-      \errorband[green!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{5}{6}
-      \addplot [thick, darkgreen,mark=square*,mark options={scale=0.7}] table [x index=0, y index=5,col sep=comma] {quad_norloss_vs_simulations.csv};
-			%%%% With IH	
-      % 2RDSA-AsymBer
-      \errorband[violet!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{11}{12}
-      \addplot [thick, violet, mark=*,mark options={scale=0.75}] table [x index=0, y index=11,col sep=comma] {quad_norloss_vs_simulations.csv};
-      \end{axis}
-      \end{tikzpicture}}\\}
-			\caption{Normalized loss vs.
-number of simulations for quadratic loss \eqref{eq:quadratic} with $\sigma=0$ for 2SPSA, 2RDSA-Unif and 2RDSA-AsymBer algorithms with/without improved Hessian estimation.
-}
-      \label{fig:QuadnorlossvssimsNoiseless} 
-\end{figure}
+\textit{\textbf{Observation 1:} Among 2RDSA schemes, 2RDSA-IH performs better than regular 2RDSA, for both perturbation choices.}
+
+\textit{\textbf{Observation 2:} 2RDSA-IH variants outperform both 2SPSA and 2SPSA-IH, with 2RDSA-IH-AsymBer performing the best overall.}

### 2017-01-05 second-order-rdsa-gen.tex
experiments added

@@
 \label{sec:2rdsa-gen}
-Our work in chapter  is centred on generalising the 2RDSA scheme of \cite{prashanth2015rdsa}.
+Our work in this chapter  is centred on generalising the 2RDSA scheme of \cite{prashanth2015rdsa}.
 In \cite{prashanth2015rdsa}, 2RDSA scheme involving only two perturbation distributions was proposed, those are uniform and asymmetric Bernoulli distributions.
@@
 
+\section{Convergence analysis}
+\label{sec:2rdsa-gen-results}
 %%%%%%%%%%%%%%%%%%%% LEMMA 
@@
 %Almost surely, as $\delta_n \rightarrow 0$, we have that
-Under (A1),(A2),(A4),(A5),(A13),(A14) for $\widehat\nabla f(x_n)$ defined according to  \eqref{eq:grad-gen} , we have a.s.
+Under (A1),(A2),(A4),(A5),(A11),(A12) for $\widehat\nabla f(x_n)$ defined according to  \eqref{eq:grad-gen} , we have a.s.
 that\footnote{Here $\widehat\nabla_i f(x_n)$ and $\nabla_i f(x_n)$ denote the $i$th coordinates in the gradient estimate $\widehat\nabla f(x_n)$ and true gradient $\nabla f(x_n)$, respectively.}
@@
 \end{align*}
-The last equality above follows from the fact that $\E\left[  d_n\left.\left(\dfrac{\xi_n^+ - \xi_n^-}{2\delta_n}\right) \right| \F_n\right]=0$ from (A2) and (A13).
+The last equality above follows from the fact that $\E\left[  d_n\left.\left(\dfrac{\xi_n^+ - \xi_n^-}{2\delta_n}\right) \right| \F_n\right]=0$ from (A2) and (A11).
 We now analyse the term on the RHS above.
@@
 \end{align}
-In the above, the first equality follows from (A14) and the last equality in \eqref{eq:l1-a} follows from $\E[(d_n^i)^2] = \lambda$ and $\E[d_n^i d_n^j] = \E[d_n^i] \E[d_n^j] = 0$ for $i\ne j$.
+In the above, the first equality follows from (A12) and the last equality in \eqref{eq:l1-a} follows from $\E[(d_n^i)^2] = \lambda$ and $\E[d_n^i d_n^j] = \E[d_n^i] \E[d_n^j] = 0$ for $i\ne j$.
 
@@
 \end{align}
-The first inequality above follows from (A1), while the second inequality follows from (A14) and considering $\eta = \max \{\eta_1^2, \eta_2\}$.
+The first inequality above follows from (A1), while the second inequality follows from (A12) and considering $\eta = \max \{\eta_1^2, \eta_2\}$.
 The claim follows by plugging \eqref{eq:l1-a} and \eqref{eq:l2} into \eqref{eq:l1} .\\
@@
 \label{lemma:2rdsa-bias}
-Under (C1)-(C3),(C5)-(C10),(C15) with $\widehat H_n$ defined according to \eqref{eq:2rdsa-estimate-gen} , we have a.s.
+Under (C1)-(C3),(C5)-(C10),(C13) with $\widehat H_n$ defined according to \eqref{eq:2rdsa-estimate-gen} , we have a.s.
 that\footnote{Here $\widehat H_n(i,j)$ and $\nabla^2_{ij}f(\cdot)$ denote the $(i,j)$th entry in the Hessian estimate $\widehat H_n$ and the true Hessian $\nabla^2 f(\cdot)$, respectively.}, for $i,j = 1,\ldots,N$,

### 2017-01-05 second-order-spsa-3.tex
experiments added

@@
- \chapter{Second-order spsa with improved hessian estimation (2SPSA-3-IH)}
+ \chapter{Second-order SPSA with improved hessian estimation (2SPSA-3-IH)}
 \label{sec:2spsa-3-ih}
@@
  \item Generate $\{\Delta_n^{i}, i=1,\ldots,N\}$,$\{\widehat\Delta_n^{i}, i=1,\ldots,N\}$ independent of $\{\Delta_m,\widehat\Delta_m, m=0,1,\ldots,n-1\}$.  
- \item For any $i=1,\ldots,N$, $\Delta_n^{i}$,$\widehat\Delta_n^{i}$  satisfy condition (C16) (most popular distribution being used is symmetric Bernoulli distribution, $\pm 1$ w.p. $1/2$).
+ \item For any $i=1,\ldots,N$, $\Delta_n^{i}$,$\widehat\Delta_n^{i}$  satisfy condition (C17) of section \ref{sec:2spsa-results},(most popular distribution being used is symmetric Bernoulli distribution, $\pm 1$ w.p.
+$1/2$).
 \begin{itemize}
@@
 
-Algorithm \ref{alg:spsa-3-ih-structure} presents the pseudocode and we describe the individual component of 2RDSA-IH below.
+Algorithm \ref{alg:spsa-3-ih-structure} presents the pseudocode and we describe the individual component of 2SPSA-3-IH below.
 \section{Function evaluations}
@@
 
-Further, $\delta_n, n\geq 0$ is a sequence of diminishing positive real numbers and $\Delta_n = (\Delta_n^1,\ldots,\Delta_n^N)\tr$ , $\widehat\Delta_n = (\widehat\Delta_n^1,\ldots,\widehat\Delta_n^N)\tr$ are the perturbation random vector at instant $n$ with $\Delta_n^i,i=1,\ldots,N$, and $\widehat\Delta_n^i,i=1,\ldots,N$ are independent identically distributed (i.i.d), mean-zero random varibles having finite inverse movements of order greater than 2 and satisfy the conditions (C16), C(21).
+Further, $\delta_n, n\geq 0$ is a sequence of diminishing positive real numbers and $\Delta_n = (\Delta_n^1,\ldots,\Delta_n^N)\tr$ , $\widehat\Delta_n = (\widehat\Delta_n^1,\ldots,\widehat\Delta_n^N)\tr$ are the perturbation random vector at instant $n$ with $\Delta_n^i,i=1,\ldots,N$, and $\widehat\Delta_n^i,i=1,\ldots,N$ are independent identically distributed (i.i.d), mean-zero random varibles having finite inverse movements of order greater than 2 and satisfy the condition (C17) of section \ref{sec:2spsa-results}.
 
@@
 
-The optimal choice for $b_n$ in \eqref{eq:2rdsa-H}  is the following:
+The optimal choice for $b_n$ in \eqref{eq:2spsa-3-ih}  is the following:
 \begin{align}
@@
 This is because the third term in \eqref{eq:spsahnhat} vanishes asymptotically, while the fourth term there dominates asymptotically.
-Moreover, the noise factors in the fourth term in \eqref{eq:spsahnhat} are bounded above due to (C21)  and independent of $n$, leaving the $\delta_n^2$ term in the denominator there.
+Moreover, the noise factors in the fourth term in \eqref{eq:spsahnhat} are bounded above due to (C22)  and independent of $n$, leaving the $\delta_n^2$ term in the denominator there.
 
@@
 The reader is referred to Section II-B of \cite{prashanth2015rdsa} for a detailed discussion of the above assumptions.
-We remark here that (C13)-(C20) are identical to that in \cite{prashanth2015rdsa}, while (C21) and (C22) introduce minor additional requirements on $\left\| \Upsilon(\overline H_n \right\|^2$ and $\delta_n$, respectively and these are inspired by \cite{spall-jacobian}.
+We remark here that (C14)-(C21) are identical to that in \cite{prashanth2015rdsa}, while (C22) and (C23) introduce minor additional requirements on $\left\| \Upsilon(\overline H_n \right\|^2$ and $\delta_n$, respectively and these are inspired by \cite{spall-jacobian}.
 
@@
 \label{lemma:2spsa-bias}
-Under (C13)-(C22), with $\widehat H_n$ defined according to  \eqref{eq:2spsa-ber} , we have a.s.
+Under (C14)-(C23), with $\widehat H_n$ defined according to  \eqref{eq:2spsa-ber} , we have a.s.
 that\footnote{Here $\widehat H_n(i,j)$ and $\nabla^2_{ij}f(\cdot)$ denote the $(i,j)$th entry in the Hessian estimate $\widehat H_n$ and the true Hessian $\nabla^2 f(\cdot)$, respectively.}, for $i,j = 1,\ldots,N$,
@@
 \label{thm:2spsa-H}
-Under (C13)-(C22), we have that 
+Under (C14)-(C23), we have that 
 $$\theta_n \rightarrow \theta^*, \overline H_n \rightarrow \nabla^2 f(\theta^*) \text{ a.s. as } n\rightarrow \infty.$$ 
@@
 \label{thm:spsaquad-bound}
-Assume (C16), (C22), (C23) and (C24) and also that the setting is noise-free. 
+Assume (C17), (C23), (C24) and (C25) and also that the setting is noise-free. 
 Let $b_n = b_0/n^r$, $n=1,2,\ldots,k$, where $1/2 < r < 1$ and $0 < b_0 \leq 1$. For notational simplicity, let $H^*=\nabla^2 f(x^*)$. Letting $\Lambda_k = \overline H_k - H^*$, we have 
@@
   \item[Step 1:] Here we prove the MSE convergence of $\overline H_k$, i.e., $\E [\Lambda_n\tr \Lambda_n] \rightarrow 0$ a.s. as $n\rightarrow \infty$.
-  \item[Step 2:] We unroll the recursion \eqref{eq:2rdsa-H} and then derive convenient representation for $\text{trace}[\E (\Lambda_n \tr \Lambda_n)]$.
-  \item[Step 3:] We derive the main result in \eqref{eq:quad-bigo} using a proof by contradiction. 
+  \item[Step 2:] We unroll the recursion \eqref{eq:2spsa-3-ih} and then derive convenient representation for $\text{trace}[\E (\Lambda_n \tr \Lambda_n)]$.
+  \item[Step 3:] We derive the main result in \eqref{eq:spsaquad-bigo} using a proof by contradiction. 
 \end{description}

### 2017-01-05 secorder-rdsa-ih.tex
experiments added

@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
-%%%%%%%%%%%%%%%%%%%%%%% Quadratic case
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-We next present a convergence rate result for the special case of a quadratic objective function under the following additional assumptions:
-\begin{enumerate}[label=(\textbf{C\arabic*}),resume]
-\item  $f$ is quadratic and $\nabla^2 f(x^*) > 0$. 
-\item The operator $\Upsilon$ is chosen such that $\E(\parallel \Upsilon(\overline H_n) - \overline H_n\parallel^2) = o(e^{-2wn^{1-r}/(1-r)})$ and $\parallel \Upsilon(H) - H \parallel^2 / (1+\parallel H \parallel^2)$ is uniformly bounded.
-\end{enumerate}
-\begin{theorem}(\textbf{Quadratic case - Convergence rate})
-\label{thm:quad-bound}
-Assume (C4), (C10), (C11) and (C12) and also that the setting is noise-free. 
-Let $b_n = b_0/n^r$, $n=1,2,\ldots,k$, where $1/2 < r < 1$ and $0 < b_0 \leq 1$. For notational simplicity, let $H^*=\nabla^2 f(x^*)$. Letting $\Lambda_k = \overline H_k - H^*$, we have 
-\begin{align}
-\text{trace}[\E (\Lambda_n \tr \Lambda_n)] = O(e^{-2b_0n^{1-r} / {1-r}}).
-\label{eq:quad-bigo}
-\end{align}
-\end{theorem}
-\begin{proof}
-Since the setting is noise-free with a quadratic objective, we can rewrite \eqref{eq:hnhat} as follows:
-\begin{align}
- \widehat H_n =    \Phi_n(\nabla^2 f(\theta_n)) &+\Psi_{n}(\nabla^2 f(\theta_n)),\label{eq:hnhat-ext}
-\end{align}
-where $\Psi_n(H)$ is defined in \eqref{eq:psi} and for any matrix $H$, 
-$$\Phi_{n}(H) = [M_n]_{D}\left(d_{n}\tr \, [H]_{D} \, d_{n}\right) +  [M_n]_{N}\left(d_{n}\tr \, [H]_{N} \, d_{n}\right).$$
-The proof involves the following steps:
-\begin{description}
-  \item[Step 1:] Here we prove the MSE convergence of $\overline H_k$, i.e., $\E [\Lambda_n\tr \Lambda_n] \rightarrow 0$ a.s. as $n\rightarrow \infty$.
-  \item[Step 2:] We unroll the recursion \eqref{eq:2rdsa-H} and then derive convenient representation for $\text{trace}[\E (\Lambda_n \tr \Lambda_n)]$.
-  \item[Step 3:] We derive the main result in \eqref{eq:quad-bigo} using a proof by contradiction. 
-\end{description}
-\noindent\textbf{Step 1: MSE convergence of $\overline H_n$} \\
-This part in exactly the same manner as part (i) in Theorem 3 of \cite{spall-jacobian}.\\
-\noindent\textbf{Step 2: Representation of $\text{trace}[\E (\Lambda_n \tr \Lambda_n)]$} \\
- From \eqref{eq:2rdsa-H} and \eqref{eq:hnhat-ext},  we have
- \begin{align}
- \Lambda_n &= \Lambda_{n-1} - b_n ( \overline H_{n-1} - \hat H_n + \hat \Psi_n) \nonumber\\
-&= (1-b_n) \Lambda_{n-1} - b_n (H^* + \hat \Psi_n - \hat H_n ) \nonumber\\&= (1-b_n) \Lambda_{n-1} - b_n (H^*+\hat \Psi_n -\Phi_n(H^*) - \Psi_n(H^*)\nonumber\\&= (1-b_n) \Lambda_{n-1} - b_n \Psi_n(\Upsilon(\overline H_{n-1})) + b_n (\Phi_n(H^*) - H^*)\nonumber\\
-&= (1-b_n) \Lambda_{n-1} - b_n\Psi_n(\Lambda_{n-1}')+ b_n (\Phi_n(H^*) - H^*),\label{eq:step1last}
- \end{align}
-where $\Lambda_n' = \Upsilon(\overline H_n) - H^*$.
-The equality in \eqref{eq:step1last} follows from \eqref{eq:psi}. 
-Unrolling the recursion \eqref{eq:step1last}, we obtain 
- \begin{align}\label{lambda-exp}
-  \Lambda_n  = \left[ \prod_{k=1}^n (1-b_k) \right]\Lambda_0 
-&- \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right] b_k \Psi_k(\Lambda_{k-1}') \nonumber\\
-&+ \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right] b_k (\Phi_k(H^*) - H^*) \hspace{0.2cm} a.s.
- \end{align}
- %%(note : $\prod_{j=n+1}^n (1-b_j) = 1$ for all n).
- %%Let us  characterise $trace[\E(\Lambda_n\tr \Lambda_n)]$ using \eqref{lambda-exp}. From independence of $d_k$ along k, \eqref{lambda-exp} represent martingale difference sequence, leading to 
-Squaring on both sides and taking expectations, we obtain
- \begin{align}
-  \E (\Lambda_n \tr \Lambda_n) =  &\left[ \prod_{k=1}^n (1-b_k) \right]^2 \E (\Lambda_0 \tr \Lambda_0) + \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2   \E (\Psi_k(\Lambda_{k-1}') \tr \Psi_k(\Lambda_{k-1}'))  \nonumber\\ &+ \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2  \times \E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*)) \nonumber\\
-&-2 \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2 \E\left(\Psi_k(\Lambda_{k-1}) \tr (\Phi_k(H^*) - H^*)\right).\label{eq:exp-lambda2}
- \end{align}
-The equality above uses the fact that $\E(\Psi_k(\Lambda_{k-1}'))=0$, $\E(\Phi_k(H^*) - H^*)=0$, which get rid of the corresponding cross terms with the first term on RHS of \eqref{lambda-exp}.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
+%%%%%%%%%%%%%%%%%%%%%%%% Quadratic case
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%We next present a convergence rate result for the special case of a quadratic objective function under the following additional assumptions:
+%\begin{enumerate}[label=(\textbf{C\arabic*}),resume]
+%\item  $f$ is quadratic and $\nabla^2 f(x^*) > 0$. 
+%\item The operator $\Upsilon$ is chosen such that $\E(\parallel \Upsilon(\overline H_n) - \overline H_n\parallel^2) = o(e^{-2wn^{1-r}/(1-r)})$ and $\parallel \Upsilon(H) - H \parallel^2 / (1+\parallel H \parallel^2)$ is uniformly bounded.
+%\end{enumerate}
+%\begin{theorem}(\textbf{Quadratic case - Convergence rate})
+%\label{thm:quad-bound}
+%Assume (C4), (C10), (C11) and (C12) and also that the setting is noise-free. 
+%Let $b_n = b_0/n^r$, $n=1,2,\ldots,k$, where $1/2 < r < 1$ and $0 < b_0 \leq 1$. For notational simplicity, let $H^*=\nabla^2 f(x^*)$. Letting $\Lambda_k = \overline H_k - H^*$, we have 
+%\begin{align}
+%\text{trace}[\E (\Lambda_n \tr \Lambda_n)] = O(e^{-2b_0n^{1-r} / {1-r}}).
+%\label{eq:quad-bigo}
+%\end{align}
+%\end{theorem}
+%\begin{proof}
+%Since the setting is noise-free with a quadratic objective, we can rewrite \eqref{eq:hnhat} as follows:
+%\begin{align}
+% \widehat H_n =    \Phi_n(\nabla^2 f(\theta_n)) &+\Psi_{n}(\nabla^2 f(\theta_n)),\label{eq:hnhat-ext}
+%\end{align}
+%where $\Psi_n(H)$ is defined in \eqref{eq:psi} and for any matrix $H$, 
+%$$\Phi_{n}(H) = [M_n]_{D}\left(d_{n}\tr \, [H]_{D} \, d_{n}\right) +  [M_n]_{N}\left(d_{n}\tr \, [H]_{N} \, d_{n}\right).$$
+%The proof involves the following steps:
+%\begin{description}
+%  \item[Step 1:] Here we prove the MSE convergence of $\overline H_k$, i.e., $\E [\Lambda_n\tr \Lambda_n] \rightarrow 0$ a.s. as $n\rightarrow \infty$.
+%  \item[Step 2:] We unroll the recursion \eqref{eq:2rdsa-H} and then derive convenient representation for $\text{trace}[\E (\Lambda_n \tr \Lambda_n)]$.
+%  \item[Step 3:] We derive the main result in \eqref{eq:quad-bigo} using a proof by contradiction. 
+%\end{description}
+%\noindent\textbf{Step 1: MSE convergence of $\overline H_n$} \\
+%This part in exactly the same manner as part (i) in Theorem 3 of \cite{spall-jacobian}.\\
+%\noindent\textbf{Step 2: Representation of $\text{trace}[\E (\Lambda_n \tr \Lambda_n)]$} \\
+% From \eqref{eq:2rdsa-H} and \eqref{eq:hnhat-ext},  we have
+% \begin{align}
+% \Lambda_n &= \Lambda_{n-1} - b_n ( \overline H_{n-1} - \hat H_n + \hat \Psi_n) \nonumber\\
+%&= (1-b_n) \Lambda_{n-1} - b_n (H^* + \hat \Psi_n - \hat H_n ) \nonumber\\&= (1-b_n) \Lambda_{n-1} - b_n (H^*+\hat \Psi_n -\Phi_n(H^*) - \Psi_n(H^*)\nonumber\\&= (1-b_n) \Lambda_{n-1} - b_n \Psi_n(\Upsilon(\overline H_{n-1})) + b_n (\Phi_n(H^*) - H^*)\nonumber\\
+%&= (1-b_n) \Lambda_{n-1} - b_n\Psi_n(\Lambda_{n-1}')+ b_n (\Phi_n(H^*) - H^*),\label{eq:step1last}
+% \end{align}
+%where $\Lambda_n' = \Upsilon(\overline H_n) - H^*$.
+%The equality in \eqref{eq:step1last} follows from \eqref{eq:psi}. 
+%Unrolling the recursion \eqref{eq:step1last}, we obtain 
+% \begin{align}\label{lambda-exp}
+%  \Lambda_n  = \left[ \prod_{k=1}^n (1-b_k) \right]\Lambda_0 
+%&- \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right] b_k \Psi_k(\Lambda_{k-1}') \nonumber\\
+%&+ \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right] b_k (\Phi_k(H^*) - H^*) \hspace{0.2cm} a.s.
+% \end{align}
+% %%(note : $\prod_{j=n+1}^n (1-b_j) = 1$ for all n).
+% %%Let us  characterise $trace[\E(\Lambda_n\tr \Lambda_n)]$ using \eqref{lambda-exp}. From independence of $d_k$ along k, \eqref{lambda-exp} represent martingale difference sequence, leading to 
+%Squaring on both sides and taking expectations, we obtain
+% \begin{align}
+%  \E (\Lambda_n \tr \Lambda_n) =  &\left[ \prod_{k=1}^n (1-b_k) \right]^2 \E (\Lambda_0 \tr \Lambda_0) + \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2   \E (\Psi_k(\Lambda_{k-1}') \tr \Psi_k(\Lambda_{k-1}'))  \nonumber\\ &+ \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2  \times \E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*)) \nonumber\\
+%&-2 \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2 \E\left(\Psi_k(\Lambda_{k-1}) \tr (\Phi_k(H^*) - H^*)\right).\label{eq:exp-lambda2}
+% \end{align}
+%The equality above uses the fact that $\E(\Psi_k(\Lambda_{k-1}'))=0$, $\E(\Phi_k(H^*) - H^*)=0$, which get rid of the corresponding cross terms with the first term on RHS of \eqref{lambda-exp}.
 Notice that there are only $n$ terms in the last term in the RHS of \eqref{eq:exp-lambda2} because $d_k$ are identically distributed for each $k$ and across $k$.
-% %
-We now characterize $\text{trace}[\E(\Lambda_n\tr \Lambda_n)]$ using \eqref{eq:exp-lambda2} as follows:
-%From the independence of $d_k$'s and $\Lambda_{k-1}'$ and using Cauchy-Schwartz inequality we get 
- \begin{align}\label{eq:trace}
- \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] = &\left[ \prod_{k=1}^n (1-b_k) \right]^2 \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  + \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2  \,\,\tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ &+ \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2   \times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right]  \nonumber\\ &- 2 \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2  \times \text{trace} \left[\E\left(\Psi_k(\Lambda_{k-1}) \tr (\Phi_k(H^*) - H^*)\right)\right].
- \end{align}
+%% %
+%We now characterize $\text{trace}[\E(\Lambda_n\tr \Lambda_n)]$ using \eqref{eq:exp-lambda2} as follows:
+%%From the independence of $d_k$'s and $\Lambda_{k-1}'$ and using Cauchy-Schwartz inequality we get 
 % \begin{align}\label{eq:trace}
-% & \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] \leq \left[ \prod_{k=1}^n (1-b_k) \right]^2 \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  \nonumber\\ &+ \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2  \,\,\tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ &+ \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2  \nonumber\\&  \hspace{0.5cm}\times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right] + \nonumber\\ &+ \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2   \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)^{1/2} \nonumber\\&  \times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right]^{1/2}.
+% \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] = &\left[ \prod_{k=1}^n (1-b_k) \right]^2 \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  + \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2  \,\,\tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ &+ \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2   \times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right]  \nonumber\\ &- 2 \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2  \times \text{trace} \left[\E\left(\Psi_k(\Lambda_{k-1}) \tr (\Phi_k(H^*) - H^*)\right)\right].
 % \end{align}
-
-As in the proof of Theorem 3 of \cite{spall-jacobian}, observe that $1-b_k = e^{-b_k}(1-O(b_k^2))$ and since $0 < b_k <1$, we have that the $O(b_k^2)$ term is strictly positive. 
-Letting $\Gamma_{ij} = \sum_{k=i}^j b_k$ with $\Gamma_{nn} = 1$ and $\beta_{kn} = \left[\prod_{i=k+1}^n (1- O(b_i^2))\right]^2$, we can simplify \eqref{eq:trace} as follows:
-\begin{align}\label{eq:wij}
- \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] = &e^{-2 \Gamma_{1n}} \beta_{0n} \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  + e^{-2 \Gamma_{1n}} \sum_{k=1}^n e^{2 \Gamma_{1k}} \beta_{kn} b_k^2  \,\,\tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ &+ e^{-2 \Gamma_{1n}} \sum_{k=1}^n e^{2 \Gamma_{1k}} \beta_{kn} b_k^2  \times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right] \nonumber\\ &- 2 e^{-2 \Gamma_{1n}} \sum_{k=1}^n e^{2 \Gamma_{1k}} \beta_{kn} b_k^2   \times \text{trace} \left[\E\left(\Psi_k(\Lambda_{k-1}) \tr (\Phi_k(H^*) - H^*)\right)\right] .
- \end{align} 
-% \begin{align}\label{eq:wij}
-% & \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] \leq e^{-2 \Gamma_{1n}} \beta_{0n} \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  \nonumber\\ &+ e^{-2 \Gamma_{1n}} \sum_{k=1}^n e^{2 \Gamma_{1k}} \beta_{kn} b_k^2  \,\,\tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ &+ e^{-2 \Gamma_{1n}} \sum_{k=1}^n e^{2 \Gamma_{1k}} \beta_{kn} b_k^2  \nonumber\\&  \hspace{0.5cm}\times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right] \nonumber\\ &+ e^{-2 \Gamma_{1n}} \sum_{k=1}^n e^{2 \Gamma_{1k}} \beta_{kn} b_k^2  \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)^{1/2} \nonumber\\&  \times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right]^{1/2}.
+% % \begin{align}\label{eq:trace}
+%% & \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] \leq \left[ \prod_{k=1}^n (1-b_k) \right]^2 \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  \nonumber\\ &+ \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2  \,\,\tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ &+ \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2  \nonumber\\&  \hspace{0.5cm}\times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right] + \nonumber\\ &+ \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2   \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)^{1/2} \nonumber\\&  \times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right]^{1/2}.
+%% \end{align}
+%
+%As in the proof of Theorem 3 of \cite{spall-jacobian}, observe that $1-b_k = e^{-b_k}(1-O(b_k^2))$ and since $0 < b_k <1$, we have that the $O(b_k^2)$ term is strictly positive. 
+%Letting $\Gamma_{ij} = \sum_{k=i}^j b_k$ with $\Gamma_{nn} = 1$ and $\beta_{kn} = \left[\prod_{i=k+1}^n (1- O(b_i^2))\right]^2$, we can simplify \eqref{eq:trace} as follows:
+%\begin{align}\label{eq:wij}
+% \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] = &e^{-2 \Gamma_{1n}} \beta_{0n} \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  + e^{-2 \Gamma_{1n}} \sum_{k=1}^n e^{2 \Gamma_{1k}} \beta_{kn} b_k^2  \,\,\tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ &+ e^{-2 \Gamma_{1n}} \sum_{k=1}^n e^{2 \Gamma_{1k}} \beta_{kn} b_k^2  \times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right] \nonumber\\ &- 2 e^{-2 \Gamma_{1n}} \sum_{k=1}^n e^{2 \Gamma_{1k}} \beta_{kn} b_k^2   \times \text{trace} \left[\E\left(\Psi_k(\Lambda_{k-1}) \tr (\Phi_k(H^*) - H^*)\right)\right] .
 % \end{align} 
-Comparing the sum with integrals, we obtain
-\begin{align*}
-\Gamma_{ij} = \int_i^j \frac{b_0}{x^r} dx + O(1) &= \left(\frac{b_0}{1-r}\right)(j^{1-r}-i^{1-r}) +O(1),
-\end{align*}
-where we have used the facts that $0 < b_k < 1, \forall k \ge 2$ and 
-$\sum_{k=i}^j b_k \to \infty$ as $j-i \to \infty$ since 
-$b_k = b_0/k^r$ with $r > 0.5$.
-Observing that $\beta_{kn}$ are uniformly upper-bounded, say by $\bar \beta_n$, we have
-\begin{align}\label{eq:thmwts}
- & \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] =  e^{-2 \Gamma_{1n}} \beta_{0n} \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  \nonumber\\ 
-&+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}  \times \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ 
-&+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}} \times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right] \nonumber\\ &- 2 \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}   \times \text{trace} \left[\E\left(\Psi_k(\Lambda_{k-1}) \tr (\Phi_k(H^*) - H^*)\right)\right].
-\end{align} 
-\begin{align}\label{eq:thmcomb}
-& \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] =  e^{-2 \Gamma_{1n}} \beta_{0n} \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  \nonumber\\ 
-&+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}  \times \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ 
-&+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}  \text{trace}\left[\E [((\Phi_k(H^*) - H^*) - 2 \Psi_k(\Lambda_{k-1}))\tr  (\Phi_k(H^*) - H^*)]\right] .
-\end{align} 
-%
+%% \begin{align}\label{eq:wij}
+%% & \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] \leq e^{-2 \Gamma_{1n}} \beta_{0n} \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  \nonumber\\ &+ e^{-2 \Gamma_{1n}} \sum_{k=1}^n e^{2 \Gamma_{1k}} \beta_{kn} b_k^2  \,\,\tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ &+ e^{-2 \Gamma_{1n}} \sum_{k=1}^n e^{2 \Gamma_{1k}} \beta_{kn} b_k^2  \nonumber\\&  \hspace{0.5cm}\times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right] \nonumber\\ &+ e^{-2 \Gamma_{1n}} \sum_{k=1}^n e^{2 \Gamma_{1k}} \beta_{kn} b_k^2  \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)^{1/2} \nonumber\\&  \times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right]^{1/2}.
+%% \end{align} 
+%Comparing the sum with integrals, we obtain
+%\begin{align*}
+%\Gamma_{ij} = \int_i^j \frac{b_0}{x^r} dx + O(1) &= \left(\frac{b_0}{1-r}\right)(j^{1-r}-i^{1-r}) +O(1),
+%\end{align*}
+%where we have used the facts that $0 < b_k < 1, \forall k \ge 2$ and 
+%$\sum_{k=i}^j b_k \to \infty$ as $j-i \to \infty$ since 
+%$b_k = b_0/k^r$ with $r > 0.5$.
+%Observing that $\beta_{kn}$ are uniformly upper-bounded, say by $\bar \beta_n$, we have
 %\begin{align}\label{eq:thmwts}
 % & \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] =  e^{-2 \Gamma_{1n}} \beta_{0n} \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  \nonumber\\ 
-%&+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}} \nonumber\\
-%& \hspace{3.5cm} \times \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ 
-%&+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}  \nonumber\\
-%&  \hspace{0.5cm}\times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right] \nonumber\\ &+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}  \nonumber\\&\hspace{3.5cm}\times \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)^{1/2}\nonumber\\&\times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right]^{1/2}.
+%&+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}  \times \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ 
+%&+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}} \times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right] \nonumber\\ &- 2 \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}   \times \text{trace} \left[\E\left(\Psi_k(\Lambda_{k-1}) \tr (\Phi_k(H^*) - H^*)\right)\right].
 %\end{align} 
-\noindent\textbf{Step 3: The big-O result on $\text{trace}[\E (\Lambda_n \tr \Lambda_n)]$ convergence} \\
-In comparison to Eq.
+%\begin{align}\label{eq:thmcomb}
+%& \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] =  e^{-2 \Gamma_{1n}} \beta_{0n} \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  \nonumber\\ 
+%&+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}  \times \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ 
+%&+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}  \text{trace}\left[\E [((\Phi_k(H^*) - H^*) - 2 \Psi_k(\Lambda_{k-1}))\tr  (\Phi_k(H^*) - H^*)]\right] .
+%\end{align} 
+%%
+%%\begin{align}\label{eq:thmwts}
+%% & \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] =  e^{-2 \Gamma_{1n}} \beta_{0n} \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  \nonumber\\ 
+%%&+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}} \nonumber\\
+%%& \hspace{3.5cm} \times \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ 
+%%&+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}  \nonumber\\
+%%&  \hspace{0.5cm}\times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right] \nonumber\\ &+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}  \nonumber\\&\hspace{3.5cm}\times \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)^{1/2}\nonumber\\&\times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right]^{1/2}.
+%%\end{align}
+%\noindent\textbf{Step 3: The big-O result on $\text{trace}[\E (\Lambda_n \tr \Lambda_n)]$ convergence} \\
+%In comparison to Eq.
 (7.6) in \cite{spall-jacobian} that corresponds to \eqref{eq:thmwts} for the N-SPSA-4 setting, there are two extra terms - the third and the fourth - in the RHS of \eqref{eq:thmwts}.
-Consider the third term on the RHS of \eqref{eq:thmwts}.
+%Consider the third term on the RHS of \eqref{eq:thmwts}.
 As a consequence of (C9) and the construction of random perturbations $d_k$, it can be seen that $\text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right]$ is uniformly bounded above, say by $\top$, independent of $k$.
 Thus, the third term is equivalent to the following:
-\begin{align}
- \bar c_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \frac{b_0^2}{k^{2 r}} \top.
-\label{eq:term3}
-\end{align}
-Now by moving out side exponential term to inside and changing the summation to integration, we can rewrite the above equation as follows
-\begin{align}
-  \overline{\overline{c_n}}  \int_1^n  \frac{1}{k^{2 r}} .
-\label{eq:term3}
-\end{align}
-Where $\overline{\overline{c_n}}$ is a constant. By above equation we can conclude that the third term is of $O(\frac{1}{k^{2 r -1}})$
-%Now, observe that $\lim_{n\rightarrow \infty} \sum_{k=1}^n \frac{1}{k^{2 r}} < \infty$ as $1/2 < r < 1$ and $e^{2 b_0 k^{1-r}/(1-r)}$ diverges as $n\rightarrow \infty$ and hence, by Kronecker's Lemma, the term in \eqref{eq:term3} vanishes asymptotically.
+%\begin{align}
+% \bar c_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \frac{b_0^2}{k^{2 r}} \top.
+%\label{eq:term3}
+%\end{align}
+%Now by moving out side exponential term to inside and changing the summation to integration, we can rewrite the above equation as follows
+%\begin{align}
+%  \overline{\overline{c_n}}  \int_1^n  \frac{1}{k^{2 r}} .
+%\label{eq:term3}
+%\end{align}
+%Where $\overline{\overline{c_n}}$ is a constant. By above equation we can conclude that the third term is of $O(\frac{1}{k^{2 r -1}})$
+%%Now, observe that $\lim_{n\rightarrow \infty} \sum_{k=1}^n \frac{1}{k^{2 r}} < \infty$ as $1/2 < r < 1$ and $e^{2 b_0 k^{1-r}/(1-r)}$ diverges as $n\rightarrow \infty$ and hence, by Kronecker's Lemma, the term in \eqref{eq:term3} vanishes asymptotically.
+%%
+%%Now consider the fourth term.
+By using the uniform upper-bound $\top$ on $\text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right]$, we can rewrite  the fourth term in \eqref{eq:thmwts} as follows:
+%% \begin{align}\label{eq:fourthterm}
+%% \bar c_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \frac{b_0^2}{k^{2 r}} \sqrt{\top} \nonumber \\ \hspace{3cm}\times \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)^{1/2},
+%% \end{align}
+%%where, as in the proof of Theorem 3 of \cite{spall-jacobian}, $\tau(\cdot)$ transforms the $\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')$ in a linear fashion and then returns the trace of the resulting $N\times N$ matrix.
 %
-%Now consider the fourth term.
-By using the uniform upper-bound $\top$ on $\text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right]$, we can rewrite  the fourth term in \eqref{eq:thmwts} as follows:
-% \begin{align}\label{eq:fourthterm}
-% \bar c_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \frac{b_0^2}{k^{2 r}} \sqrt{\top} \nonumber \\ \hspace{3cm}\times \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)^{1/2},
-% \end{align}
-%where, as in the proof of Theorem 3 of \cite{spall-jacobian}, $\tau(\cdot)$ transforms the $\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')$ in a linear fashion and then returns the trace of the resulting $N\times N$ matrix.
-
-\textbf{Part (iii) needed fix} 
-%of Theorem 3 in \cite{spall-jacobian} uses a proof by contradiction to show that the first and second terms in \eqref{eq:thmwts} are both of the order $O(e^{-2b_0n^{1-r} / {1-r}})$.
+%\textbf{Part (iii) needed fix} 
+%%of Theorem 3 in \cite{spall-jacobian} uses a proof by contradiction to show that the first and second terms in \eqref{eq:thmwts} are both of the order $O(e^{-2b_0n^{1-r} / {1-r}})$.
 By a completely parallel argument to that used for the second term in \cite{spall-jacobian}, it can be seen that the fourth term in our setting vanishes at a faster rate than the second term.
 The claim follows.
-\end{proof}
-
-\section{Simulation experiments}
-\label{sec:expts}
-\subsection{Implementation}
-We test the performance of 2RDSA-Unif, 2RDSA-AsymBer and 2SPSA, with/without improved Hessian estimation. 
-2SPSA algorithm uses Bernoulli $\pm 1$-valued perturbations, while 2RDSA/2RDSA-IH come in two variants - one that uses $U[-1,1]$ distributed perturbations (referred to as 2RDSA-Unif/2RDSA-IH-Unif) and the other that uses asymmetric Bernoulli perturbations (referred to as 2RDSA-AsymBer/2RDSA-IH-AsymBer)\footnote{The implementation is available at \url{https://github.com/prashla/RDSA/archive/master.zip}.}.
-
-For the empirical evaluations, we use the following two loss functions in $N=10$ dimensions:
-\paragraph{Quadratic loss}
-\begin{align}
-f(x) = \theta\tr A \theta+ b\tr \theta.\label{eq:quadratic}
-\end{align} 
-The optimum $\theta^*$ for the above $f$ is such that each coordinate of $\theta^*$ is $-0.9091$, with $f(\theta^*) = -4.55$.
-
-\paragraph{Fourth-order loss}
-\begin{align} 
-f(x) = \theta\tr A\tr A \theta + 0.1 \sum_{j=1}^N (A\theta)^3_j + 0.01 \sum_{j=1}^N (A\theta)^4_j.\label{eq:4thorder}
- \end{align} 
-The optimum $\theta^*$ for above $f$ is $\theta^*=0$, with $f(\theta^*) = 0$.
-
-In both functions, $A$ is such that $NA$ is an upper triangular matrix with each entry one, $b$ is the $N$-dimensional vector of ones and the noise structure is similar to that used in \cite{spall_adaptive}.
-For any $\theta$, the noise is $[\theta\tr, 1]z$, where $z \approx \N(0,\sigma^2 I_{11\times11})$.
-We perform experiments for noisy as well as noise-less settings, with $\sigma=0.1$ for the noisy case.
-
-For all algorithms, we set $\delta_n = 3.8/n^{0.101}$ and $a_n = 1/n^{0.6}$, while $b_n$ are set according to \eqref{eq:wieghts}.
-These choices have been used  for 2SPSA implementations before (see \cite{spall_adaptive}) and have demonstrated good finite-sample performance empirically, while satisfying the theoretical requirements needed for asymptotic convergence.
-For all the algorithms, the initial point $\theta_0$ is the $N$-dimensional vector of ones.
-For both 2SPSA and 2RDSA/2RDSA-IH, an initial $20\%$ of the  simulation budget was used up by 1SPSA/1RDSA and the resulting iterate was used to initialize 2SPSA/2RDSA.
-The distribution parameter $\epsilon$ is set to $0.0001$ for 2RDSA and to $0.01$ for 1RDSA.
-
-\subsection{Results}
-We use normalized loss and normalized MSE (NMSE) as performance metrics for evaluating the algorithms. 
-NMSE is the ratio $\l \theta_{n_\text{end}} - \theta^* \r^2 / \l \theta_0 - \theta^*\r^2$, while normalized loss is the ratio $f(\theta_{n_\text{end}})/f(\theta_0)$.
-Here $n_\text{end}$ denotes the iteration number when the algorithm stopped updating its parameter.
-Note that $n_\text{end}$ is a function of the simulation budget.
-2RDSA/2RDSA-IH use only three simulations per-iteration and hence, $n_\text{end}$ is $1/3$rd of the simulation budget, while it is $1/4$th of the simulation budget for 2SPSA, since the latter algorithm uses four simulations per-iteration.
-
-\begin{table}
-\centering
- \caption{Normalized loss values for fourth-order  objective \eqref{eq:4thorder} with and without noise: standard error from $500$ replications shown after $\pm$}
-\label{tab:norloss-4thf}
-\begin{tabular}{|c|c|c|}
-\toprule
-\rowcolor{gray!20}
-\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0.1$}}}\\[1em]
-%&&&\\
-\midrule
-  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
-  & & \textbf{ estimation} \\
- \midrule
-\textbf{2SPSA} & $0.132 \pm 0.0267$ & $0.104 \pm 0.0355$\\
-&&\\
-\textbf{2RDSA-Unif} &$0.115 \pm 0.0214$ & $0.0271 \pm 0.0538$\\ 
-&&\\
-\textbf{2RDSA-AsymBer}& $0.0471 \pm 0.021$& $\bm{0.0099 \pm 0.0014}$\\
- \bottomrule
- %%%%%%%%%%%%2nd ordr
-\rowcolor{gray!20}
-\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0$}}}\\[1em]
-%&&&\\
-\midrule
-  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
-  & & \textbf{ estimation} \\
- \midrule
-\textbf{2SPSA} & $0.0795 \pm 0.0234$ & $0.0628 \pm 0.0234$\\
-&&\\
-\textbf{2RDSA-Unif} &$0.0813 \pm 0.0275$ & $0.0214 \pm 0.00376$\\ 
-&&\\
-\textbf{2RDSA-AsymBer}& $0.0199 \pm 0.0114$& $\bm{0.0098 \pm 0.00147}$\\
- \bottomrule
-\end{tabular}
-\end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%% Quadratic
-\begin{table}
-\centering
- \caption{Normalized loss values for quadratic objective \eqref{eq:quadratic} with and without noise: standard error from $500$ replications shown after $\pm$}
-\label{tab:norloss-quadratic}
-\adjustbox{max height=\dimexpr\textheight-10cm\relax,
-           max width=\textwidth}{
-\begin{tabular}{|c|c|c|}
-\toprule
-\rowcolor{gray!20}
-\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0.1$}}}\\[1em]
-%&&&\\
-\midrule
-  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
-  & & \textbf{ estimation} \\
- \midrule
-\textbf{2SPSA} & $-0.0062 \pm 0.1164$ & $-0.1229 \pm 0.1374$\\
-&&\\
-\textbf{2RDSA-Unif} &$0.0485 \pm 0.1465$ & $-0.259 \pm 0.0398$\\ 
-&&\\
-\textbf{2RDSA-AsymBer}& $-0.2564 \pm 0.068$& $\bm{-0.2877 \pm 0.0051}$\\
- \bottomrule
- %%%%%%%%%%%%2nd ordr
-\rowcolor{gray!20}
-\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0$}}}\\[1em]
-%&&&\\
-\midrule
-  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
-  & & \textbf{ estimation} \\
- \midrule
-\textbf{2SPSA} & $-0.0785 \pm 0.1178$ & $-0.1716 \pm 0.1339$\\
-&&\\
-\textbf{2RDSA-Unif} &$0.0326 \pm 0.1599$ & $-0.2672 \pm 0.0299$\\ 
-&&\\
-\textbf{2RDSA-AsymBer}& $-0.2777 \pm 0.0488$& $\bm{-0.2881 \pm 0.0012}$\\
- \bottomrule
-\end{tabular}}
-
-\end{table}
-
-%%%%%%%%%%%%%% NMSE for quadratic
-\begin{table}
-\centering
- \caption{NMSE values for quadratic objective \eqref{eq:quadratic} with and without noise: standard error from $500$ replications shown after $\pm$}
-\label{tab:nmse-quadratic}
-\adjustbox{max height=\dimexpr\textheight-10cm\relax,
-           max width=\textwidth}{
-\begin{tabular}{|c|c|c|}
-\toprule
-\rowcolor{gray!20}
-\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0.1$}}}\\[1em]
-%&&&\\
-\midrule
-  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
-  & & \textbf{ estimation} \\
- \midrule
-\textbf{2SPSA} & $0.9491 \pm 0.0131$ & $0.5495 \pm 0.0217$\\
-&&\\
-\textbf{2RDSA-Unif} &$1.0073 \pm 0.0140$ & $0.1953 \pm 0.0095$\\ 
-&&\\
-\textbf{2RDSA-AsymBer}& $0.1667 \pm 0.0095$& $\bm{0.0324 \pm 0.0007}$\\
- \bottomrule
- %%%%%%%%%%%%2nd ordr
-\rowcolor{gray!20}
-\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0$}}}\\[1em]
-%&&&\\
-\midrule
-  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
-  & & \textbf{ estimation} \\
- \midrule
-\textbf{2SPSA} & $0.7325 \pm 0.0180$ & $0.3939 \pm 0.0230$\\
-&&\\
-\textbf{2RDSA-Unif} &$0.9834 \pm 0.0170$ & $0.1623 \pm 0.0086$\\ 
-&&\\
-\textbf{2RDSA-AsymBer}& $0.0686 \pm 0.0078$& $\bm{0.0316 \pm 0.0006}$\\
- \bottomrule
-\end{tabular}}
-\end{table}
-
-\newcommand{\errorband}[5][]{ % x column, y column, error column, optional argument for setting style of the area plot
-\pgfplotstableread[col sep=comma, skip first n=2]{#2}\datatable
-    % Lower bound (invisible plot)
-    \addplot [draw=none, stack plots=y, forget plot] table [
-        x={#3},
-        y expr=\thisrow{#4}-2*\thisrow{#5}
-    ] {\datatable};
-
-    % Stack twice the error, draw as area plot
-    \addplot [draw=none, fill=gray!40, stack plots=y, area legend, #1] table [
-        x={#3},
-        y expr=4*\thisrow{#5}
-    ] {\datatable} \closedcycle;
-
-    % Reset stack using invisible plot
-    \addplot [forget plot, stack plots=y,draw=none] table [x={#3}, y expr=-(\thisrow{#4}+2*\thisrow{#5})] {\datatable};
-}
-
- \begin{figure}
-    \centering
-\tabl{c}{\scalebox{0.9}{\begin{tikzpicture}
-      \begin{axis}[
-	xlabel={number of simulations},
-	ylabel={Normalized loss},
-       legend entries={
-	 ,2SPSA,,2SPSA-IH,,2RDSA-Unif,,2RDSA-Unif-IH,,2RDSA-AsymBer,,2RDSA-AsymBer-IH },
-        %legend pos=outer north east,
-			legend style={
-at={(0,0)},
-anchor=north,at={(axis description cs:0.5,-0.2)},legend columns=2}	]
-      % 2SPSA
-      \errorband[red!50!white, opacity=0.3]{norloss_vs_simulations_noisycase.csv}{0}{1}{2}
-      \addplot [thick, red,mark=square*,mark options={scale=0.7}] table [x index=0, y index=1,col sep=comma] {norloss_vs_simulations_noisycase.csv};
-			%%% With IH
-      % 2SPSA
-      \errorband[darkgray!50!white, opacity=0.3]{norloss_vs_simulations_noisycase.csv}{0}{7}{8}
-      \addplot [thick, darkgray,mark=*,mark options={scale=0.75}] table [x index=0, y index=7,col sep=comma] {norloss_vs_simulations_noisycase.csv};
-
-      % 2RDSA-Unif
-      \errorband[blue!50!white, opacity=0.3]{norloss_vs_simulations_noisycase.csv}{0}{3}{4}
-      \addplot [thick, blue,mark=square*,mark options={scale=0.7}] table [x index=0, y index=3,col sep=comma] {norloss_vs_simulations_noisycase.csv};
-			%%% With IH
-      % 2RDSA-Unif
-      \errorband[magenta!50!white, opacity=0.3]{norloss_vs_simulations_noisycase.csv}{0}{9}{10}
-      \addplot [thick, magenta,mark=*,mark options={scale=0.75}] table [x index=0, y index=9,col sep=comma] {norloss_vs_simulations_noisycase.csv};
-
-      % 2RDSA-AsymBer
-      \errorband[green!50!white, opacity=0.3]{norloss_vs_simulations_noisycase.csv}{0}{5}{6}
-      \addplot [thick, darkgreen,mark=square*,mark options={scale=0.7}] table [x index=0, y index=5,col sep=comma] {norloss_vs_simulations_noisycase.csv};
-			%%% With IH	
-      % 2RDSA-AsymBer
-      \errorband[violet!50!white, opacity=0.3]{norloss_vs_simulations_noisycase.csv}{0}{11}{12}
-      \addplot [thick, violet, mark=*,mark options={scale=0.75}] table [x index=0, y index=11,col sep=comma] {norloss_vs_simulations_noisycase.csv};
-      \end{axis}
-      \end{tikzpicture}}\\}
-			\caption{Normalized loss vs.
-number of simulations for fourth-order loss \eqref{eq:4thorder} with $\sigma=0.1$ for 2SPSA, 2RDSA-Unif and 2RDSA-AsymBer algorithms with/without improved Hessian estimation: bands around the curves represent standard error from $500$ replications.
-}
-      \label{fig:norlossvssims} 
-\end{figure}
-
-Tables \ref{tab:norloss-4thf}--\ref{tab:norloss-quadratic} present the normalized loss values observed for the three algorithms - 2SPSA, 2RDSA-Unif and 2RDSA-AsymBer - with/without improved Hessian estimation scheme and for the fourth-order and quadratic loss functions, respectively.
-Table \ref{tab:nmse-quadratic} presents the NMSE values obtained for the aforementioned algorithms with the quadratic loss.
-The results in Tables \ref{tab:norloss-4thf}--\ref{tab:nmse-quadratic} are obtained after running all the algorithms with a budget of $10000$ function evaluations.
- Figure \ref{fig:norlossvssims} plots the normalized loss as a function of the simulation budget with the fourth-order loss objective with $\sigma=0.1$ 
-%  (see Figures \ref{fig:norlossvssimsNoiseless}--\ref{fig:norlossvssimsNoiseless} in the appendix for similar results with $\sigma = 0$ for fourth-order loss and $\sigma=0.1$ for quadratic loss). 
- From the results in Tables \ref{tab:norloss-4thf}--\ref{tab:nmse-quadratic} and Fig \ref{fig:norlossvssims}, we make the following observations:
-
-\textit{\textbf{Observation 1:} Among 2RDSA schemes, 2RDSA-IH performs better than regular 2RDSA, for both perturbation choices.}
-
-\textit{\textbf{Observation 2:} 2RDSA-IH variants outperform both 2SPSA and 2SPSA-IH, with 2RDSA-IH-AsymBer performing the best overall.}
-
-\begin{figure}[h]
-    \centering
-		\begin{tabular}{c}
-		\begin{subfigure}{0.5\textwidth}
-\tabl{c}{\scalebox{0.8}{\begin{tikzpicture}
-      \begin{axis}[
-	xlabel={number of simulations},
-	ylabel={Normalized loss},
-       legend entries={
-	 ,2SPSA,,2SPSA-IH,,2RDSA-Unif,,2RDSA-Unif-IH,,2RDSA-AsymBer,,2RDSA-AsymBer-IH },
-        %legend pos=outer north east,
-			legend style={
-at={(0,0)},
-anchor=north,at={(axis description cs:0.5,-0.2)},legend columns=2}	]
-      % 2SPSA
-      \errorband[red!50!white, opacity=0.3]{norloss_vs_simulations.csv}{0}{1}{2}
-      \addplot [thick, red,mark=square*,mark options={scale=0.7}] table [x index=0, y index=1,col sep=comma] {norloss_vs_simulations.csv};
-			%%% With IH
-      % 2SPSA
-      \errorband[darkgray!50!white, opacity=0.3]{norloss_vs_simulations.csv}{0}{7}{8}
-      \addplot [thick, darkgray,mark=*,mark options={scale=0.75}] table [x index=0, y index=7,col sep=comma] {norloss_vs_simulations.csv};
-
-      % 2RDSA-Unif
-      \errorband[blue!50!white, opacity=0.3]{norloss_vs_simulations.csv}{0}{3}{4}
-      \addplot [thick, blue,mark=square*,mark options={scale=0.7}] table [x index=0, y index=3,col sep=comma] {norloss_vs_simulations.csv};
-			%%% With IH
-      % 2RDSA-Unif
-      \errorband[magenta!50!white, opacity=0.3]{norloss_vs_simulations.csv}{0}{9}{10}
-      \addplot [thick, magenta,mark=*,mark options={scale=0.75}] table [x index=0, y index=9,col sep=comma] {norloss_vs_simulations.csv};
-
-      % 2RDSA-AsymBer
-      \errorband[green!50!white, opacity=0.3]{norloss_vs_simulations.csv}{0}{5}{6}
-      \addplot [thick, darkgreen,mark=square*,mark options={scale=0.7}] table [x index=0, y index=5,col sep=comma] {norloss_vs_simulations.csv};
-			%%% With IH	
-      % 2RDSA-AsymBer
-      \errorband[violet!50!white, opacity=0.3]{norloss_vs_simulations.csv}{0}{11}{12}
-      \addplot [thick, violet, mark=*,mark options={scale=0.75}] table [x index=0, y index=11,col sep=comma] {norloss_vs_simulations.csv};
-      \end{axis}
-      \end{tikzpicture}}\\}
-			\caption{Fourth-order loss \eqref{eq:4thorder} with $\sigma=0$. }
-      \label{fig:norlossvssimsNoiseless} 
-\end{subfigure} 
-\\
-\\
-%%%%%%% Norloss-Quadratic-noisy %%%%%%%%%%%%%%
-\begin{subfigure}{0.5\textwidth}
-\tabl{c}{\scalebox{0.8}{\begin{tikzpicture}
-      \begin{axis}[
-	xlabel={number of simulations},
-	ylabel={Normalized loss},
-       legend entries={
-	 ,2SPSA,,2SPSA-IH,,2RDSA-Unif,,2RDSA-Unif-IH,,2RDSA-AsymBer,,2RDSA-AsymBer-IH },
-        %legend pos=outer north east,
-			legend style={
-at={(0,0)},
-anchor=north,at={(axis description cs:0.5,-0.2)},legend columns=2}	]
-      % 2SPSA
-      \errorband[red!50!white, opacity=0.3]{quad_norloss_vs_simulations_noisycase.csv}{0}{1}{2}
-      \addplot [thick, red,mark=square*,mark options={scale=0.7}] table [x index=0, y index=1,col sep=comma] {quad_norloss_vs_simulations_noisycase.csv};
-			%%% With IH
-      % 2SPSA
-      \errorband[darkgray!50!white, opacity=0.3]{quad_norloss_vs_simulations_noisycase.csv}{0}{7}{8}
-      \addplot [thick, darkgray,mark=*,mark options={scale=0.75}] table [x index=0, y index=7,col sep=comma] {quad_norloss_vs_simulations_noisycase.csv};
-
-      % 2RDSA-Unif
-      \errorband[blue!50!white, opacity=0.3]{quad_norloss_vs_simulations_noisycase.csv}{0}{3}{4}
-      \addplot [thick, blue,mark=square*,mark options={scale=0.7}] table [x index=0, y index=3,col sep=comma] {quad_norloss_vs_simulations_noisycase.csv};
-			%%% With IH
-      % 2RDSA-Unif
-      \errorband[magenta!50!white, opacity=0.3]{quad_norloss_vs_simulations_noisycase.csv}{0}{9}{10}
-      \addplot [thick, magenta,mark=*,mark options={scale=0.75}] table [x index=0, y index=9,col sep=comma] {quad_norloss_vs_simulations_noisycase.csv};
-
-      % 2RDSA-AsymBer
-      \errorband[green!50!white, opacity=0.3]{quad_norloss_vs_simulations_noisycase.csv}{0}{5}{6}
-      \addplot [thick, darkgreen,mark=square*,mark options={scale=0.7}] table [x index=0, y index=5,col sep=comma] {quad_norloss_vs_simulations_noisycase.csv};
-			%%% With IH	
-      % 2RDSA-AsymBer
-      \errorband[violet!50!white, opacity=0.3]{quad_norloss_vs_simulations_noisycase.csv}{0}{11}{12}
-      \addplot [thick, violet, mark=*,mark options={scale=0.75}] table [x index=0, y index=11,col sep=comma] {quad_norloss_vs_simulations_noisycase.csv};
-      \end{axis}
-      \end{tikzpicture}}\\}
-			\caption{Quadratic loss \eqref{eq:quadratic} with $\sigma=0.1$. }
-      \label{fig:QuadnorlossvssimsNoisy} 
-\end{subfigure} 
-		\end{tabular}
-			\caption{Normalized loss vs. number of simulations in two different loss settings for all the algorithms. }
-      \label{fig:extras} 
-		\end{figure}
-
-%%%%%% Norloss-Quadratic-noisefree %%%%%%%%%%%%%%
-\begin{figure}
-    \centering
-\tabl{c}{\scalebox{0.9}{\begin{tikzpicture}
-      \begin{axis}[
-	xlabel={number of simulations},
-	ylabel={Normalized loss},
-       legend entries={
-	 ,2SPSA,,2SPSA-IH,,2RDSA-Unif,,2RDSA-Unif-IH,,2RDSA-AsymBer,,2RDSA-AsymBer-IH },
-        %%legend pos=outer north east,
-			legend style={
-at={(0,0)},
-anchor=north,at={(axis description cs:0.5,-0.2)},legend columns=2}	]
-      % 2SPSA
-      \errorband[red!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{1}{2}
-      \addplot [thick, red,mark=square*,mark options={scale=0.7}] table [x index=0, y index=1,col sep=comma] {quad_norloss_vs_simulations.csv};
-			%%%% With IH
-      %2SPSA
-      \errorband[darkgray!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{7}{8}
-      \addplot [thick, darkgray,mark=*,mark options={scale=0.75}] table [x index=0, y index=7,col sep=comma] {quad_norloss_vs_simulations.csv};
-			%
-      % 2RDSA-Unif
-      \errorband[blue!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{3}{4}
-      \addplot [thick, blue,mark=square*,mark options={scale=0.7}] table [x index=0, y index=3,col sep=comma] {quad_norloss_vs_simulations.csv};
-			%%%% With IH
-      % 2RDSA-Unif
-      \errorband[magenta!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{9}{10}
-      \addplot [thick, magenta,mark=*,mark options={scale=0.75}] table [x index=0, y index=9,col sep=comma] {quad_norloss_vs_simulations.csv};
-
-      % 2RDSA-AsymBer
-      \errorband[green!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{5}{6}
-      \addplot [thick, darkgreen,mark=square*,mark options={scale=0.7}] table [x index=0, y index=5,col sep=comma] {quad_norloss_vs_simulations.csv};
-			%%%% With IH	
-      % 2RDSA-AsymBer
-      \errorband[violet!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{11}{12}
-      \addplot [thick, violet, mark=*,mark options={scale=0.75}] table [x index=0, y index=11,col sep=comma] {quad_norloss_vs_simulations.csv};
-      \end{axis}
-      \end{tikzpicture}}\\}
-			\caption{Normalized loss vs.
-number of simulations for quadratic loss \eqref{eq:quadratic} with $\sigma=0$ for 2SPSA, 2RDSA-Unif and 2RDSA-AsymBer algorithms with/without improved Hessian estimation.
-}
-      \label{fig:QuadnorlossvssimsNoiseless} 
-\end{figure}
+%\end{proof}

### 2017-01-05 experiments.tex
added spsa-3 tables

@@
 We test the performance of 2RDSA-Unif, 2RDSA-AsymBer, 2SPSA-3 and 2SPSA, with/without improved Hessian estimation. 
-2SPSA and 2SPSA-3 algorithm uses Bernoulli $\pm 1$-valued perturbations, while 2RDSA/2RDSA-IH come in two variants - one that uses $U[-1,1]$ distributed perturbations (referred to as 2RDSA-Unif/2RDSA-IH-Unif) and the other that uses asymmetric Bernoulli perturbations (referred to as 2RDSA-AsymBer/2RDSA-IH-AsymBer)\footnote{The implementation is available at \url{https://github.com/prashla/RDSA/archive/master.zip}.}.
+2SPSA and 2SPSA-3 algorithms use Bernoulli $\pm 1$-valued perturbations, while 2RDSA/2RDSA-IH come in two variants - one that uses $U[-1,1]$ distributed perturbations (referred to as 2RDSA-Unif/2RDSA-IH-Unif) and the other that uses asymmetric Bernoulli perturbations (referred to as 2RDSA-AsymBer/2RDSA-IH-AsymBer)\footnote{The implementation is available at \url{https://github.com/prashla/RDSA/archive/master.zip}.}.
 
@@
 \label{tab:norloss-4thf}
-\scalebox{0.9}{
+\scalebox{0.8}{
 \begin{tabular}{|c|c|c|}
@@
 \textbf{2SPSA} & $0.132 \pm 0.0267$ & $0.104 \pm 0.0355$\\
+&&\\
+\textbf{2SPSA-3} & $0.080 \pm 0.0053$ & $0.043 \pm 0.0015$\\
 &&\\
@@
 &&\\
+\textbf{2SPSA-3} & $0.0338 \pm 0.0001$ & $0.0315 \pm 0.0007$\\
+&&\\
 \textbf{2RDSA-Unif} &$0.0813 \pm 0.0275$ & $0.0214 \pm 0.00376$\\ 
@@
 \label{tab:norloss-quadratic}
-\scalebox{0.9}{
+\scalebox{0.8}{
 \begin{tabular}{|c|c|c|}
@@
 \textbf{2SPSA} & $-0.0062 \pm 0.1164$ & $-0.1229 \pm 0.1374$\\
+&&\\
+\textbf{2SPSA-3} & $-0.0161 \pm 0.0955$ & $-0.2748 \pm 0.0629$\\
 &&\\
@@
 &&\\
+\textbf{2SPSA-3} & $-0.0667 \pm 0.0602$ & $-0.2793 \pm 0.0625$\\
+&&\\
 \textbf{2RDSA-Unif} &$0.0326 \pm 0.1599$ & $-0.2672 \pm 0.0299$\\ 
@@
 
-\end{table}
-
-%%%%%%%%%%%%%% NMSE for quadratic
-\begin{table}
-\centering
- \caption{NMSE values for quadratic objective \eqref{eq:quadratic} with and without noise: standard error from $500$ replications shown after $\pm$}
-\label{tab:nmse-quadratic}
-\scalebox{0.9}{
-\begin{tabular}{|c|c|c|}
-\toprule
-\rowcolor{gray!20}
-\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0.1$}}}\\[1em]
-%&&&\\
-\midrule
-  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
-  & & \textbf{ estimation} \\
- \midrule
-\textbf{2SPSA} & $0.9491 \pm 0.0131$ & $0.5495 \pm 0.0217$\\
-&&\\
-\textbf{2RDSA-Unif} &$1.0073 \pm 0.0140$ & $0.1953 \pm 0.0095$\\ 
-&&\\
-\textbf{2RDSA-AsymBer}& $0.1667 \pm 0.0095$& $\bm{0.0324 \pm 0.0007}$\\
- \bottomrule
- %%%%%%%%%%%%2nd ordr
-\rowcolor{gray!20}
-\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0$}}}\\[1em]
-%&&&\\
-\midrule
-  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
-  & & \textbf{ estimation} \\
- \midrule
-\textbf{2SPSA} & $0.7325 \pm 0.0180$ & $0.3939 \pm 0.0230$\\
-&&\\
-\textbf{2RDSA-Unif} &$0.9834 \pm 0.0170$ & $0.1623 \pm 0.0086$\\ 
-&&\\
-\textbf{2RDSA-AsymBer}& $0.0686 \pm 0.0078$& $\bm{0.0316 \pm 0.0006}$\\
- \bottomrule
-\end{tabular}}
 \end{table}
@@
 
-Tables \ref{tab:norloss-4thf}--\ref{tab:norloss-quadratic} present the normalized loss values observed for the three algorithms - 2SPSA, 2RDSA-Unif and 2RDSA-AsymBer - with/without improved Hessian estimation scheme and for the fourth-order and quadratic loss functions, respectively.
+%%%%%%%%%%%%%% NMSE for quadratic
+\begin{table}
+\centering
+ \caption{NMSE values for quadratic objective \eqref{eq:quadratic} with and without noise: standard error from $500$ replications shown after $\pm$}
+\label{tab:nmse-quadratic}
+%\scalebox{0.8}{
+\begin{tabular}{|c|c|c|}
+\toprule
+\rowcolor{gray!20}
+\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0.1$}}}\\[1em]
+%&&&\\
+\midrule
+  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
+  & & \textbf{ estimation} \\
+ \midrule
+\textbf{2SPSA} & $0.9491 \pm 0.0131$ & $0.5495 \pm 0.0217$\\
+&&\\
+\textbf{2SPSA-3} & $0.8378 \pm 0.0179$ & $0.1045 \pm 0.0005$\\
+&&\\
+\textbf{2RDSA-Unif} &$1.0073 \pm 0.0140$ & $0.1953 \pm 0.0095$\\ 
+&&\\
+\textbf{2RDSA-AsymBer}& $0.1667 \pm 0.0095$& $\bm{0.0324 \pm 0.0007}$\\
+ \bottomrule
+ %%%%%%%%%%%%2nd ordr
+\rowcolor{gray!20}
+\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0$}}}\\[1em]
+%&&&\\
+\midrule
+  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
+  & & \textbf{ estimation} \\
+ \midrule
+\textbf{2SPSA} & $0.7325 \pm 0.0180$ & $0.3939 \pm 0.0230$\\
+&&\\
+\textbf{2SPSA-3} & $0.6661 \pm 0.01554$ & $0.0684 \pm 0.0006$\\
+&&\\
+\textbf{2RDSA-Unif} &$0.9834 \pm 0.0170$ & $0.1623 \pm 0.0086$\\ 
+&&\\
+\textbf{2RDSA-AsymBer}& $0.0686 \pm 0.0078$& $\bm{0.0316 \pm 0.0006}$\\
+ \bottomrule
+\end{tabular}
+\end{table}
+
+Tables \ref{tab:norloss-4thf}--\ref{tab:norloss-quadratic} present the normalized loss values observed for the four algorithms - 2SPSA, 2SPSA-3, 2RDSA-Unif and 2RDSA-AsymBer - with/without improved Hessian estimation scheme and for the fourth-order and quadratic loss functions, respectively.
 Table \ref{tab:nmse-quadratic} presents the NMSE values obtained for the aforementioned algorithms with the quadratic loss.
 The results in Tables \ref{tab:norloss-4thf}--\ref{tab:nmse-quadratic} are obtained after running all the algorithms with a budget of $10000$ function evaluations.
- Figure \ref{fig:norlossvssims} plots the normalized loss as a function of the simulation budget with the fourth-order loss objective with $\sigma=0.1$ 
-%  (see Figures \ref{fig:norlossvssimsNoiseless}--\ref{fig:norlossvssimsNoiseless} in the appendix for similar results with $\sigma = 0$ for fourth-order loss and $\sigma=0.1$ for quadratic loss). 
+
+ Figure \ref{fig:norlossvssims} plots the normalized loss as a function of the simulation budget with the fourth-order loss objective with $\sigma=0.1$.
+Figure \ref{fig:QuadnorlossvssimsNoiseless} plots the normalized loss as a function of the simulation budget with the quadratic loss objective with $\sigma=0$.
+See Figures \ref{fig:norlossvssimsNoiseless}--\ref{fig:norlossvssimsNoiseless}  for similar results with $\sigma = 0$ for fourth-order loss and $\sigma=0.1$ for quadratic loss.
  From the results in Tables \ref{tab:norloss-4thf}--\ref{tab:nmse-quadratic} and Fig \ref{fig:norlossvssims}, we make the following observations:

### 2017-01-05 intro.tex
experiments added
added spsa-3 tables

@@
 In such cases it makes difficult for the search procedure to update the search parameter with out estimating the cost over infinite long trajectory.
+This kind of objective functions are very common in reinforcement learning applications.
 Another subclassifications of optimization problems is based on feasible region $\Theta$, see \eqref{eq:opt}.
@@
 In Chapter~\ref{sec:2rdsa-gen}, we describe the generalised RDSA scheme for both the gradient and Hessian estimates and also present the theoretical convergence results as well.
-The work related to improving the Hessian estimation scheme 2SPSA-3 was presented in Chapter~\ref{sec:2spsa-3-ih} and finally, in  Chapter~\ref{sec:conclusions}, the concluding remarks were provided.
+The work related to improving the Hessian estimation scheme 2SPSA-3 was presented in Chapter~\ref{sec:2spsa-3-ih}.
+In Chapter~\ref{sec:expts}, we provide simulation experiments for all the methods proposed in the previous chapters and finally, in  Chapter~\ref{sec:conclusions}, the concluding remarks and future directions were provided.

### 2017-01-06 experiments.tex
version submitted to sir

@@
 &&\\
-\textbf{2SPSA-3} & $0.080 \pm 0.0053$ & $0.043 \pm 0.0015$\\
+\textbf{2SPSA-3} & $0.0893 \pm 0.001$ & $0.0457 \pm 0.0004$\\
 &&\\
@@
 
- Figure \ref{fig:norlossvssims} plots the normalized loss as a function of the simulation budget with the fourth-order loss objective with $\sigma=0.1$.
-Figure \ref{fig:QuadnorlossvssimsNoiseless} plots the normalized loss as a function of the simulation budget with the quadratic loss objective with $\sigma=0$.
-See Figures \ref{fig:norlossvssimsNoiseless}--\ref{fig:norlossvssimsNoiseless}  for similar results with $\sigma = 0$ for fourth-order loss and $\sigma=0.1$ for quadratic loss.
- From the results in Tables \ref{tab:norloss-4thf}--\ref{tab:nmse-quadratic} and Fig \ref{fig:norlossvssims}, we make the following observations:
+ Figures \ref{fig:norlossvssims}, \ref{fig:norlossvssimsspsa-3} plots the normalized loss as a function of the simulation budget with the fourth-order loss objective with $\sigma=0.1$ for algorithms 2RDSA and 2SPSA-3 respectively when compared with 2SPSA.
+Figures \ref{fig:QuadnorlossvssimsNoiseless}, \ref{fig:QuadnorlossvssimsNoiselessspsa3} plots the normalized loss as a function of the simulation budget with the quadratic loss objective with $\sigma=0$ for algorithms 2RDSA and 2SPSA-3 respectively when compared with 2SPSA.
+See Figures \ref{fig:norlossvssimsNoiseless}-- \ref{fig:QuadnorlossvssimsNoisy} and \ref{fig:norlossvssimsNoiselessspsa3}-- \ref{fig:QuadnorlossvssimsNoisyspsa3} for similar results with $\sigma = 0$ for fourth-order loss and $\sigma=0.1$ for quadratic loss.
+ From the results in Tables \ref{tab:norloss-4thf}--\ref{tab:nmse-quadratic} and Figures, we make the following observations:
 
- \textit{\textbf{Observation 1:} Among 2RDSA schemes, 2RDSA-IH performs better than regular 2RDSA, for both perturbation choices.}
+ \textit{\textbf{Observation 1:} Schemes with improved Hessian estimation performs better than their respective regular schemes.}
 
-\textit{\textbf{Observation 2:} 2RDSA-IH variants outperform both 2SPSA and 2SPSA-IH, with 2RDSA-IH-AsymBer performing the best overall.}
+\textit{\textbf{Observation 2:} 2RDSA-IH variants outperform both 2SPSA and 2SPSA-IH.}
+
+\textit{\textbf{Observation 3:} 2SPSA-3-IH variants outperform both 2SPSA and 2SPSA-IH.}
+
+\textit{\textbf{Observation 4:} 2RDSA-IH-AsymBer performing the best overall.}
+
+\begin{figure}
+    \centering
+\tabl{c}{\scalebox{0.9}{\begin{tikzpicture}
+      \begin{axis}[
+	xlabel={number of simulations},
+	ylabel={Normalized loss},
+       legend entries={
+	 ,2SPSA,,2SPSA-IH,,2SPSA-3,,2SPSA-3-IH },
+        %legend pos=outer north east,
+			legend style={
+at={(0,0)},
+anchor=north,at={(axis description cs:0.5,-0.2)},legend columns=2}	]
+      % 2SPSA
+      \errorband[red!50!white, opacity=0.3]{norloss_vs_simulations_noisycase.csv}{0}{1}{2}
+      \addplot [thick, red,mark=square*,mark options={scale=0.7}] table [x index=0, y index=1,col sep=comma] {norloss_vs_simulations_noisycase_spsa3.csv};
+			%%% With IH
+      % 2SPSA
+      \errorband[darkgray!50!white, opacity=0.3]{norloss_vs_simulations_noisycase.csv}{0}{5}{6}
+      \addplot [thick, darkgray,mark=*,mark options={scale=0.75}] table [x index=0, y index=5,col sep=comma] {norloss_vs_simulations_noisycase_spsa3.csv};
+
+      % 2SPSA-3
+      \errorband[blue!50!white, opacity=0.3]{norloss_vs_simulations_noisycase.csv}{0}{3}{4}
+      \addplot [thick, blue,mark=square*,mark options={scale=0.7}] table [x index=0, y index=3,col sep=comma] {norloss_vs_simulations_noisycase_spsa3.csv};
+			%%% With IH
+      % 2SPSA-3
+      \errorband[magenta!50!white, opacity=0.3]{norloss_vs_simulations_noisycase.csv}{0}{7}{8}
+      \addplot [thick, magenta,mark=*,mark options={scale=0.75}] table [x index=0, y index=7,col sep=comma] {norloss_vs_simulations_noisycase_spsa3.csv};
+
+            \end{axis}
+      \end{tikzpicture}}\\}
+			\caption{Normalized loss vs.
+number of simulations for fourth-order loss \eqref{eq:4thorder} with $\sigma=0.1$ for 2SPSA, 2SPSA-3 algorithms with/without improved Hessian estimation: bands around the curves represent standard error from $500$ replications.
+}
+      \label{fig:norlossvssimsspsa-3} 
+\end{figure}
+
+%%%%%% Norloss-Quadratic-noisefree %%%%%%%%%%%%%%
+\begin{figure}
+    \centering
+\tabl{c}{\scalebox{0.9}{\begin{tikzpicture}
+      \begin{axis}[
+	xlabel={number of simulations},
+	ylabel={Normalized loss},
+       legend entries={
+	 ,2SPSA,,2SPSA-IH,,2SPSA-3,,2SPSA-3-IH  },
+        %%legend pos=outer north east,
+			legend style={
+at={(0,0)},
+anchor=north,at={(axis description cs:0.5,-0.2)},legend columns=2}	]
+      % 2SPSA
+      \errorband[red!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{1}{2}
+      \addplot [thick, red,mark=square*,mark options={scale=0.7}] table [x index=0, y index=1,col sep=comma] {quad_norloss_vs_simulations_spsa3.csv};
+			%%%% With IH
+      %2SPSA
+      \errorband[darkgray!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{5}{6}
+      \addplot [thick, darkgray,mark=*,mark options={scale=0.75}] table [x index=0, y index=5,col sep=comma] {quad_norloss_vs_simulations_spsa3.csv};
+			%
+      % 2SPSA-3
+      \errorband[blue!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{3}{4}
+      \addplot [thick, blue,mark=square*,mark options={scale=0.7}] table [x index=0, y index=3,col sep=comma] {quad_norloss_vs_simulations_spsa3.csv};
+			%%%% With IH
+      % 2SPSA-3
+      \errorband[magenta!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{7}{8}
+      \addplot [thick, magenta,mark=*,mark options={scale=0.75}] table [x index=0, y index=7,col sep=comma]  {quad_norloss_vs_simulations_spsa3.csv};
+
+           \end{axis}
+      \end{tikzpicture}}\\}
+			\caption{Normalized loss vs. number of simulations for quadratic loss \eqref{eq:quadratic} with $\sigma=0$ for 2SPSA, 2SPSA-3 algorithms with/without improved Hessian estimation. }
+      \label{fig:QuadnorlossvssimsNoiselessspsa3} 
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+		\begin{tabular}{c}
+		\begin{subfigure}{0.5\textwidth}
+\tabl{c}{\scalebox{0.8}{\begin{tikzpicture}
+      \begin{axis}[
+	xlabel={number of simulations},
+	ylabel={Normalized loss},
+       legend entries={
+	 ,2SPSA,,2SPSA-IH,,2SPSA-3,,2SPSA-3-IH },
+        %legend pos=outer north east,
+			legend style={
+at={(0,0)},
+anchor=north,at={(axis description cs:0.5,-0.2)},legend columns=2}	]
+      % 2SPSA
+      \errorband[red!50!white, opacity=0.3]{norloss_vs_simulations.csv}{0}{1}{2}
+      \addplot [thick, red,mark=square*,mark options={scale=0.7}] table [x index=0, y index=1,col sep=comma] {norloss_vs_simulations_spsa3.csv};
+			%%% With IH
+      % 2SPSA
+      \errorband[darkgray!50!white, opacity=0.3]{norloss_vs_simulations.csv}{0}{5}{6}
+      \addplot [thick, darkgray,mark=*,mark options={scale=0.75}] table [x index=0, y index=5,col sep=comma]  {norloss_vs_simulations_spsa3.csv};
+
+      % 2RDSA-Unif
+      \errorband[blue!50!white, opacity=0.3]{norloss_vs_simulations.csv}{0}{3}{4}
+      \addplot [thick, blue,mark=square*,mark options={scale=0.7}] table [x index=0, y index=3,col sep=comma] {norloss_vs_simulations_spsa3.csv};
+			%%% With IH
+      % 2RDSA-Unif
+      \errorband[magenta!50!white, opacity=0.3]{norloss_vs_simulations.csv}{0}{7}{8}
+      \addplot [thick, magenta,mark=*,mark options={scale=0.75}] table [x index=0, y index=7,col sep=comma]  {norloss_vs_simulations_spsa3.csv};
+
+            \end{axis}
+      \end{tikzpicture}}\\}
+			\caption{Fourth-order loss \eqref{eq:4thorder} with $\sigma=0$. }
+      \label{fig:norlossvssimsNoiselessspsa3} 
+\end{subfigure} 
+\\
+\\
+%%%%%%% Norloss-Quadratic-noisy %%%%%%%%%%%%%%
+\begin{subfigure}{0.5\textwidth}
+\tabl{c}{\scalebox{0.9}{\begin{tikzpicture}
+      \begin{axis}[
+	xlabel={number of simulations},
+	ylabel={Normalized loss},
+       legend entries={
+	 ,2SPSA,,2SPSA-IH,,2SPSA-3,,2SPSA-3-IH },
+        %legend pos=outer north east,
+			legend style={
+at={(0,0)},
+anchor=north,at={(axis description cs:0.5,-0.2)},legend columns=2}	]
+      % 2SPSA
+      \errorband[red!50!white, opacity=0.3]{quad_norloss_vs_simulations_noisycase.csv}{0}{1}{2}
+      \addplot [thick, red,mark=square*,mark options={scale=0.7}] table [x index=0, y index=1,col sep=comma] {quad_norloss_vs_simulations_noisycase_spsa3.csv};
+			%%% With IH
+      % 2SPSA
+      \errorband[darkgray!50!white, opacity=0.3]{quad_norloss_vs_simulations_noisycase.csv}{0}{5}{6}
+      \addplot [thick, darkgray,mark=*,mark options={scale=0.75}] table [x index=0, y index=5,col sep=comma]  {quad_norloss_vs_simulations_noisycase_spsa3.csv};
+
+      % 2RDSA-Unif
+      \errorband[blue!50!white, opacity=0.3]{quad_norloss_vs_simulations_noisycase.csv}{0}{3}{4}
+      \addplot [thick, blue,mark=square*,mark options={scale=0.7}] table [x index=0, y index=3,col sep=comma] {quad_norloss_vs_simulations_noisycase_spsa3.csv};
+			%%% With IH
+      % 2RDSA-Unif
+      \errorband[magenta!50!white, opacity=0.3]{quad_norloss_vs_simulations_noisycase.csv}{0}{7}{8}
+      \addplot [thick, magenta,mark=*,mark options={scale=0.75}] table [x index=0, y index=7,col sep=comma] {quad_norloss_vs_simulations_noisycase_spsa3.csv};
+
+            \end{axis}
+      \end{tikzpicture}}\\}
+			\caption{Quadratic loss \eqref{eq:quadratic} with $\sigma=0.1$. }
+      \label{fig:QuadnorlossvssimsNoisyspsa3} 
+\end{subfigure} 
+		\end{tabular}
+			\caption{Normalized loss vs. number of simulations in two different loss settings for  2SPSA, 2SPSA-3 algorithms. }
+      \label{fig:extras} 
+		\end{figure}

### 2017-01-06 intro.tex
version submitted to sir

@@
 However, perturbations have to satisfy the i.i.d, mean-zero assumptions.
-Apart from these common assumptions  one additional assumption like difference between the second moment square  and fourth moment should be non zero for Hessian estimation is also required.
+Apart from these common assumptions  one additional assumption, which is the difference between the second moment square  and fourth moment should be non zero for Hessian estimation is also required.
 This generalisation of distributions of random variables for perturbations is important because it enhances the scope of improving the performance of the 1RDSA and 2RDSA schemes by incorporating several other distributions.

### 2017-01-13 abstractShort.tex
sir corrections first intro only

@@
 continuously valued set. One can use information on the gradient/Hessian
-of the objective to aid the search process. However, due to the lack of
+of the objective to aid the search process. However, due to lack of
 knowledge of the noise distribution, one needs to estimate the
@@
 
-We propose enhancements to the Hessian estimation scheme used in two recently proposed stochastic Newton methods, based on the ideas of random directions stochastic approximation (2RDSA) \cite{prashanth2015rdsa} and simultaneous perturbation stochastic approximation (2SPSA-3\footnote{The 3 in the abberviation of the algorithm is used to indicate that the algorithms in \cite{bhatnagar2015simultaneous} require 3 function evaluations per iteration.}) \cite{bhatnagar2015simultaneous}, respectively.
+We propose enhancements to the Hessian estimation scheme used in two recently proposed stochastic Newton methods, based on the ideas of random directions stochastic approximation (2RDSA) \cite{prashanth2015rdsa} and simultaneous perturbation stochastic approximation (2SPSA-3\footnote{The 3 in the abbereviation of the algorithm is used to indicate that the algorithm in \cite{bhatnagar2015simultaneous} requires 3 function evaluations per iteration.}) \cite{bhatnagar2015simultaneous}, respectively.
 The proposed scheme, inspired by \cite{spall-jacobian}, reduces the error in the Hessian estimate by
@@
 \end{inparaenum}
-We prove that 2RDSA and 2SPSA-3 with our Hessian improvement scheme converges asymptotically to the true Hessian.
+We prove that both 2RDSA and 2SPSA-3 with our Hessian improvement scheme converges asymptotically to the true Hessian.
  %Moreover, for the special of a quadratic objective, we provide a convergence rate result for 2RDSA-IH that is of the same order - in the number of iterations of Hessian recursion - as that of 2SPSA with improved Hessian estimation (2SPSA-IH) \cite{spall-jacobian}.
-The advantage with 2RDSA and 2SPSA-3  is that it requires only 75\% of the simulation cost per-iteration for 2SPSA with improved Hessian estimation (2SPSA-IH) \cite{spall-jacobian}.
+The key advantage with 2RDSA and 2SPSA-3  is that they require only 75\% of the simulation cost per-iteration for 2SPSA with improved Hessian estimation (2SPSA-IH) \cite{spall-jacobian}.
 Numerical experiments show that 2RDSA-IH outperforms both 2SPSA-IH and 2RDSA without the improved Hessian estimation scheme.

### 2017-01-13 intro.tex
sir corrections first intro only

--- 2016-12-24 intro.tex
@@
 \chapter{Introduction}
-Optimization problems can be classified into several branches depending upon the nature of the problem.
-Consider the following optimization problem which is to find a $\theta^{*}$ that minimizes the objective function $f(\theta)$:
+Optimization problems can be classified into various categories depending upon the nature of the problem.
+Consider the  problem of finding a $\theta^{*}$ that minimizes the objective function $f(\theta)$:
 %\begin{equation} 
@@
 %\end{equation}
-where $f \colon \mathbb{R}^N \to \mathbb{R}$ is called the objective function, $\theta$ is tunable N-dimensional parameter and $\Theta \subseteq \mathbb{R}^N$ is the constraint set in which $\theta$ takes values.
+where $f \colon \mathbb{R}^N \to \mathbb{R}$ is called the objective function, $\theta$ is a tunable $N$-dimensional parameter and $\Theta \subseteq \mathbb{R}^N$ is the constraint set in which $\theta$ takes values.
 \section{Classification of optimization problems based on objective function}\label{sc:objclass}
@@
 In the real world unfortunately, many problems do not fall in this class.
-However, Optimization problems involving uncertainties are very common in a variety of engineering disciplines such as transportation systems, manufacturing, networks, healthcare and finance.
+However, optimization problems involving uncertainties are very common in a variety of engineering disciplines such as transportation systems, manufacturing, networks, healthcare and finance.
 
-The setting in \emph{stochastic optimization}, however presumes that we have little knowledge on the structure of $f$ and moreover $f$ cannot be obtained directly, but rather is an expectation of another quantity $h(\theta,\xi)$, to which we have access, i.e.,
+The setting in \emph{stochastic optimization}, however, presumes that we have little knowledge on the structure of $f$ and moreover $f$ cannot be obtained directly, but rather is an expectation of another quantity $h(\theta,\xi)$, to which we have access, i.e.,
 \begin{align}\label{eq:stopt}
@@
 \end{align}
- where $\xi$ comprises the randomness in the system and one is allowed to observe only these $h(\theta,\xi)$ samples, largely  because one dose not have access to distribution of noise $\xi$.
-This kind of optimization problems are more challenging because of the added complexity of not knowing $f$ and to find $\theta^{*}$ only on the basis of aforementioned noisy samples.
-The large number of input variables and the lack of precise system model may prohibit analytical solution approaches and a viable alternative is to employ a simulation-based optimization approach.
+ where $\xi$ comprises the randomness in the system and one is allowed to observe only the $h(\theta,\xi)$ samples, largely  because one does not have access to the distribution of noise $\xi$.
+These kinds of optimization problems are more challenging because of the added complexity of not knowing $f$ and to find $\theta^{*}$ only on the basis of the aforementioned noisy samples.
+The large number of input variables and the lack of a precise system model may prohibit analytical solution approaches and a viable alternative is to employ a simulation-based optimization approach.
 As illustrated in Figure \ref{fig:so}, the idea here is to simulate a few times the stochastic system under consideration until a good enough solution is obtained.
@@
 
-Suppose the objective function $f$ has the form $f(\theta) = \sum_{i=1}^{n} E[h_{i}(X_{i})]$ , where $n$ denotes the number of stages and $X_{i}$ is the state of the underlying process in stage $i$ and $h_{i}$ denotes a stage and state dependent cost function.
-Then this type of optimization problems are called \emph{multi-stage problems}.
-Let $\theta = (\theta_{1},\ldots,\theta_{n})^{T}$,  with each $\theta_{j}$ being scalar and Let $X_{i}$ depend on $\theta_{1},\ldots,\theta_{i}$.
+Suppose the objective function $f$ has the form $f(\theta) = \sum_{i=1}^{n} E[h_{i}(X_{i})]$, where $n$ denotes the number of stages, $X_{i}$ is the state of the underlying process in stage $i$ and $h_{i}$ denotes a stage and state dependent cost function.
+Then such optimization problems are called \emph{multi-stage problems}.
+Let $\theta = (\theta_{1},\ldots,\theta_{n})^{T}$,  with each $\theta_{j}$ being a scalar and let $X_{i}$ depend on $\theta_{1},\ldots,\theta_{i}$.
 The idea is that optimization can be done one stage at a time over $n$ stages after observing the state $X_{i}$ in each stage $i$.
-The value $\theta_{i}$ in stage $i$ has bearing on the cost of all subsequent stages $i+1,...,n$.
+The value $\theta_{i}$ in stage $i$ has a bearing on the cost of all subsequent stages $i+1,...,n$.
 This is  a problem of dynamic optimization.
-Approaches such as Dynamic programming can be used to solve this optimization problem.\par
-Now we consider sub-class of multi-stage problems having infinite number of stages and having \emph{long-run average cost function} as objective. Objective function here is of the form:
+Approaches such as dynamic programming can be used to solve this optimization problem.\par
+One may also consider a sub-class of multi-stage problems having an infinite number of stages and with the \emph{long-run average cost function} as objective.
+The objective function here is of the form:
 \begin{align}\label{eq:long}
@@
 \end{align}
-where $X_{i}$ is defined same as before and we assume it depends on parameter $\theta$.
-Objective function \eqref{eq:long} in most cases is not known analytically.
-In such cases it makes difficult for the search procedure to update the search parameter with out estimating the cost over infinite long trajectory.
-Another subclassifications of optimization problems is based on feasible region $\Theta$, see \eqref{eq:opt}.
+where $X_{i}$ is defined  as before and is a function of the parameter $\theta$.
+The objective function \eqref{eq:long} in most cases is not known analytically.
+In such cases it becomes difficult for the search procedure to update the search parameter without estimating the cost over an infinitly long trajectory.
+Such  objective functions are  common in reinforcement learning applications.
+Another sub-class of optimization problems are those based on the feasible region $\Theta$, see \eqref{eq:opt}.
 
-\section{Classification of optimization problems based on feasible region}
-Consider the basic optimization problem in \eqref{eq:opt}. Optimization problems can be classified based on the structure of feasible region $\Theta$. Where set $\Theta$ can be of following types:
+\section{Classification of optimization problems based on constrained feasible region}
+Consider the basic optimization problem in \eqref{eq:opt}. The problem depends critically on the structure of the feasible region $\Theta$, where $\Theta$ can be of the following types:
 
@@
 \end{inparaenum}
-Hence corresponding optimization problems are called discrete, continuous, hybrid optimization problems.
-These subclassifications are common to both deterministic and stochastic optimization problems presented in section \ref{sc:objclass}.
+The corresponding optimization problems are called discrete, continuous or hybrid optimization problems.
+These classifications are common to both deterministic and stochastic optimization problems presented in section \ref{sc:objclass}.
 
 \section{Optimization methods via simulation }\label{sc:dis}
-Optimization via simulation is fast developing area for both research and practice.
-As mentioned before, for stochastic optimization problems simulation is a viable approach compared to solving them analytically.
-Multiple simulations replications must be performed in order to get good estimate of $E_{\xi}[h(\theta_i,\xi)]$, which is function value at $\theta_i$.
-The standard approach to estimate is by the sample mean
+Optimization via simulation is a fast developing area for both researchers and practioners.
+As mentioned before, for stochastic optimization problems, simulation is a viable approach compared.
+Multiple simulation replications must be performed in order to get a good estimate of $E_{\xi}[h(\theta_i,\xi)]$, which is the function value at parameter $\theta_i$.
+The standard approach to estimate this is via the sample mean
 \begin{align}\label{eq:sm}
-\overline f(\theta_i) = \frac{1}{n} ~  \sum_{j=1}^{n} h(\theta,\xi_j),
+\overline {f_n} (\theta_i) = \frac{1}{n} ~  \sum_{j=1}^{n} h(\theta,\xi_j),
 \end{align}
-where $n$ is the number of simulation replications and $\xi_j$ is the j-th sample of randomness.
-As $n \to \infty$, $\overline f(\theta_i) \to E_{\xi}[h(\theta_i,\xi)]$.
-Primary concerns in stochastic optimization via simulation are
+where $n$ is the number of simulation replications and $\xi_j$ is the $j$th sample of randomness.
+As $n \to \infty$, $\overline{ f_n} (\theta_i) \to E_{\xi}[h(\theta_i,\xi)]$.
+The primary concerns in stochastic optimization via simulation are
 \begin{inparaenum}[\bfseries (i)]
-	\item $n$ must be large enough to get good estimate of  $E_{\xi}[h(\theta_i,\xi)]$, which has ultimate impact on the final solution that the optimization procedure need to find; and
-	\item The above procedure must be applied for many different parameters ,i.e., $\theta$'s in order to find best $\theta$.
+	\item $n$ must be large enough to get a good estimate of  $E_{\xi}[h(\theta_i,\xi)]$, which has an impact on the final solution that the optimization procedure needs to find; and
+	\item The above procedure must be applied for many different parameters, i.e., $\theta$'s in order to find the best $\theta$.
 \end{inparaenum}
-Hence one of the key criteria for comparing different simulation-based optimization approaches is by total number of simulations performed by them to attain specific function value.
-Note that different approaches may take different simulations per single iteration.
+Hence one of the key criteria for comparing different simulation-based optimization approaches is through the total number of simulations performed by them to attain a specific function value.
+Note that different approaches may require different number of simulations per  iteration.
 In this section we present methods only for Continuous and Discrete optimization via simulation problems which often occur in practise.
@@
 \subsection{Discrete optimization via simulation}
-Optimization problems for which $\theta \in \Theta$ is discrete and integer ordered and $\Theta$ is a subset of $N$-dimensional integers are known as Discrete optimization via simulation problems.
-In the discrete case $\theta$ can take only countable (finite or infinite) number of values.
-Discrete optimization problems are some times called as \emph{combinatorial optimization} and have applications in resource allocation, network routing, policy planning etc.
+Optimization problems for which $\Theta$ is a discrete set are known as discrete optimization via simulation problems.
+In the discrete case $\theta$, can take only countable (finite or infinite) number of values.
+Discrete optimization problems are some times take the form of \emph{combinatorial optimization} problems and they have applications in resource allocation, network routing, policy planning etc.
 In this section we consider the case of $\Theta$ being finite i.e., $\Theta = \{\theta_{1},\theta_{2},\ldots,\theta_{k}\}$ where $k \in \mathbb{N}$.
-Discrete optimization problems are divided into two sub-classes when feasible region is  either small (often less than 100 ) or large.
-we present optimization methods in literature separately for each class.\\
+Discrete optimization problems are divided into two sub-classes when the feasible region is  either small (often less than 100) or is large.
+We present optimization methods in the literature separately for each class.\\
 %\begin{enumerate}
@@
 In this case we can simulate all possible solutions and select the best among them.
-However, unlike in deterministic optimization, simulating only once for each parameter is not enough since the objective function is noisy in stochastic optimization setting.
-Hence the main question is how to conduct multiple simulation effectively for each parameter in order to determine best $\theta$.
+However, unlike in deterministic optimization, simulating the system only once for each parameter is not enough since the objective function is noisy in the stochastic optimization setting.
+Hence the main question is how to conduct multiple simulations effectively for each parameter in order to determine the  best parameter $\theta$.
 Two popular methods in literature for this case are Ranking and Selection (R\&S)\cite{bech} and Multiple comparison procedures (MCP) \cite{mcp}.
-These methods are well suited for simulation since under lying assumptions such as normality and independence of observations can be met by careful selection procedures.
-Traditional methods are applicable for maximum size of 20 while some recent techniques allow more. 
+These methods are well suited for simulation since  the underlying assumptions such as normality and independence of observations can be met through careful selection procedures.
+Traditional methods in this setting are applicable for a maximum set size of 20 while some recent techniques allow more. 
 \begin{enumerate}
@@
 R\&S procedures are statistical procedures specially developed to select the best parameter or subset that contains the best parameter.
-Majority of work in R\&S is classified into  two categories.
-The first one is called \emph{Indifferent-zone ranking}\cite{bech,izrs} which is aimed  to select the best and another one is \emph{Subset selection}\cite{ssrs} where the procedures are designed to find subset consisting of best.
-Again there are two approaches for selecting the best problem : \emph{the frequentist approach} and \emph{ Bayesian approach}.
-One of the popular streams of Bayesian approach procedures is  optimal computing budget allocation (OCBA) procedures \cite{ocbabook,chen2000} in which simulation budget is allocated in a manner that maximises the posterior probability of correct selection.
+Majority of the work in R\&S is classified into  two categories.
+The first  is called \emph{Indifferent-zone ranking}\cite{bech,izrs} which is aimed  to select the best parameter  and the other one is \emph{Subset selection}\cite{ssrs} where the procedures are designed to find a subset of parameters consisting of the best parameter.
+Again there are two approaches for selecting the best parameter : \emph{the frequentist approach} and \emph{ Bayesian approach}.
+One of the popular approaches in the  Bayesian stream is the optimal computing budget allocation (OCBA) procedure \cite{ocbabook,chen2000} where the simulation budget is allocated in a manner that maximizes the posterior probability of correct selection.
 This method is widely considered the best procedure for small scale discrete optimization.
 
-\item \textbf {MCPs}
-Like R\&S, MCPs attempt to identify the best parameter.
-But MCPs approach optimization problem as a statistical inference problem and do not guarantee a decision.
-Three main classes of MCPs that are used in practise are \emph{Multiple Comparison approach} (MCA) , \emph{Multiple Comparison with the Best} (MCB) and \emph{Multiple Comparisons with Control} (MCC).
+\item \textbf {MCP}
+Like R\&S, MCP attempts to identify the best parameter.
+But MCPs approach the optimization problem as a statistical inference problem and do not guarantee a decision.
+Three main classes of MCP that are used in practise are \emph{Multiple Comparison approach} (MCA) , \emph{Multiple Comparison with the Best} (MCB) and \emph{Multiple Comparisons with Control} (MCC).
 The most popular among these is MCB.
-In particular, MCB looks at $f(\theta_{j}) - opt_{i \neq j} f(\theta_{i})$ for $j = 1,2,\ldots,k$ to determine that $j_{*}$ with $f(\theta_{j_{*}}) - opt_{ i \neq j_{*}} f(\theta_{i}) > 0$.
-Simultaneous confidence intervals $f(\theta_{j}) - opt_{i \neq j} f(\theta_{i})$ for $j = 1,2,\ldots,k$ can be used to determine $j_{*}$ by looking  for the confidence interval with lower confidence limit that is zero.
+In particular, MCB looks at $f(\theta_{j}) - opt_{i \neq j} f(\theta_{i})$ for $j = 1,2,\ldots,k$ to determine  $j_{*}$ such that $f(\theta_{j_{*}}) - opt_{ i \neq j_{*}} f(\theta_{i}) > 0$.
+Simultaneous confidence intervals $f(\theta_{j}) - opt_{i \neq j} f(\theta_{i})$ for $j = 1,2,\ldots,k$ can be used to determine $j_{*}$ by looking  for the confidence interval with the lower confidence limit of zero.
 \end{enumerate}
-For a comprehensive treatment of these methods one can refer \cite{swish,swish2}.
+For a comprehensive treatment of these methods, see \cite{swish,swish2}.
 \subsubsection{$\Theta$ is large}
-When $\Theta$ is large enumeration becomes too expensive to conduct.
-Some sort of search techniques like the ones in deterministic optimization must be applied to avoid simulating for all the parameters while ensuring a high chance of finding the best or good parameters.
+When $\Theta$ is large, simulating for each parameter value is expensive.
+Some sort of search techniques like the ones for deterministic optimization must be applied to avoid simulating for all the parameters while ensuring a high chance of finding the best or good parameters.
 Some approaches in this category include the following :
@@
 \item \emph{Model-based approaches:} Iterative algorithms using statistical methods search the feasible region to improve upon the candidate parameter.
-One can  use \emph{gradient/Hessian information} with respect to parameter to help the search similar to deterministic case.
-But due to lack of availability of function analytical form and distribution of noise one needs to estimate the gradient/Hessian from samples of function $f$.
-However gradient/Hessian  estimation could be  quite noisy due to stochastic nature involved  in these problems.
-There are several approaches  which are used in continuous case can also be used in this setting to estimate the gradient/Hessian when simulation model is treated as black box.
+One can  use \emph{gradient/Hessian information} with respect to the parameter to help the search similar to deterministic case.
+But due to lack of knowledge of the function in analytical form and distribution of noise one needs to estimate the gradient/Hessian from samples of the function $f$.
+However, gradient/Hessian  estimation could be  quite noisy due to the stochastic nature involved  in these problems.
+Various approaches  that are used in the continuous case can also be used in this setting to estimate the gradient/Hessian when the simulation model is treated as a black box.
 We will present detailed literature survey of these approaches in sections \ref{sec:gradest} and \ref{sec:hesest}.
@@
 They include approaches such as genetic algorithms, evolutionary algorithms, simulated annealing \cite{simann}, tabu search \cite{tabu} and cross entropy \cite{crossent}.
-Most of these iterative algorithms start with an initial population of parameters and in each iteration elite parameters are selected from previous population and generates  better parameter population as search progresses.
+Most of these iterative algorithms start with an initial population of parameters and in each iteration elite parameters are selected from the previous population and a better parameter population is generated as search progresses.
 \end{enumerate}
@@
 Optimization problems for which $\theta \in \Theta$ is a vector of continuous decision variables and $\Theta$ is a convex subset of $\mathcal{R}^N$ are called continuous optimization via simulation  problems.
-In these problems $\theta$ can take uncountable number of values.
-These kind of problems frequently occur in applications such as model fitting, adaptive control, neural network training, signal processing etc.
+In these problems $\theta$  takes an uncountable number of values.
+These kinds of problems frequently occur in applications such as model fitting, adaptive control, neural network training, signal processing etc.
 A natural solution approach is to devise an algorithm that incrementally updates the parameter, say $\theta_n$, in the descent direction using the gradient and/or Hessian of the objective $f$.
@@
 where $\theta_{n}$ is the solution found at iteration $n$.
-$ \widehat{\nabla} f(\theta_{n})$ is an estimate of the gradient $ {\nabla} f(\theta_{n})$ and $\{a_{n}\}$ is a sequence of positive reals satisfy following properties: $\sum_{n=1}^{\infty} a_{n}  = \infty$,  $\sum_{n = 1}^{\infty} a_{n}^{2} < \infty$, and $\Gamma_{\Theta}$ denotes the projection of a point outside $\Theta$ onto $\Theta$.
-Under appropriate conditions, as number of iterations goes to infinity one can guarantee convergence to the local minimum with probability one.
-The SA algorithm can be considered as a stochastic analog of gradient decent method which seeks the next solution along negative gradient direction.
-The main difference between the SA algorithm and steepest descent algorithm is even though gradient estimate in former algorithm is a noisy estimate it requires only weak assumptions (like $E \left[\widehat{\nabla} f(\theta_{n})\right] - \nabla f(\theta_{n}) \to 0 $ at a certain rate) for converging to local minimum.
-In practice, the performance of the SA algorithm is quite sensitive to the sequence $\{a_{n}\}$ for example in the case $ a_n = a/n$, the convergence is highly dependent on the choice of $a$.
-However, in practice, one can only obtain estimates of the function $f$ through black-box simulation and the challenge is to estimate the gradient and/or Hessian of $f$ from function samples.
-In the following sections we will present a brief survey of existing methods to estimate gradient and Hessian from function samples.
+Also, $ \widehat{\nabla} f(\theta_{n})$ is an estimate of the gradient $ {\nabla} f(\theta_{n})$ and $\{a_{n}\}$ is a sequence of positive reals satisfying the following properties: $\sum_{n=1}^{\infty} a_{n}  = \infty$,  $\sum_{n = 1}^{\infty} a_{n}^{2} < \infty$, and $\Gamma_{\Theta}$ denotes the projection operator.
+Under appropriate conditions, as the number of iterations goes to infinity one can guarantee convergence to the local minimum with probability one.
+The SA algorithm can be considered as a stochastic analog of the gradient decent method which seeks the next solution along the negative gradient direction.
+The main difference between the SA algorithm and steepest descent algorithm is even though the gradient estimate in the former algorithm is a noisy estimate it requires only weak assumptions (like $E \left[\widehat{\nabla} f(\theta_{n})\right] - \nabla f(\theta_{n}) \to 0 $ at a certain rate) for convergence to local minimum.
+In practice, the performance of the SA algorithm is quite sensitive to the sequence $\{a_{n}\}$.
+For example in the case of $ a_n = a/n$, the convergence is highly dependent on the choice of $a$.
+However, in practice, one can only obtain estimates of the function $f$ through black-box simulation and the challenge is to estimate the gradient and/or Hessian of $f$ through function samples.
+In the following sections we will present a brief survey of the existing methods to estimate the gradient and Hessian from function samples.
 %There are several Gradient estimation methods present in the literature.
@@
 \subsection{Finite-difference stochastic approximation (FDSA)}
-When the simulation model is treated as a block box, the traditional means of forming the estimate of gradient is by using finite-difference stochastic approximation (FDSA) method.
+When the simulation model is treated as a block box, the traditional means of forming the estimate of gradient is by using the finite-difference stochastic approximation (FDSA) method.
 There are mainly two variations in the FDSA scheme.
-First one is  two-sided gradient approximation which involve function measurements $f(\theta_n + \delta_n e_i)$ and $f(\theta_n - \delta_n e_i)$, where $e_i$ denotes $i$th column of a identity matrix of size $N \times N$.
+First one is  two-sided gradient approximation which involves function measurements $f(\theta_n + \delta_n e_i)$ and $f(\theta_n - \delta_n e_i)$, where $e_i$ denotes $i$th column of the identity matrix of size $N \times N$.
 Denote these respective values by $y_{ni}^+$ and $y_{ni}^-$, i.e.,
@@
 In the above, we assume the noise vector $(\xi_{ni}^+ - \xi_{ni}^-, i=1,2,\ldots,N)\tr$ is a martingale difference sequence for every $n \ge 0$, the sequence of the perturbation constants $\{\delta_n, n\ge 0\}$ is a positive and asymptotically vanishing sequence.
-The gradient estimate of objective function using this scheme is given by
+The gradient estimate of the objective function using this scheme is given by
 \begin{align}\label{eq:twofdsa}
@@
 \end{array}
-\right)
+\right).
 \end{align}
-Second one is  one-sided gradient approximations which involve function measurements $f(\theta_n+\delta_n e_i)$ and $f(\theta_n)$.
+The second approach is of one-sided gradient approximation  involves function measurements $f(\theta_n+\delta_n e_i)$ and $f(\theta_n)$.
 Let us denote these values by  $y_{ni}^+$, $y_{n}$  respectively, i.e.,
 $y_{ni}^+ = f(\theta_n+\delta_n e_i) + \xi_{ni}^+$, $y_{n} = f(\theta_n) + \xi_n$,  where we assume the noise terms $(\xi_{ni}^+ - \xi_{n}, i=1,2,\ldots,N)\tr$ satisfy the martingale difference sequence property for $n \ge 0$.
-The gradient estimate of objective function using this scheme is given by
+The gradient estimate of the objective function using this scheme is given by
 \begin{align}\label{eq:onefdsa}
@@
 \end{array}
-\right)
+\right).
 \end{align}
-It is important to determine conditions under which $\theta_n$ converges to $\theta^*$, when one uses estimates shown in \eqref{eq:twofdsa} or \eqref{eq:onefdsa} used in \eqref{saalg}.
-The convergence theory for FDSA algorithm is similar to convergence theory of Robbins and Monro root finding SA algorithm \cite{rm}.
-However difficulties arise due to a bias in gradient approximation, i.e.
-$\E \left[\widehat{\nabla} f(\theta_{n})|\F_n \right] - \nabla f(\theta_{n}) $, where $\F_n = \sigma(x_m, m\le n)$ denotes the underlying sigma-field.
-The conditions required for showing convergence of this algorithm is as follows
+It is important to determine conditions under which $\theta_n$ converges to $\theta^*$, when one uses estimates shown in \eqref{eq:twofdsa} or \eqref{eq:onefdsa}.
+The convergence theory for FDSA algorithm is similar to the convergence theory of Robbins and Monro root finding SA algorithm \cite{rm}.
+However, difficulties arise due to a bias in gradient approximation, i.e, $\E \left[\widehat{\nabla} f(\theta_{n})|\F_n \right] - \nabla f(\theta_{n}) $, where $\F_n = \sigma(\theta_m, m\le n)$ denotes the underlying sigma-field.
+The conditions required for showing convergence of this algorithm are as follows:
 \begin{enumerate}[label=(\textbf{A\arabic*})]
-\item $f:\R^N\rightarrow \R$ is three-times continuously differentiable\footnote{Here $\nabla^3 f(\theta) = \dfrac{\partial^3 f (\theta)}{\partial \theta\tr \partial \theta\tr \partial \theta\tr}$ denotes the third derivate of $f$ at $\theta$ and $\nabla^3_{i_1 i_2 i_3} f(\theta)$ denotes the $(i_1 i_2 i_3)$th entry of $\nabla^3 f(\theta)$, for $i_1, i_2, i_3=1,\ldots, N$.}  with $\left|\nabla^3_{i_1 i_2 i_3} f(\theta) \right| < \alpha_0 < \infty$, for $i_1, i_2, i_3=1,\ldots, N$ and for all $\theta\in \R^N$.
+\item $f:\R^N\rightarrow \R$ is three-times continuously differentiable\footnote{Here $\nabla^3 f(\theta) = \dfrac{\partial^3 f (\theta)}{\partial \theta\tr \partial \theta\tr \partial \theta\tr}$ denotes the third derivate of $f$ at $\theta$ and $\nabla^3_{i_1 i_2 i_3} f(\theta)$ denotes the $(i_1 ,i_2, i_3)$th entry of $\nabla^3 f(\theta)$, for $i_1, i_2, i_3=1,\ldots, N$.}  with $\left|\nabla^3_{i_1 i_2 i_3} f(\theta) \right| < \alpha_0 < \infty$, for $i_1, i_2, i_3=1,\ldots, N$ and for all $\theta\in \R^N$.
 \item $\{\xi_n^+,\xi_n^-, n=1,2,\ldots\}$ satisfy $\E\left[\left.\xi_n^+ - \xi_n^- \right| \F_n\right] = 0$.
@@
 \end{enumerate}
-Note that two-sided gradient estimation scheme requires $2N$ simulations of objective function in order to obtain single gradient estimate, where as one-sided gradient estimation scheme requires only $N+1$ simulations, where $N$ is the dimension of the vector $\theta$.
-Dependence on $N$ for number of simulations to get one gradient estimate makes this procedure highly computationally expensive when $N$ is large.
-Hence one requires gradient estimate procedure that is independent of dimension of $\theta$.
+Note that the  two-sided gradient estimation scheme requires $2N$ simulations of the objective function in order to obtain a single gradient estimate, whereas the one-sided gradient estimation scheme requires only $N+1$ simulations, where $N$ is the dimension of the vector $\theta$.
+A requirement of number of simulations to be proportional to $N$ in order to get one gradient estimate makes this procedure highly computationally expensive when $N$ is large.
+Hence one requires a gradient estimation procedure that is independent of the dimension of $\theta$.
 \subsection{Simultaneous perturbation stochastic approximation (SPSA)}
@@
 The first-order SPSA algorithm, henceforth referred to as 1SPSA, was proposed in \cite{spall2005introduction}.
-1SPSA scheme for approximating gradient requires only two simulations of function measurements irrespective of parameter dimension.
-The idea in this scheme is to perturb all components of parameter randomly for getting two function measurements.
-The function measurements required for this scheme corresponds to $\theta_n + \delta_n \Delta_n$ and $\theta_n - \delta_n \Delta_n$, respectively, where $\Delta_n = (\Delta_{n1},\ldots,\Delta_{nN})\tr$ is any vector consists of i.i.d, mean-zero, symmetric random variables whose inverse moments are bounded.
-The simplest and most commonly used perturbation distribution of random variables being used are symmetric Bernoulli distribution with $\Delta_{ni} = \pm 1 \,\,w.p.\,\,1/2$, $i=1,\ldots,N,\,\, n\ge 0$.
-The gradient estimate of objective function using this scheme is as follows
+The 1SPSA scheme for approximating gradient requires only two simulations or function measurements irrespective of the parameter dimension.
+The idea in this scheme is to use two function measurements by perturbing all components of the parameter randomly.
+The function measurements required for this scheme correspond to $\theta_n + \delta_n \Delta_n$ and $\theta_n - \delta_n \Delta_n$, respectively, where $\Delta_n = (\Delta_{n1},\ldots,\Delta_{nN})\tr$ is any vector consisting of i.i.d, mean-zero, symmetric random variables whose inverse moments are bounded.
+The simplest and most commonly used perturbation distribution of random variables being used are the symmetric Bernoulli random variables $\Delta_{ni} = \pm 1 \,\,w.p.\,\,1/2$, $i=1,\ldots,N,\,\, n\ge 0$.
+The gradient estimate of objective function using this scheme is as follows:
 \begin{align}\label{eq:spsa}
@@
 \end{array}
-\right)
+\right).
 \end{align}
-The reason why 1SPSA algorithm is valid gradient estimation scheme can be seen easily from Taylor expansions of $f(\theta_n+\delta_n \Delta_n$ and $f(\theta_n-\delta_n \Delta_n)$.
+The reason why 1SPSA algorithm is a valid gradient estimation scheme can be seen easily from Taylor expansions of $f(\theta_n+\delta_n \Delta_n$ and $f(\theta_n-\delta_n \Delta_n)$ as under:
 \begin{align*}
-&f(\theta_n + \delta_n \Delta_n) = f(\theta_n) + \delta_n \Delta_n\tr \nabla f(\theta_n) + \frac{\delta_n^2}{2} \Delta_n\tr \nabla^2 f(\theta_n) \Delta_n + o(\delta_n^2)\\
-&f(\theta_n - \delta_n \Delta_n) = f(\theta_n) - \delta_n \Delta_n\tr \nabla f(\theta_n) + \frac{\delta_n^2}{2} \Delta_n\tr \nabla^2 f(\theta_n) \Delta_n + o(\delta_n^2)
+&f(\theta_n + \delta_n \Delta_n) = f(\theta_n) + \delta_n \Delta_n\tr \nabla f(\theta_n) + \frac{\delta_n^2}{2} \Delta_n\tr \nabla^2 f(\theta_n) \Delta_n + o(\delta_n^2),\\
+&f(\theta_n - \delta_n \Delta_n) = f(\theta_n) - \delta_n \Delta_n\tr \nabla f(\theta_n) + \frac{\delta_n^2}{2} \Delta_n\tr \nabla^2 f(\theta_n) \Delta_n + o(\delta_n^2).
 \end{align*}
-From above two equations, the $i$th component of gradient approximation \eqref{eq:spsa} is given by
+From the above two equations, the $i$th component of the gradient approximation \eqref{eq:spsa} is given by
 \begin{align}\label{eq:spsaexp}
-\frac{f(\theta_n + \delta_n \Delta_n) - \theta_n - \delta_n \Delta_n)}{2\delta_n \Delta_{ni}} = \nabla_i f(\theta_n) + \sum\limits_{j=1, j \neq i}^N \frac{\Delta_{nj} }{\Delta_{ni}}  \nabla^2_{j} f(\theta_n)  + o(\delta_n)
+\frac{f(\theta_n + \delta_n \Delta_n) - f(\theta_n - \delta_n \Delta_n)}{2\delta_n \Delta_{ni}} = \nabla_i f(\theta_n) + \sum\limits_{j=1, j \neq i}^N \frac{\Delta_{nj} }{\Delta_{ni}}  \nabla^2_{j} f(\theta_n)  + o(\delta_n).
 \end{align}
@@
 \begin{align}\label{eq:spsacondmean}
-\E\left[\frac{f(\theta_n + \delta_n \Delta_n) - \theta_n - \delta_n \Delta_n)}{2\delta_n \Delta_{ni}} | \F_n \right] = \nabla_i f(\theta_n) + o(\delta_n)
+\E\left[\frac{f(\theta_n + \delta_n \Delta_n) - f(\theta_n - \delta_n \Delta_n)}{2\delta_n \Delta_{ni}} | \F_n \right] = \nabla_i f(\theta_n) + o(\delta_n).
 \end{align}
 
-The asymptotic convergence of 1SPSA algorithm requires one extra assumption other than the conditions specified in FDSA scheme above.
-The extra condition for convergence of $\theta_n$ to $\theta^*$ as $n \to \infty$ is as follows
+The asymptotic convergence of 1SPSA algorithm requires one extra assumption other than the conditions specified in the FDSA scheme above.
+The extra condition for convergence of $\theta_n$ to $\theta^*$  is as follows:
 \begin{enumerate}[label=(\textbf{A\arabic*}),resume]
@@
 \end{enumerate}
-Since 1SPSA randomly perturbing the parameter vector $\theta$, number of function measurements required is two, irrespective of the dimension of $\theta$, see that the numerator of \eqref{eq:spsa} is same in all $N$ components of $\widehat{\nabla} f(\theta_{n})$.
-However, Inspite of the fact that the 1SPSA requires only two function measurements compared to  $2N$  measurements required by FDSA, the asymptotic convergence rate is the same as the FDSA scheme.
+Since 1SPSA randomly perturbs the parameter vector $\theta$, the number of function measurements required is two, irrespective of the dimension of $\theta$, see that the numerator of \eqref{eq:spsa} is the same in all $N$ components of $\widehat{\nabla} f(\theta_{n})$.
+However, inspite of the fact that the 1SPSA requires only two function measurements compared to  $2N$  measurements required by FDSA, the asymptotic convergence rate is the same as the FDSA scheme.
 \subsection{Random directions stochastic approximation (RDSA)}
-A closely related algorithm to SPSA is random directions stochastic approximation (RDSA) \cite[pp.~58-60]{kushcla}.
+A closely related algorithm to SPSA is the random directions stochastic approximation (RDSA) \cite[pp.~58-60]{kushcla}.
 The gradient estimate in RDSA differs from that in SPSA, both in the construction as well as in the choice of random perturbations.
@@
 \label{eq:grad-unif}
-\widehat\nabla f(\theta_n) = d_n \left[ \dfrac{y_n^+ - y_n^-}{2\delta_n}\right].
+\widehat\nabla f(\theta_n) = d_n \left[ \dfrac{y_n^+ - y_n^-}{2\delta_n}\right],
 \end{align}
-where $y_{n}^+ = f(\theta_n+\delta_n d_n) + \xi_{n}^+,\quad y_{n}^- = f(\theta_n-\delta_n d_n) + \xi_{n}^-$ ,$d_n = (d_n^i,\ldots,d_n^N)\tr$, and $d_n^i$, $ i=1,\ldots,N$ are i.i.d random perturbations distributed uniformly on $N$-dimensional sphere with radius 1.
-Like 1SPSA, 1RDSA also requires only two function measurements irrespective of dimension of parameter $\theta$.
-The asymptotic convergence of 1RDSA algorithm also requires one extra assumption other than the conditions specified in FDSA scheme. 
+where $y_{n}^+ = f(\theta_n+\delta_n d_n) + \xi_{n}^+,\quad y_{n}^- = f(\theta_n-\delta_n d_n) + \xi_{n}^-$, $d_n = (d_n^i,\ldots,d_n^N)\tr$, and $d_n^i$, $ i=1,\ldots,N$ are i.i.d random perturbations distributed uniformly on the $N$-dimensional sphere with radius 1.
+Like 1SPSA, 1RDSA also requires only two function measurements irrespective of the dimension of parameter $\theta$.
+The asymptotic convergence of 1RDSA algorithm also requires one extra assumption other than the conditions specified in the FDSA scheme. 
 \begin{enumerate}[label=(\textbf{A\arabic*}),resume]
@@
 \end{enumerate}
-Recent work in \cite{prashanth2015rdsa} attempts to bridge the gap between 1RDSA and 1SPSA in terms of improving the performance by incorporating random perturbations based on an asymmetric Bernoulli
-distribution as well as those with the  i.i.d uniform distribution. However,  1SPSA was found to be still marginally better than 1RDSA.  
+Recent work in \cite{prashanth2015rdsa} attempts to bridge the gap between 1RDSA and 1SPSA in terms of improving the performance by incorporating random perturbations based on a certain asymmetric Bernoulli
+distribution as well as another with the  i.i.d uniform distribution. However,  1SPSA was found to be still marginally better than 1RDSA.  
 % On the other hand, results in \cite{prashanth2015rdsa} show that a second order RDSA approach (referred to as 2RDSA hereafter) can considerably outperform the corresponding higher order SPSA algorithm \cite{spall_adaptive} (referred to as 2SPSA hereafter).
 \section{Hessian estimation}\label{sec:hesest}
-Stochastic Newton methods can counter the ill-conditioning of the objective $f$ as they incorporate second-order information into the update iteration given by   
+Stochastic Newton methods can counter the ill-conditioning problem of the objective $f$ as they incorporate second-order information into the update iteration given by   
 \begin{align}
@@
 \end{align}
-where $a_n$ is the step-size that satisfies standard stochastic approximation conditions (see (C5) in Section \ref{sec:2rdsa-results}), $\widehat\nabla f(\theta_n)$ and $\overline H_n$ are estimates of the gradient and Hessian, respectively.
+where $a_n$ is the step-size that satisfies standard stochastic approximation conditions (see (C5) in Section \ref{sec:2rdsa-results}), and $\widehat\nabla f(\theta_n)$ and $\overline H_n$ are estimates of the gradient and Hessian, respectively.
 Thus, \eqref{eq:newton} can be considered as  the stochastic version of the well-known Newton method for optimization.
@@
 \subsection{Second-order SPSA (2SPSA)}
-The number of samples per-iteration for estimating Hessian was brought down to four, irrespective of dimension $N$, by the second-order SPSA algorithm (henceforth referred to as 2SPSA).
+The number of samples per-iteration for estimating Hessian was brought down to four, regardless of the dimension $N$, by using the second-order SPSA algorithm (henceforth referred to as 2SPSA).
 The Hessian estimator is projected to the space of positive definite and symmetric matrices at each iterate for the algorithm to progress along a descent direction.
-In this scheme two independent perturbation sequences $\Delta_n = (\Delta_{n1},\ldots,\Delta_{nN})\tr$, $\widehat \Delta_n = (\widehat \Delta_{n1},\ldots,\widehat \Delta_{nN})\tr$, each consists of random variables satisfying conditions specified in 1SPSA scheme are used.
-The 2SPSA algorithm obtains four function samples $y_n^{++}$, $y_n^+$,  $y_n^{-+}$ and $y_n^{-}$ at $\theta_n + \delta_n \Delta_n + \widehat \delta_n \widehat \Delta_n $, $\theta_n +\delta_n \Delta_n$, $\theta_n - \delta_n \Delta_n + \widehat \delta_n \widehat \Delta_n$ and $\theta_n - \delta_n \Delta_n$, where  the sequences of the perturbation constants $\{\delta_n, n\ge 0\}, \{\widehat \delta_n, n\ge 0\}$ is a positive and asymptotically vanishing sequence and
+In this scheme two independent perturbation sequences $\Delta_n = (\Delta_{n1},\ldots,\Delta_{nN})\tr$, $\widehat \Delta_n = (\widehat \Delta_{n1},\ldots,\widehat \Delta_{nN})\tr$, each consisting of random variables satisfying conditions specified in the 1SPSA scheme are used.
+The 2SPSA algorithm obtains four function samples $y_n^{++}$, $y_n^+$,  $y_n^{-+}$ and $y_n^{-}$ at $\theta_n + \delta_n \Delta_n + \widehat \delta_n \widehat \Delta_n $, $\theta_n +\delta_n \Delta_n$, $\theta_n - \delta_n \Delta_n + \widehat \delta_n \widehat \Delta_n$ and $\theta_n - \delta_n \Delta_n$, where  the sequences of the perturbation constants $\{\delta_n, n\ge 0\}, \{\widehat \delta_n, n\ge 0\}$ are individually a positive and asymptotically vanishing sequences and
 the random perturbations $\Delta_n, \widehat \Delta_n$ are such that $\{\Delta_{ni},\widehat\Delta_{ni}, i=1,\ldots,N, n=1,2,\ldots\}$ are i.i.d.
 and independent of the noise sequence.
-And $y_n^{++} = f(\theta_n + \delta_n \Delta_n + \widehat \delta_n \widehat \Delta_n ) + \xi_n^{++}$, $y_n^+ = f(\theta_n+\delta_n d_n) + \xi_n^+$, $y_n^{-+} = f(\theta_n - \delta_n \Delta_n + \widehat \delta_n \widehat \Delta_n) + \xi_n^{+-}$ and $y_n^- = f(\theta_n-\delta_n d_n) + \xi_n^-$, where the noise terms $\xi_n^{++}, \xi_n^+, \xi_n^{-+}, \xi_n^-$ satisfy $\E\left[\left.\xi_n^{++} - \xi_n^+ - \xi_n^{-+} +  \xi_n^-\right| \F_n\right] = 0$ with $\F_n = \sigma(\theta_m,m\le n)$ denoting the underlying sigma-field.
-The $i,j$th entry of the Hessian estimate $\widehat H_n$ in this case is given by
+Further, $y_n^{++} = f(\theta_n + \delta_n \Delta_n + \widehat \delta_n \widehat \Delta_n ) + \xi_n^{++}$, $y_n^+ = f(\theta_n+\delta_n \Delta_n) + \xi_n^+$, $y_n^{-+} = f(\theta_n - \delta_n \Delta_n + \widehat \delta_n \widehat \Delta_n) + \xi_n^{+-}$ and $y_n^- = f(\theta_n-\delta_n \Delta_n) + \xi_n^-$, where the noise terms $\xi_n^{++}, \xi_n^+, \xi_n^{-+}, \xi_n^-$ satisfy $\E\left[\left.\xi_n^{++} - \xi_n^+ - \xi_n^{-+} +  \xi_n^-\right| \F_n\right] = 0$ with $\F_n = \sigma(\theta_m,m\le n)$ denoting the underlying sigma-field.
+The $(i,j)$th entry of the Hessian estimate $\widehat H_n$ in this case is given by
 \begin{align}
 \label{eq:2spsahhat}
-&\left(\widehat H_n\right)_{ij} = \left[\frac{1}{\Delta_{ni}\widehat \Delta_{nj}} + \frac{1}{\Delta_{nj}\widehat \Delta_{ni}} \right] \left[\dfrac{y_n^{++} - y_n^+ - y_n^{-+} + y_n^-}{4 \delta_n \widehat \delta_n}\right],
+&\left(\widehat H_n\right)_{ij} = \left[\frac{1}{\Delta_{ni}\widehat \Delta_{nj}} + \frac{1}{\Delta_{nj}\widehat \Delta_{ni}} \right] \left[\dfrac{y_n^{++} - y_n^+ - y_n^{-+} + y_n^-}{4 \delta_n \widehat \delta_n}\right].
 \end{align}
@@
 \begin{itemize}
- \item $\widehat\nabla f(\theta_n)$ is the estimate of $\nabla f(\theta_n)$ and this corresponds to \eqref{eq:spsa} 
+ \item $\widehat\nabla f(\theta_n)$ as in \eqref{eq:spsa} is the estimate of $\nabla f(\theta_n)$ and this corresponds to \eqref{eq:spsa} 
   \item $\widehat H_n$ is an estimate of the true Hessian ${\nabla}^2 f(\cdot)$ at $\theta_n$. 
@@
 \section{Contributions of this thesis}
-The following are the brief description of the contributions of this thesis:
+The following is a brief description of the contributions of this thesis:
 \begin{enumerate}
@@
 In \cite{prashanth2015rdsa}, 2RDSA scheme involving only two perturbation distributions was proposed, those are uniform and asymmetric Bernoulli distributions.
-In chapter \ref{sec:2rdsa-gen} , we proposed gradient and Hessian estimation schemes which are independent of the perturbation distributions.
+In chapter \ref{sec:2rdsa-gen}, we have proposed gradient and Hessian estimation schemes which are independent of the perturbation distributions.
 However, perturbations have to satisfy the i.i.d, mean-zero assumptions.
-Apart from these common assumptions  one additional assumption like difference between the second moment square  and fourth moment should be non zero for Hessian estimation is also required.
+Apart from these common assumptions  one additional assumption, which is the difference between the squared second moment  and fourth moment should be non-zero for Hessian estimation is also required.
 This generalisation of distributions of random variables for perturbations is important because it enhances the scope of improving the performance of the 1RDSA and 2RDSA schemes by incorporating several other distributions.
-Some well known distribution that can used for these schemes as a result of generalisation are Normal (0,1) and Cauchy distributions.
-\item Our work in \ref{sec:2spsa-3-ih} is centred on improving the second order SPSA scheme with three simulations ( referred to as 2SPSA-3 hereafter), which was proposed in  \cite{bhatnagar2015simultaneous}.
-This improvements are as a result of incorporating the zero mean feedback term in Hessian estimate and optimizing the step-sizes used in the Hessian estimate recursion.
+Some well known distribution that can be used for these schemes as a result of the generalization are Normal (0,1) and Cauchy distributions.
+\item Our work in chapter \ref{sec:2spsa-3-ih} is centred on improving the second order SPSA scheme with three simulations (referred to as 2SPSA-3 hereafter), which was proposed in  \cite{bhatnagar2015simultaneous}.
+These improvements arise as a result of incorporating the zero-mean feedback term in Hessian estimate and optimizing the step-sizes used in the Hessian estimate recursion.
 Though these ideas are same as the ideas presented in this section for 2RDSA scheme, the contribution is non-trivial because the derivation of feedback term and Hessian estimate is entirely different from 2SPSA, 2RDSA improvements.
@@
 The rest of the thesis is organised as follows: In Chapter~\ref{sec:2rdsa-ih}, we describe the improved Hessian estimation scheme, which is incorporated into the  2RDSA algorithm from \cite{prashanth2015rdsa}.
-We also presented the theoretical results for the 2RDSA algorithm with  improved Hessian estimation and the results from numerical experiments.
+We also present the theoretical results for the 2RDSA algorithm with  improved Hessian estimation and the results from numerical experiments.
 In Chapter~\ref{sec:2rdsa-gen}, we describe the generalised RDSA scheme for both the gradient and Hessian estimates and also present the theoretical convergence results as well.
-The work related to improving the Hessian estimation scheme 2SPSA-3 was presented in Chapter~\ref{sec:2spsa-3-ih} and finally, in  Chapter~\ref{sec:conclusions}, the concluding remarks were provided.
+The work related to improving the Hessian estimation scheme 2SPSA-3 is presented in Chapter~\ref{sec:2spsa-3-ih}.
+In Chapter~\ref{sec:expts}, we provide simulation experiments for all the methods proposed in the earlier chapters and finally, in  Chapter~\ref{sec:conclusions}, the concluding remarks and future directions are provided.

### 2017-01-13 second-order-rdsa-gen.tex
sir corrections first intro only

@@
 
-The 2RDSA-IH algorithm obtains three function samples $y_n$, $y_n^+$ and $y_n^-$ at $\theta_n$, $\theta_n+\delta_n d_n$ and $\theta_n - \delta_n d_n$, respectively, i.e.,
+The generalised 2RDSA algorithm obtains three function samples $y_n$, $y_n^+$ and $y_n^-$ at $\theta_n$, $\theta_n+\delta_n d_n$ and $\theta_n - \delta_n d_n$, respectively, i.e.,
 $y_n = f(\theta_n) + \xi_n$, $y_n^+ = f(\theta_n+\delta_n d_n) + \xi_n^+$ and 
@@
 \section{Generalised RDSA gradient estimate}
-The RDSA estimate of the gradient $\nabla f(\theta_n)$ is given by
+The generalised RDSA estimate of the gradient $\nabla f(\theta_n)$ is given by
 \begin{align}

### 2017-01-17 second-order-rdsa-gen.tex
sir corrections 2,3chap only

@@
 Our work in this chapter  is centred on generalising the 2RDSA scheme of \cite{prashanth2015rdsa}.
-In \cite{prashanth2015rdsa}, 2RDSA scheme involving only two perturbation distributions was proposed, those are uniform and asymmetric Bernoulli distributions.
+In \cite{prashanth2015rdsa}, the RDSA scheme involving only two perturbation distributions was proposed, those are uniform and asymmetric Bernoulli distributions.
 We propose gradient and Hessian estimation schemes which are independent of the perturbation distributions.
 However, perturbations have to satisfy the i.i.d, mean-zero assumptions.
-Apart from these common assumptions  one additional assumption like difference between the second moment square  and fourth moment should be non zero for Hessian estimation is also required.
+Apart from these common assumptions,  an additional assumption required is that the difference between the squared second moment  and fourth moment should be non zero for the Hessian estimation.
 This generalisation of distributions of random variables for perturbations is important because it enhances the scope of improving the performance of the 1RDSA and 2RDSA schemes by incorporating several other distributions.
-Some well known distribution that can used for these schemes as a result of generalisation are Normal (0,1) and Cauchy distributions.
-We will prove the  gradient and Hessian estimates as a result of generalisation are indeed asymptotically unbiased estimates.
- The improved Hessian estimation scheme proposed in chapter \ref{sec:2rdsa-ih} for 2RDSA algorithm with uniform and asymmetric Bernoulli distributions can be easily extended to the generalised RDSA scheme.
+Some well known distribution that can be used for these schemes as a result of the generalisation are Normal (0,1) and Cauchy distributions.
+We will prove that  the  gradient and Hessian estimates as a result of the generalisation are indeed asymptotically unbiased.
+ The improved Hessian estimation scheme proposed in chapter \ref{sec:2rdsa-ih} for the  2RDSA algorithm with uniform and asymmetric Bernoulli distributions can be easily extended to the generalised RDSA scheme.
 
@@
 Let $\delta_n, n\geq 0$ denote a sequence of diminishing positive real numbers and $d_n = (d_n^1,\ldots,d_n^N)\tr$ denote a random perturbation vector at instant $n$,
-where the perturbations $\{d_n^i, i=1,\ldots,N, n=1,2,\ldots\}$ are any i.i.d., mean zero random varibles. 
+where the perturbations $\{d_n^i, i=1,\ldots,N, n=1,2,\ldots\}$ are any i.i.d., mean zero random variables. 
 %\begin{equation}
@@
 \item For some $\alpha_1, \alpha_2,\zeta >0$ and for all $n$, 
-$\E \left|\xi_n^{\pm}\right|^{2+\zeta} \le \alpha_1$, $\E \left|f(x_n\pm \delta_n d_n)\right|^{2+\zeta} \le \alpha_2$. 
+$\E \left|\xi_n^{\pm}\right|^{2+\zeta} \le \alpha_1$, $\E \left|f(\theta_n\pm \delta_n d_n)\right|^{2+\zeta} \le \alpha_2$. 
 \item $\{d_n^i, i=1,\ldots,N, n=1,2,\ldots\}$ are i.i.d., zero-mean and independent of $\F_n$.
@@
 %Almost surely, as $\delta_n \rightarrow 0$, we have that
-Under (A1),(A2),(A4),(A5),(A11),(A12) for $\widehat\nabla f(x_n)$ defined according to  \eqref{eq:grad-gen} , we have a.s.
-that\footnote{Here $\widehat\nabla_i f(x_n)$ and $\nabla_i f(x_n)$ denote the $i$th coordinates in the gradient estimate $\widehat\nabla f(x_n)$ and true gradient $\nabla f(x_n)$, respectively.}
+Under (A1),(A2),(A4),(A5),(A11),(A12) for $\widehat\nabla f(\theta_n)$ defined according to  \eqref{eq:grad-gen} , we have a.s.
+that\footnote{Here $\widehat\nabla_i f(\theta_n)$ and $\nabla_i f(\theta_n)$ denote the $i$th coordinates in the gradient estimate $\widehat\nabla f(\theta_n)$ and true gradient $\nabla f(\theta_n)$, respectively.}
 \begin{align}
- \left| \E\left[\left.\widehat\nabla_i f(x_n)\right| \F_n \right] - \nabla_i f(x_n)\right| = O(\delta_n^2),  \quad \text{ for } i=1,\ldots,N.
+ \left| \E\left[\left.\widehat\nabla_i f(\theta_n)\right| \F_n \right] - \nabla_i f(\theta_n)\right| = O(\delta_n^2),  \quad \text{ for } i=1,\ldots,N.
 \end{align} 
@@
 &\E\left[\left.\dfrac{y_n^+ - y_n^-}{2\delta_n} \right| \F_n\right] \\
-%= &\E\left[d_n\left.\left(\dfrac{f(x_n+\delta_n d_n) - f(x_n-\delta_n d_n)}{2\delta_n}\right) \right| \F_n\right] + \E\left[  d_n\left.\left(\dfrac{\xi_n^+ - \xi_n^-}{2\delta_n}\right) \right| \F_n\right]\\
-= & \E\left[d_n\left.\left(\dfrac{f(x_n+\delta_n d_n) - f(x_n-\delta_n d_n)}{2\delta_n}\right) \right| \F_n\right].
+%= &\E\left[d_n\left.\left(\dfrac{f(\theta_n+\delta_n d_n) - f(\theta_n-\delta_n d_n)}{2\delta_n}\right) \right| \F_n\right] + \E\left[  d_n\left.\left(\dfrac{\xi_n^+ - \xi_n^-}{2\delta_n}\right) \right| \F_n\right]\\
+= & \E\left[d_n\left.\left(\dfrac{f(\theta_n+\delta_n d_n) - f(\theta_n-\delta_n d_n)}{2\delta_n}\right) \right| \F_n\right].
 \end{align*}
@@
 \begin{align*}
-f(x_n \pm \delta_n d_n) =& f(x_n) \pm \delta_n d_n\tr \nabla f(x_n) + \frac{\delta_n^2}{2} d_n\tr \nabla^2 f(x_n) d_n \\
-& \pm  \frac{\delta_n^3}{6} \nabla^3 f(\tilde  x_n^+)(d_n \otimes d_n \otimes d_n),
+f(\theta_n \pm \delta_n d_n) =& f(\theta_n) \pm \delta_n d_n\tr \nabla f(\theta_n) + \frac{\delta_n^2}{2} d_n\tr \nabla^2 f(\theta_n) d_n \\
+& \pm  \frac{\delta_n^3}{6} \nabla^3 f(\tilde  \theta_n^+)(d_n \otimes d_n \otimes d_n),
 \end{align*}
-where $\otimes$ denotes the Kronecker product and $\tilde x_n^+$ (resp. $\tilde x_n^-$) are on the line segment between $x_n$ and $(x_n + \delta_n d_n)$ (resp. $(x_n - \delta_n d_n)$).
+where $\otimes$ denotes the Kronecker product and $\tilde \theta_n^+$ (resp.
+$\tilde \theta_n^-$) are on the line segment between $\theta_n$ and $(\theta_n + \delta_n d_n)$ (resp.
+$(\theta_n - \delta_n d_n)$).
 Hence,
 \begin{align}
-&\E\left[d_n\left.\left(\dfrac{f(x_n+\delta_n d_n) - f(x_n-\delta_n d_n)}{2\delta_n}\right)\right| \F_n \right] \nonumber\\
-&= \E\left[d_n d_n\tr \left.\nabla f(x_n)\right| \F_n\right]\nonumber \\
-& +   \E\left[\left.\frac{\delta_n^2}{12} d_n (\nabla^3 f(\tilde  x_n^+)+\nabla^3 f(\tilde  x_n^-))(d_n \otimes d_n \otimes d_n)\right| \F_n\right]. \label{eq:l1}
+&\E\left[d_n\left.\left(\dfrac{f(\theta_n+\delta_n d_n) - f(\theta_n-\delta_n d_n)}{2\delta_n}\right)\right| \F_n \right] \nonumber\\
+&= \E\left[d_n d_n\tr \left.\nabla f(\theta_n)\right| \F_n\right]\nonumber \\
+& +   \E\left[\left.\frac{\delta_n^2}{12} d_n (\nabla^3 f(\tilde  \theta_n^+)+\nabla^3 f(\tilde  \theta_n^-))(d_n \otimes d_n \otimes d_n)\right| \F_n\right]. \label{eq:l1}
 \end{align}
@@
 \begin{align}
-\E\left[d_n d_n\tr \left.\nabla f(x_n)\right| \F_n\right] = &\E\left[d_n d_n\tr \right] \nabla f(x_n) \nonumber\\
+\E\left[d_n d_n\tr \left.\nabla f(\theta_n)\right| \F_n\right] = &\E\left[d_n d_n\tr \right] \nabla f(\theta_n) \nonumber\\
 % = & \E\left[
@@
 %\right]\nabla f(x_n) 
-= &  \lambda \nabla f(x_n). \label{eq:l1-a}
+= &  \lambda \nabla f(\theta_n). \label{eq:l1-a}
 \end{align}
@@
 \begin{align}
-&\E\left[\left.\frac{\delta_n^2}{12} d_n^l (\nabla^3 f(\tilde  x_n^+)+\nabla^3 f(\tilde  x_n^-))(d_n \otimes d_n \otimes d_n)\right| \F_n\right]\nonumber\\
+&\E\left[\left.\frac{\delta_n^2}{12} d_n^l (\nabla^3 f(\tilde  \theta_n^+)+\nabla^3 f(\tilde  \theta_n^-))(d_n \otimes d_n \otimes d_n)\right| \F_n\right]\nonumber\\
 \le & \dfrac{\alpha_0 \delta_n^2}{6} \sum_{i_1=1}^N\sum_{i_2=1}^N\sum_{i_3=1}^N \E\left( d_n^l d_n^{i_1} d_n^{i_2} d_n^{i_3}\right)\nonumber\\
@@
 
-\begin{enumerate}[label=(\textbf{C\arabic*}),resume]
+\begin{enumerate}[label=(\textbf{A\arabic*}),resume]
 \item $\{d_n^i, i=1,\ldots,N, n=1,2,\ldots\}$ are i.i.d., zero-mean and independent of $\F_n$.
@@
 \begin{lemma}(\textbf{Bias in Hessian estimate})
-\label{lemma:2rdsa-bias}
+\label{lemma:2rdsa-gen-bias}
 Under (C1)-(C3),(C5)-(C10),(C13) with $\widehat H_n$ defined according to \eqref{eq:2rdsa-estimate-gen} , we have a.s.
@@
 We use the proof techniques of \cite{prashanth2015rdsa} (see Lemma 4) in order to prove the main claim here.
-However, we provide proof for general RDSA scheme in which perturbations need not be asymmetric Bernoulli or uniform as in the case of \cite{prashanth2015rdsa}.
+However, we provide the proof for a general RDSA scheme in which perturbations need not be asymmetric Bernoulli or uniform as in the case of \cite{prashanth2015rdsa}.
 
@@
 \end{align*}
-The fourth-order term in each of the expansions above can be shown to be of order $O(\delta_n^4)$ using (C1) and arguments similar to that in Lemma 3.1. Hence,
+The fourth-order term in each of the expansions above can be shown to be of order $O(\delta_n^4)$ using (C1) and arguments similar to those in Lemma 3.1. Hence,
 \begin{align*}
@@
 \dfrac{1}{\kappa} \E\left(\left. (d_n^l)^2 \sum\limits_{i=1}^N (d_n^i)^2 \nabla^2_{ii} f(\theta_n) \right| \F_n\right)
-= & \dfrac{1}{\kappa} \E\left((d_n^l)^4 \nabla^2_{ll} f(\theta_n) + \sum\limits_{i=1,i\ne l}^N (d_n^l)^2(d_n^i)^2 \nabla^2_{ii} f(\theta_n) \right)\nonumber\\
+= & \dfrac{1}{\kappa} \E\left((d_n^l)^4 \nabla^2_{ll} f(\theta_n)| \F_n\right) + \dfrac{1}{\kappa} \E\left(\left.\sum\limits_{i=1,i\ne l}^N (d_n^l)^2(d_n^i)^2 \nabla^2_{ii} f(\theta_n)\right|\F_n\right)\nonumber\\
+= & \dfrac{1}{\kappa} \E\left((d_n^l)^4| \F_n\right) \nabla^2_{ll} f(\theta_n) + \dfrac{1}{\kappa} \sum\limits_{i=1,i\ne l}^N \E\left(\left.(d_n^l)^2(d_n^i)^2\right| \F_n\right) \nabla^2_{ii} f(\theta_n) \nonumber\\
+= & \dfrac{1}{\kappa} \E\left((d_n^l)^4\right) \nabla^2_{ll} f(\theta_n) + \dfrac{1}{\kappa} \sum\limits_{i=1,i\ne l}^N \E\left((d_n^l)^2(d_n^i)^2\right) \nabla^2_{ii} f(\theta_n) \nonumber\\
 = & \dfrac{1}{\kappa} \left( \tau \nabla^2_{ll} f(\theta_n) + \lambda^2 \sum\limits_{i=1,i\ne l}^N  \nabla^2_{ii} f(\theta_n) \right), \text{ a.s.}
 \end{align} 
-For the second equality above, we have used the fact that $\E[(d_n^l)^4] =  \tau$ and $\E[(d_n^l)^2 (d_n^i)^2] = \E[(d_n^l)^2] \E[(d_n^i)^2] = \lambda^2$, $\forall l \ne i$.
+For the second equality above, we have used the fact that $\F_n = \sigma(\theta_m,m\le n)$.
+For the third equality above, we have used the fact that $d_n$ is independent of $\F_n$, for all $n$.
+And for the last equality above, we have used the fact that $\E[(d_n^l)^4] =  \tau$ and $\E[(d_n^l)^2 (d_n^i)^2] = \E[(d_n^l)^2] \E[(d_n^i)^2] = \lambda^2$, $\forall l \ne i$.
 
@@
 \begin{align*}
-\frac{1}{\kappa} (\tau - \lambda^2) \nabla^2_{ll} f(\theta_n)
+\frac{1}{\kappa} (\tau - \lambda^2) \nabla^2_{ll} f(\theta_n).
 \end{align*} 

### 2017-01-17 secorder-rdsa-ih.tex
sir corrections 2,3chap only

@@
 $\overline H_n$ is an estimate of the true Hessian ${\nabla}^2 f(\cdot)$, $\Upsilon(\cdot)$ projects any matrix onto the set of positive definite matrices and $\{a_n, n\ge 0\}$ is a step-size sequence that satisfies standard stochastic approximation conditions.
-There are standard procedures such as Cholesky factorization, see \cite{bert22}, for projecting a given square matrix to set of positive definite matrices.
+There are standard procedures such as Cholesky factorization, see \cite{bert22}, for projecting a given square matrix to the set of positive definite matrices.
 Moreover, in the vicinity of a local minimum, one expects the Hessian to be positive definite.
@@
 
-On the other hand, $\widehat H_n$ is identical to that in 2RDSA, i.e., it estimates the true Hessian in each iteration using $3$ function evaluations. %Further, $\widehat H_0 = I$. 
-For the sake of completeness, we provide below the construction for $\widehat\nabla f(\theta_n)$ and $\widehat H_n$ using asymmetric Bernoulli perturbations, before we present the feedback term that reduces the error in $\widehat H_n$.
+On the other hand, $\widehat H_n$ is identical to that in 2RDSA, i.e., it estimates the true Hessian in each iteration using three function evaluations. %Further, $\widehat H_0 = I$. 
+For the sake of completeness, we first provide below the construction for $\widehat\nabla f(\theta_n)$ and $\widehat H_n$ using asymmetric Bernoulli perturbations, and subsequently we present the feedback term that reduces the error in $\widehat H_n$.
 
@@
  \item Generate $\{d_n^{i}, i=1,\ldots,N\}$, independent of $\{d_m, m=0,1,\ldots,n-1\}$.  
- \item For any $i=1,\ldots,N$, $d_n^{i}$ is distributed either as an asymmetric Bernoulli (see \eqref{eq:det-proj}) or Uniform $U[-\eta,\eta]$ for some $\eta >0$ (see Remark \ref{remark:unif}).
+ \item For any $i=1,\ldots,N$, $d_n^{i}$ is distributed according to either  asymmetric Bernoulli (see \eqref{eq:det-proj}) or Uniform $U[-\eta,\eta]$ distributions for some $\eta >0$ (see Remark \ref{remark:unif}).
 \begin{itemize}
@@
 \theta_{n+1} = \,& \theta_n - a_n \Upsilon(\overline H_n)^{-1}\widehat\nabla f(\theta_n),\\
-\overline H_n =\,& (1-b_{n})  \overline H_{n-1} + b_{n} ( \widehat H_n - \widehat \Psi_n).
+\overline H_n =\,& (1-b_{n})  \overline H_{n-1} + b_{n} ( \widehat H_n - \widehat \Psi_n),
 \end{align*}
@@
 
-Algorithm \ref{alg:structure} presents the pseudocode and we describe the individual component of 2RDSA-IH below.
+Algorithm \ref{alg:structure} presents the pseudocode and we describe the individual components of 2RDSA-IH below.
 \section{Function evaluations}
@@
 M_n \times\right.\right.\left.\left.\left(\sum\limits_{i=1}^{N-1} (d_n^i)^2 \nabla^2_{ii} f(\theta_n) + 2\sum\limits_{i=1}^N\sum\limits_{j=i+1}^N d_n^i d_n^j \nabla^2_{ij} f(\theta_n) \right)\right| \F_n\right].
-\label{eq:h1}
+\label{eq:h1-ch2}
 \end{align}
-In analyzing the $l$th diagonal term of the above expression, the following zero-mean term appears (see the proof of Lemma 4 in \cite{prashanth2015rdsa}):
+In analyzing the $l$th diagonal term in the above expression, the following zero-mean term appears (see the proof of Lemma 4 in \cite{prashanth2015rdsa}):
 \begin{align}
@@
 
-In analyzing the off-diagonal term ($(k,l)$ where $k < l$) of \eqref{eq:h1}, the following zero-mean term appears:
+In analyzing the off-diagonal term ($(k,l)$ where $k < l$) of \eqref{eq:h1-ch2}, the following zero-mean term appears:
 \begin{align}
@@
 We make the same assumptions as those used in the analysis of \cite{prashanth2015rdsa}, with a few minor alterations. The assumptions are listed below:
-\begin{enumerate}[label=(\textbf{C\arabic*})]
+\begin{enumerate}[label=(\textbf{A\arabic*}),resume]
 \item  The function
-$f$ is four-times differentiable\footnote{Here $\nabla^4 f(\theta) = \dfrac{\partial^4 f (\theta)}{\partial \theta\tr \partial \theta\tr \partial \theta\tr \partial \theta\tr}$ denotes the fourth derivate of $f$ at $\theta$ and $\nabla^4_{i_1 i_2 i_3 i_4} f(\theta)$ denotes the $(i_1 i_2 i_3 i_4)$th entry of $\nabla^4 f(\theta)$, for $i_1, i_2, i_3,i_4=1,\ldots, N$.} with $\left|\nabla^4_{i_1 i_2 i_3 i_4} f(\theta) \right| < \infty$, for $i_1, i_2, i_3,i_4=1,\ldots, N$ and for all $\theta\in \R^N$.
+$f$ is four-times differentiable\footnote{Here $\nabla^4 f(\theta) = \dfrac{\partial^4 f (\theta)}{\partial \theta\tr \partial \theta\tr \partial \theta\tr \partial \theta\tr}$ denotes the fourth derivate of $f$ at $\theta$ and $\nabla^4_{i_1 i_2 i_3 i_4} f(\theta)$ denotes the $(i_1, i_2, i_3, i_4)$th entry of $\nabla^4 f(\theta)$, for $i_1, i_2, i_3,i_4=1,\ldots, N$.} with $\left|\nabla^4_{i_1 i_2 i_3 i_4} f(\theta) \right| < \infty$, for $i_1, i_2, i_3,i_4=1,\ldots, N$ and for all $\theta\in \R^N$.
 
@@
 \item For each $i=1,\ldots,N$ and any $\rho>0$, 
-$P(\{ \bar f_{ni} (\theta_n) \ge 0 \text{ i.o}\} \cap \{ \bar f_{ni} (\theta_n) < 0 \text{ i.o}\} \mid \{ |\theta_{ni} - x^*_i| \ge \rho\quad \forall n\}) =0.$
+$P(\{ \bar f_{ni} (\theta_n) \ge 0 \text{ i.o}\} \cap \{ \bar f_{ni} (\theta_n) < 0 \text{ i.o}\} \mid \{ |\theta_{ni} - \theta^*_i| \ge \rho\quad \forall n\}) =0.$
 
@@
 \item For any $\tau >0$ and nonempty $S \subseteq \{1,\ldots,N\}$, there exists a $\rho'(\tau,S)>\tau$ such that 
-$$ \limsup_{n\rightarrow \infty} \left| \dfrac{\sum_{i \notin S} (x-x^*)_i \bar f_{ni}(x)}{\sum_{i \in S} (x-x^*)_i \bar f_{ni}(x)}               \right| < 1 \text{ a.s.}$$
-for all $|(x-x^*)_i| < \tau$ when $i \notin S$ and $|(x-x^*)_i| \ge \rho'(\tau,S)$ when $i\in S$.
+$$ \limsup_{n\rightarrow \infty} \left| \dfrac{\sum_{i \notin S} (\theta-\theta^*)_i \bar f_{ni}(\theta)}{\sum_{i \in S} (\theta-\theta^*)_i \bar f_{ni}(\theta)}               \right| < 1 \text{ a.s.}$$
+for all $|(\theta-\theta^*)_i| < \tau$ when $i \notin S$ and $|(\theta-\theta^*)_i| \ge \rho'(\tau,S)$ when $i\in S$.
 \item For some $\alpha_0, \alpha_1>0$ and for all $n$, $\E {\xi_n}^{2} \le \alpha_0$, $\E {\xi_n^{\pm}}^{2} \le \alpha_0$, $\E f(\theta_n)^{2} \le \alpha_1$,  $\E f(\theta_n\pm \delta_n d_n)^{2} \le \alpha_1$ and $\E \left(\left\| \Upsilon(\overline H_n) \right\|^2 \mid \F_n\right) \le \alpha_1$.
@@
 The reader is referred to Section II-B of \cite{prashanth2015rdsa} for a detailed discussion of the above assumptions.
-We remark here that (C1)-(C8) are identical to that in \cite{prashanth2015rdsa}, while (C9) and (C10) introduce minor additional requirements on $\left\| \Upsilon(\overline H_n \right\|^2$ and $\delta_n$, respectively and these are inspired by \cite{spall-jacobian}.
+We remark here that (C1)-(C8) are identical to assumptions in \cite{prashanth2015rdsa}, while (C9) and (C10) introduce minor additional requirements on $\left\| \Upsilon(\overline H_n \right\|^2$ and $\delta_n$, respectively and these are inspired from \cite{spall-jacobian}.
 
@@
 Under (C1)-(C10), with $\widehat H_n$ defined according to \eqref{eq:2rdsa-estimate-ber}, we have a.s.
-that\footnote{Here $\widehat H_n(i,j)$ and $\nabla^2_{ij}f(\cdot)$ denote the $(i,j)$th entry in the Hessian estimate $\widehat H_n$ and the true Hessian $\nabla^2 f(\cdot)$, respectively.}, for $i,j = 1,\ldots,N$,
+that\footnote{Here $\widehat H_n(i,j)$ and $\nabla^2_{ij}f(\cdot)$ denote the $(i,j)$th entries in the Hessian estimate $\widehat H_n$ and the true Hessian $\nabla^2 f(\cdot)$, respectively.}, for $i,j = 1,\ldots,N$,
 \begin{align}
@@
 $$\theta_n \rightarrow \theta^*, \overline H_n \rightarrow \nabla^2 f(\theta^*) \text{ a.s. as } n\rightarrow \infty.$$ 
-In the above, $\theta_n$ and $\overline H_n$ are updated according to \eqref{eq:e2rdsa} and \eqref{eq:2rdsa-H}, respectively, $\widehat H_n$ defined according to \eqref{eq:2rdsa-estimate-ber} and the step-sizes $b_n$ are chosen as suggested in \eqref{eq:wieghts}.
+In the above, $\theta_n$ and $\overline H_n$ are updated according to \eqref{eq:e2rdsa} and \eqref{eq:2rdsa-H}, respectively, $\widehat H_n$ is defined according to \eqref{eq:2rdsa-estimate-ber} and the step-sizes $b_n$ are chosen as suggested in \eqref{eq:wieghts}.
 \end{theorem}
@@
 
-By a completely parallel argument to that used in the proof of Theorem 1 in \cite{spall-jacobian}, we obtain: For any $i,j = 1,\ldots,N$,
+By a completely similar argument to that used in the proof of Theorem 1 in \cite{spall-jacobian}, we obtain: For any $i,j = 1,\ldots,N$,
 \begin{align*}
@@
 From Lemma \ref{lemma:2rdsa-bias}, we have 
-$$ \E[ \widehat H_k \mid \F_k] = \nabla^2 f(\theta_n) + O(\delta_n^2) \text{ a.s.}$$
+$$ \E[ \widehat H_k \mid \F_k] = \nabla^2 f(\theta_k) + O(\delta_k^2) \text{ a.s.}$$
 Since the Hessian is continuous near $\theta_n$ and $\theta_n$ converges almost surely to $\theta^*$, we have
@@
 \sum_{k=0}^n \dfrac{\delta_k^4 \left(\E(\widehat H_k \mid \F_k)\right)}{\sum_{i=0}^n \delta_i^4} 
-=&\sum_{k=0}^n \dfrac{\delta_k^4 \left(\nabla^2 f(\theta_n) + O(\delta_n^2)\right)}{\sum_{i=0}^n \delta_i^4}\\
-=&\sum_{k=0}^n \dfrac{\delta_k^4 \left(\nabla^2 f(x^*) + o(1)\right)}{\sum_{i=0}^n \delta_i^4}\\
-&\rightarrow \nabla^2 f(x^*) \text{ a.s. as } n \rightarrow \infty.
+=&\sum_{k=0}^n \dfrac{\delta_k^4 \left(\nabla^2 f(\theta_k) + O(\delta_k^2)\right)}{\sum_{i=0}^n \delta_i^4}\\
+=&\sum_{k=0}^n \dfrac{\delta_k^4 \left(\nabla^2 f(\theta^*) + o(1)\right)}{\sum_{i=0}^n \delta_i^4}\\
+&\rightarrow \nabla^2 f(\theta^*) \text{ a.s. as } n \rightarrow \infty.
 \end{align*}

### 2017-01-17 intro.tex
counter conditions problem resoved

@@
 The conditions required for showing convergence of this algorithm are as follows:
-\begin{enumerate}[label=(\textbf{A\arabic*})]
+\begin{pvn}
 \item $f:\R^N\rightarrow \R$ is three-times continuously differentiable\footnote{Here $\nabla^3 f(\theta) = \dfrac{\partial^3 f (\theta)}{\partial \theta\tr \partial \theta\tr \partial \theta\tr}$ denotes the third derivate of $f$ at $\theta$ and $\nabla^3_{i_1 i_2 i_3} f(\theta)$ denotes the $(i_1 ,i_2, i_3)$th entry of $\nabla^3 f(\theta)$, for $i_1, i_2, i_3=1,\ldots, N$.}  with $\left|\nabla^3_{i_1 i_2 i_3} f(\theta) \right| < \alpha_0 < \infty$, for $i_1, i_2, i_3=1,\ldots, N$ and for all $\theta\in \R^N$.
@@
 \item $\sup_n \left\| \theta_n \right\| < \infty$ w.p. $1$.
-\end{enumerate}
+\end{pvn}
 Note that the  two-sided gradient estimation scheme requires $2N$ simulations of the objective function in order to obtain a single gradient estimate, whereas the one-sided gradient estimation scheme requires only $N+1$ simulations, where $N$ is the dimension of the vector $\theta$.
@@
 The extra condition for convergence of $\theta_n$ to $\theta^*$  is as follows:
-\begin{enumerate}[label=(\textbf{A\arabic*}),resume]
+\begin{pvn}
 \item $\{\Delta_{ni}, i=1,\ldots,N, n=1,2,\ldots\}$ are i.i.d., independent of $\F_n$ and for some $\alpha_0 > 0$ and $\forall n$, $\E (\Delta_{ni})^{-2} \le \alpha_0$, for $i = 1,2,\ldots,N$.
-\end{enumerate}
+\end{pvn}
 Since 1SPSA randomly perturbs the parameter vector $\theta$, the number of function measurements required is two, irrespective of the dimension of $\theta$, see that the numerator of \eqref{eq:spsa} is the same in all $N$ components of $\widehat{\nabla} f(\theta_{n})$.
@@
 The asymptotic convergence of 1RDSA algorithm also requires one extra assumption other than the conditions specified in the FDSA scheme. 
-\begin{enumerate}[label=(\textbf{A\arabic*}),resume]
+\begin{pvn}
 \item $\{ d_{n}^i, i=1,\ldots,N, n=1,2,\ldots\}$ are i.i.d., independent of $\F_n$, and $\forall n$, $\E (d_{n} d_{n}\tr) = I_{N \times N}$, and for some $\alpha_0 > 0$, $\E {d_{n}^i}^2 f(\theta_n \pm \delta_n d_n) \le \alpha_0$ for $i=1,2,\ldots,N$.
-\end{enumerate}
+\end{pvn}
 Recent work in \cite{prashanth2015rdsa} attempts to bridge the gap between 1RDSA and 1SPSA in terms of improving the performance by incorporating random perturbations based on a certain asymmetric Bernoulli
@@
 \end{align}
-where $a_n$ is the step-size that satisfies standard stochastic approximation conditions (see (C5) in Section \ref{sec:2rdsa-results}), and $\widehat\nabla f(\theta_n)$ and $\overline H_n$ are estimates of the gradient and Hessian, respectively.
+where $a_n$ is the step-size that satisfies standard stochastic approximation conditions (see (A12) in Section \ref{sec:2rdsa-results}), and $\widehat\nabla f(\theta_n)$ and $\overline H_n$ are estimates of the gradient and Hessian, respectively.
 Thus, \eqref{eq:newton} can be considered as  the stochastic version of the well-known Newton method for optimization.
@@
  \item $\Upsilon$ is an operator that projects a matrix onto the set of positive definite matrices.
-Update \eqref{eq:2spsa-H} does not necessarily ensure that $\overline H_n$ is invertible and without $\Upsilon$, the parameter update \eqref{eq:2spsa} may not move along a descent direction - see conditions (C7)  in Section \ref{sec:2rdsa-results}  for the precise requirements on the matrix projection operator.
+Update \eqref{eq:2spsa-H} does not necessarily ensure that $\overline H_n$ is invertible and without $\Upsilon$, the parameter update \eqref{eq:2spsa} may not move along a descent direction - see conditions (A14)  in Section \ref{sec:2rdsa-results}  for the precise requirements on the matrix projection operator.
 \end{itemize}

### 2017-01-17 second-order-rdsa-gen.tex
counter conditions problem resoved

@@
 %%%%%%%%%%%%%%%%%%%% LEMMA 
-\begin{enumerate}[label=(\textbf{A\arabic*}),resume]
+\begin{pvn}
 \item For some $\alpha_1, \alpha_2,\zeta >0$ and for all $n$, 
@@
 For some $\eta_1,\eta_2,\zeta >0$ and for all $n$, $\E(d_n^i)^{2+\zeta} \le \eta_1,\E(d_n^i)^{4+\zeta} \le \eta_2$.
-\end{enumerate}
+\end{pvn}
 
@@
 %Almost surely, as $\delta_n \rightarrow 0$, we have that
-Under (A1),(A2),(A4),(A5),(A11),(A12) for $\widehat\nabla f(\theta_n)$ defined according to  \eqref{eq:grad-gen} , we have a.s.
+Under (A1),(A2),(A4),(A5),(A18),(A19) for $\widehat\nabla f(\theta_n)$ defined according to  \eqref{eq:grad-gen} , we have a.s.
 that\footnote{Here $\widehat\nabla_i f(\theta_n)$ and $\nabla_i f(\theta_n)$ denote the $i$th coordinates in the gradient estimate $\widehat\nabla f(\theta_n)$ and true gradient $\nabla f(\theta_n)$, respectively.}
@@
 \end{align*}
-The last equality above follows from the fact that $\E\left[  d_n\left.\left(\dfrac{\xi_n^+ - \xi_n^-}{2\delta_n}\right) \right| \F_n\right]=0$ from (A2) and (A11).
+The last equality above follows from the fact that $\E\left[  d_n\left.\left(\dfrac{\xi_n^+ - \xi_n^-}{2\delta_n}\right) \right| \F_n\right]=0$ from (A2) and (A18).
 We now analyse the term on the RHS above.
@@
 \end{align}
-In the above, the first equality follows from (A12) and the last equality in \eqref{eq:l1-a} follows from $\E[(d_n^i)^2] = \lambda$ and $\E[d_n^i d_n^j] = \E[d_n^i] \E[d_n^j] = 0$ for $i\ne j$.
+In the above, the first equality follows from (A19) and the last equality in \eqref{eq:l1-a} follows from $\E[(d_n^i)^2] = \lambda$ and $\E[d_n^i d_n^j] = \E[d_n^i] \E[d_n^j] = 0$ for $i\ne j$.
 
@@
 \end{align}
-The first inequality above follows from (A1), while the second inequality follows from (A12) and considering $\eta = \max \{\eta_1^2, \eta_2\}$.
+The first inequality above follows from (A1), while the second inequality follows from (A19) and considering $\eta = \max \{\eta_1^2, \eta_2\}$.
 The claim follows by plugging \eqref{eq:l1-a} and \eqref{eq:l2} into \eqref{eq:l1} .\\
 \end{proof}
-
-\begin{enumerate}[label=(\textbf{A\arabic*}),resume]
-\item $\{d_n^i, i=1,\ldots,N, n=1,2,\ldots\}$ are i.i.d., zero-mean and independent of $\F_n$.
+%
+%\begin{pvn}
+%\item $\{d_n^i, i=1,\ldots,N, n=1,2,\ldots\}$ are i.i.d., zero-mean and independent of $\F_n$.
 For some $\eta_1,\eta_2,\zeta >0$ and for all $n$, $\E(d_n^i)^{2+\zeta} \le \eta_1,\E(d_n^i)^{4+\zeta} \le \eta_2$.
-\end{enumerate}
+%\end{pvn}
+%
 
@@
 \label{lemma:2rdsa-gen-bias}
-Under (C1)-(C3),(C5)-(C10),(C13) with $\widehat H_n$ defined according to \eqref{eq:2rdsa-estimate-gen} , we have a.s.
+Under (A8)-(A10),(A12)-(A17),(A19) with $\widehat H_n$ defined according to \eqref{eq:2rdsa-estimate-gen} , we have a.s.
 that\footnote{Here $\widehat H_n(i,j)$ and $\nabla^2_{ij}f(\cdot)$ denote the $(i,j)$th entry in the Hessian estimate $\widehat H_n$ and the true Hessian $\nabla^2 f(\cdot)$, respectively.}, for $i,j = 1,\ldots,N$,
@@
 \end{align*}
-The fourth-order term in each of the expansions above can be shown to be of order $O(\delta_n^4)$ using (C1) and arguments similar to those in Lemma 3.1. Hence,
+The fourth-order term in each of the expansions above can be shown to be of order $O(\delta_n^4)$ using (A8) and arguments similar to those in Lemma 3.1. Hence,
 \begin{align*}
@@
 \end{align*}
-Now, taking the conditional expectation of the Hessian estimate $\widehat{H_n}$ and observing that $\E[\xi_n^+ + \xi_n^- - 2\xi_n \mid \F_n, d_n] = 0$ by (C3), we obtain the following:
+Now, taking the conditional expectation of the Hessian estimate $\widehat{H_n}$ and observing that $\E[\xi_n^+ + \xi_n^- - 2\xi_n \mid \F_n, d_n] = 0$ by (A10), we obtain the following:
 \begin{align}

### 2017-01-17 second-order-spsa-3.tex
sir corrections first intro only
sir corrections 2,3chap only
counter conditions problem resoved

@@
- \chapter{Second-order SPSA with improved hessian estimation (2SPSA-3-IH)}
+ \chapter{Second-order SPSA-3 with improved hessian estimation (2SPSA-3-IH)}
 \label{sec:2spsa-3-ih}
-The second-order SPSA-3 with improved Hessian estimate performs an update iteration similar to that of 2RDSA-IH proposed in section \ref{sec:2rdsa-ih} (See \eqref{eq:e2rdsa} and \eqref{eq:2rdsa-H}).
+The second-order SPSA-3 with improved Hessian estimate performs an update iteration similar to that of 2RDSA-IH proposed in chapter \ref{sec:2rdsa-ih} (See \eqref{eq:e2rdsa} and \eqref{eq:2rdsa-H}).
 The main difference in the iterates is that 2SPSA-3-IH uses gradient and Hessian estimates according to \eqref{eq:grad-spsa} and \eqref{eq:2spsa-ber}.
-In this section we present the improvements to 2SPSA-3 Hessian recursion \eqref{eq:2spsa-3-ih} by incorporating the zero-mean feedback term and general step-size.
-Note that though these ideas of improving the Hessian recursion is similar to the ideas presented in section \ref{sec:2rdsa-ih} for 2RDSA-IH, the derivation of feedback term does not follow from it.
+In this chapter we present the improvements to 2SPSA-3 Hessian recursion \eqref{eq:2spsa-3-ih} by incorporating a zero-mean feedback term and a general step-size.
+Note that though these ideas of improving the Hessian recursion are similar to the ideas presented in chapter \ref{sec:2rdsa-ih} for 2RDSA-IH, the derivation of the feedback term does not directly follow.
 We show that the proposed improvements to Hessian estimation in 2SPSA-3 are such  that the resulting 2SPSA-3-IH algorithm is provably convergent, in particular, the Hessian estimate $\overline H_n$ of 2SPSA-3-IH converges almost surely to the true Hessian.
-Moreover, we show, for the special case of a quadratic objective in noise-free setting, that 2SPSA-3 scheme along with proposed improvement to Hessian estimation (henceforth referred to as 2SPSA-3-IH) results in a convergence rate that is on par with the corresponding rate for 2SPSA with Hessian estimation improvements (2SPSA-IH) \cite{spall-jacobian}.
+Moreover, we show, for the special case of a quadratic objective in noise-free setting, that the 2SPSA-3 scheme along with proposed improvement to Hessian estimation (henceforth referred to as 2SPSA-3-IH) results in a convergence rate that is on par with the corresponding rate for 2SPSA with Hessian estimation improvements (2SPSA-IH) \cite{spall-jacobian}.
 The advantage with 2SPSA-3-IH is that it requires only 75\% of the simulation cost per-iteration for 2SPSA-IH.
@@
 
-Algorithm \ref{alg:spsa-3-ih-structure} presents the pseudocode and we describe the individual component of 2SPSA-3-IH below.
+Algorithm \ref{alg:spsa-3-ih-structure} presents the pseudocode and we describe the individual components of 2SPSA-3-IH below.
 \section{Function evaluations}
@@
 We make the same assumptions as those used in the analysis of \cite{prashanth2015rdsa}, with a few minor alterations. The assumptions are listed below:
-\begin{enumerate}[label=(\textbf{C\arabic*}),resume]
+\begin{pvn}
 \item  The function
@@
 \item  $\delta_n = \frac{\delta_0}{(n+1)^{\varsigma}}$, where $\delta_0 > 0$ and $0 < \varsigma \le 1/8$.
-\end{enumerate}
+\end{pvn}
 The reader is referred to Section II-B of \cite{prashanth2015rdsa} for a detailed discussion of the above assumptions.
@@
 We next present a convergence rate result for the special case of a quadratic objective function under the following additional assumptions:
-\begin{enumerate}[label=(\textbf{C\arabic*}),resume]
+\begin{pvn}
 \item  $f$ is quadratic and $\nabla^2 f(x^*) > 0$. 
 \item The operator $\Upsilon$ is chosen such that $\E(\parallel \Upsilon(\overline H_n) - \overline H_n\parallel^2) = o(e^{-2wn^{1-r}/(1-r)})$ and $\parallel \Upsilon(H) - H \parallel^2 / (1+\parallel H \parallel^2)$ is uniformly bounded.
-\end{enumerate}
+\end{pvn}
 \begin{theorem}(\textbf{2SPSA-3-IH Quadratic case - Convergence rate})

### 2017-01-17 secorder-rdsa-ih.tex
counter conditions problem resoved

@@
 This is because the third term in \eqref{eq:hnhat} vanishes asymptotically, while the fourth term there dominates asymptotically.
-Moreover, the noise factors in the fourth term in \eqref{eq:hnhat} are bounded above due to (C9) and independent of $n$, leaving the $\delta_n^2$ term in the denominator there.
+Moreover, the noise factors in the fourth term in \eqref{eq:hnhat} are bounded above due to (A16) and independent of $n$, leaving the $\delta_n^2$ term in the denominator there.
 
@@
 We make the same assumptions as those used in the analysis of \cite{prashanth2015rdsa}, with a few minor alterations. The assumptions are listed below:
-\begin{enumerate}[label=(\textbf{A\arabic*}),resume]
+\begin{pvn}
 \item  The function
@@
 \item  $\delta_n = \frac{\delta_0}{(n+1)^{\varsigma}}$, where $\delta_0 > 0$ and $0 < \varsigma \le 1/8$.
-\end{enumerate}
+\end{pvn}
 The reader is referred to Section II-B of \cite{prashanth2015rdsa} for a detailed discussion of the above assumptions.
-We remark here that (C1)-(C8) are identical to assumptions in \cite{prashanth2015rdsa}, while (C9) and (C10) introduce minor additional requirements on $\left\| \Upsilon(\overline H_n \right\|^2$ and $\delta_n$, respectively and these are inspired from \cite{spall-jacobian}.
+We remark here that (A8)-(A15) are identical to assumptions in \cite{prashanth2015rdsa}, while (A16) and (A17) introduce minor additional requirements on $\left\| \Upsilon(\overline H_n \right\|^2$ and $\delta_n$, respectively and these are inspired from \cite{spall-jacobian}.
 
@@
 \label{lemma:2rdsa-bias}
-Under (C1)-(C10), with $\widehat H_n$ defined according to \eqref{eq:2rdsa-estimate-ber}, we have a.s.
+Under (A8)-(A17), with $\widehat H_n$ defined according to \eqref{eq:2rdsa-estimate-ber}, we have a.s.
 that\footnote{Here $\widehat H_n(i,j)$ and $\nabla^2_{ij}f(\cdot)$ denote the $(i,j)$th entries in the Hessian estimate $\widehat H_n$ and the true Hessian $\nabla^2 f(\cdot)$, respectively.}, for $i,j = 1,\ldots,N$,
@@
 \label{thm:2rdsa-H}
-Under (C1)-(C10), we have that 
+Under (A8)-(A17), we have that 
 $$\theta_n \rightarrow \theta^*, \overline H_n \rightarrow \nabla^2 f(\theta^*) \text{ a.s. as } n\rightarrow \infty.$$ 

### 2017-01-21 intro.tex
quad thm part3 completed

@@
 Since 1SPSA randomly perturbs the parameter vector $\theta$, the number of function measurements required is two, irrespective of the dimension of $\theta$, see that the numerator of \eqref{eq:spsa} is the same in all $N$ components of $\widehat{\nabla} f(\theta_{n})$.
-However, inspite of the fact that the 1SPSA requires only two function measurements compared to  $2N$  measurements required by FDSA, the asymptotic convergence rate is the same as the FDSA scheme.
+The efficiency of 1SPSA depends on the shape of the objective function $f$, $\{a_n\}$, $\{\delta_n\}$ and the distribution of perturbation vectors, i.e., $\Delta_n$ $\forall n$.
+In \cite{spall}, in addition to establishing formal convergence of $\theta_n \to \theta^*$, it was shown that probability distribution of appropriately scaled $\theta_n$ is approximately normal  (with some mean and covariance matrix) for larger $n$.
+The general metric which is used for evaluating these class of algorithms is asymptotic mean squared error (AMSE) defined as $\E\l \theta_{n} - \theta^* \r^2$.
+The relative efficiency of 1SPSA when compared to FDSA can be shown when we compare the similar asymptotic normality result of FDSA.
+In \cite{spall}, it was shown that under reasonably general conditions, and when $\{a_n\}$, $\{\delta_n\}$ are chosen according to the guidelines in \cite{spall1998implementation} then 1SPSA and FDSA gives same AMSE for given number of iterations.
+Instead of looking the
+\begin{align}\label{eq:amse}
+\frac{\text{AMSE of 1SPSA}}{\text{AMSE of FDSA}}, 
+\end{align}
+an equivalent way of looking the relative AMSE is to compare the number of function measurements required by FDSA to achieve same AMSE as 1SPSA which can be computed by equating the \eqref{eq:amse} to one.
+\begin{align}\label{eq:amsenom}
+ \frac{\text{No.of function measurements for 1SPSA }}{\text{No.of function measurements for FDSA}}=  \frac{1}{N}.
+\end{align}
+
+%However, inspite of the fact that the 1SPSA requires only two function measurements compared to  $2N$  measurements required by FDSA, the asymptotic convergence rate is the same as the FDSA scheme.
 \subsection{Random directions stochastic approximation (RDSA)}

### 2017-01-21 second-order-spsa-3.tex
quad thm part3 completed

@@
  \item Generate $\{\Delta_n^{i}, i=1,\ldots,N\}$,$\{\widehat\Delta_n^{i}, i=1,\ldots,N\}$ independent of $\{\Delta_m,\widehat\Delta_m, m=0,1,\ldots,n-1\}$.  
- \item For any $i=1,\ldots,N$, $\Delta_n^{i}$,$\widehat\Delta_n^{i}$  satisfy condition (C17) of section \ref{sec:2spsa-results},(most popular distribution being used is symmetric Bernoulli distribution, $\pm 1$ w.p.
+ \item For any $i=1,\ldots,N$, $\Delta_n^{i}$,$\widehat\Delta_n^{i}$  satisfy condition (C17) of chapter \ref{sec:2spsa-results},(most popular distribution being used is symmetric Bernoulli distribution, $\pm 1$ w.p.
 $1/2$).
@@
 \theta_{n+1} = \,& \theta_n - a_n \Upsilon(\overline H_n)^{-1}\widehat\nabla f(\theta_n),\\
-\overline H_n =\,& (1-b_{n})  \overline H_{n-1} + b_{n} ( \widehat H_n - \widehat \Psi_n).\label{eq:2spsa-3-ih}
+\overline H_n =\,& (1-b_{n})  \overline H_{n-1} + b_{n} ( \widehat H_n - \widehat \Psi_n),\label{eq:2spsa-3-ih}
 \end{align}
@@
 
-Further, $\delta_n, n\geq 0$ is a sequence of diminishing positive real numbers and $\Delta_n = (\Delta_n^1,\ldots,\Delta_n^N)\tr$ , $\widehat\Delta_n = (\widehat\Delta_n^1,\ldots,\widehat\Delta_n^N)\tr$ are the perturbation random vector at instant $n$ with $\Delta_n^i,i=1,\ldots,N$, and $\widehat\Delta_n^i,i=1,\ldots,N$ are independent identically distributed (i.i.d), mean-zero random varibles having finite inverse movements of order greater than 2 and satisfy the condition (C17) of section \ref{sec:2spsa-results}.
+Further, $\delta_n, n\geq 0$ is a sequence of diminishing positive real numbers and 
+$\Delta_n = (\Delta_n^1,\ldots,\Delta_n^N)\tr$ , $\widehat\Delta_n = (\widehat\Delta_n^1,\ldots,\widehat\Delta_n^N)\tr$ are the perturbation random vectors at instant $n$ with $\Delta_n^i,i=1,\ldots,N$, and $\widehat\Delta_n^i,i=1,\ldots,N$ being independent identically distributed (i.i.d), mean-zero random variables having finite inverse moments of order greater than 2 and satisfying the condition (C17) of section \ref{sec:2spsa-results}.
 
@@
 \section{Hessian estimate}
-The $i,j$th entry of the Hessian estimate in this case is given by
+The $(i,j)$th entry of the Hessian estimate in this case is given by
 \begin{align}
 \label{eq:2spsa-ber}
-&\left(\widehat H_n\right)_{ij} =  \left(\dfrac{y_n^+ + y_n^- - 2 y_n}{2\delta_n^2 \Delta_n^{(i)} \widehat \Delta_n^{(j)}}\right), 
+&\left(\widehat H_n\right)_{ij} =  \left(\dfrac{y_n^+ + y_n^- - 2 y_n}{2\delta_n^2 \Delta_n^{(i)} \widehat \Delta_n^{(j)}}\right).
 \end{align}
@@
 \section{Feedback term $\widehat \Psi_n$}
-The $i,j$th term of the Hessian estimate $\widehat H_n$ can be simplified as follows: 
+The $(i,j)$th term of the Hessian estimate $\widehat H_n$ can be simplified as follows: 
 \begin{align}
@@
 \begin{align}
-\E\left[\left.
-\frac{(\Delta_n+ \widehat \Delta_n)\tr \nabla^2 f(\theta_n) (\Delta_n+ \widehat \Delta_n)}{2\Delta_n^{(i)} \widehat \Delta_n^{(j)}}  \right | \F_n\right] &=\E\left[\sum\limits_{l=1}^N\sum\limits_{m=1}^N \frac{\Delta_n^l \nabla^2_{lm} f(\theta_n) \Delta_n^m}{2\Delta_n^i \widehat\Delta_n^j} + \sum\limits_{l=1}^N\sum\limits_{m=1}^N \frac{\Delta_n^l \nabla^2_{lm} f(\theta_n) \widehat\Delta_n^m}{\Delta_n^i \widehat\Delta_n^j}  \right.\nonumber \\&\hspace{2.5cm}+\left.\sum\limits_{l=1}^N\sum\limits_{m=1}^N \left.
-\frac{\widehat\Delta_n^l \nabla^2_{lm} f(\theta_n)\widehat \Delta_n^m}{2\Delta_n^i \widehat\Delta_n^j}\right| \F_n\right].\label{eq:2spsaexp}
+&\E\left[\left.
+\frac{(\Delta_n+ \widehat \Delta_n)\tr \nabla^2 f(\theta_n) (\Delta_n+ \widehat \Delta_n)}{2\Delta_n^{(i)} \widehat \Delta_n^{(j)}}  \right | \F_n\right] =\E\left[\sum\limits_{l=1}^N\sum\limits_{m=1}^N \frac{\Delta_n^{(l)} \nabla^2_{lm} f(\theta_n) \Delta_n^{(m)}}{2\Delta_n^{(i)} \widehat\Delta_n^{(j)}} \right.\nonumber \\&\hspace{2.5cm}+\left.
+\sum\limits_{l=1}^N\sum\limits_{m=1}^N \frac{\Delta_n^{(l)} \nabla^2_{lm} f(\theta_n) \widehat\Delta_n^{(m)}}{\Delta_n^{(i)} \widehat\Delta_n^{(j)}} +\sum\limits_{l=1}^N\sum\limits_{m=1}^N \left.
+\frac{\widehat\Delta_n^{(l)} \nabla^2_{lm} f(\theta_n)\widehat \Delta_n^{(m)}}{2\Delta_n^{(i)} \widehat\Delta_n^{(j)}}\right| \F_n\right].\label{eq:2spsaexp}
 \end{align}
-In analysing the RHS of the above expression, the first an the third terms are mean-zero (See Proposition 4.2 in \cite{bhatnagar2015simultaneous}):
+In analysing the RHS of the above expression, the first and the third terms are mean-zero (See Proposition 4.2 in \cite{bhatnagar2015simultaneous}):
 \begin{align}
-\E\left[\left.\sum\limits_{l=1}^N\sum\limits_{m=1}^N \frac{\Delta_n^l \nabla^2_{lm} f(\theta_n) \Delta_n^m}{\Delta_n^i \widehat\Delta_n^j} \right | \F_n\right] = 0 \hspace{0.2cm} and \hspace{0.2cm}\E\left[\left.\sum\limits_{l=1}^N\sum\limits_{m=1}^N \frac{\widehat\Delta_n^l \nabla^2_{lm} f(\theta_n) \widehat\Delta_n^m}{\Delta_n^i \widehat\Delta_n^j }\right | \F_n\right] = 0.
+\E\left[\left.\sum\limits_{l=1}^N\sum\limits_{m=1}^N \frac{\Delta_n^{(l)} \nabla^2_{lm} f(\theta_n) \Delta_n^{(m)}}{\Delta_n^{(i)} \widehat\Delta_n^{(j)}} \right | \F_n\right] = 0 \hspace{0.2cm} and \hspace{0.2cm}\E\left[\left.\sum\limits_{l=1}^N\sum\limits_{m=1}^N \frac{\widehat\Delta_n^{(l)} \nabla^2_{lm} f(\theta_n) \widehat\Delta_n^{(m)}}{\Delta_n^{(i)} \widehat\Delta_n^{(j)} }\right | \F_n\right] = 0.
 \end{align}
@@
 \begin{align}
-\Psi_n^1(\nabla^2 f(\theta_n)) =\frac{1}{2} M_n \left[\Delta_n\tr \nabla^2 f(\theta_n) \Delta_n + \widehat\Delta_n\tr \nabla^2 f(\theta_n) \widehat\Delta_n\right].
+\Psi_n^1(\nabla^2 f(\theta_n)) =\frac{1}{2} M_n \left[\Delta_n\tr \nabla^2 f(\theta_n) \Delta_n + \widehat\Delta_n\tr \nabla^2 f(\theta_n) \widehat\Delta_n\right],
 \end{align}
-where $M_n = [1/\Delta_n^1, \ldots ,1/\Delta_n^N]\tr [1/\Delta_n^1, \ldots ,1/\Delta_n^N]$.
-Now consider the second term in RHS of the \eqref{eq:2spsaexp}. It can be re-written as follows:
+where $M_n = [1/\Delta_n^{(1)}, \ldots ,1/\Delta_n^{(N)}]\tr [1/\Delta_n^{(1)}, \ldots ,1/\Delta_n^{(N)}]$.
+Now consider the second term in the RHS of \eqref{eq:2spsaexp}. It can be re-written as follows:
 \begin{align}
 \E\left[\left.
-\sum\limits_{l=1}^N\sum\limits_{m=1}^N \frac{\Delta_n^l \nabla^2_{lm} f(\theta_n) \widehat\Delta_n^m}{\Delta_n^i \widehat\Delta_n^j} \right | \F_n\right] = \E\left[\left.\nabla^2_{ij} f(\theta_n) + \frac{1}{\Delta_n^i \widehat\Delta_n^j} \mathop{\sum_{l=1}^N \sum_{m=1}^{N}}_{lm \neq ij} \Delta_n^l \nabla^2_{lm} f(\theta_n) \widehat\Delta_n^m \right | \F_n\right]
+\sum\limits_{l=1}^N\sum\limits_{m=1}^N \frac{\Delta_n^{(l)} \nabla^2_{lm} f(\theta_n) \widehat\Delta_n^{(m)}}{\Delta_n^{(i)} \widehat\Delta_n^{(j)}} \right | \F_n\right] = \E\left[\left.\nabla^2_{ij} f(\theta_n) + \frac{1}{\Delta_n^{(i)} \widehat\Delta_n^{(j)}} \mathop{\sum_{l=1}^N \sum_{m=1}^{N}}_{lm \neq ij} \Delta_n^{(l)} \nabla^2_{lm} f(\theta_n) \widehat\Delta_n^{(m)} \right | \F_n\right]
 \end{align}
@@
 \begin{align}
-\E\left[\left. \frac{1}{\Delta_n^i \widehat\Delta_n^j} \mathop{\sum_{l=1}^N \sum_{m=1}^{N}}_{lm \neq ij} \Delta_n^l \nabla^2_{lm} f(\theta_n) \widehat\Delta_n^m \right | \F_n\right] = 0.
+\E\left[\left.
+\frac{1}{\Delta_n^{(i)} \widehat\Delta_n^{(j)}} \mathop{\sum_{l=1}^N \sum_{m=1}^{N}}_{lm \neq ij} \Delta_n^{(l)} \nabla^2_{lm} f(\theta_n) \widehat\Delta_n^{(m)} \right | \F_n\right] = 0.
 \end{align}
-The term on the LHS above, can be denoted by $\Psi_n^2(\nabla^2 f(\theta_n))$, can be written in the matrix form as follows:
+The term on the LHS above, can be denoted by $\Psi_n^2(\nabla^2 f(\theta_n))$, and can be written in  matrix form as follows:
 \begin{align}
@@
 \end{align}
-where $N_n,\widehat N_n$ are defined as follows $N_n = \Delta_n \left[\frac{1}{\Delta_n^1},\ldots,\frac{1}{\Delta_n^N}\right] - I_N$ and $\widehat N_n = \widehat \Delta_n \left [\frac{1}{\widehat\Delta_n^1},\ldots,\frac{1}{\widehat \Delta_n^N}\right] - I_N$ and $I_N$ is  identity matrix of size $N \times N$.
+where $N_n,\widehat N_n$ are defined as follows: $N_n = \Delta_n \left[\frac{1}{\Delta_n^{(1)}},\ldots,\frac{1}{\Delta_n^{(N)}}\right] - I_N$ and $\widehat N_n = \widehat \Delta_n \left [\frac{1}{\widehat\Delta_n^{(1)}},\ldots,\frac{1}{\widehat \Delta_n^{(N)}}\right] - I_N$ and $I_N$ is  identity matrix of size $N \times N$.
 From the foregoing, the per-iteration Hessian estimate $\widehat H_n$ can be re-written as follows:
@@
 \begin{align}
-\sum \limits_{i=0}^{n} (\tilde b_k)^2 \delta_i^{-4}, \text{ subject to} \label{eq:spsawn-opt}\\
+\sum \limits_{i=0}^{n} (\tilde b_i)^2 \delta_i^{-4}, \text{ subject to} \label{eq:spsawn-opt}\\
 \tilde b_i \geq 0 ~\forall i \text{ and }\sum \limits_{i=0}^{n} \tilde b_i = 1.
@@
 \label{eq:spsahess}
-\overline H_n = \sum\limits_{i=0}^{n} \tilde b_k(\widehat H_i -\widehat \Psi_i).
+\overline H_n = \sum\limits_{i=0}^{n} \tilde b_i(\widehat H_i -\widehat \Psi_i).
 \end{align}
@@
 \item  The function
-$f$ is four-times differentiable\footnote{Here $\nabla^4 f(\theta) = \dfrac{\partial^4 f (\theta)}{\partial \theta\tr \partial \theta\tr \partial \theta\tr \partial \theta\tr}$ denotes the fourth derivate of $f$ at $\theta$ and $\nabla^4_{i_1 i_2 i_3 i_4} f(\theta)$ denotes the $(i_1 i_2 i_3 i_4)$th entry of $\nabla^4 f(\theta)$, for $i_1, i_2, i_3,i_4=1,\ldots, N$.} with $\left|\nabla^4_{i_1 i_2 i_3 i_4} f(\theta) \right| < \infty$, for $i_1, i_2, i_3,i_4=1,\ldots, N$ and for all $\theta\in \R^N$.
+$f$ is four-times differentiable\footnote{Here $\nabla^4 f(\theta) = \dfrac{\partial^4 f (\theta)}{\partial \theta\tr \partial \theta\tr \partial \theta\tr \partial \theta\tr}$ denotes the fourth derivate of $f$ at $\theta$ and $\nabla^4_{i_1, i_2, i_3, i_4} f(\theta)$ denotes the $(i_1, i_2, i_3, i_4)$th entry of $\nabla^4 f(\theta)$, for $i_1, i_2, i_3,i_4=1,\ldots, N$.} with $\left|\nabla^4_{i_1 i_2 i_3 i_4} f(\theta) \right| < \infty$, for $i_1, i_2, i_3,i_4=1,\ldots, N$ and for all $\theta\in \R^N$.
 
@@
 
-\item For each $n$ and all $x$, there exists a $\rho>0$ not dependent on $n$ and $\theta$, such that $(\theta-\theta^*)\tr \bar f_n(\theta) \ge \rho \left\| \theta_n - \theta\right\|$, where $\bar f_n(\theta) = \Upsilon(\overline H_n)^{-1} \nabla f(\theta)$.
+\item For each $n$ and all $\theta$, there exists a $\rho>0$ not dependent on $n$ and $\theta$, such that $(\theta-\theta^*)\tr \bar f_n(\theta) \ge \rho \left\| \theta_n - \theta\right\|$, where $\bar f_n(\theta) = \Upsilon(\overline H_n)^{-1} \nabla f(\theta)$.
 
 \item $\{\xi_n, \xi_n^+,\xi_n^-, n=1,2,\ldots\}$ are such that, for all $n$, $\E\left[\left.
-\xi_n^+ + \xi_n^- - 2 \xi_n \right| \F_n, \Delta_n, \widehat \Delta_n \right] = 0$, where $\mathcal{F}_n = \sigma(\theta_m,m < n)$ denotes the underlying sigma-field..
+\xi_n^+ + \xi_n^- - 2 \xi_n \right| \F_n, \Delta_n, \widehat \Delta_n \right] = 0$, where $\mathcal{F}_n = \sigma(\theta_m,m < n)$ denotes the underlying sigma-field.
 
@@
 The reader is referred to Section II-B of \cite{prashanth2015rdsa} for a detailed discussion of the above assumptions.
-We remark here that (C14)-(C21) are identical to that in \cite{prashanth2015rdsa}, while (C22) and (C23) introduce minor additional requirements on $\left\| \Upsilon(\overline H_n \right\|^2$ and $\delta_n$, respectively and these are inspired by \cite{spall-jacobian}.
+We remark here that (C14)-(C21) are identical to those in \cite{prashanth2015rdsa}, while (C22) and (C23) introduce minor additional requirements on $\left\| \Upsilon(\overline H_n) \right\|^2$ and $\delta_n$, respectively and these are inspired from \cite{spall-jacobian}.
 
@@
 Under (C14)-(C23), with $\widehat H_n$ defined according to  \eqref{eq:2spsa-ber} , we have a.s.
-that\footnote{Here $\widehat H_n(i,j)$ and $\nabla^2_{ij}f(\cdot)$ denote the $(i,j)$th entry in the Hessian estimate $\widehat H_n$ and the true Hessian $\nabla^2 f(\cdot)$, respectively.}, for $i,j = 1,\ldots,N$,
+that\footnote{Here $\widehat H_n(i,j)$ and $\nabla^2_{i,j}f(\cdot)$ denote the $(i,j)$th entry in the Hessian estimate $\widehat H_n$ and the true Hessian $\nabla^2 f(\cdot)$, respectively.}, for $i,j = 1,\ldots,N$,
 \begin{align}
@@
 \begin{proof}
-This proof is similar to the proof of the theorem \ref{thm:2rdsa-H}
+The proof is similar to the proof of the Theorem \ref{thm:2rdsa-H} and is therefore omitted.
 \end{proof}
@@
   \item[Step 1:] Here we prove the MSE convergence of $\overline H_k$, i.e., $\E [\Lambda_n\tr \Lambda_n] \rightarrow 0$ a.s. as $n\rightarrow \infty$.
-  \item[Step 2:] We unroll the recursion \eqref{eq:2spsa-3-ih} and then derive convenient representation for $\text{trace}[\E (\Lambda_n \tr \Lambda_n)]$.
+  \item[Step 2:] We unroll the recursion \eqref{eq:2spsa-3-ih} and then derive a convenient representation for the  $\text{trace}[\E (\Lambda_n \tr \Lambda_n)]$.
   \item[Step 3:] We derive the main result in \eqref{eq:spsaquad-bigo} using a proof by contradiction. 
@@
 \noindent\textbf{Step 2: Representation of $\text{trace}[\E (\Lambda_n \tr \Lambda_n)]$} \\
-This part in exactly the same manner as part (ii) in Theorem 3 of \cite{spall-jacobian}.\\
+ From \eqref{eq:2spsa-3-ih} and \eqref{eq:spsahnhat-ext},  we have
+ \begin{align}
+ \Lambda_n &= \Lambda_{n-1} - b_n ( \overline H_{n-1} - \hat H_n + \hat \Psi_n) \nonumber\\
+&= (1-b_n) \Lambda_{n-1} - b_n (H^* + \hat \Psi_n - \hat H_n ) \nonumber\\&= (1-b_n) \Lambda_{n-1} - b_n (\hat \Psi_n - \Psi_n(H^*)) \nonumber\\&= (1-b_n) \Lambda_{n-1} - b_n \Psi_n(\Upsilon(\overline H_{n-1})) + b_n \Psi_n(H^*) \nonumber\\
+&= (1-b_n) \Lambda_{n-1} - b_n\Psi_n(\Lambda_{n-1}'),\label{eq:step1last}
+ \end{align}
+where $\Lambda_n' = \Upsilon(\overline H_n) - H^*$.
+The equality in \eqref{eq:step1last} follows from \eqref{eq:psi}. 
+Unrolling the recursion \eqref{eq:step1last}, we obtain 
+ \begin{align}\label{lambda-exp}
+  \Lambda_n  = \left[ \prod_{k=1}^n (1-b_k) \right]\Lambda_0 
+&- \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right] b_k \Psi_k(\Lambda_{k-1}')  \hspace{0.2cm} a.s.
+ \end{align}
+ %%(note : $\prod_{j=n+1}^n (1-b_j) = 1$ for all n).
+ %%Let us  characterise $trace[\E(\Lambda_n\tr \Lambda_n)]$ using \eqref{lambda-exp}. From independence of $d_k$ along k, \eqref{lambda-exp} represent martingale difference sequence, leading to 
+Squaring on both sides and taking expectations, we obtain
+ \begin{align}
+  \E (\Lambda_n \tr \Lambda_n) =  &\left[ \prod_{k=1}^n (1-b_k) \right]^2 \E (\Lambda_0 \tr \Lambda_0) + \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2   \E (\Psi_k(\Lambda_{k-1}') \tr \Psi_k(\Lambda_{k-1}')).\label{eq:exp-lambda2}
+ \end{align}
+The equality above uses the fact that $\E(\Psi_k(\Lambda_{k-1}'))=0$, which get rid of the corresponding cross terms with the first term on RHS of \eqref{lambda-exp}. % %
+We now characterize $\text{trace}[\E(\Lambda_n\tr \Lambda_n)]$ using \eqref{eq:exp-lambda2} as follows:
+%From the independence of $d_k$'s and $\Lambda_{k-1}'$ and using Cauchy-Schwartz inequality we get 
+ \begin{align}\label{eq:trace}
+ \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] = &\left[ \prod_{k=1}^n (1-b_k) \right]^2 \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  + \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2  \,\,\tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right).
+ \end{align}
+ % \begin{align}\label{eq:trace}
+% & \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] \leq \left[ \prod_{k=1}^n (1-b_k) \right]^2 \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  \nonumber\\ &+ \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2  \,\,\tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ &+ \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2  \nonumber\\&  \hspace{0.5cm}\times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right] + \nonumber\\ &+ \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2   \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)^{1/2} \nonumber\\&  \times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right]^{1/2}.
+% \end{align}
+Where, as in the proof of Theorem 3 of \cite{spall-jacobian}, $\tau(\cdot)$ transforms the $\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')$ in a  fashion and then returns the trace of the resulting $N\times N$ matrix.
+As in the proof of Theorem 3 of \cite{spall-jacobian}, observe that $1-b_k = e^{-b_k}(1-O(b_k^2))$ and since $0 < b_k <1$, we have that the $O(b_k^2)$ term is strictly positive. 
+Letting $\Gamma_{ij} = \sum_{k=i}^j b_k$ with $\Gamma_{nn} = 1$ and $\beta_{kn} = \left[\prod_{i=k+1}^n (1- O(b_i^2))\right]^2$, we can simplify \eqref{eq:trace} as follows:
+\begin{align}\label{eq:wij}
+ \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] = &e^{-2 \Gamma_{1n}} \beta_{0n} \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  + e^{-2 \Gamma_{1n}} \sum_{k=1}^n e^{2 \Gamma_{1k}} \beta_{kn} b_k^2  \,\,\tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right).
+ \end{align} 
+% \begin{align}\label{eq:wij}
+% & \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] \leq e^{-2 \Gamma_{1n}} \beta_{0n} \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  \nonumber\\ &+ e^{-2 \Gamma_{1n}} \sum_{k=1}^n e^{2 \Gamma_{1k}} \beta_{kn} b_k^2  \,\,\tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ &+ e^{-2 \Gamma_{1n}} \sum_{k=1}^n e^{2 \Gamma_{1k}} \beta_{kn} b_k^2  \nonumber\\&  \hspace{0.5cm}\times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right] \nonumber\\ &+ e^{-2 \Gamma_{1n}} \sum_{k=1}^n e^{2 \Gamma_{1k}} \beta_{kn} b_k^2  \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)^{1/2} \nonumber\\&  \times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right]^{1/2}.
+% \end{align} 
+Comparing the sum with integrals, we obtain
+\begin{align*}
+\Gamma_{ij} = \int_i^j \frac{b_0}{x^r} dx + O(1) &= \left(\frac{b_0}{1-r}\right)(j^{1-r}-i^{1-r}) +O(1),
+\end{align*}
+where we have used the facts that $0 < b_k < 1, \forall k \ge 2$ and 
+$\sum_{k=i}^j b_k \to \infty$ as $j-i \to \infty$ since 
+$b_k = b_0/k^r$ with $r > 0.5$.
+Observing that $\beta_{kn}$ are uniformly upper-bounded, say by $\bar \beta_n$, we have
+\begin{align}\label{eq:thmwts}
+ & \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] =  e^{-2 \Gamma_{1n}} \beta_{0n} \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  \nonumber\\ 
+&\hspace{3cm}+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}  \times \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right).
+\end{align} 
+%\begin{align}\label{eq:thmcomb}
+%& \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] =  e^{-2 \Gamma_{1n}} \beta_{0n} \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  \nonumber\\ 
+%&+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}  \times \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ 
+%&+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}  \text{trace}\left[\E [((\Phi_k(H^*) - H^*) - 2 \Psi_k(\Lambda_{k-1}))\tr  (\Phi_k(H^*) - H^*)]\right] .
+%\end{align} 
+%
+%%\begin{align}\label{eq:thmwts}
+%% & \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] =  e^{-2 \Gamma_{1n}} \beta_{0n} \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  \nonumber\\ 
+%%&+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}} \nonumber\\
+%%& \hspace{3.5cm} \times \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ 
+%%&+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}  \nonumber\\
+%%&  \hspace{0.5cm}\times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right] \nonumber\\ &+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}  \nonumber\\&\hspace{3.5cm}\times \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)^{1/2}\nonumber\\&\times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right]^{1/2}.
+%%\end{align}
+%\noindent\textbf{Step 3: The big-O result on $\text{trace}[\E (\Lambda_n \tr \Lambda_n)]$ convergence} \\
+%In comparison to Eq.
+(7.6) in \cite{spall-jacobian} that corresponds to \eqref{eq:thmwts} for the N-SPSA-4 setting, there are two extra terms - the third and the fourth - in the RHS of \eqref{eq:thmwts}.
+%Consider the third term on the RHS of \eqref{eq:thmwts}.
+As a consequence of (C9) and the construction of random perturbations $d_k$, it can be seen that $\text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right]$ is uniformly bounded above, say by $\top$, independent of $k$.
+Thus, the third term is equivalent to the following:
+%\begin{align}
+% \bar c_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \frac{b_0^2}{k^{2 r}} \top.
+%\label{eq:term3}
+%\end{align}
+%Now by moving out side exponential term to inside and changing the summation to integration, we can rewrite the above equation as follows
+%\begin{align}
+%  \overline{\overline{c_n}}  \int_1^n  \frac{1}{k^{2 r}} .
+%\label{eq:term3}
+%\end{align}
+%Where $\overline{\overline{c_n}}$ is a constant. By above equation we can conclude that the third term is of $O(\frac{1}{k^{2 r -1}})$
+%%Now, observe that $\lim_{n\rightarrow \infty} \sum_{k=1}^n \frac{1}{k^{2 r}} < \infty$ as $1/2 < r < 1$ and $e^{2 b_0 k^{1-r}/(1-r)}$ diverges as $n\rightarrow \infty$ and hence, by Kronecker's Lemma, the term in \eqref{eq:term3} vanishes asymptotically.
+%%
+%%Now consider the fourth term.
+By using the uniform upper-bound $\top$ on $\text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right]$, we can rewrite  the fourth term in \eqref{eq:thmwts} as follows:
+%% \begin{align}\label{eq:fourthterm}
+%% \bar c_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \frac{b_0^2}{k^{2 r}} \sqrt{\top} \nonumber \\ \hspace{3cm}\times \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)^{1/2},
+%% \end{align}
+%%where, as in the proof of Theorem 3 of \cite{spall-jacobian}, $\tau(\cdot)$ transforms the $\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')$ in a  fashion and then returns the trace of the resulting $N\times N$ matrix.
+
 \noindent\textbf{Step 3: The big-O result on $\text{trace}[\E (\Lambda_n \tr \Lambda_n)]$ convergence} \\
-This part in exactly the same manner as part (iii) in Theorem 3 of \cite{spall-jacobian}.\\
+We will use similar argument in part (iii) in Theorem 3 of \cite{spall-jacobian} to show the claim \eqref{eq:spsaquad-bigo}.
+The idea here is to show that the contradiction exists if RHS of \eqref{eq:spsaquad-bigo} exhibits  any slower rate of decay.
+Let us assume the following slower rate of decay for the LHS of the \eqref{eq:spsaquad-bigo}.
+\begin{align}
+\text{trace}[\E (\Lambda_n \tr \Lambda_n)] = d(n) e^{-2b_0n^{1-r} / {1-r}}.
+\label{eq:spsaquad-bigocont}
+\end{align}
+Where $d(n) = o(e^{2b_0n^{1-r} / {1-r}})$.
+Now we need to show that there is contradiction to show the claim.
+By using the series of arguments used in part (iii) in Theorem 3 of \cite{spall-jacobian} and applying the condition (A31) we get the following
+\begin{align}
+\E (\Lambda_{n}' \otimes \Lambda_{n}') = \E (\Lambda_{n} \otimes \Lambda_{n}) + o(d(n) e^{-2b_0n^{1-r} / {1-r}}).
+\label{eq:spsaquad-bigocont2}
+\end{align}
+Now from the above equation we can write 
+\begin{align}
+\E (\Lambda_{n-1}' \otimes \Lambda_{n-1}') = O (\text{trace}[\E (\Lambda_{n-1} \tr \Lambda_{n-1})]). 
+\label{eq:spsaquad-bigocont3}
+\end{align}
+
+Now we can use the this result in RHS of the equation \eqref{eq:thmwts} to show the contradiction.
+We consider the two cases to show the contradiction.
+One is where $\frac{d(n)}{n^{2r}} = O(1)$ and another is $lim sup_{n \to \infty} \frac{d(n)}{n^{2r}} = \infty$.
+We show contradiction for the both cases separately.
+Let us consider the first case $\frac{d(n)}{n^{2r}} = O(1)$, then
+\begin{align}
+\text{RHS of \eqref{eq:thmwts}} &= O(1) e^{-2 \Gamma_{1n}} +  O(1) e^{-2 b_0 n^{1-r}/(1-r)} \sum_{i=1}^n \frac{d(i)}{i^{p}} \frac{b_0^2}{i^{2r-p}}\nonumber \\
+&= e^{-2b_0n^{1-r} / {1-r}} \left[ O(1) + O(1) \sum_{i =1}^n \frac{1}{i^{2r-p}}\right].
+\label{eq:spsaquad-bigocont4}
+\end{align}
+where there exist  $0 < p \le 2r$, $0 < \epsilon < 2r -1$ such that $\frac{d(n)}{i^{p}} = O(1)$ and $lim sup_{n \to \infty} \frac{d(n)}{n^{p-\epsilon}} = \infty$.
+In \eqref{eq:spsaquad-bigocont4}, the first equality uses \eqref{eq:spsaquad-bigocont} and \eqref{eq:spsaquad-bigocont3}.
+
+Now there exists a subsequence $\{i_1,\ldots,i_n\}$ such that $ \frac{d(i_n)}{i_n^{p-\epsilon}} = \infty$ as $n \to \infty$.
+Now LHS of \eqref{eq:thmwts} satisfies $LHS(i_n) \ge C i_n^{p-\epsilon} e^{-2b_0 i_n^{1-r} / {1-r}}$ for some $C > 0$.
+Now we have $\frac{LHS(i_n)}{RHS(i_n)} \to 0$ because  $0 < \epsilon < 2r -1$.
+It is contradiction since $LHS(n) = RHS(n)$ for all $n$.
+
+Now consider the second case where $lim sup_{n \to \infty} \frac{d(n)}{n^{2r}} = \infty$.
+There exists a subsequence $\{i_1,\ldots,i_n\}$ such that $ \frac{d(i_n)}{i_n^{2r}} \ge \frac{d(i)}{i^{2r}}$ as $i \le i_n$.
+Then for some $C >0$ the RHS of \eqref{eq:thmwts} satisfies the following
+\begin{align}
+\text{RHS}(i_n)  \text{of}  \eqref{eq:thmwts} &= O(1) e^{-2 \Gamma_{1i_n}} +  O(1) e^{-2 b_0 i_n^{1-r}/(1-r)} \left[\sum_{i=1}^{i_n} e^{2 b_0 i^{1-r}/(1-r)} \frac{b_0^2}{i^{2r}} \E (\Lambda_{n} \otimes \Lambda_{n}) \right]\nonumber \\
+&\le C e^{-2b_0n^{1-r} / {1-r}} \left[ 1 + O(1) \sum_{i =1}^{i_n} \frac{d(i)}{i^{2r}}\right]\nonumber \\
+&\le C e^{-2b_0n^{1-r} / {1-r}} \left[ 1 + \frac{d(i_n)}{{i_n}^{2r}}\right].
+\label{eq:spsaquad-bigocont5}
+\end{align}
+From the equation \eqref{eq:spsaquad-bigocont}, the  LHS of \eqref{eq:thmwts} is $d(i_n) e^{-2b_0 i_n^{1-r} / {1-r}}$.
+By substituting \eqref{eq:spsaquad-bigocont} and \eqref{eq:spsaquad-bigocont5}, in $\frac{RHS{i_n}}{LHS{i_n}}$ then $\frac{RHS{i_n}}{LHS{i_n}} \to 0$ as $n \to \infty$.
+It is contradiction since $LHS(n) = RHS(n)$ for all $n$.
+Hence the claim $\text{trace}[\E (\Lambda_n \tr \Lambda_n)] = O(e^{-2b_0n^{1-r} / {1-r}})$ follows.
 \end{proof}

### 2017-01-21 secorder-rdsa-ih.tex
quad thm part3 completed

@@
 The reader is referred to Section II-B of \cite{prashanth2015rdsa} for a detailed discussion of the above assumptions.
-We remark here that (A8)-(A15) are identical to assumptions in \cite{prashanth2015rdsa}, while (A16) and (A17) introduce minor additional requirements on $\left\| \Upsilon(\overline H_n \right\|^2$ and $\delta_n$, respectively and these are inspired from \cite{spall-jacobian}.
+We remark here that (A8)-(A15) are identical to assumptions in \cite{prashanth2015rdsa}, while (A16) and (A17) introduce minor additional requirements on $\left\| \Upsilon(\overline H_n )\right\|^2$ and $\delta_n$, respectively and these are inspired from \cite{spall-jacobian}.
 
@@
 \end{align*}
-The last step above follows from Toeplitz Lemma (see p. 89 of \cite{lahaprobability}) after observing that $\sum_{i=0}^n \delta_i^4 \rightarrow \infty$ due to (C10). 
+The last step above follows from Toeplitz Lemma (see p. 89 of \cite{lahaprobability}) after observing that $\sum_{i=0}^n \delta_i^4 \rightarrow \infty$ due to (A17). 
 The main claim now follows since 

### 2017-01-21 second-order-spsa-3.tex
chapter 4 conditions changed

@@
  \item Generate $\{\Delta_n^{i}, i=1,\ldots,N\}$,$\{\widehat\Delta_n^{i}, i=1,\ldots,N\}$ independent of $\{\Delta_m,\widehat\Delta_m, m=0,1,\ldots,n-1\}$.  
- \item For any $i=1,\ldots,N$, $\Delta_n^{i}$,$\widehat\Delta_n^{i}$  satisfy condition (C17) of chapter \ref{sec:2spsa-results},(most popular distribution being used is symmetric Bernoulli distribution, $\pm 1$ w.p.
+ \item For any $i=1,\ldots,N$, $\Delta_n^{i}$,$\widehat\Delta_n^{i}$  satisfy condition (A23) of chapter \ref{sec:2spsa-results},(most popular distribution being used is symmetric Bernoulli distribution, $\pm 1$ w.p.
 $1/2$).
@@
 Further, $\delta_n, n\geq 0$ is a sequence of diminishing positive real numbers and 
-$\Delta_n = (\Delta_n^1,\ldots,\Delta_n^N)\tr$ , $\widehat\Delta_n = (\widehat\Delta_n^1,\ldots,\widehat\Delta_n^N)\tr$ are the perturbation random vectors at instant $n$ with $\Delta_n^i,i=1,\ldots,N$, and $\widehat\Delta_n^i,i=1,\ldots,N$ being independent identically distributed (i.i.d), mean-zero random variables having finite inverse moments of order greater than 2 and satisfying the condition (C17) of section \ref{sec:2spsa-results}.
+$\Delta_n = (\Delta_n^1,\ldots,\Delta_n^N)\tr$, $\widehat\Delta_n = (\widehat\Delta_n^1,\ldots,\widehat\Delta_n^N)\tr$ are the perturbation random vectors at instant $n$ with $\Delta_n^i,i=1,\ldots,N$, and $\widehat\Delta_n^i,i=1,\ldots,N$ being independent identically distributed (i.i.d), mean-zero random variables having finite inverse moments of order greater than 2 and satisfying the condition (A23) of section \ref{sec:2spsa-results}.
 
@@
 This is because the third term in \eqref{eq:spsahnhat} vanishes asymptotically, while the fourth term there dominates asymptotically.
-Moreover, the noise factors in the fourth term in \eqref{eq:spsahnhat} are bounded above due to (C22)  and independent of $n$, leaving the $\delta_n^2$ term in the denominator there.
+Moreover, the noise factors in the fourth term in \eqref{eq:spsahnhat} are bounded above due to (A28)  and independent of $n$, leaving the $\delta_n^2$ term in the denominator there.
 
@@
 The reader is referred to Section II-B of \cite{prashanth2015rdsa} for a detailed discussion of the above assumptions.
-We remark here that (C14)-(C21) are identical to those in \cite{prashanth2015rdsa}, while (C22) and (C23) introduce minor additional requirements on $\left\| \Upsilon(\overline H_n) \right\|^2$ and $\delta_n$, respectively and these are inspired from \cite{spall-jacobian}.
+We remark here that (A20)-(A27) are identical to those in \cite{prashanth2015rdsa}, while (A28) and (A29) introduce minor additional requirements on $\left\| \Upsilon(\overline H_n) \right\|^2$ and $\delta_n$, respectively and these are inspired from \cite{spall-jacobian}.
 
@@
 \label{lemma:2spsa-bias}
-Under (C14)-(C23), with $\widehat H_n$ defined according to  \eqref{eq:2spsa-ber} , we have a.s.
+Under (A20)-(A29), with $\widehat H_n$ defined according to  \eqref{eq:2spsa-ber} , we have a.s.
 that\footnote{Here $\widehat H_n(i,j)$ and $\nabla^2_{i,j}f(\cdot)$ denote the $(i,j)$th entry in the Hessian estimate $\widehat H_n$ and the true Hessian $\nabla^2 f(\cdot)$, respectively.}, for $i,j = 1,\ldots,N$,
@@
 \label{thm:2spsa-H}
-Under (C14)-(C23), we have that 
+Under (A20)-(A29), we have that 
 $$\theta_n \rightarrow \theta^*, \overline H_n \rightarrow \nabla^2 f(\theta^*) \text{ a.s. as } n\rightarrow \infty.$$ 
@@
 \item  $f$ is quadratic and $\nabla^2 f(x^*) > 0$. 
-\item The operator $\Upsilon$ is chosen such that $\E(\parallel \Upsilon(\overline H_n) - \overline H_n\parallel^2) = o(e^{-2wn^{1-r}/(1-r)})$ and $\parallel \Upsilon(H) - H \parallel^2 / (1+\parallel H \parallel^2)$ is uniformly bounded.
+\item The operator $\Upsilon$ is chosen such that $\E(\parallel \Upsilon(\overline H_n) - \overline H_n\parallel^2) = o(e^{-2wn^{1-r}/(1-r)})$ and \\$\parallel \Upsilon(H) - H \parallel^2 / (1+\parallel H \parallel^2)$ is uniformly bounded.
 \end{pvn}
@@
 \label{thm:spsaquad-bound}
-Assume (C17), (C23), (C24) and (C25) and also that the setting is noise-free. 
+Assume (A23), (A29), (A30) and (A31) and also that the setting is noise-free. 
 Let $b_n = b_0/n^r$, $n=1,2,\ldots,k$, where $1/2 < r < 1$ and $0 < b_0 \leq 1$. For notational simplicity, let $H^*=\nabla^2 f(x^*)$. Letting $\Lambda_k = \overline H_k - H^*$, we have 
@@
 This part in exactly the same manner as part (i) in Theorem 3 of \cite{spall-jacobian}.\\
+By rewriting the \eqref{eq:2spsa-3-ih} in the following form
+\begin{align}\label{eq:hhatexpanded}
+\overline H_{n} &= \overline H_{n-1} - b_{n} ( \overline H_{n-1} - \widehat H_n + \widehat \Psi_n)\nonumber\\
+&= \overline H_{n-1} - b_{n} ( \overline H_{n-1} - H^* + \widehat \Psi_n - \Psi_n (H^*))
+\end{align}
+The above equation can be thought as a stochastic approximation analogue of \eqref{eq:2spsa-3-ih}.
 \noindent\textbf{Step 2: Representation of $\text{trace}[\E (\Lambda_n \tr \Lambda_n)]$} \\
@@
 From the equation \eqref{eq:spsaquad-bigocont}, the  LHS of \eqref{eq:thmwts} is $d(i_n) e^{-2b_0 i_n^{1-r} / {1-r}}$.
-By substituting \eqref{eq:spsaquad-bigocont} and \eqref{eq:spsaquad-bigocont5}, in $\frac{RHS{i_n}}{LHS{i_n}}$ then $\frac{RHS{i_n}}{LHS{i_n}} \to 0$ as $n \to \infty$.
+By substituting \eqref{eq:spsaquad-bigocont} and \eqref{eq:spsaquad-bigocont5}, in $\frac{RHS(i_n)}{LHS(i_n)}$ then $\frac{RHS(i_n)}{LHS(i_n)} \to 0$ as $n \to \infty$.
 It is contradiction since $LHS(n) = RHS(n)$ for all $n$.

### 2017-01-21 second-order-spsa-3.tex
quad thm chapter4 part1 slightly done

@@
 \begin{align}
- \widehat H_n =    \nabla^2 f(\theta_n) +\Psi_{n}(\nabla^2 f(\theta_n)),\label{eq:spsahnhat-ext}
+ \widehat H_n &=    \nabla^2 f(\theta_n) +\Psi_{n}(\nabla^2 f(\theta_n))\nonumber\\
+ &=    \nabla^2 f(\theta^*) +\Psi_{n}(\nabla^2 f(\theta_n))\nonumber\\
+ &=    H^* +\Psi_{n}(H^* ),\label{eq:spsahnhat-ext}
 \end{align}
@@
 The above equation can be thought as a stochastic approximation analogue of \eqref{eq:2spsa-3-ih}.
+The above stochastic approximation algorithm aimed at finding the roots of the equation $H-H^*=0$.
+In \eqref{eq:hhatexpanded}, $\E (\widehat \Psi_n - \Psi_n (H^*) | \F_n) = 0 \,\,\ a.s$.
+Given a Lyapunov function, by showing the conditions required to show   the mean square convergence of the \eqref{eq:hhatexpanded} from the \cite{nevel1973stochastic}, pp.
+92-94.
+From part (i) in Theorem 3 of \cite{spall-jacobian}, one can check that our algorithm indeed satisfy those conditions,  which in turn implies the $\E [\Lambda_n\tr \Lambda_n] \rightarrow 0$ a.s.
+as $n\rightarrow \infty$.
+
 \noindent\textbf{Step 2: Representation of $\text{trace}[\E (\Lambda_n \tr \Lambda_n)]$} \\

### 2017-01-22 intro.tex
changes almost done

@@
 In \cite{spall}, in addition to establishing formal convergence of $\theta_n \to \theta^*$, it was shown that probability distribution of appropriately scaled $\theta_n$ is approximately normal  (with some mean and covariance matrix) for larger $n$.
-The general metric which is used for evaluating these class of algorithms is asymptotic mean squared error (AMSE) defined as $\E\l \theta_{n} - \theta^* \r^2$.
+The general metric which is used for evaluating these class of algorithms is asymptotic mean squared error (AMSE), defined as follows  $\E\l \theta_{n} - \theta^* \r^2$.
 The relative efficiency of 1SPSA when compared to FDSA can be shown when we compare the similar asymptotic normality result of FDSA.
-In \cite{spall}, it was shown that under reasonably general conditions, and when $\{a_n\}$, $\{\delta_n\}$ are chosen according to the guidelines in \cite{spall1998implementation} then 1SPSA and FDSA gives same AMSE for given number of iterations.
-Instead of looking the
+That is
 \begin{align}\label{eq:amse}
@@
 \end{align}
-an equivalent way of looking the relative AMSE is to compare the number of function measurements required by FDSA to achieve same AMSE as 1SPSA which can be computed by equating the \eqref{eq:amse} to one.
+In \cite{spall}, it was shown that under reasonably general conditions, and when $\{a_n\}$, $\{\delta_n\}$ are chosen according to the guidelines in \cite{spall1998implementation} then 1SPSA and FDSA gives same AMSE when both the algorithms are ran for the same number of iterations.
+Instead of looking \eqref{eq:amse}, an equivalent way of looking the relative AMSE is to compare the number of function measurements required by FDSA to achieve same AMSE as 1SPSA.
+One can compute it  by equating the \eqref{eq:amse} to one.
+Then it will result the following:
 \begin{align}\label{eq:amsenom}

### 2017-01-22 second-order-spsa-3.tex
changes almost done

@@
  \widehat H_n &=    \nabla^2 f(\theta_n) +\Psi_{n}(\nabla^2 f(\theta_n))\nonumber\\
- &=    \nabla^2 f(\theta^*) +\Psi_{n}(\nabla^2 f(\theta_n))\nonumber\\
+ &=    \nabla^2 f(\theta^*) +\Psi_{n}(\nabla^2 f(\theta^*))\nonumber\\
  &=    H^* +\Psi_{n}(H^* ),\label{eq:spsahnhat-ext}
@@
 \overline H_{n} &= \overline H_{n-1} - b_{n} ( \overline H_{n-1} - \widehat H_n + \widehat \Psi_n)\nonumber\\
-&= \overline H_{n-1} - b_{n} ( \overline H_{n-1} - H^* + \widehat \Psi_n - \Psi_n (H^*))
+&= \overline H_{n-1} - b_{n} ( \overline H_{n-1} - H^* + \widehat \Psi_n - \Psi_n (H^*)).
 \end{align}
@@
 In \eqref{eq:hhatexpanded}, $\E (\widehat \Psi_n - \Psi_n (H^*) | \F_n) = 0 \,\,\ a.s$.
-Given a Lyapunov function, by showing the conditions required to show   the mean square convergence of the \eqref{eq:hhatexpanded} from the \cite{nevel1973stochastic}, pp.
-92-94.
-From part (i) in Theorem 3 of \cite{spall-jacobian}, one can check that our algorithm indeed satisfy those conditions,  which in turn implies the $\E [\Lambda_n\tr \Lambda_n] \rightarrow 0$ a.s.
-as $n\rightarrow \infty$.
+Given a Lyapunov function, one can  the conditions required to show  the mean square convergence of the \eqref{eq:hhatexpanded} from the \cite{nevel1973stochastic}, pp.92-94.
+It is easy to see that $V(H) = \frac{1}{2} (H-H^*)\tr(H-H^*)$ indeed serves as Lyapunov function.
+From part (i) in Theorem 3 of \cite{spall-jacobian}, one can check that our algorithm indeed satisfy those conditions,  which in turn implies the $\E [\Lambda_n\tr \Lambda_n] \rightarrow 0 \,\,\,a.s.$ as $n\rightarrow \infty$.
 
@@
 \begin{align}
-\text{RHS}(i_n)  \text{of}  \eqref{eq:thmwts} &= O(1) e^{-2 \Gamma_{1i_n}} +  O(1) e^{-2 b_0 i_n^{1-r}/(1-r)} \left[\sum_{i=1}^{i_n} e^{2 b_0 i^{1-r}/(1-r)} \frac{b_0^2}{i^{2r}} \E (\Lambda_{n} \otimes \Lambda_{n}) \right]\nonumber \\
+\text{RHS}(i_n)\,\,\,  \text{of}  \,\,\,\eqref{eq:thmwts} &= O(1) e^{-2 \Gamma_{1i_n}} +  O(1) e^{-2 b_0 i_n^{1-r}/(1-r)} \left[\sum_{i=1}^{i_n} e^{2 b_0 i^{1-r}/(1-r)} \frac{b_0^2}{i^{2r}} \E (\Lambda_{n} \otimes \Lambda_{n}) \right]\nonumber \\
 &\le C e^{-2b_0n^{1-r} / {1-r}} \left[ 1 + O(1) \sum_{i =1}^{i_n} \frac{d(i)}{i^{2r}}\right]\nonumber \\
@@
 \end{align}
-From the equation \eqref{eq:spsaquad-bigocont}, the  LHS of \eqref{eq:thmwts} is $d(i_n) e^{-2b_0 i_n^{1-r} / {1-r}}$.
+Where $\text{RHS}(i_n)$ denotes, RHS after the algorithm is iterated for $i_n$ times.
+From the equation \eqref{eq:spsaquad-bigocont}, the  LHS of \eqref{eq:thmwts} after $i_n$ iterations is $d(i_n) e^{-2b_0 i_n^{1-r} / {1-r}}$.
 By substituting \eqref{eq:spsaquad-bigocont} and \eqref{eq:spsaquad-bigocont5}, in $\frac{RHS(i_n)}{LHS(i_n)}$ then $\frac{RHS(i_n)}{LHS(i_n)} \to 0$ as $n \to \infty$.

### 2017-01-23 conc.tex
experiments sir corrections done

--- 2016-12-24 conc.tex
@@
 \label{sec:conclusions}
-We presented an improved Hessian estimation scheme for the 2RDSA algorithm \cite{prashanth2015rdsa}.
+In this thesis, we presented an improved Hessian estimation scheme for the 2RDSA algorithm \cite{prashanth2015rdsa}.
 The proposed scheme was shown to be provably convergent to the true Hessian.
-We proposed generalisation of RDSA algorithm proposed in \cite{prashanth2015rdsa}.
-As a result of generalisation we are able to use much larger class of distributions  for perturbation random variables.
+We proposed generalisations of the RDSA algorithm proposed in \cite{prashanth2015rdsa}.
+As a result of the generalisation, we are able to use a much larger class of distributions  for the perturbation random variables.
 The proposed generalisation inherits all the improved Hessian estimation  schemes proposed earlier.
-We also presented an improved Hessian estimation scheme for 2SPSA-3 algorithm \cite{bhatnagar2015simultaneous} and as a special case, for a quadratic objective resulted in a  convergence rate that is on par with the corresponding rate for 2SPSA with Hessian estimation improvements (2SPSA-IH) \cite{spall-jacobian}.
-The advantage with 2RDSA-IH, 2SPSA-3-IH is that it requires only 75\% of the simulation cost per-iteration for 2SPSA with Hessian estimation improvements (2SPSA-IH) \cite{spall-jacobian}.
-Numerical experiments demonstrated that 2RDSA-IH, 2SPSA-3-IH outperforms both 2SPSA-IH and 2SPSA without the improved Hessian estimation scheme.
-They also indicate that schemes with improved Hessian estimation outperforms their respective regular schemes without the improved Hessian estimation scheme.
-2RDSA-IH with asymmetric Bernoulli distribution performing the best overall.
+We also presented an improved Hessian estimation scheme for the 2SPSA-3 algorithm \cite{bhatnagar2015simultaneous} and for the special case of a quadratic objective, it resulted in a  convergence rate that is on par with the corresponding rate for 2SPSA with Hessian estimation improvements (2SPSA-IH) \cite{spall-jacobian}.
+The advantage with 2RDSA-IH, 2SPSA-3-IH is that these schemes it require only 75\% of the simulation cost per-iteration for 2SPSA with Hessian estimation improvements (2SPSA-IH) \cite{spall-jacobian}.
+Numerical experiments demonstrated that 2RDSA-IH, 2SPSA-3-IH outperform both 2SPSA-IH and 2SPSA without the improved Hessian estimation procesdure.
+They also indicate that schemes with improved Hessian estimation outperform the respective regular schemes without the improved Hessian estimation procedure.
+2RDSA-IH with asymmetric Bernoulli distribution is seen to perform the best overall.
 
-As future work, it would be interesting to derive finite time bounds that show a lower Hessian estimation error for 2RDSA-IH when compared to 2RDSA and 2SPSA.
-Stochastic Newton methods  methods converge faster and  are often more accurate than simple gradient search schemes.
+As future work, it would be interesting to look into the following directions:
+\begin{enumerate}
+\item Derive finite time bounds that show a lower Hessian estimation error for 2RDSA-IH when compared to 2RDSA and 2SPSA. 
+\item Stochastic Newton methods converge faster and  are often more accurate than simple gradient search schemes.
 However they are computationally much costlier than first-order methods.
-Since first-order methods exhibits slower convergence rate and heavily depended on the step-size selection one has to resort to second-order methods.
-As a future work, it would be interesting to look at the methods like momentum descent, conjugate gradient methods, Hessian free optimization schemes which result in faster convergence for the first-order methods in deterministic optimization case.
-It would be interesting to extend those methods to stochastic optimization case and prove that they exhibits  faster convergence than first-oder methods.
-From application point of view it would be interesting to apply these methods to some interesting real world applications.
+Since first-order methods exhibit slower convergence rate and heavily depend on the step-size selection one has to resort to second-order methods for improved performance.
+As a future work, it would be interesting to look at  methods such as momentum descent, conjugate gradient as well as, Hessian free optimization approaches which result in faster convergence for the first-order methods and are known to be computationally cheaper than second-order methods in the deterministic optimization scenario.
+It would be interesting to extend those methods to the stochastic optimization setting and prove that they exhibit  faster convergence than first-order methods and result in lower computational cost.
+\item In the methods described in this thesis the perturbations used are random variables.
+In \cite{bhatnagar2003two}, for the first-order SPSA method,  construction of two different deterministic  perturbation schemes was proposed and  under certain conditions on the perturbation vectors and noise sequences, the algorithm is shown to converge.
+The idea behind using deterministic perturbations is to reduce the bias in the gradient estimates.
+It would be interesting to extend this construction to the RDSA scheme as well.
+And moreover this kind of construction for second-order methods is not available for both SPSA and RDSA.
+One can use the above ideas and show significant improvements to the existing methods.
+\item From an application point of view it would be interesting to apply these methods to some interesting real world applications.
+It would be interesting to explore applications of  these methods  to the design of reinforcement learning algorithms.
+\end{enumerate}

### 2017-01-23 experiments.tex
experiments sir corrections done

@@
  \end{align} 
-The optimum $\theta^*$ for above $f$ is $\theta^*=0$, with $f(\theta^*) = 0$.
+The optimum $\theta^*$ for the above $f$ is $\theta^*=0$, with $f(\theta^*) = 0$.
 
@@
 \centering
- \caption{Normalized loss values for fourth-order  objective \eqref{eq:4thorder} with and without noise: standard error from $500$ replications shown after $\pm$}
+ \caption{Normalized loss values for fourth-order  objective \eqref{eq:4thorder} with and without noise: standard error from $500$ is replications shown after $\pm$.}
 \label{tab:norloss-4thf}
@@
 \centering
- \caption{Normalized loss values for quadratic objective \eqref{eq:quadratic} with and without noise: standard error from $500$ replications shown after $\pm$}
+ \caption{Normalized loss values for quadratic objective \eqref{eq:quadratic} with and without noise: standard error from $500$ replications is shown after $\pm$.}
 \label{tab:norloss-quadratic}
@@
 \centering
- \caption{NMSE values for quadratic objective \eqref{eq:quadratic} with and without noise: standard error from $500$ replications shown after $\pm$}
+ \caption{NMSE values for quadratic objective \eqref{eq:quadratic} with and without noise: standard error from $500$ replications is shown after $\pm$.}
 \label{tab:nmse-quadratic}
@@
 Table \ref{tab:nmse-quadratic} presents the NMSE values obtained for the aforementioned algorithms with the quadratic loss.
-The results in Tables \ref{tab:norloss-4thf}--\ref{tab:nmse-quadratic} are obtained after running all the algorithms with a budget of $10000$ function evaluations.
+The results in Tables \ref{tab:norloss-4thf}--\ref{tab:nmse-quadratic} are obtained after running all the algorithms with a budget of $10,000$ function evaluations.
 
- Figures \ref{fig:norlossvssims}, \ref{fig:norlossvssimsspsa-3} plots the normalized loss as a function of the simulation budget with the fourth-order loss objective with $\sigma=0.1$ for algorithms 2RDSA and 2SPSA-3 respectively when compared with 2SPSA.
-Figures \ref{fig:QuadnorlossvssimsNoiseless}, \ref{fig:QuadnorlossvssimsNoiselessspsa3} plots the normalized loss as a function of the simulation budget with the quadratic loss objective with $\sigma=0$ for algorithms 2RDSA and 2SPSA-3 respectively when compared with 2SPSA.
+ Figures \ref{fig:norlossvssims}, \ref{fig:norlossvssimsspsa-3} plot the normalized loss as a function of the simulation budget with the fourth-order loss objective with $\sigma=0.1$ for algorithms 2RDSA and 2SPSA-3 respectively when compared with 2SPSA.
+Figures \ref{fig:QuadnorlossvssimsNoiseless}, \ref{fig:QuadnorlossvssimsNoiselessspsa3} plot the normalized loss as a function of the simulation budget with the quadratic loss objective with $\sigma=0$ for algorithms 2RDSA and 2SPSA-3 respectively when compared with 2SPSA.
 See Figures \ref{fig:norlossvssimsNoiseless}-- \ref{fig:QuadnorlossvssimsNoisy} and \ref{fig:norlossvssimsNoiselessspsa3}-- \ref{fig:QuadnorlossvssimsNoisyspsa3} for similar results with $\sigma = 0$ for fourth-order loss and $\sigma=0.1$ for quadratic loss.
@@
 
- \textit{\textbf{Observation 1:} Schemes with improved Hessian estimation performs better than their respective regular schemes.}
+ \textit{\textbf{Observation 1:} Schemes with improved Hessian estimation perform better than their respective regular schemes.}
 
@@
 
-\textit{\textbf{Observation 4:} 2RDSA-IH-AsymBer performing the best overall.}
+\textit{\textbf{Observation 4:} 2RDSA-IH-AsymBer perform the best overall.}
 

### 2017-01-23 second-order-spsa-3.tex
experiments sir corrections done

@@
 \noindent\textbf{Step 1: MSE convergence of $\overline H_n$} \\
-This part in exactly the same manner as part (i) in Theorem 3 of \cite{spall-jacobian}.\\
+%This part in exactly the same manner as part (i) in Theorem 3 of \cite{spall-jacobian}.\\
 By rewriting the \eqref{eq:2spsa-3-ih} in the following form
@@
 In \eqref{eq:hhatexpanded}, $\E (\widehat \Psi_n - \Psi_n (H^*) | \F_n) = 0 \,\,\ a.s$.
-Given a Lyapunov function, one can  the conditions required to show  the mean square convergence of the \eqref{eq:hhatexpanded} from the \cite{nevel1973stochastic}, pp.92-94.
+Given a Lyapunov function, one can  get the conditions required to show  the mean square convergence of the \eqref{eq:hhatexpanded} from the \cite{nevel1973stochastic}, pp.92-94.
 It is easy to see that $V(H) = \frac{1}{2} (H-H^*)\tr(H-H^*)$ indeed serves as Lyapunov function.

### 2017-01-23 abstract.tex
sir corrections + synopsis + abstract

--- 2017-01-13 abstractShort.tex
@@
+%\documentclass[oneside,8pt,onecolumn,openright]{IIScthesisPSnPDF}
+\documentclass[12pt]{article}
+
+\usepackage{paralist}
+\let\labelindent\relax
+\usepackage{adjustbox}
+\usepackage{siunitx}
+\usepackage{booktabs}
+\usepackage{url}
+\usepackage{hyperref}
+
+\usepackage{xcolor}
+\usepackage{colortbl}
+\usepackage{enumitem}
+\usepackage{amsmath}
+
+\usepackage{bm}
+\usepackage{tikz}
+\usepackage{pgfplots}
+\usepackage{pgfplotstable}
+\pgfplotsset{compat=1.12}
+\usepackage{todonotes}
+\usepackage{nicefrac}
+\usetikzlibrary{positioning,arrows,calc,patterns,backgrounds}
+\usetikzlibrary{shapes}
+\usetikzlibrary{fit}
+\usepackage{multirow}
+
+\usepackage{makecell}
+\usepackage{caption}
+\usepackage{subcaption}
+
+\usepackage{amssymb}
+\usepackage{amsmath}
+\usepackage{latexsym}
+\usepackage{multicol}
+\usepackage{booktabs}
+\usepackage{multirow}
+\usepackage{fancyhdr}
+\usepackage[numbers]{natbib}
+\usepackage{cleveref}
+\usepackage{wrapfig}
+\usepackage{algorithm}
+\usepackage{algorithmic}
+\usepackage{enumerate}
+\usepackage{enumitem}
+\usepackage{cancel}
+\usepackage{mathrsfs}
+% \usepackage{sfgame}
+\usepackage{color}
+\usepackage[compact]{titlesec}
+\usepackage{mdwlist}
+\usepackage{varwidth}
+\usepackage[vertfit]{breakurl}
+\usepackage{datetime}
+\usepackage{pdfpages}
+% \usepackage{biblatex}
+\usepackage{setspace}
+\title{\textbf{ABSTRACT}}
+\author{}
+\date{}
+\begin{document}
+%\Huge
+%\begin{center}
+%Panel of Experts
+%\end{center}
+\maketitle
+\begin{tabular}{ll}
+Name of the Candidate & \textbf{Danda Sai Koti Reddy}\\ 
+SR Number &\textbf{04-04-00-10-21-14-1-11609} \\
+Title of the Thesis & \textbf{Stochastic Newton methods with   } \\
+ & \textbf{ enhanced Hessian estimation.}\\
+ Research Supervisor  & \textbf{Prof. Shalabh Bhatnagar} \\
+Degree Registered  &  \textbf{M.Sc(engg)} \\
+Department  & \textbf{Computer Science and Automation} \\
+Institute  & \textbf{Indian Institute of Science, Bangalore }
+\end{tabular}
+
+\vspace{1.5cm}
+
 Optimization problems involving uncertainties are common in a variety
@@
 system under consideration while updating the system parameters until a
-good enough solution is obtained.
+good enough solution is obtained.\\
 
@@
 second-order information into the stochastic updates. Stochastic Newton
-methods are often more accurate than simple gradient search schemes.
+methods are often more accurate than simple gradient search schemes.\\
 
@@
 Numerical experiments show that 2RDSA-IH outperforms both 2SPSA-IH and 2RDSA without the improved Hessian estimation scheme.
+
+\bibliographystyle{plain}
+\bibliography{refThesis}
+
+\end{document}

### 2017-01-23 intro.tex
sir corrections + synopsis + abstract

@@
 \begin{pvn}
-\item $\{\Delta_{ni}, i=1,\ldots,N, n=1,2,\ldots\}$ are i.i.d., independent of $\F_n$ and for some $\alpha_0 > 0$ and $\forall n$, $\E (\Delta_{ni})^{-2} \le \alpha_0$, for $i = 1,2,\ldots,N$.
+\item $\{\Delta_{ni}, i=1,\ldots,N, n=1,2,\ldots\}$ are i.i.d., independent of $\F_n$ and for some $\alpha_3 > 0$ and $\forall n$, $\E (\Delta_{ni})^{-2} \le \alpha_3$, for $i = 1,2,\ldots,N$.
 \end{pvn}
@@
 In \cite{spall}, in addition to establishing formal convergence of $\theta_n \to \theta^*$, it was shown that probability distribution of appropriately scaled $\theta_n$ is approximately normal  (with some mean and covariance matrix) for larger $n$.
-The general metric which is used for evaluating these class of algorithms is asymptotic mean squared error (AMSE), defined as follows  $\E\l \theta_{n} - \theta^* \r^2$.
-The relative efficiency of 1SPSA when compared to FDSA can be shown when we compare the similar asymptotic normality result of FDSA.
-That is
+The general metric which is used for evaluating these classes of algorithms is asymptotic mean squared error (AMSE), defined as $\E\l \theta_{n} - \theta^* \r^2$.
+The relative efficiency of 1SPSA when compared to FDSA can be shown when we compare the similar asymptotic normality results of FDSA.
+That is,
 \begin{align}\label{eq:amse}
-\frac{\text{AMSE of 1SPSA}}{\text{AMSE of FDSA}},
+\frac{\text{AMSE of 1SPSA}}{\text{AMSE of FDSA}}.
 \end{align}
-In \cite{spall}, it was shown that under reasonably general conditions, and when $\{a_n\}$, $\{\delta_n\}$ are chosen according to the guidelines in \cite{spall1998implementation} then 1SPSA and FDSA gives same AMSE when both the algorithms are ran for the same number of iterations.
-Instead of looking \eqref{eq:amse}, an equivalent way of looking the relative AMSE is to compare the number of function measurements required by FDSA to achieve same AMSE as 1SPSA.
+In \cite{spall}, it was shown that under reasonably general conditions, and when $\{a_n\}$, $\{\delta_n\}$ are chosen according to the guidelines in \cite{spall1998implementation}, 1SPSA and FDSA gives the same AMSE when both the algorithms are run for the same number of iterations.
+Instead of looking at  \eqref{eq:amse}, an equivalent way of looking at the relative AMSE is to compare the number of function measurements required by FDSA to achieve the same AMSE as 1SPSA.
 One can compute it  by equating the \eqref{eq:amse} to one.
-Then it will result the following:
+Then it will result in the following:
 \begin{align}\label{eq:amsenom}
@@
 \begin{pvn}
-\item $\{ d_{n}^i, i=1,\ldots,N, n=1,2,\ldots\}$ are i.i.d., independent of $\F_n$, and $\forall n$, $\E (d_{n} d_{n}\tr) = I_{N \times N}$, and for some $\alpha_0 > 0$, $\E {d_{n}^i}^2 f(\theta_n \pm \delta_n d_n) \le \alpha_0$ for $i=1,2,\ldots,N$.
+\item $\{ d_{n}^i, i=1,\ldots,N, n=1,2,\ldots\}$ are i.i.d., independent of $\F_n$, and $\forall n$, $\E (d_{n} d_{n}\tr) = I_{N \times N}$, and for some $\alpha_4 > 0$, $\E {d_{n}^i}^2 f(\theta_n \pm \delta_n d_n) \le \alpha_4$ for $i=1,2,\ldots,N$.
 \end{pvn}
@@
 This generalisation of distributions of random variables for perturbations is important because it enhances the scope of improving the performance of the 1RDSA and 2RDSA schemes by incorporating several other distributions.
-Some well known distribution that can be used for these schemes as a result of the generalization are Normal (0,1) and Cauchy distributions.
+Some well known distribution that can be used for these schemes as a result of the generalization are Normal (0,1) and truncated Cauchy distributions.
 \item Our work in chapter \ref{sec:2spsa-3-ih} is centred on improving the second order SPSA scheme with three simulations (referred to as 2SPSA-3 hereafter), which was proposed in  \cite{bhatnagar2015simultaneous}.

### 2017-01-23 second-order-rdsa-gen.tex
sir corrections + synopsis + abstract

@@
 This generalisation of distributions of random variables for perturbations is important because it enhances the scope of improving the performance of the 1RDSA and 2RDSA schemes by incorporating several other distributions.
-Some well known distribution that can be used for these schemes as a result of the generalisation are Normal (0,1) and Cauchy distributions.
+Some well known distribution that can be used for these schemes as a result of the generalisation are Normal (0,1) and truncated Cauchy distributions.
 We will prove that  the  gradient and Hessian estimates as a result of the generalisation are indeed asymptotically unbiased.
@@
 %Almost surely, as $\delta_n \rightarrow 0$, we have that
-Under (A1),(A2),(A4),(A5),(A18),(A19) for $\widehat\nabla f(\theta_n)$ defined according to  \eqref{eq:grad-gen} , we have a.s.
+Under (A1),(A4),(A5),(A18),(A19) for $\widehat\nabla f(\theta_n)$ defined according to  \eqref{eq:grad-gen} , we have a.s.
 that\footnote{Here $\widehat\nabla_i f(\theta_n)$ and $\nabla_i f(\theta_n)$ denote the $i$th coordinates in the gradient estimate $\widehat\nabla f(\theta_n)$ and true gradient $\nabla f(\theta_n)$, respectively.}
@@
 \begin{align*}
-&\E\left[\left.\dfrac{y_n^+ - y_n^-}{2\delta_n} \right| \F_n\right] \\
+\E\left[\left. d_n\left(\dfrac{y_n^+ - y_n^-}{2\delta_n}\right) \right| \F_n\right] 
 %= &\E\left[d_n\left.\left(\dfrac{f(\theta_n+\delta_n d_n) - f(\theta_n-\delta_n d_n)}{2\delta_n}\right) \right| \F_n\right] + \E\left[  d_n\left.\left(\dfrac{\xi_n^+ - \xi_n^-}{2\delta_n}\right) \right| \F_n\right]\\
-= & \E\left[d_n\left.\left(\dfrac{f(\theta_n+\delta_n d_n) - f(\theta_n-\delta_n d_n)}{2\delta_n}\right) \right| \F_n\right].
+=  \E\left[d_n\left.\left(\dfrac{f(\theta_n+\delta_n d_n) - f(\theta_n-\delta_n d_n)}{2\delta_n}\right) \right| \F_n\right].
 \end{align*}
@@
 \begin{align*}
-f(\theta_n \pm \delta_n d_n) =& f(\theta_n) \pm \delta_n d_n\tr \nabla f(\theta_n) + \frac{\delta_n^2}{2} d_n\tr \nabla^2 f(\theta_n) d_n \\
-& \pm  \frac{\delta_n^3}{6} \nabla^3 f(\tilde  \theta_n^+)(d_n \otimes d_n \otimes d_n),
+f(\theta_n \pm \delta_n d_n) = f(\theta_n) \pm \delta_n d_n\tr \nabla f(\theta_n) + \frac{\delta_n^2}{2} d_n\tr \nabla^2 f(\theta_n) d_n 
+ \pm  \frac{\delta_n^3}{6} \nabla^3 f(\tilde  \theta_n^+)(d_n \otimes d_n \otimes d_n),
 \end{align*}
@@
 \begin{align}
-&\E\left[d_n\left.\left(\dfrac{f(\theta_n+\delta_n d_n) - f(\theta_n-\delta_n d_n)}{2\delta_n}\right)\right| \F_n \right] \nonumber\\
-&= \E\left[d_n d_n\tr \left.\nabla f(\theta_n)\right| \F_n\right]\nonumber \\
-& +   \E\left[\left.\frac{\delta_n^2}{12} d_n (\nabla^3 f(\tilde  \theta_n^+)+\nabla^3 f(\tilde  \theta_n^-))(d_n \otimes d_n \otimes d_n)\right| \F_n\right]. \label{eq:l1}
+&\E\left[d_n\left.\left(\dfrac{f(\theta_n+\delta_n d_n) - f(\theta_n-\delta_n d_n)}{2\delta_n}\right)\right| \F_n \right] 
+= \E\left[d_n d_n\tr \left.\nabla f(\theta_n)\right| \F_n\right]\nonumber \\
+& \hspace{5.5cm}+   \E\left[\left.\frac{\delta_n^2}{12} d_n (\nabla^3 f(\tilde  \theta_n^+)+\nabla^3 f(\tilde  \theta_n^-))(d_n \otimes d_n \otimes d_n)\right| \F_n\right]. \label{eq:l1}
 \end{align}
@@
 \begin{align}
-&\E\left[\left.\frac{\delta_n^2}{12} d_n^l (\nabla^3 f(\tilde  \theta_n^+)+\nabla^3 f(\tilde  \theta_n^-))(d_n \otimes d_n \otimes d_n)\right| \F_n\right]\nonumber\\
-\le & \dfrac{\alpha_0 \delta_n^2}{6} \sum_{i_1=1}^N\sum_{i_2=1}^N\sum_{i_3=1}^N \E\left( d_n^l d_n^{i_1} d_n^{i_2} d_n^{i_3}\right)\nonumber\\
+\left|\E\left[\left.\frac{\delta_n^2}{12} d_n^l (\nabla^3 f(\tilde  \theta_n^+)+\nabla^3 f(\tilde  \theta_n^-))(d_n \otimes d_n \otimes d_n)\right| \F_n\right] \right| \le& \dfrac{\alpha_0 \delta_n^2}{6} \sum_{i_1=1}^N\sum_{i_2=1}^N\sum_{i_3=1}^N \E \left|\left( d_n^l d_n^{i_1} d_n^{i_2} d_n^{i_3}\right)\right|\nonumber\\
 \le & \dfrac{\alpha_0 \delta_n^2\eta N^3}{6}.\label{eq:l2}

### 2017-01-23 second-order-spsa-3.tex
sir corrections + synopsis + abstract

--- 2017-01-22 second-order-spsa-3.tex
@@
 
-\item $\{\Delta_n^i,\widehat\Delta_n^i, i=1,\ldots,N, n=1,2,\ldots\}$ are i.i.d independent of $\F_n$ and for some $\alpha_0 > 0$ and for all $n,l$ $|\Delta_{nl}|\le \alpha_0$, $\Delta_{nl}$ is symmetrically distributed about 0, $\Delta_{nl}$ are mutually independent across $n$ and $l$ and they satisfy $\E (\Delta_{nl}^{-2})$,$\E (\widehat\Delta_{nl}^{-2}) \le \alpha_0$.
+\item $\{\Delta_n^i,\widehat\Delta_n^i, i=1,\ldots,N, n=1,2,\ldots\}$ are i.i.d independent of $\F_n$ and for some $\alpha_3 > 0$ and for all $n,l$ $|\Delta_{nl}|\le \alpha_3$, $\Delta_{nl}$ is symmetrically distributed about 0, $\Delta_{nl}$ are mutually independent across $n$ and $l$ and they satisfy $\E (\Delta_{nl}^{-2})$, $\E (\widehat\Delta_{nl}^{-2}) \le \alpha_3$.
 
@@
 for all $|(\theta-\theta^*)_i| < \tau$ when $i \notin S$ and $|(\theta-\theta^*)_i| \ge \rho'(\tau,S)$ when $i\in S$.
-\item For some $\alpha_0, \alpha_1, \alpha_2 >0$ and for all $n,l,m$, $\E {\xi_n}^{2} \le \alpha_0$, $\E {\xi_n^{\pm}}^{2} \le \alpha_0$, $\E f(\theta_n)^{2} \le \alpha_1$,  $\E f(\theta_n + \delta_n \Delta_n + \delta_n \widehat \Delta_n)^{2},\E f(\theta_n - \delta_n \Delta_n - \delta_n \widehat \Delta_n)^{2} \le \alpha_1$,
+\item For some $\alpha_7, \alpha_8, \alpha_9 >0$ and for all $n,l,m$, $\E {\xi_n}^{2} \le \alpha_7$, $\E {\xi_n^{\pm}}^{2} \le \alpha_7$, $\E f(\theta_n)^{2} \le \alpha_8$,  $\E f(\theta_n + \delta_n \Delta_n + \delta_n \widehat \Delta_n)^{2},\E f(\theta_n - \delta_n \Delta_n - \delta_n \widehat \Delta_n)^{2} \le \alpha_8$,
 %$\E \left[ f(\theta_n + \delta_n \Delta_n + \delta_n \widehat \Delta_n)^{2}/ (\Delta_{nl}\widehat\Delta_{nm})^2 \mid \F_n\right]$,
-$\E \left[| f(\theta_n + \delta_n \Delta_n + \delta_n \widehat \Delta_n)/ (\Delta_{nl}\widehat\Delta_{nm})|^{2+\alpha_2} \mid \F_n\right]$,
+$\E \left[| f(\theta_n + \delta_n \Delta_n + \delta_n \widehat \Delta_n)/ (\Delta_{nl}\widehat\Delta_{nm})|^{2+\alpha_9} \mid \F_n\right]$,
 %$\E \left[ f(\theta_n - \delta_n \Delta_n - \delta_n \widehat \Delta_n)^{2}/ (\Delta_{nl}\widehat\Delta_{nm})^2 \mid \F_n\right]$,
-$\E \left[ |f(\theta_n - \delta_n \Delta_n - \delta_n \widehat \Delta_n)/ (\Delta_{nl}\widehat\Delta_{nm})|^{2+\alpha_2} \mid \F_n\right]$,
- $\E\left[\left. (\xi_n^+ + \xi_n^- - 2 \xi_n)^2/(\Delta_{nl}\widehat\Delta_{nm})^2 \right| \F_n \right] \le \alpha_1$ 
-and $\E \left(\left\| \Upsilon(\overline H_n) \right\|^2 \mid \F_n\right) \le \alpha_1$. 
+$\E \left[ |f(\theta_n - \delta_n \Delta_n - \delta_n \widehat \Delta_n)/ (\Delta_{nl}\widehat\Delta_{nm})|^{2+\alpha_9} \mid \F_n\right]$,
+ $\E\left[\left. (\xi_n^+ + \xi_n^- - 2 \xi_n)^2/(\Delta_{nl}\widehat\Delta_{nm})^2 \right| \F_n \right] \le \alpha_8$ 
+and $\E \left(\left\| \Upsilon(\overline H_n) \right\|^2 \mid \F_n\right) \le \alpha_8$. 
 \item  $\delta_n = \frac{\delta_0}{(n+1)^{\varsigma}}$, where $\delta_0 > 0$ and $0 < \varsigma \le 1/8$.
@@
 \begin{pvn}
-\item  $f$ is quadratic and $\nabla^2 f(x^*) > 0$. 
+\item  $f$ is quadratic and $\nabla^2 f(\theta^*) > 0$. 
 \item The operator $\Upsilon$ is chosen such that $\E(\parallel \Upsilon(\overline H_n) - \overline H_n\parallel^2) = o(e^{-2wn^{1-r}/(1-r)})$ and \\$\parallel \Upsilon(H) - H \parallel^2 / (1+\parallel H \parallel^2)$ is uniformly bounded.
@@
 Assume (A23), (A29), (A30) and (A31) and also that the setting is noise-free. 
-Let $b_n = b_0/n^r$, $n=1,2,\ldots,k$, where $1/2 < r < 1$ and $0 < b_0 \leq 1$. For notational simplicity, let $H^*=\nabla^2 f(x^*)$. Letting $\Lambda_k = \overline H_k - H^*$, we have 
+Let $b_n = b_0/n^r$, $n=1,2,\ldots,k$, where $1/2 < r < 1$ and $0 < b_0 \leq 1$. For notational simplicity, let $H^*=\nabla^2 f(\theta^*)$. Letting $\Lambda_k = \overline H_k - H^*$, we have 
 \begin{align}
@@
 \noindent\textbf{Step 1: MSE convergence of $\overline H_n$} \\
-This part in exactly the same manner as part (i) in Theorem 3 of \cite{spall-jacobian}.\\
+This part of the proof follows along similar lines as  part (i) in Theorem 3 of \cite{spall-jacobian}.\\
 By rewriting the \eqref{eq:2spsa-3-ih} in the following form
@@
 The above equation can be thought as a stochastic approximation analogue of \eqref{eq:2spsa-3-ih}.
-The above stochastic approximation algorithm aimed at finding the roots of the equation $H-H^*=0$.
+The above stochastic approximation algorithm is aimed at finding the roots of the equation $H-H^*=0$.
 In \eqref{eq:hhatexpanded}, $\E (\widehat \Psi_n - \Psi_n (H^*) | \F_n) = 0 \,\,\ a.s$.
-Given a Lyapunov function, one can  the conditions required to show  the mean square convergence of the \eqref{eq:hhatexpanded} from the \cite{nevel1973stochastic}, pp.92-94.
+Given a Lyapunov function, one can  get the conditions required to show  the mean square convergence of the \eqref{eq:hhatexpanded} from \cite{nevel1973stochastic}, pp.92-94.
 It is easy to see that $V(H) = \frac{1}{2} (H-H^*)\tr(H-H^*)$ indeed serves as Lyapunov function.
-From part (i) in Theorem 3 of \cite{spall-jacobian}, one can check that our algorithm indeed satisfy those conditions,  which in turn implies the $\E [\Lambda_n\tr \Lambda_n] \rightarrow 0 \,\,\,a.s.$ as $n\rightarrow \infty$.
+From part (i) in Theorem 3 of \cite{spall-jacobian}, one can check that our algorithm indeed satisfies those conditions,  which in turn implies that $\E [\Lambda_n\tr \Lambda_n] \rightarrow 0 \,\,\,a.s.$ as $n\rightarrow \infty$.
 
@@
  %%Let us  characterise $trace[\E(\Lambda_n\tr \Lambda_n)]$ using \eqref{lambda-exp}. From independence of $d_k$ along k, \eqref{lambda-exp} represent martingale difference sequence, leading to 
-Squaring on both sides and taking expectations, we obtain
+It follows from \eqref{lambda-exp} that 
  \begin{align}
@@
  \end{align}
-The equality above uses the fact that $\E(\Psi_k(\Lambda_{k-1}'))=0$, which get rid of the corresponding cross terms with the first term on RHS of \eqref{lambda-exp}. % %
+The equality above uses the fact that $\E(\Psi_k(\Lambda_{k-1}'))=0$, which gets rid of the corresponding cross terms with the first term on RHS of \eqref{lambda-exp}. % %
 We now characterize $\text{trace}[\E(\Lambda_n\tr \Lambda_n)]$ using \eqref{eq:exp-lambda2} as follows:
@@
 % \end{align}
-Where, as in the proof of Theorem 3 of \cite{spall-jacobian}, $\tau(\cdot)$ transforms the $\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')$ in a  fashion and then returns the trace of the resulting $N\times N$ matrix.
+whereas in the proof of Theorem 3 of \cite{spall-jacobian}, $\tau(\cdot)$ transforms the $\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')$ in a  fashion and then returns the trace of the resulting $N\times N$ matrix.
 As in the proof of Theorem 3 of \cite{spall-jacobian}, observe that $1-b_k = e^{-b_k}(1-O(b_k^2))$ and since $0 < b_k <1$, we have that the $O(b_k^2)$ term is strictly positive. 
@@
 The idea here is to show that the contradiction exists if RHS of \eqref{eq:spsaquad-bigo} exhibits  any slower rate of decay.
-Let us assume the following slower rate of decay for the LHS of the \eqref{eq:spsaquad-bigo}.
+Let us assume the following slower rate of decay for the LHS of \eqref{eq:spsaquad-bigo}.
 \begin{align}
-\text{trace}[\E (\Lambda_n \tr \Lambda_n)] = d(n) e^{-2b_0n^{1-r} / {1-r}}.
+\text{trace}[\E (\Lambda_n \tr \Lambda_n)] = d(n) e^{-2b_0n^{1-r} / {1-r}},
 \label{eq:spsaquad-bigocont}
 \end{align}
-Where $d(n) = o(e^{2b_0n^{1-r} / {1-r}})$.
-Now we need to show that there is contradiction to show the claim.
-By using the series of arguments used in part (iii) in Theorem 3 of \cite{spall-jacobian} and applying the condition (A31) we get the following
+where $d(n) = o(e^{2b_0n^{1-r} / {1-r}})$.
+Now we will to show that there is a contradiction.
+By using the series of arguments used in part (iii) in Theorem 3 of \cite{spall-jacobian} and applying the condition (A31) we get the following:
 \begin{align}
@@
 
-Now we can use the this result in RHS of the equation \eqref{eq:thmwts} to show the contradiction.
-We consider the two cases to show the contradiction.
-One is where $\frac{d(n)}{n^{2r}} = O(1)$ and another is $lim sup_{n \to \infty} \frac{d(n)}{n^{2r}} = \infty$.
+Now we can use the result in the RHS of equation \eqref{eq:thmwts} to show the contradiction.
+We consider two cases to show the contradiction.
+The first is where $\frac{d(n)}{n^{2r}} = O(1)$ and the other is $lim sup_{n \to \infty} \frac{d(n)}{n^{2r}} = \infty$.
 We show contradiction for the both cases separately.
-Let us consider the first case $\frac{d(n)}{n^{2r}} = O(1)$, then
+Let us consider the first case, i.e., $\frac{d(n)}{n^{2r}} = O(1)$, Then
 \begin{align}
 \text{RHS of \eqref{eq:thmwts}} &= O(1) e^{-2 \Gamma_{1n}} +  O(1) e^{-2 b_0 n^{1-r}/(1-r)} \sum_{i=1}^n \frac{d(i)}{i^{p}} \frac{b_0^2}{i^{2r-p}}\nonumber \\
-&= e^{-2b_0n^{1-r} / {1-r}} \left[ O(1) + O(1) \sum_{i =1}^n \frac{1}{i^{2r-p}}\right].
+&= e^{-2b_0n^{1-r} / {1-r}} \left[ O(1) + O(1) \sum_{i =1}^n \frac{1}{i^{2r-p}}\right],
 \label{eq:spsaquad-bigocont4}
@@
 Now we have $\frac{LHS(i_n)}{RHS(i_n)} \to 0$ because  $0 < \epsilon < 2r -1$.
-It is contradiction since $LHS(n) = RHS(n)$ for all $n$.
+This leads to a contradiction since $LHS(n) = RHS(n)$ for all $n$.
 
@@
 There exists a subsequence $\{i_1,\ldots,i_n\}$ such that $ \frac{d(i_n)}{i_n^{2r}} \ge \frac{d(i)}{i^{2r}}$ as $i \le i_n$.
-Then for some $C >0$ the RHS of \eqref{eq:thmwts} satisfies the following
+Then for some $C >0$ the RHS of \eqref{eq:thmwts} satisfies the following:
 \begin{align}
@@
 &\le C e^{-2b_0n^{1-r} / {1-r}} \left[ 1 + O(1) \sum_{i =1}^{i_n} \frac{d(i)}{i^{2r}}\right]\nonumber \\
-&\le C e^{-2b_0n^{1-r} / {1-r}} \left[ 1 + \frac{d(i_n)}{{i_n}^{2r}}\right].
+&\le C e^{-2b_0n^{1-r} / {1-r}} \left[ 1 + \frac{d(i_n)}{{i_n}^{2r}}\right],
 \label{eq:spsaquad-bigocont5}
 \end{align}
-Where $\text{RHS}(i_n)$ denotes, RHS after the algorithm is iterated for $i_n$ times.
+where $\text{RHS}(i_n)$ denotes, RHS after the algorithm is iterated for $i_n$ times.
 From the equation \eqref{eq:spsaquad-bigocont}, the  LHS of \eqref{eq:thmwts} after $i_n$ iterations is $d(i_n) e^{-2b_0 i_n^{1-r} / {1-r}}$.
-By substituting \eqref{eq:spsaquad-bigocont} and \eqref{eq:spsaquad-bigocont5}, in $\frac{RHS(i_n)}{LHS(i_n)}$ then $\frac{RHS(i_n)}{LHS(i_n)} \to 0$ as $n \to \infty$.
-It is contradiction since $LHS(n) = RHS(n)$ for all $n$.
+By substituting \eqref{eq:spsaquad-bigocont} and \eqref{eq:spsaquad-bigocont5}, in $\frac{RHS(i_n)}{LHS(i_n)}$, it turns out that $\frac{RHS(i_n)}{LHS(i_n)} \to 0$ as $n \to \infty$.
+This results in a contradiction since $LHS(n) = RHS(n)$ for all $n$.
 Hence the claim $\text{trace}[\E (\Lambda_n \tr \Lambda_n)] = O(e^{-2b_0n^{1-r} / {1-r}})$ follows.

### 2017-01-23 secorder-rdsa-ih.tex
sir corrections + synopsis + abstract

@@
 for all $|(\theta-\theta^*)_i| < \tau$ when $i \notin S$ and $|(\theta-\theta^*)_i| \ge \rho'(\tau,S)$ when $i\in S$.
-\item For some $\alpha_0, \alpha_1>0$ and for all $n$, $\E {\xi_n}^{2} \le \alpha_0$, $\E {\xi_n^{\pm}}^{2} \le \alpha_0$, $\E f(\theta_n)^{2} \le \alpha_1$,  $\E f(\theta_n\pm \delta_n d_n)^{2} \le \alpha_1$ and $\E \left(\left\| \Upsilon(\overline H_n) \right\|^2 \mid \F_n\right) \le \alpha_1$.
+\item For some $\alpha_5, \alpha_6>0$ and for all $n$, $\E {\xi_n}^{2} \le \alpha_5$, $\E {\xi_n^{\pm}}^{2} \le \alpha_5$, $\E f(\theta_n)^{2} \le \alpha_6$,  $\E f(\theta_n\pm \delta_n d_n)^{2} \le \alpha_6$ and $\E \left(\left\| \Upsilon(\overline H_n) \right\|^2 \mid \F_n\right) \le \alpha_6$.
 \item  $\delta_n = \frac{\delta_0}{(n+1)^{\varsigma}}$, where $\delta_0 > 0$ and $0 < \varsigma \le 1/8$.

### 2017-01-23 synopsis.tex
sir corrections + synopsis + abstract

--- 2017-01-23 abstract.tex
@@
 \let\labelindent\relax
-\usepackage{adjustbox}
+
 \usepackage{siunitx}
@@
 \usepackage{hyperref}
-
 \usepackage{xcolor}
 \usepackage{colortbl}
-\usepackage{enumitem}
+%\usepackage{enumitem}
 \usepackage{amsmath}
-
 \usepackage{bm}
@@
 \usepackage{multirow}
-
 \usepackage{makecell}
@@
 \usepackage{subcaption}
+
+\newcommand{\tabl}[2]{\begin{tabular}{#1} #2 \end{tabular}}
+\newcommand{\ith}[2][th]{$#2^{\text{#1}}$}
+
+%\newcommand{\cP}{\mathcal{P}}
+%\newcommand{\cS}{\mathcal{S}}
+%\newcommand{\R}{\mathbb{R}}
+%\newcommand{\E}{\mathbb{E}}
+%\renewcommand{\S}{\mathcal{S}}
+%\newcommand{\D}{\mathcal{D}}
+%\newcommand{\A}{\mathcal{A}}
+%\newcommand{\B}{\mathcal{B}}
+%\newcommand{\X}{\mathcal{X}}
+\newcommand{\F}{\mathcal{F}}
+%\newcommand{\G}{\mathcal{G}}
+\newcommand{\N}{\mathcal{N}}
+%\newcommand{\Z}{\mathcal{Z}}
+%\newcommand{\T}{\mathcal{T}}
+%\renewcommand{\O}{\mathcal{O}}
+%\renewcommand{\P}{\mathbb{P}}
+
+\newcommand{\tr}{^\mathsf{\scriptscriptstyle T}}
+\newcommand{\tpi}{\Pi}
+\newcommand{\tgamma}{\tilde{\gamma}}
+\newcommand{\ttheta}{\tilde{\theta}}
+\newcommand{\tz}{\tilde{z}}
+
+\renewcommand{\l}{\left\|}
+\renewcommand{\r}{\right\|}
+\newcommand{\vml}{\left\|}
+\newcommand{\vmr}{\right\| _{\Psi}}
+\newcommand{\ml}{\left\|}
+\newcommand{\mr}{\right\|_2}
+
+\newcommand{\e}{\exp}
 
@@
 \usepackage{enumerate}
-\usepackage{enumitem}
 \usepackage{cancel}
@@
 \usepackage{setspace}
-\title{\textbf{ABSTRACT}}
+\title{\textbf{SYNOPSIS}}
 \author{}
@@
 
-\vspace{1.5cm}
+\vspace{2cm}
+
+\begin{center}
+
+\textbf{\large{Introduction}}
+\end{center}
 
@@
 
+\vspace{1cm}
+
+\begin{center}
+
+\textbf{\large{Outline and Contributions of the Thesis}}
+\end{center}
+
+\textit{Chapter 1: Introduction} \\
+
+This chapter introduces the problems we address in the thesis.
+\vspace{0.5cm}
+
+\textit{Chapter 2: Second-order RDSA with improved hessian estimation (2RDSA-IH) } \\
+
+Our work in this chapter  is centred on improving the 2RDSA scheme of \cite{prashanth2015rdsa} by 
+\begin{enumerate}
+\item reducing the error in the Hessian estimate through a feedback term; and
+\item optimizing the step-sizes used in the Hessian estimation recursion, again with the objective of improving the quality of the Hessian estimate.
+\end{enumerate}
+
+Items (1) and (2) are inspired by the corresponding improvements to the Hessian estimation recursion in the enhanced 2SPSA from \cite{spall-jacobian}.
+We shall refer to the latter algorithm as 2SPSA-IH.
+While item (2) above is a relatively straightforward migration to the 2RDSA setting, item (1) is a non-trivial contribution, primarily because the Hessian estimate in 2RDSA is entirely different from that in 2SPSA and the feedback term that we incorporate in 2RDSA to improve the Hessian estimate neither correlates with that in 2SPSA nor follows from the analysis in \cite{spall-jacobian}.
+The advantage with 2RDSA scheme along with proposed improvement to Hessian estimation (henceforth referred to as 2RDSA-IH) is that it requires only 75\% of the simulation cost per-iteration for 2SPSA-IH.
+
+We establish that the proposed improvements to Hessian estimation in 2RDSA are such  that the resulting 2RDSA-IH algorithm is provably convergent, in particular, the Hessian estimate $\overline H_n$ of 2RDSA-IH converges almost surely to the true Hessian.
+%Moreover, we show, for the special case of a quadratic objective, that	2RDSA-IH results in a convergence rate that is on par with the corresponding rate for 2SPSA with Hessian estimation improvements (2SPSA-IH) \cite{spall-jacobian}.
+Further, we show empirically that 2RDSA-IH outperforms both 2SPSA-IH of \cite{spall-jacobian} and regular 2RDSA of \cite{prashanth2015rdsa}.
+Our contribution is important because 2RDSA-IH, like 2RDSA, has lower simulation cost per iteration than 2SPSA and unlike 2RDSA, has an improved Hessian estimation scheme.\\
+
+\textit{Chapter 3: Generalised RDSA}\\
+
+Our work in this chapter is centred on generalising the 2RDSA scheme of \cite{prashanth2015rdsa}.
+In \cite{prashanth2015rdsa}, 2RDSA scheme involving only two perturbation distributions was proposed, those are uniform and asymmetric Bernoulli distributions.
+In this chapter, we have proposed gradient and Hessian estimation schemes which are independent of the perturbation distributions.
+However, perturbations have to satisfy the i.i.d, mean-zero assumptions.
+Apart from these common assumptions  one additional assumption, which is the difference between the squared second moment  and fourth moment should be non-zero for Hessian estimation is also required.
+This generalisation of distributions of random variables for perturbations is important because it enhances the scope of improving the performance of the 1RDSA and 2RDSA schemes by incorporating several other distributions.
+Some well known distribution that can be used for these schemes as a result of the generalization are Normal (0,1) and truncated Cauchy distributions.\\
+
+ \textit{Chapter 4: Second-order SPSA-3 with improved hessian estimation (2SPSA-3-IH)}\\
+
+Our work in this chapter is centred on improving the second order SPSA scheme with three simulations (referred to as 2SPSA-3 hereafter), which was proposed in  \cite{bhatnagar2015simultaneous}.
+These improvements arise as a result of incorporating the zero-mean feedback term in Hessian estimate and optimizing the step-sizes used in the Hessian estimate recursion.
+Though these ideas are same as the ideas presented in this section for 2RDSA scheme, the contribution is non-trivial because the derivation of feedback term and Hessian estimate is entirely different from 2SPSA, 2RDSA improvements.
+Moreover, we show, for the special case of a quadratic objective and when there is no noise, that 2SPSA-3 scheme along with proposed improvement to Hessian estimation (henceforth referred to as 2SPSA-3-IH) results in a convergence rate that is on par with the corresponding rate for 2SPSA with Hessian estimation improvements (2SPSA-IH) \cite{spall-jacobian}.
+The advantage with 2SPSA-3-IH is that it requires only 75\% of the simulation cost per-iteration for 2SPSA-IH.
+\\
+
+\textit{Chapter 5: Simulation experiments}\\
+
+Our work in this chapter is centred on testing the performance of the several algorithms proposed in earlier chapters along with the existing algorithms.
+For the empirical evaluations, we use the two loss functions.
+The first one is quadratic  loss  function and the other is Fourth-order loss function.
+We perform experiments for both the noisy and noise-less settings.
+We use normalized loss and normalized MSE (NMSE) as performance metrics for evaluating the algorithms.
+NMSE is the ratio $\l \theta_{n_\text{end}} - \theta^* \r^2 / \l \theta_0 - \theta^*\r^2$, while normalized loss is the ratio $f(\theta_{n_\text{end}})/f(\theta_0)$.
+Here $n_\text{end}$ denotes the iteration number when the algorithm stopped updating its parameter.
+From the results, we make the following observations:
+
+ \textit{\textbf{Observation 1:} Schemes with improved Hessian estimation perform better than their respective regular schemes.}
+
+\textit{\textbf{Observation 2:} 2RDSA-IH variants outperform both 2SPSA and 2SPSA-IH.}
+
+\textit{\textbf{Observation 3:} 2SPSA-3-IH variants outperform both 2SPSA and 2SPSA-IH.}
+
+\textit{\textbf{Observation 4:} 2RDSA-IH-AsymBer perform the best overall.}\\
+
+ \textit{Chapter 6: Conclusions and Future work}\\
+
+Finally, we summarize our contributions of the thesis and present a few interesting directions for future work.
+
 \bibliographystyle{plain}

### 2017-01-26 synopsis.tex
final changes

@@
 
-Our work in this chapter is centred on testing the performance of the several algorithms proposed in earlier chapters along with the existing algorithms.
-For the empirical evaluations, we use the two loss functions.
-The first one is quadratic  loss  function and the other is Fourth-order loss function.
+Our work in this chapter is centred on testing the performance of the algorithms we proposed in the earlier chapters along with the existing algorithms in the literature.
+For the empirical evaluations, we use two loss functions.
+The first one is quadratic  loss  function and the other is a fourth-order loss function.
 We perform experiments for both the noisy and noise-less settings.
-We use normalized loss and normalized MSE (NMSE) as performance metrics for evaluating the algorithms.
+We use normalized loss and normalized mean squared error (NMSE) as performance metrics for evaluating the algorithms.
 NMSE is the ratio $\l \theta_{n_\text{end}} - \theta^* \r^2 / \l \theta_0 - \theta^*\r^2$, while normalized loss is the ratio $f(\theta_{n_\text{end}})/f(\theta_0)$.
-Here $n_\text{end}$ denotes the iteration number when the algorithm stopped updating its parameter.
+Here $n_\text{end}$ denotes the iteration number when the algorithm stop updating its parameter.
 From the results, we make the following observations:

### 2017-04-27 intro.tex
final changes
review corrections

@@
 Optimization via simulation is a fast developing area for both researchers and practioners.
-As mentioned before, for stochastic optimization problems, simulation is a viable approach compared.
+As mentioned before, for stochastic optimization problems, simulation is a viable approach.
 Multiple simulation replications must be performed in order to get a good estimate of $E_{\xi}[h(\theta_i,\xi)]$, which is the function value at parameter $\theta_i$.
@@
 Stochastic Newton methods are often more accurate than simple gradient search schemes.
+A key early work in stochastic Newton methods  is given in \cite{anbar1978stochastic}.
+A sleeker `ODE' based convergence argument might be facilitated by \cite{smale1976convergent}.
 %which are sensitive to the choice of the constant $a_0$ in the canonical step-size, $a_n= a_0/n$.

### 2017-05-01 conc.tex
review corrections
thesis final corrections

@@
 It would be interesting to extend those methods to the stochastic optimization setting and prove that they exhibit  faster convergence than first-order methods and result in lower computational cost.
+\item Newton scheme is not popular due to expensive matrix inversion at each parameter update step even in deterministic scenario.
+Quasi-Newton schemes given by the Broyden family may be used instead.
+As a future work, it would be interesting to develop stochastic approximation versions of the quasi-Newton schemes.
 \item In the methods described in this thesis the perturbations used are random variables.

### 2017-05-01 intro.tex
thesis final corrections

@@
 Stochastic Newton methods are often more accurate than simple gradient search schemes.
-A key early work in stochastic Newton methods  is given in \cite{anbar1978stochastic}.
-A sleeker `ODE' based convergence argument might be facilitated by \cite{smale1976convergent}.
+A key early work in stochastic Newton methods  is given in \cite{anbar1978stochastic} and a sleek `ODE' based convergence argument is provided in \cite{smale1976convergent}.
 %which are sensitive to the choice of the constant $a_0$ in the canonical step-size, $a_n= a_0/n$.

### 2017-05-01 response.tex
response to reviwer comments

%\documentclass[oneside,8pt,onecolumn,openright]{IIScthesisPSnPDF}
\documentclass[12pt]{article}

\usepackage{paralist}
\let\labelindent\relax

\usepackage{siunitx}
\usepackage{booktabs}
\usepackage{url}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{colortbl}
%\usepackage{enumitem}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\pgfplotsset{compat=1.12}
\usepackage{todonotes}
\usepackage{nicefrac}
\usetikzlibrary{positioning,arrows,calc,patterns,backgrounds}
\usetikzlibrary{shapes}
\usetikzlibrary{fit}
\usepackage{multirow}
\usepackage{makecell}
\usepackage{caption}
\usepackage{subcaption}

\newcommand{\tabl}[2]{\begin{tabular}{#1} #2 \end{tabular}}
\newcommand{\ith}[2][th]{$#2^{\text{#1}}$}

%\newcommand{\cP}{\mathcal{P}}
%\newcommand{\cS}{\mathcal{S}}
%\newcommand{\R}{\mathbb{R}}
%\newcommand{\E}{\mathbb{E}}
%\renewcommand{\S}{\mathcal{S}}
%\newcommand{\D}{\mathcal{D}}
%\newcommand{\A}{\mathcal{A}}
%\newcommand{\B}{\mathcal{B}}
%\newcommand{\X}{\mathcal{X}}
\newcommand{\F}{\mathcal{F}}
%\newcommand{\G}{\mathcal{G}}
\newcommand{\N}{\mathcal{N}}
%\newcommand{\Z}{\mathcal{Z}}
%\newcommand{\T}{\mathcal{T}}
%\renewcommand{\O}{\mathcal{O}}
%\renewcommand{\P}{\mathbb{P}}

\newcommand{\tr}{^\mathsf{\scriptscriptstyle T}}
\newcommand{\tpi}{\Pi}
\newcommand{\tgamma}{\tilde{\gamma}}
\newcommand{\ttheta}{\tilde{\theta}}
\newcommand{\tz}{\tilde{z}}

\renewcommand{\l}{\left\|}
\renewcommand{\r}{\right\|}
\newcommand{\vml}{\left\|}
\newcommand{\vmr}{\right\| _{\Psi}}
\newcommand{\ml}{\left\|}
\newcommand{\mr}{\right\|_2}

\newcommand{\e}{\exp}

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{latexsym}
\usepackage{multicol}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{fancyhdr}
\usepackage[numbers]{natbib}
\usepackage{cleveref}
\usepackage{wrapfig}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{enumerate}
\usepackage{cancel}
\usepackage{mathrsfs}
% \usepackage{sfgame}
\usepackage{color}
\usepackage[compact]{titlesec}
\usepackage{mdwlist}
\usepackage{varwidth}
\usepackage[vertfit]{breakurl}
\usepackage{datetime}
\usepackage{pdfpages}
% \usepackage{biblatex}
\usepackage{setspace}
\title{\textbf{Response to Reviewer Comments}}
\author{}
\date{}
\begin{document}
%\Huge
%\begin{center}
%Panel of Experts
%\end{center}
\maketitle
\begin{tabular}{ll}
Name of the Candidate & \textbf{Danda Sai Koti Reddy}\\ 
SR Number &\textbf{04-04-00-10-21-14-1-11609} \\
Title of the Thesis & \textbf{Stochastic Newton methods with   } \\
 & \textbf{ enhanced Hessian estimation.}\\
 Research Supervisor  & \textbf{Prof. Shalabh Bhatnagar} \\
Degree Registered  &  \textbf{M.Sc(engg)} \\
Department  & \textbf{Computer Science and Automation} \\
Institute  & \textbf{Indian Institute of Science, Bangalore }
\end{tabular}

\vspace{2cm}

\section{Reviewer Suggestions}
The questions are presented in \textit{italicized} font and the answers in normal font. The response to the reviewer questions are as under.\\

\textit{\textbf{Suggestion (i): } A key early work in this direction merits mention: \\  \\Anbar, Dan. ``A stochastic Newton-Raphson method." Journal of Statistical Planning and Inference 2.2 (1978): 153-163.\\ \\ A sleeker `ODE' based convergent argument might be facilitated by:\\ \\  Smale, Steve. ``A convergent process of price adjustment and global Newton methods." Journal of Mathematical Economics 3.2 (1976): 107-120.\\ \\}
 \textbf{Response:}\\
We have mentioned the references given in suggestion in the section 1.5, titled ``Hessian estimation". Which is given as below \\ \\
\textbf{1.5 Hessian estimation} \\
Stochastic Newton methods can counter the ill-conditioning problem of the objective $f$ as they incorporate second-order information into the update iteration given by   
\begin{align}
\label{eq:newton}
\theta_{n+1} = \theta_n - a_n (\overline H_n)^{-1}\widehat\nabla f(\theta_n), 
\end{align}
where $a_n$ is the step-size that satisfies standard stochastic approximation conditions (see (A12) in Section 2.6), and $\widehat\nabla f(\theta_n)$ and $\overline H_n$ are estimates of the gradient and Hessian, respectively. Thus, \eqref{eq:newton} can be considered as  the stochastic version of the well-known Newton method for optimization. Stochastic Newton methods are often more accurate than simple gradient search schemes. A key early work in stochastic Newton methods  is given in \cite{anbar1978stochastic} and a sleek `ODE' based convergence argument is provided in \cite{smale1976convergent}.\\ \\
\textit{\textbf{Suggestion (ii):} I would also appreciate any comments on the following issue: Even in the deterministic scenario, Newton scheme is not popular due to the expensive matrix inversion, and quasi-Newton schemes given by the Broyden family are used instead. Why not develope stochastic approximation versions of the latter?  }\\ \\
 \textbf{Response:}\\
 We have mentioned this as an interesting future direction in chapter 6, titled ``Conclusions and Future work". Which is given below \\
 3. Newton scheme is not popular due to expensive matrix inversion at each parameter update step even in deterministic scenario. Quasi-Newton schemes given by the Broyden family may be used instead. As a future work, it would be interesting to develop stochastic approximation versions of the quasi-Newton schemes.

\bibliographystyle{plain}
\bibliography{refThesis}

\end{document}
